### 2015-11-10 1_finding_groups/finding_groups.Rmd
init

---
title: "Business Analytics 1 : Finding Groups Within the Data"
author: "Ming-Yu Liu"
output: 
  html_document: 
    number_sections: yes
    theme: united
    toc: yes
---

## Background Information

This module will go over examples of how identifying groups within data allows business to improve efficiency. Finding groups within our data allows us to strike the right balance between similarities and differences. The reason behind this is we want to treat similar cases with similar appoaches to benefit from the economics of scale. On the other hand, treating different scenarios with different ways that are more suitable to the current situations will improve your actions' effectiveness. The following documentation will cover two examples of where this notion can be use in the business world.

> Note that this documentation assumes you're familiar with R and hierarchical clustering, thus technical details are not explained.

## Example 1 : Supply Chain Management

Our concern as a supply chain manager is that we want to organize products' stock more efficiently. To be explicit, we don't really need to store lots of stocks for every single product as this increases inventory costs. On the other hand, we also wish to maintain the stock amount of each product at a suitable level, so that we will be able to deliver the product on time when demanded. Thus, the goal for this task is to identify groups of products' stocks that "behaves" the same and can be treated in a similar fashion.

The common approach to solve this issue is to analyze the question along two dimensions, where you have the average daily sales of your stock keeping unit, or so called "SKU" on one axis and the "volatility" of each SKU's average daily sales.

- **SKU** : Note that in the field of inventory SKU refers to a specific item stored to a specific "location". The key word here is "location". In a retail network, the same product, let's say TV can have multiple SKU. Examples of this is : 
    - There are multiple warehouses where this TV can be stored. This is the most typical case where the same TV will have one SKU per warehouse. 
    - Even items that are at the same warehouses can have multiple SKUs, because they are sold at different locations ( e.g. retail stores at City A, shoppinig malls at City B and so on ).
    - Also a product may have many variants based on attributes such as size, color or conditioning. This will also lead to multiple SKUs for the same given product.

- **volaility** : This can be measured by the [coefficient of variation](https://en.wikipedia.org/wiki/Coefficient_of_variation), which is simply the standard deviation of sales divided by the mean sales for each SKU.

Let's read in the dataset to get a feel of what it looks like, before going further.

```{r, message=FALSE}

# environment setting
library(dplyr)
library(caret)
library(ggplot2)
library(reshape2)
setwd("/Users/ethen/Business-Analytics/1_finding_groups")
rm( list = ls() )

data_sku <- read.csv("data/DATA_2.01_SKU.csv")
head(data_sku)

```

There are `r nrow(data_sku)` rows and 2 variables within this dataset, namely :    
- `ADS` average daily sales.  
- `CV` coefficient of variations.

As there is only two variables (columns), we can easily visualize it and look at their relationships.

```{r}

ggplot( data_sku, aes( CV, ADS ) ) + geom_point() + 
labs( title = "SKU Example", y = "Average Daily Sales", x = "Coefficient of Variation" ) + 
geom_hline( yintercept = 4, color = "blue" ) + 
geom_vline( xintercept = .2, color = "blue" )

```

As you'll notice from the plot, there are three distinct groups, separated by the horizontal and vertical line on the graph. In practice, however, there're only a few situations where you can define how many distinct groups are there in the data through straightforward visualization. Because, often times, you'll have more than two variables that you may consider for your clustering. Therefore, we'll now use the more rigorous approach, the hierarchical clustering algorithm to group our datasets.

One precaution, from the data (or from the visualization) you’ll notice that two variables are recorded with different units. Therefore when input variables differ by orders of magnitudes, performing some kind of normalization is often times required to make them comparable. Here we’ll apply the widely used z-score normalization using the built in `scale` function, where given a input variable you subtract every row element with its mean and then divide it by its standard deviation.

```{r}

# hierarchical clustering, with k (clustering number) = 3 
d <- dist( scale(data_sku), method = "euclidean" )
cluster <- hclust( d, method = "ward.D" )
( data_sku$groups  <- as.factor( cutree( cluster, k = 3 ) ) )

```

Now that we have obtain the grouping of the data, let's visualize the grouping to see confirm that the algorithm's result matches our intuition.

```{r}

ggplot( data_sku, aes( CV, ADS, color = groups ) ) + geom_point() + 
labs( title = "SKU Example", y = "Average Daily Sales", x = "Coefficient of Variation" )

```

After we've identify the grouping the next step is to transform these insights into actions, that is what should we do with the patterns we found.

- group 1 : SKUs with low sales and high variability. We'll call them "crickets" (giving groups names, so business audiences will be more likely to remember them), because not only are the sales for these products low, but can also jump unexpectedly. These stocks will be "made to order". 
    - **Make to Order** : This strategy wait until an order comes in before starting the production process. Since the sales are small, in any case, it's not really efficient to prepare production too long in advance. So we want to reduce the risks by producing the goods only if the order is made by the end of the chain.

- group 2 : SKUs with high sales and low variability. We'll call them "horses", because these products are strong and reliable. These stocks will be "made to stock". 
    - **Make to Stock** : This production strategy forecasts demand to determine how much stock should be produced. If demand for the product can be accurately forecasted, the MTS strategy can be an efficient choice. This will be reasonable for this case, because the sales for these products are expected to be high and the risk of the inaccurate forecast is low (coefficient of variation).

- group 3 : SKUs with high sales and high variability. We'll call them "white bulls", because despite of its possible high sales, it is difficult to control. These SKUs will be treated on a case by case basis. Things may be a bit difficult to anticipate, but their returns may be high.

This draws an end to the first example.

## Example 2 : Humance Resources

For the next example, suppose you're an HR manager of a big consulting company, and that you're concerned by the high number of employees leaving the firm. In practice, you cannot follow-up with each one of them too often as that would be very time consuming, hence, this is also a case where you have to rely on an HR analytic solution to understand what are the most frequent situations explaining why an employee decides to leave and discover the actions we can take in order to retain them.

Now let's say your company has collected a bunch of data from employees that have already left.

```{r}

data_hr <- read.csv("data/DATA_2.02_HR.csv")
str(data_hr)

```

This dataset contains `r nrow(data_hr)` observations and 6 variables, each representing :

- `S` The satisfaction level on a scale of 0 to 1.   
- `LPE` Last project evaluation by a client on a scale of 0 to 1.   
- `NP` Represents the number of projects worked on by employee in the last 12 month.  
- `ANH` Average number of hours worked in the last 12 month for that employee.  
- `TIC` The amount of time the emplyee spent in the company, measured in years.  
- `Newborn` This variable will take the value 1 if the employee had a newborn within the last 12 month and 0 otherwise.

There're two lessons we can learn from this example.

1. This is the kind of data where we can't use visualization to determine the grouping number before applying the clustering algorithm like we did in the first example, because now we have the total of 6 variables to consider.

2. Before applying the clustering algorithm, we should determine which variables to include. If the dataset contains no missing values, as a data scientist, we're often used to doing this by looking at the correlations between the variables ( of course there're more complex ways to do this ). Then suppose the correlations between variables A and B are high, then we can choose drop one of them, because after obtaining the results we can simply use the variables we've retained to properly assume the ones we dropped. We can easily achieve this step by using the `findCorrelation` function from the caret package. Where, given a correlation matrix , this function will print out highly correlated attributes.

```{r}
# scale the data (you don't really need to do this when comparing correlations)
# this is used for latter plotting and applying clustering
hr_scaled <- data.frame( scale(data_hr) )

# find variables that have correlations higher than 0.8 and print out variable names
findCorrelation( cor(hr_scaled), cutoff = 0.8, names = TRUE )

```

In this case, the function suggests that we remove the `NP` attribute as it correlates highly with the `LPE` attribute. But here, it might be a bad idea to use the correlation method to determine which variables to exclude as neither of them are not normally distributed ( the default method of the cor function that does the correlation calculation is Pearson correlation and this method is sensitive to non-normality in the variables ).

```{r, message=FALSE}

ggplot( melt( hr_scaled[ , 2:3 ] ), aes( value, fill = variable ) ) + 
geom_histogram( alpha = .4, position = "identity" )

```

So next, we'll look at this feature selection approach from the business point of view. In this case, we should note that the `S` variable, level of satisfaction is actually a consequence of everything else. We cannot really act on it directly, it has to be seen as a consequence and not as the driver of managerial impact.

Knowing this by intuition, we'll can remove this variable before conducting the clustering. This time we'll choose the clustering number to be 4, or you can call it segmentation number if you wish. This number is decided by plotting the dendogram of the clustering result. Not shown here in the report since it's a bit awful looking, which is a common problem of dendogram when it comes to bigger dataset. We'll also calculate the median of each variable for each of the four groups to discover patterns and count the proportion size of each group. Note that we will also be including the `S` variable for the aggregation.

```{r}

d <- dist( hr_scaled[ , -1 ], method = "euclidean" ) 
cluster <- hclust( d, method = "ward.D" )
# dendogram
# plot(cluster)

data_hr$groups <- cutree( cluster, k = 4 )

# median aggregation 
hr_agg <- aggregate( .~ groups, data = data_hr, FUN = median )

# count the cluster proportion size 
tl <- table(data_hr$groups)
hr_agg$proportion <- tl / sum(tl)

# order by proportion 
mutate( hr_agg, proportion = tl / sum(tl) ) %>%
arrange( desc(proportion) )

```

From the result of this table, you can see that :

- group 1 : We can see from the low quantity of the four variables `LPE` last project evaluation by a client, `NP` number of projects, `ANH` average number of working hours and `TIC` total time spent in the company. That these are probably the "low performance" ones and they are still quite junior in the company, thus it may not be our top priority to try and retain them.

- group 3 : This is the second row, don't get confused. Notice the consideraly high `LPE`, `NP`, `ANH`, `TIC` and a low satisfaction level `S` suggests that these people are the ones that are over-utilized. Simply that they are doing a good job, but are "burned out". Given this result, as a HR manager these kind of people should be your first priority. Next time you see that a high performing employee is working too much, you should be able to anticipate this situation and helped those employees to take a step back proactively.

- group 2 : This group is a bit tricky. You'll notice that similar to group 3, where they have a high `LPE`, `NP`, `ANH`, `TIC` levels, from the high rating of the `S` variable, it seems that they're still quite satisfied with their work. Well, what can do to retain these "high-performing" employees that are happy but want to leave nonetheless. For these people we can assume that we're given them too little cash or the projects that were assigned to them are simply too easy. Then we can try to give them a raise, a promotion or more challenging projects.

- group 4 : For this final group, this is assume to be a "miscellaneous" segment. As it doesn't have any obvious characteristic apart from the `Newborn` column indicating that they the employee had a newborn within the last 12 month and a straightforward way to retain these employees is to provide parental leave or on-site child care. But notice from the `proportion` column that since it's a relatively small group, given the limited resources we have, it's not really a priority either.

When presenting the upper result table, one visualization you can use is the radar chart. However since the current version of the ggplot2 library still does not support actual radar chart with a simple function call (as least not to my knowledge) and each variable of the data is recorded with different scales, therefore, the visualization after using the max-min normalization method may look a bit hacky, as with this normalization, there will be values of 0 ( A different kind of normalization is used here because the scale function produces negative numbers and that does not sound very intuitive for radar charts ).

```{r, fig.height=8, fig.width=12  }

Normalize <- function(x) ( x - min(x) ) / ( max(x) - min(x) )

# exclude the proportion, groups and store the groups
hr_agg_2 <- hr_agg[ , c( -1, -8 ) ]
groups <- hr_agg$groups

# transform to long format for plotting
hr_agg_melt <- apply( hr_agg_2, 2, Normalize ) %>%
			   data.frame() %>%
			   mutate( groups = groups ) %>%
               melt( id.vars = "groups" )

# radar chart
ggplot( hr_agg_melt, aes( x = variable  ) ) + 
geom_bar( aes( y = value, fill = groups ), stat = "identity" ) + 
coord_polar() + 
facet_wrap( ~ groups ) +
ggtitle( "HR" ) + 
theme( legend.position = "none",  
	   axis.text.y = element_blank(), 
	   axis.text.x = element_text( color = "black" ), 
	   axis.ticks  = element_blank(),
	   axis.title  = element_blank() )

```

End of the second example.

## Takeaways

1. There are different perspective (intuition, business sense, technical approaches) as to which variable should have or should not have been included in the dataset prior to running the clustering method. As clustering algorithms are so categorized under unsupervised algorithms, there're really no "correct" ouput, so choose your variables or features wisely ! Suppose that for some reason you think that it's nonsense to exclude the `S` variable : the satisfaction level, like we did in this documentation, you should add it back and run the clustering, though you'll obtain similar grouping results for this dataset, and when you realize that whether you include a variable or not does not affect the result, you should exclude it as this lessen the calculation loading for the algorithm and your computer.

2. After you obtain the results, giving visualizations and naming your segmentations in a self explanatory and meaningful way can make your results more convincing. We've heard like a billion times already that a picture is worth thousands of words. And more importantly, you should always ask yourself, given the results what are you going to do with it, how can this result help my business goals ?

3. Following up on the second point about interpreting the result. It may be the case that even if we've identify different groups, but they'll be treated similarly. Also there will be times when we can't transform every results into immediate actions, times where the events may be exogenous and can't be explained. If this group occurs with a high proportion, we may want to collect additional data to try to understand this specific case.

4. For clustering methods, the number of segmentation is a parameter that has to be user-specified before running the algorithm. In this module this number is chosen mostly by common sense and business criteria. If you're interested in a more technical way of choosing the suitable number and evaluate the stability of your clustering result, you can refer to another documentation [here](http://ethen8181.github.io/machine-learning/clustering/clustering.html).

This wraps up our example of where clustering algorithms can be used to improve your businesses efficiency. You can find all the code to this documentation [here](https://github.com/ethen8181/Business-Analytics/blob/master/1_finding_groups/finding_groups.R).

## R Session Information

```{r}

sessionInfo()

```

### 2015-11-10 1_finding_groups/finding_groups.R

@@ -1,0 +1,114 @@
+#### finding similar groups
+
+# environment setting
+library(caret)
+library(dplyr)
+library(ggplot2)
+library(reshape2)
+
+setwd("/Users/ethen/Business-Analytics/1_finding_groups/data")
+rm( list = ls() )
+
+# -- [example 1 : supply chain] -----------------------------------------------------
+data_sku <- read.csv("DATA_2.01_SKU.csv")
+head(data_sku)
+
+# figure 1
+ggplot( data_sku, aes( CV, ADS ) ) + 
+geom_point() + 
+labs( title = "SKU Example", y = "Average Daily Sales", x = "Coefficient of Variation" ) + 
+geom_hline( yintercept = 4, color = "blue" ) + 
+geom_vline( xintercept = .2, color = "blue" )
+
+# hierarchical clustering 
+d <- dist( scale(data_sku), method = "euclidean" )
+cluster <- hclust( d, method = "ward.D" )
+data_sku$groups  <- as.factor( cutree( cluster, k = 3 ) )
+
+# figure 2
+ggplot( data_sku, aes( CV, ADS, color = groups ) ) + 
+geom_point() + 
+labs( title = "SKU Example", y = "Average Daily Sales", x = "Coefficient of Variation" )
+
+
+# -- [example 2 : HR]---------------------------------------------------------------
+data_hr <- read.csv("DATA_2.02_HR.csv")
+str(data_hr)
+
+# correlation, method 1
+# convert the correlation of oneself to 0
+# look that which index is highly correlated, or i.e. often appears together
+M <- abs( cor(data_hr) )
+diag(M) <- 0 
+which( M > 0.8, arr.ind = TRUE )
+
+# correlation, method 2
+hr_scaled <- data.frame( scale(data_hr) )
+findCorrelation( cor(hr_scaled), cutoff = 0.8, names = TRUE )
+
+
+# normality plot 
+ggplot( melt( hr_scaled[ , 2:3 ] ), aes( value, fill = variable ) ) + 
+geom_histogram( alpha = .4, position = "identity" )
+
+
+# clustering 
+d <- dist( hr_scaled[ , -1 ], method = "euclidean" ) 
+cluster <- hclust( d, method = "ward.D" )
+plot(cluster)
+data_hr$groups <- cutree( cluster, k = 4 )
+
+# aggregation 
+hr_agg <- aggregate( .~ groups, data = data_hr, FUN = median )
+
+# proportion size
+tl <- table(data_hr$groups)
[…]

### 2015-12-02 1_finding_groups/finding_groups.Rmd
half done

@@
 author: "Ming-Yu Liu"
+date: "November 11, 2015"
 output: 
@@
 
-Let's read in the dataset to get a feel of what it looks like, before going further.
+Let's read in the dataset to get a feel of what it looks like, before going further. Oh, and the link to the datasets used in this documentation is located at the bottom of the report.
 
@@
 - group 3 : SKUs with high sales and high variability.
-We'll call them "white bulls", because despite of its possible high sales, it is difficult to control.
+We'll call them "wild bulls", because despite of its possible high sales, it is difficult to control.
 These SKUs will be treated on a case by case basis.
@@
 When presenting the upper result table, one visualization you can use is the radar chart.
-However since the current version of the ggplot2 library still does not support actual radar chart with a simple function call (as least not to my knowledge) and each variable of the data is recorded with different scales, therefore, the visualization after using the max-min normalization method may look a bit hacky, as with this normalization, there will be values of 0 ( A different kind of normalization is used here because the scale function produces negative numbers and that does not sound very intuitive for radar charts ).
+However since the current version of the ggplot2 library still does not support actual radar chart with a simple function call (as least not to my knowledge) and each variable of the data is recorded with different scales.
+Therefore, the visualization after using the max-min normalization method on the recorded score may look a bit hacky ( the `scale` function produces negative numbers and that does not sound very intuitive for radar charts, thus a different kind of normalization is used here ).
+As with this normalization, there will be values of 0.
 
@@
 This wraps up our example of where clustering algorithms can be used to improve your businesses efficiency.
-You can find all the code to this documentation [here](https://github.com/ethen8181/Business-Analytics/blob/master/1_finding_groups/finding_groups.R).
+You can find all the code (finding_groups.R) and datasets (in the data folder) used in this documentation [here](https://github.com/ethen8181/Business-Analytics/tree/master/1_finding_groups).
+
+Any comments or suggestions is appreciated. That includes bugs or typos report. You can file these new issues [here](https://github.com/ethen8181/Business-Analytics/issues).
 

### 2015-12-02 2_bike_sharing/bike_sharing.Rmd
half done

---
title: "Kaggle Competition : Bike Sharing"
author: "Ming-Yu Liu"
date: "December 1, 2015"
output: 
  html_document: 
    highlight: haddock
    number_sections: yes
    theme: cerulean
    toc: yes
---

> Still far away from the leaderboard

## Background

This documentation is for the bike sharing Kaggle competition. The overall task is to use a hourly bike rental data spanning two years to train a machine learning algorithm and predict the total count of bikes rented during each hour covered by the test set. More detailed information, datasets and variables explanations can be found at this [link](https://www.kaggle.com/c/bike-sharing-demand).

We'll load in the training dataset and give a brief explanations of the variables that may not seem so intuitive.

```{r, message=FALSE, warning=FALSE}

library(dplyr)
library(tidyr)
library(caret)
library(rpart)
library(party)
library(scales)
library(ggplot2)
library(gridExtra)
library(lubridate)
library(data.table)
setwd("/Users/ethen/Business-Analytics/2_bike_sharing/data")

data <- fread("train.csv")
data

```

Dependent variables :

- `registered` Number of registered user.
- `casual` Number of non-registered user.
- `count` Number of total rentals (registered + casual).

Independent variables :

- `temp` Temperature in Celsius.
- `atemp` "feels like" temperature in Celsius.
- `weather` Has four different values ranging from 1 to 4. 1 : Clear, 2 : Cloudy, 3 : Little rain, 4 : Heavy rain.
- `workingday` 1 indicates that the day is neither a weekend nor holiday and 0 otherwise, same goes for `holiday`.

## Exploratory Analysis

Before diving straight into machine learning algorithms, we'll perform some exploratory analysis to get to know the distribution of our data better. This allows us to :

- Check for missing values, and outliers, so that we can know whether they'll be needing transformations or included (excluded) for the modeling. 
- Discover the relationships between target (dependent) variables or predictors (independent), which can be used for feature selection and engineering.
- In our mind, there might be a lot of straightforward factors that may influence the demand of bikes, we can seize this chance to confirm if the patterns discovered in the data meets our original hypothesis.

Let's start with a quick check to see if there's any NA value (missing value) for each column.

```{r}

sapply( data, function(y) sum( is.na(y) ) )

```

No missing values, good! Next we'll plot a histogram for all the numeric independent variables to visualize and understand its distribution.

```{r, message = FALSE, warning = FALSE, fig.height = 7, fig.width = 9 }

independent <- c( "casual", "registered", "count" )
columns <- which( !names(data) %in% independent & sapply( data, is.numeric ) )

histogram <- lapply( names(columns), function(column)
{
	ggplot( data[ , column, with = FALSE ], aes_string( column ) ) + 
	geom_histogram()
})
do.call( grid.arrange, histogram )

```

From the visualization, we can see that :

- `season` Has four categories ranging from 1 to 4 that are equally distributed.
- `holiday` Consists mainly of 0s tells you most days are not considered as holidays.
- `weather` Shows that the weather for most of the days are clear (1).
- `workday` consists primarily of 1s tells you most days are neither weekends or holidays.
- Four variables including `temp`, `atemp`, `humidity` and `windspeed` are continuous variables and their values are not heavily skewed.

Third, we'll check the correlation between the continuous variables and the independent variables.

```{r}

cor( select( data, casual, registered, temp:windspeed ) ) %>%
findCorrelation( cutoff = .85, name = TRUE )

```

The result indicates that `atemp` is highly correlated with one of the other variables ( above the cutoff threshold .85 ), we can consider removing one of them from the model. You can check the correlation matrix, it'll show you that it is actually highly correlated with the `temp` variable.

Now that we've had the glimpse at our data, we'll start checking some hypotheses that comes into our mind. For the following hypotheses, we'll simualtaneously check to see if there's a difference in the patterns for registered users and non-registered users.

**1. Hourly trend: The demand for bikes should be higher for timespans that people get to and off work, shouldn't it?**

```{r, fig.height = 6, fig.width = 10 }

# convert datetime column to date type
data[ , datetime := ymd_hms(datetime) ]

# extract the datetime; and casual, registered column 
subdata1 <- select( data, datetime, casual, registered ) %>%
		    tbl_dt() %>%
		    mutate( datetime = hour(datetime) ) %>%
		    gather( "user", "count", -1 )

# box plot for registered and unregisted users
ggplot( subdata1, aes( datetime, count, group = datetime, fill = user ) ) + 
geom_boxplot() + 
facet_wrap( ~ user ) +
scale_fill_discrete( limit = c( "registered", "casual" ) ) + 
guides( fill = FALSE ) + 
ggtitle( "2011-2012 User's Hourly Bike Demand" )

```

To be clear, the "dots" at the end of the boxplot represent outliers, which is determined by checking if a data point is: less than Q1 - 1.5 * IQR or greater than Q3 + 1.5 * IQR, where IQR stands for interquartile range = Q3 - Q1. And the line goes to the first data point before the the cut-off. The fact that there're so many outliers shown in this boxplot infers that the `count` variable may requires some transformation (e.g. log transformation).

From the boxplot, clearly the *registered* users' ( boxes in red ) behavior can be sloppily categorized into three groups

- High: 7-9 and 17-19 hours.
- Average: 10-16 hours.
- Low: 0-6 and 20-24 hours.

**2. Daily trend: Is the demand for bikes higher on weekdays or weekends?**

```{r}

subdata2 <- select( data, datetime, casual, registered ) %>%
		    tbl_dt() %>%
		    mutate( datetime = lubridate::wday( datetime, label = TRUE ) ) %>%
		    gather( "user", "count", -1 ) %>%
		    group_by( datetime, user ) %>%
			summarise( count = sum(count) )

ggplot( subdata2, aes( datetime, count, group = user, color = user ) ) + 
geom_line( size = 1 ) + geom_point( size = 3 ) + 
scale_y_continuous( label = comma ) +
scale_colour_discrete( limit = c( "registered", "casual" ) ) + 
ggtitle( "2011-2012 User's Daily Bike Demand" )

```

The plot tells us that registered users' use bikes more often on weekdays, while non-registered users' behaviors are completely the opposite. Their demand increases over the weekend.

**3. Trend of bike demand over year: Is the market size of bike sharing increasing or decreasing?**

```{r, message=FALSE, warning=FALSE}

subdata3 <- select( data, datetime, casual, registered ) %>%
			tbl_dt() %>%
		    mutate( datetime = year(datetime) ) %>%
			gather( "user", "count", -1 ) %>% 
			group_by( datetime, user ) %>%
			summarise( count = sum(count) )

ggplot( subdata3, aes( datetime, count, fill = user ) ) + 
geom_bar( stat = "identity", position = "dodge", width = .6 ) + 
geom_text( aes( label = count ), position = position_dodge(.6), vjust = -.6, size = 3 ) +
scale_x_continuous( breaks = unique(subdata3$datetime) ) + 
scale_y_continuous( label = comma ) +
scale_fill_discrete( limit = c( "registered", "casual" ) ) +
ggtitle( "2011-2012 Bike Demand of Users" )

```

From the year 2011 to 2012, the demand for bike sharing is steadily increasing, especially for registered users.

**Section Conclusion:**

From the exploratory analysis section we can already tell that bike using behaviors of the registered and non-registered users are totally different.

## Feature Engineering

To be specific, here's a list of what we'll be doing in the feature engineering section:

1. We know from the exploratory analysis section that bike demands will increase over time, and that Split the `datetime` variable into four variables. `hour`, `weekday`, `year` and `month`.
2. Combine `holiday` and `workingday` into one new variable indicating whether the day its a holiday, weekend or workday.
4. Log transformation of dependent variables, `casual` and `registered`.
5. Convert discrete variables into factors.

**Step 1:**

- Add the hour, weekday, year, month column. 
- Split the original data into train / test set of 80 / 20 percent and remove the original dataset. This way we'll train the model using the training set and evaluate performance on test set. 
- Define the independent variables that will be used in the final model, things will be more clear after looking further down the documentation.

```{r}

data[ , c( "hour", "weekday", "year", "month" ) := list( hour(datetime), 
												   		 lubridate::wday(datetime),
												   		 year(datetime),
												   		 month(datetime) ) ]
# split 
set.seed(1234) 
train_index <- createDataPartition( data$count, p = .8, list = FALSE )
data_train  <- data[ train_index, ]
data_test   <- data[ -train_index, ]
rm(data)

# list of the final independent variables 
variables <- c( "weather", "temp", "humidity", "windspeed", 
				"type", "year", "month", "hour", "weekday" )

```

**Step 2 ~ 4:**

As for the rest of the steps, we'll first print out the [`ExtractFeatures`][ExtractFeatures] function that does the feature engineering for our dataset and explain along the way.

```{r}

ExtractFeatures <- function( data, data.type )
{

	# 2. convert holiday and workingday into one variable
	#    indicating whether its a holiday, weekend and workday
	data$type <- with( data,

		ifelse( holiday == 1, "holiday",
		ifelse( holiday == 0 & workingday == 0, "weekend", "workday" ) ) 
	)

	# 3. keep the needed variables, and convert discrete variables into factors
	# if it is the original training data, log transformation of dependent variables 
	if( data.type == "train" )
	{
		columns <- which( colnames(data) %in% c( "casual", "registered", variables ) )
		data <- select( data, columns )

		# log transformation
		data <- transform( data, casual     = log( casual + 1 ),
					   		 	 registered = log( registered + 1 ) )

	}else # "test"
	{
		columns <- which( colnames(data) %in% variables )
		data <- select( data, columns )	
	}

	# 4. convert to factor 
	factor_col <- which( colnames(data) %in% 
						 c( "weather", "hour", "year", "month", "weekday", "type" ) )

	data <- modifyList( data, lapply( select( data, factor_col ), as.factor ) )

	return(data)
}

count <- data_test$count
data_train <- ExtractFeatures( data = data_train, data.type = "train" )
data_test  <- ExtractFeatures( data = data_test , data.type = "train" )

```

## Model Training

```{r, cache=TRUE}

# 10-fold cross validation 
control <- trainControl( method = "cv", number = 10 )

# evaluation measure 
RMSLE <- function( y, pred )
{
	sqrt( sum( ( log( y + 1) - log( pred + 1 ) )^2 ) / length(y) )
}

# formula 
formula_casual 	  <- as.formula( paste( "casual ~ ", paste( variables, collapse = " + " ) ) )
formula_registerd <- as.formula( paste( "registered ~ ", paste( variables, collapse = " + " ) ) )

model_ctree1 <- train( formula_casual, data 	 = data_train, 
									   trControl = control, 
									   method 	 = "ctree" )

model_ctree2 <- train( formula_registerd, data   	= data_train, 
									      trControl = control, 
									      method    = "ctree" )

result_ctree1 <- predict( model_ctree1, newdata = data_test )
result_ctree2 <- predict( model_ctree2, newdata = data_test )

predict_count <- ( exp(result_ctree1) - 1 ) + ( exp(result_ctree2) - 1 )
RMSLE( count, predict_count )

```

```{r}

# -------------------------------------------------------------------------
# predicting on the test set 
setwd("/Users/ethen/Business-Analytics/2_bike_sharing/data")
test1 <- fread("test.csv")
test1[ , datetime := ymd_hms(datetime) ]
test1[ , c( "hour", "weekday", "year", "month" ) := list( hour(datetime), 
												   		  lubridate::wday(datetime),
												   		  year(datetime),
												   		  month(datetime) ) ]
test2 <- ExtractFeatures( data = test1, data.type = "test" )

result1 <- predict( model_ctree1, newdata = test2 )
result2 <- predict( model_ctree2, newdata = test2 )

predict_count <- ( exp(result1) - 1 ) + ( exp(result2) - 1 )

result <- data.frame( datetime = test1$datetime, count = predict_count )
write.csv( result, "result.csv", row.names = FALSE )

```

## R session Information

```{r}

sessionInfo()

```

[ExtractFeatures]:

### 2015-12-02 3_marketing_analysis/marketing_analysis.Rmd
half done

---
title: "Marketing Analysis"
author: "Ming-Yu Liu"
date: "November 30, 2015"
output: 
  html_document: 
    highlight: haddock
    number_sections: yes
    theme: united
    toc: yes
---

## Background Information

The domain of marketing analysis is huge, topics can range from social network analysis, real time bidding, campaign optimization and so on. Though at the heart of the marketing lies some techniques that address fundamental questions that businesses all wish to be to able answer well.

1. **Segmentation:** Allows businesses to identify who are my customers or so called target audiences? 
2. **Scoring Models:** Targeting the right customers, that is which customers should I spend the most marketing budget on?
3. **Customer Life Time Value:** What's the future value of my customers? That way the company will be able to concentrate the most on those that will be worth the most.

### 2015-12-02 1_finding_groups/finding_groups.R

@@ -9,7 +9,10 @@ library(reshape2)
 setwd("/Users/ethen/Business-Analytics/1_finding_groups/data")
 rm( list = ls() )
 
-# -- [example 1 : supply chain] -----------------------------------------------------
+# ----------------------------------------------------------
+#              example 1 : supply chain
+# ----------------------------------------------------------
+
 data_sku <- read.csv("DATA_2.01_SKU.csv")
 head(data_sku)
 
@@ -30,8 +33,9 @@ ggplot( data_sku, aes( CV, ADS, color = groups ) ) +
 geom_point() + 
 labs( title = "SKU Example", y = "Average Daily Sales", x = "Coefficient of Variation" )
 
-
-# -- [example 2 : HR]---------------------------------------------------------------
+# ----------------------------------------------------------
+#              example 2 : HR
+# ----------------------------------------------------------
 data_hr <- read.csv("DATA_2.02_HR.csv")
 str(data_hr)
 
@@ -73,6 +77,10 @@ hr_agg <- mutate( hr_agg, proportion = tl / sum(tl) ) %>%
 # -------------------------------------------------------------------------------------------------------
 # radar chart ??
 # http://stackoverflow.com/questions/29452431/how-to-plot-a-radar-chart-in-ggplot2-or-r
+# http://artax.karlin.mff.cuni.cz/r-help/library/fmsb/html/radarchart.html
+# http://stackoverflow.com/questions/24248161/plot-a-radar-chart-for-each-row-in-a-data-frame-with-r
+# rChart library
+# http://stackoverflow.com/questions/28486186/modify-existing-function-for-a-radar-plot-in-r
 
 Normalize <- function(x) ( x - min(x) ) / ( max(x) - min(x) ) 
 
@@ -101,14 +109,8 @@ theme( legend.position = "none",
 # ----------------------------------------------------------------------------------------------------------
 # testing code, not implemented
 
-# adding white borders
-variable <- unique(hr_agg_melt$variable)
-border_df <- data.frame( variable = rep( variable, each = 11 ) ) 
-border_df$border <- rep( seq( 0, 1, .1 ), length(variable) )
-geom_bar( data = border_df, aes( y = border ), stat = "identity", fill = NA, color = "white" ) +
-
-
 # correlation plot
 library(GGally)
 sample <- dput( data_hr[ 1:10, -c(5,6) ] )
 ggcorr( sample, nbreaks = 5 )
+

### 2015-12-02 2_bike_sharing/bike_sharing.R

@@ -1,0 +1,292 @@
+
+# kaggle bike sharing 
+# https://www.kaggle.com/c/bike-sharing-demand
+
+# objective : predict the total count of bikes rented during each hour covered by the test set
+
+# dependent variables :
+# registered : number of registered user
+# casual : number of non-registered user 
+# count : number of total rentals
+
+library(dplyr)
+library(tidyr)
+library(caret)
+library(rpart)
+library(party)
+library(scales)
+library(ggplot2)
+library(gridExtra)
+library(lubridate)
+library(data.table)
+setwd("/Users/ethen/Business-Analytics/2_bike_sharing/data")
+
+data <- fread("train.csv")
+
+
+# ----------------------------------------------------------------------------
+# 						Exploratory Data Analysis 
+# ----------------------------------------------------------------------------
+
+# two main goals of exploratoy data analysis
+# 1. knowing the distribution of our data more. Check for missing values, 
+# 	 and outliers, so that we can know whether they'll be needing transformations
+# 	 or included (excluded) for the modeling 
+# 2. discover the relationships between target (dependent) variables or predictors
+# 	 (independent), which can be used for feature selection
+
+# --------------------------------------------------------
+# a quick check on if there's any na value for each column
+sapply( data, function(y) sum( is.na(y) ) )
+
+# plot a histogram for numeric independent variables to visualize and understand the distribution 
+independent <- c( "casual", "registered", "count" )
+columns <- which( !names(data) %in% independent & sapply( data, is.numeric ) )
+
+histogram <- lapply( names(columns), function(column)
+{
+	ggplot( data[ , column, with = FALSE ], aes_string( column ) ) + 
+	geom_histogram()
+})
+do.call( grid.arrange, histogram )
+
+# from the visualization 
+# we can see that four variables including temp, atemp, humidity and windspeed are continuous variables
+# and values are normally distributed
+# the weather variable shows that most of the days are clear (1)
+# the holiday variable consists mainly of 0 tells you most days are not considered as holidays
+# and most value of 1 for workday tells you most days are neither weekends or holidays
+# Season has four categories ranging from 1 to 4 that are equally distributed
+
+
[…]

### 2015-12-02 3_marketing_analysis/marketing_analysis.R

@@ -1,0 +1,66 @@
+# marketing analysis
+
+
+library(dplyr)
+library(scales)
+library(ggplot2)
+library(lubridate)
+library(data.table)
+setwd("/Users/ethen/Business-Analytics/2_marketing_analysis")
+
+# customer purchase record for a retailer
+data <- fread( "purchase.txt", sep = "\t", header = FALSE  )
+
+# ----------------------------------------------------------------------------
+# 										Data Cleaning 
+# ----------------------------------------------------------------------------
+
+# 1. name the column variables :
+# there are three columns for this dataset
+# namely, customer id, purchase amount and date of purchase ( year-month-day ) 
+setnames( data, c( "customer_id", "purchase_amount", "date_of_purchase" ) )
+
+# 2. convert date_of_purchase column to date type
+data[ , date_of_purchase := ymd(date_of_purchase) ]
+
+# 3. extract the year of purchase 
+data[ , year_of_purchase := year(date_of_purchase) ]
+
+summary(data)
+# from the dataset, we can see that this dataset consists of customer purchase
+# from January 2005 to the last day of 2015. and the mean of the year_of_purchase
+# tells that most purchases have been made during 2011.
+
+
+# obtain the 
+# 1. number of purchases per year
+# 2. average purchase amount per year
+# 3. total purchase amounts per year
+# data.table allows you to compute all that in one concatenated command 
+
+explore <- data[ , .( counter 	 = .N, 
+					  avg_amount = mean(purchase_amount),
+					  sum_amount = sum(purchase_amount) ), by = year_of_purchase ] %>%
+		   arrange( year_of_purchase )
+
+ggplot( explore, aes( year_of_purchase, sum_amount ) ) + 
+geom_bar( stat = "identity", fill = "lightblue", color = "black" ) + 
+scale_y_continuous( label = comma )
+ggtitle( "Total Purchase Amount 2005-2015" )
+# as you can see from the bar plot there's a nice and positive trend in terms 
+# of how much money is spent for this retailer 
+
+# using sql-like command
+# GROUP BY 1 :group by the first column regardless of what it's called
+library(sqldf)
+sqldf( "SELECT 	 year_of_purchase, 
+			     COUNT(year_of_purchase) AS 'counter',
+			     AVG(purchase_amount) AS 'avg_amount',
+			     SUM(purchase_amount) AS 'sum_amount'
+		FROM     data
+		GROUP BY 1" )
+
+
+
+
+

### 2015-12-02 4_titanic/titanic.R

@@ -1,0 +1,268 @@
+# kaggle titanic  
+# https://www.kaggle.com/c/titanic/data
+
+library(mice)
+library(caret)
+library(dplyr)
+library(rpart)
+library(party)
+library(stringr)
+library(ggplot2)
+library(data.table)
+library(randomForest)
+setwd("/Users/ethen/Desktop")
+
+data <- fread( "train.csv", stringsAsFactors = FALSE )
+# @Parch : Number of Siblings/Spouses Aboard
+# @Sibsp : Number of Parents/Children Aboard
+
+# percentage of people that have survived 
+prop.table( table( data$Survived ) )
+
+# unboard passenger's gender and survived distribution
+prop.table( with( data, table( Sex, Survived ) ), 1 )
+
+# confirmed that 74 percent of the female survived 
+
+# ------------------------------------------------------------------
+# 						Exploratory Analysis 
+# ------------------------------------------------------------------
+
+# 1. missing value (numeric)
+col_numeric <- which( sapply( data, is.numeric ) )
+
+Missing <- function(x) sum( is.na(x) ) / length(x) * 100
+
+missing_df <- apply( data %>% select(col_numeric), 2, Missing )
+missing_df <- data.frame( variable   = names(missing_df),
+						  percentage = missing_df ) %>%
+			  arrange( desc(percentage) )
+
+# visualize the percentage of missing values 
+ggplot( missing_df, aes( reorder( variable, -percentage ), percentage ) ) + 
+geom_bar( stat = "identity", fill = "skyblue3" ) + 
+labs( x = "variables", title = "Variables with Missing Values" )
+
+# count the percentage of columns that are missing for each observations
+# remove the observations that have too large of a proportion 
+table( apply( data %>% select(col_numeric), 1, Missing ) )
+
+
+# 2. Extract the title info from the Name information and replace the old one  
+
+# split the name on both comma and period symbol, remove the white space
+# at front
+data$Name <- lapply( data$Name, function(x)
+{
+	string <- strsplit( x, split = "[,.]" )[[1]][2] %>% str_trim()
+}) %>% unlist()
+
+# group rare factor variables together with Other  
+table(data$Name)
+
+data$Name[ !data$Name %in% c( "Master", "Miss", "Mr", "Mrs" ) ] <- "Other"
+
+
+
+# ------------------------------------------------------------------------
+#							Feature Engineering 
[…]

### 2015-12-06 1_finding_groups/finding_groups.Rmd
market analysis -1

@@
 
+> This documentation assumes you're familiar with R and hierarchical clustering, thus technical details are not explained.
+
 ## Background Information
 
-This module will go over examples of how identifying groups within data allows business to improve efficiency.
+This documentation will go over two examples of how identifying groups within data allows business to improve efficiency.
 Finding groups within our data allows us to strike the right balance between similarities and differences.
@@
 The following documentation will cover two examples of where this notion can be use in the business world.
-
-> Note that this documentation assumes you're familiar with R and hierarchical clustering, thus technical details are not explained.
 

### 2015-12-06 2_bike_sharing/bike_sharing.Rmd
market analysis -1

@@
 
-The plot tells us that registered users' use bikes more often on weekdays, while non-registered users' behaviors are completely the opposite. Their demand increases over the weekend.
+The plot tells us that registered users use bikes more often on weekdays, while non-registered users' behaviors are completely the opposite. Their demand increases over the weekend.
 
@@
 
-As for the rest of the steps, we'll first print out the [`ExtractFeatures`][ExtractFeatures] function that does the feature engineering for our dataset and explain along the way.
+We'll combine the rest of the  feature engineering steps into one single  [`ExtractFeatures`][ExtractFeatures] function.
+Note that the actual test set does not include the `casual` and `registered` column.
+Hence, for this function, we also defined a data.type parameter so that we can use it for both training and test set.
 
@@
 {
-
 	# 2. convert holiday and workingday into one variable
@@
 
+# glimpse at the data
+data_train
+
 ```
+
+Now that we have preprocessed our dataset, we're ready to apply the machine learning algorithms.
 

### 2015-12-06 2_bike_sharing/slide.Rmd
market analysis -1

---
title: "Predicting Bike Demand and Segmenting Users to Further Maximize Profit"
author: "Ming-Yu Liu"
date: "December 4, 2015"
output: pdf_document
---

```{r, echo = FALSE, message = FALSE, warning = FALSE}
library(dplyr)
library(tidyr)
library(scales)
library(ggplot2)
library(gridExtra)
library(lubridate)
library(data.table)
setwd("/Users/ethen/Business-Analytics/2_bike_sharing")
data <- fread("data/train.csv")
data[ , datetime := ymd_hms(datetime) ]

demand <- select( data, datetime, casual, registered ) %>%
		  tbl_dt() %>%
		  mutate( datetime = year(datetime) ) %>%
	      gather( "user", "count", -1 ) %>%
		  group_by( datetime ) %>%
		  summarise( count = sum(count) )

```

The total demand for bikes has increased significantly through out the last year. Rising from `r demand$datetime[1]`'s `r demand$count[1]` to `r demand$datetime[2]`'s `r demand$count[2]`. Thus we should capitalize on visable growth by effective pricing and market segmentation.

**Findings: **

There are different daily bike-usage behaviors between registered and casaul users. As shown below :

```{r, echo=FALSE, fig.height=5, fig.width=8}

subdata2 <- select( data, datetime, casual, registered ) %>%
		    tbl_dt() %>%
		    mutate( datetime = lubridate::wday( datetime, label = TRUE ) ) %>%
		    gather( "user", "count", -1 ) %>%
		    group_by( datetime, user ) %>%
			summarise( count = sum(count) )

# registered and unregisted users
ggplot( subdata2, aes( datetime, count, group = user, color = user ) ) + 
geom_line( size = 1 ) + geom_point( size = 3 ) + 
scale_y_continuous( label = comma ) +
scale_colour_discrete( limit = c( "registered", "casual" ) ) + 
ggtitle( "2011-2012 User's Daily Bike Demand" )

```

Daily total bike demand plot infers that registered users use bikes more often on weekdays, while casual users' demand are higher on weekends.

**Suggested Actions: **

- Investigate potential likeings more different users to back up dynamic campaign placement on bikes towards different users
- Different pricing strategy for various timespan.
- Predict bike demands for different types of users in different locations. That way, not only can we maximize our profit and customer satisfaction by providing enough bikes at the right moment and right timing, but also identify locations of demands for different users and adjust bike downtime (for maintenance) accordingly.

### 2015-12-06 3_marketing_analysis/marketing_analysis.Rmd
market analysis -1

@@
 3. **Customer Life Time Value:** What's the future value of my customers? That way the company will be able to concentrate the most on those that will be worth the most.
+
+## Segmentation
+
+Each company may have stored a large pool of data for each of their customers based on hundreds of indicators.
+Though rich in quantity, it is often hard to directly use it to obtain the insights needed.
+Thus, segmentation is implemented to transform your data into something that is clear and usable.
+
+A more intuitive way when referring to the purpose of segmentation is: You can't treat every single one of your customers the same way, charge them the same price or communicate in the same benefits.
+But as that being said, you can't treat each and every customer individually as that will be way too costly.
+So a good segmentation is to find the right balance between simplifying it to make it usable and not simplifying it much so that it is still valuable.
+
+Knowing what segmentation does, we now have to choose the variables that are suitable to perform the segmentation on.
+While the suitable variables will still differ according to the goals that you what to accomplish, there are three variables that have proven are often times useful to predict customer profitability.
+Namely:
+
+- **Recency:** Indicates when the customer made his/her last purchase.
+If a customer has lapsed and has not made a purchase for a long period of time he/she may have lost interest.
+Or worst, switch to your competition.
+- **Frequency:** Refers to the number of purchases made in the past. More could possibly means that the more likely additional purchases will occur in the future.
+- **Monetary Value:** The amount of money spent on average at each purchase occasion. Obviously, the more money a customer spent the more valuable he/she is to the company.
+
+Segmentation performed by these three varaibles are often times referred to as RFM segmentation.
+
+After saying that much, let's start looking at our dataset at hand, a selling customer purchase record for a retailer.
+Note that the original text file data does not contain a header.
+And the three columns represents customer id, purchase amount and date of purchase ( in year-month-day format ) respectively.
+
+```{r, warning=FALSE, message=FALSE}
+
+# environment setting
+library(dplyr)
+library(ggplot2)
+library(gridExtra)
+library(lubridate)
+library(data.table)
+setwd("/Users/ethen/Business-Analytics/3_marketing_analysis")
+
+# load data, reset column names and convert to date type 
+data <- fread( "purchase.txt", sep = "\t", header = FALSE  )
+setnames( data, c( "customer_id", "purchase_amount", "purchase_date" ) )
+data[ , purchase_date := ymd(purchase_date) ]          
+data
+
+```
+
+Since recency value isn't built-in with the dataset we'll have to do it ourselves.
+We're going to compute the time difference between the `purchase_date` and the very last day (plus 1 day) recorded in the dataset.
+After obtaining the day difference we'll convert it back to numeric type.
+
+```{r}
+
+last <- max(data$purchase_date) + days(1)
+data[ , recency := difftime( last, purchase_date ) %>% as.numeric() ]
+summary(data)
+
+```
+
+Using a quick summary function on the dataset, we can see that it consists of customer purchase
+from January 2005 to the last day of 2015. The mean of the `purchase_year` tells that most purchases have been made during 2011 and the mean recency value is around 1632.
+
+Next, we'll compute the RFM value for each customer.
+Despite there's `r nrow(data)` rows in the dataset, there's actually only `r length( unique(data$customer_id) )` unique customers.
+Data.table allows you to compute all that in one concatenated command.
+After computing the value, histogram can be used to look at the value's distribution.
+
+```{r, message=FALSE, fig.width=10}
+
+rfm1 <- data[ , .( recency 	 = min(recency),
+				   frequency  = .N, 
+				   avg_amount = mean(purchase_amount) ), by = customer_id ]
+
+# distibution
+grid.arrange( ggplot( rfm1, aes( avg_amount ) ) + geom_histogram(),
+			  ggplot( rfm1, aes( frequency )  ) + geom_histogram(), ncol = 2 )
+
+```
+
+`avg_amount`'s extremely skewed histogram tells us that most people spent very few for each purchase and probably requires some tuning.
+Because for example, someone who spends 15 dollars on average generates three times the revenue for the company compared with someone who only spends 5 dollars.
+But what about two customers who spent 310 and 320 dollars respectively?
+You'll notice the difference in purchase amounts for are 10 dollars for both occasions, meaningly that they will treated the same in distance computations if we do not do something about them.
+The same goes for `frequency`.
+
+Next, since our rfm variables are recorded in completely different units (days, dollars), to group customers into different segmentations, one thing we have to do is to transform our variables.
+Only by standardizing or normalizing the scale, will make them comparable.
+The complete transformation steps are listed in the following :
+
+1. `customer_id` Remove it and store it elsewhere for now as they're meaningless in terms of segmentation.
+2. `avg_amount` and `frequency` Take logarithm value to make them more normally distributed.
+3. Scale each column to make them comparable.
+
+```{r}
+
+# store the original one  
+rfm2 <- copy(rfm1)
+
+# 1. 
+customer_id <- rfm2$customer_id
+rfm2[ , customer_id := NULL ]
+
+# 2. 
+rfm2[ , `:=`( avg_amount = log(avg_amount),
+		   	  frequency  = log(frequency) ) ]
+
+# 3.
+rfm2[ , names(rfm2) := lapply( .SD, scale ) ]
+rfm2
+
+```

### 2015-12-06 2_bike_sharing/bike_sharing.R

@@ -1,4 +1,3 @@
-
 # kaggle bike sharing 
 # https://www.kaggle.com/c/bike-sharing-demand
 
@@ -111,7 +110,15 @@ scale_colour_discrete( limit = c( "registered", "casual" ) ) +
 ggtitle( "2011-2012 User's Daily Bike Demand" )
 
 
-# 3. trend of bike demand over year 
+# total bike demand 
+demand <- select( data, datetime, casual, registered ) %>%
+		  tbl_dt() %>%
+		  mutate( datetime = year(datetime) ) %>%
+	      gather( "user", "count", -1 ) %>%
+		  group_by( datetime ) %>%
+		  summarise( count = sum(count) )
+
+# 3. trend of bike demand over year for casual and registered users
 subdata3 <- select( data, datetime, casual, registered ) %>%
 			tbl_dt() %>%
 		    mutate( datetime = year(datetime) ) %>%
@@ -141,7 +148,7 @@ ggtitle( "2011-2012 Bike Demand of Users" )
 # ----------------------------------------------------------------------------
 # 						Feature Engineering
 # ----------------------------------------------------------------------------
-data[ , datetime := ymd_hms(datetime) ]
+# data[ , datetime := ymd_hms(datetime) ]
 # add a hour, weekday, year, month column 
 data[ , c( "hour", "weekday", "year", "month" ) := list( hour(datetime), 
 												   		 lubridate::wday(datetime),
@@ -209,7 +216,7 @@ data_train <- ExtractFeatures( data = data_train, data.type = "train" )
 data_test  <- ExtractFeatures( data = data_test , data.type = "train" )
 
 # 10-fold cross validation 
-control <- trainControl( method = "cv", number = 10 )
+control <- trainControl( method = "cv", number = 6 )
 
 # evaluation measure 
 RMSLE <- function( y, pred )
@@ -233,9 +240,35 @@ model_ctree2 <- train( formula_registerd, data   	= data_train,
 result_ctree1 <- predict( model_ctree1, newdata = data_test )
 result_ctree2 <- predict( model_ctree2, newdata = data_test )
 
-predict_count <- ( exp(result_ctree1) - 1 ) + ( exp(result_ctree2) - 1 )
-RMSLE( count, predict_count )
+prediction_ctree <- ( exp(result_ctree1) - 1 ) + ( exp(result_ctree2) - 1 )
[…]

### 2015-12-06 3_marketing_analysis/marketing_analysis.R

@@ -4,33 +4,28 @@
 library(dplyr)
 library(scales)
 library(ggplot2)
+library(gridExtra)
 library(lubridate)
 library(data.table)
-setwd("/Users/ethen/Business-Analytics/2_marketing_analysis")
+setwd("/Users/ethen/Business-Analytics/3_marketing_analysis")
 
 # customer purchase record for a retailer
 data <- fread( "purchase.txt", sep = "\t", header = FALSE  )
 
 # ----------------------------------------------------------------------------
-# 										Data Cleaning 
+# 										Exploring around, not included  
 # ----------------------------------------------------------------------------
 
 # 1. name the column variables :
 # there are three columns for this dataset
 # namely, customer id, purchase amount and date of purchase ( year-month-day ) 
-setnames( data, c( "customer_id", "purchase_amount", "date_of_purchase" ) )
+setnames( data, c( "customer_id", "purchase_amount", "purchase_date" ) )
 
 # 2. convert date_of_purchase column to date type
-data[ , date_of_purchase := ymd(date_of_purchase) ]
+data[ , purchase_date := ymd(purchase_date) ]
 
 # 3. extract the year of purchase 
-data[ , year_of_purchase := year(date_of_purchase) ]
-
-summary(data)
-# from the dataset, we can see that this dataset consists of customer purchase
-# from January 2005 to the last day of 2015. and the mean of the year_of_purchase
-# tells that most purchases have been made during 2011.
-
+data[ , purchase_year := year(purchase_date) ]
 
 # obtain the 
 # 1. number of purchases per year
@@ -40,12 +35,12 @@ summary(data)
 
 explore <- data[ , .( counter 	 = .N, 
 					  avg_amount = mean(purchase_amount),
-					  sum_amount = sum(purchase_amount) ), by = year_of_purchase ] %>%
-		   arrange( year_of_purchase )
+					  sum_amount = sum(purchase_amount) ), by = purchase_year ] %>%
+		   arrange( purchase_year )
 
-ggplot( explore, aes( year_of_purchase, sum_amount ) ) + 
+ggplot( explore, aes( purchase_year, sum_amount ) ) + 
 geom_bar( stat = "identity", fill = "lightblue", color = "black" ) + 
-scale_y_continuous( label = comma )
+scale_y_continuous( label = comma ) + 
 ggtitle( "Total Purchase Amount 2005-2015" )
[…]

### 2015-12-06 4_titanic/titanic.R

@@ -10,7 +10,7 @@ library(stringr)
 library(ggplot2)
 library(data.table)
 library(randomForest)
-setwd("/Users/ethen/Desktop")
+setwd("/Users/ethen/Business-Analytics/4_titanic")
 
 data <- fread( "train.csv", stringsAsFactors = FALSE )
 # @Parch : Number of Siblings/Spouses Aboard
@@ -47,6 +47,7 @@ labs( x = "variables", title = "Variables with Missing Values" )
 # remove the observations that have too large of a proportion 
 table( apply( data %>% select(col_numeric), 1, Missing ) )
 
+# Embark also has missing values ! 
 
 # 2. Extract the title info from the Name information and replace the old one  
 
@@ -92,7 +93,8 @@ ExtractFeatures <- function(data)
 	variable <- c( "Ticket", "Cabin" )
 	data <- data %>% 
 			select( -one_of(variable) ) %>%
-			mutate( Sex 	 = as.factor(Sex), 
+			mutate( Sex 	 = as.factor(Sex),
+					Name     = as.factor(Name),
 					Pclass   = as.factor(Pclass),
 					Embarked = as.factor(Embarked),
 					Survived = as.factor(Survived) )
@@ -110,7 +112,7 @@ ExtractFeatures <- function(data)
 	# combine the categorical data back 
 	data <- cbind( data_complete, data %>% select( -col_numeric ) )
 
-	return( data.table(data) )
+	return( data.table( data[ complete.cases(data), ] ) )
 }
 
 data <- ExtractFeatures(data)
@@ -122,13 +124,14 @@ data <- ExtractFeatures(data)
 # ------------------------------------------------------------------------
 
 # split into 80 / 20 percent and evaluate prediction accuracy on test set 
+set.seed(4321)
 train_index <- createDataPartition( data$Survived, p = .8, list = FALSE )
 data_train  <- data[ train_index, ]
 data_test   <- data[ -train_index, ]
 
 # define training control to be 10-fold cross validation
 # classProbs = TRUE 
-control <- trainControl( method = "cv", number = 10 )
+control <- trainControl( method = "cv", number = 6 )
 
 # initialize the method 
 methods <- c( "rpart", "glm", "ctree" )
@@ -148,15 +151,18 @@ TrainModel <- function( method )
 
 		# convert result to numeric (from factor)
 		result <- predict( model, newdata = data_test )	
[…]

### 2015-12-12 2_marketing_analysis/marketing_analysis.Rmd
update work

--- 2015-12-06 3_marketing_analysis/marketing_analysis.Rmd
@@
 A more intuitive way when referring to the purpose of segmentation is: You can't treat every single one of your customers the same way, charge them the same price or communicate in the same benefits.
-But as that being said, you can't treat each and every customer individually as that will be way too costly.
+But as that being said, you can't treat each and every customer individually, as that will be way too costly.
 So a good segmentation is to find the right balance between simplifying it to make it usable and not simplifying it much so that it is still valuable.
@@
 library(data.table)
-setwd("/Users/ethen/Business-Analytics/3_marketing_analysis")
+setwd("/Users/ethen/Business-Analytics/2_marketing_analysis")
 
@@
 				   avg_amount = mean(purchase_amount) ), by = customer_id ]
-
 # distibution

### 2015-12-12 2_bike_sharing/bike_sharing.R

@@ -1,327 +1,0 @@
-# kaggle bike sharing 
-# https://www.kaggle.com/c/bike-sharing-demand
-
-# objective : predict the total count of bikes rented during each hour covered by the test set
-
-# dependent variables :
-# registered : number of registered user
-# casual : number of non-registered user 
-# count : number of total rentals
-
-library(dplyr)
-library(tidyr)
-library(caret)
-library(rpart)
-library(party)
-library(scales)
-library(ggplot2)
-library(gridExtra)
-library(lubridate)
-library(data.table)
-setwd("/Users/ethen/Business-Analytics/2_bike_sharing/data")
-
-data <- fread("train.csv")
-
-
-# ----------------------------------------------------------------------------
-# 						Exploratory Data Analysis 
-# ----------------------------------------------------------------------------
-
-# two main goals of exploratoy data analysis
-# 1. knowing the distribution of our data more. Check for missing values, 
-# 	 and outliers, so that we can know whether they'll be needing transformations
-# 	 or included (excluded) for the modeling 
-# 2. discover the relationships between target (dependent) variables or predictors
-# 	 (independent), which can be used for feature selection
-
-# --------------------------------------------------------
-# a quick check on if there's any na value for each column
-sapply( data, function(y) sum( is.na(y) ) )
-
-# plot a histogram for numeric independent variables to visualize and understand the distribution 
-independent <- c( "casual", "registered", "count" )
-columns <- which( !names(data) %in% independent & sapply( data, is.numeric ) )
-
-histogram <- lapply( names(columns), function(column)
-{
-	ggplot( data[ , column, with = FALSE ], aes_string( column ) ) + 
-	geom_histogram()
-})
-do.call( grid.arrange, histogram )
-
-# from the visualization 
-# we can see that four variables including temp, atemp, humidity and windspeed are continuous variables
-# and values are normally distributed
-# the weather variable shows that most of the days are clear (1)
-# the holiday variable consists mainly of 0 tells you most days are not considered as holidays
-# and most value of 1 for workday tells you most days are neither weekends or holidays
[…]

### 2015-12-12 2_marketing_analysis/marketing_analysis.R

@@ -1,0 +1,309 @@
+# marketing analysis
+
+library(dplyr)
+library(scales)
+library(ggplot2)
+library(gridExtra)
+library(lubridate)
+library(data.table)
+setwd("/Users/ethen/Business-Analytics/2_marketing_analysis")
+
+# customer purchase record for a retailer
+data <- fread( "purchase.txt", sep = "\t", header = FALSE )
+
+# ----------------------------------------------------------------------------
+# 										Exploring around, not included  
+# ----------------------------------------------------------------------------
+
+# 1. name the column variables :
+# there are three columns for this dataset
+# namely, customer id, purchase amount and date of purchase ( year-month-day ) 
+setnames( data, c( "customer_id", "purchase_amount", "purchase_date" ) )
+
+# 2. convert date_of_purchase column to date type
+data[ , purchase_date := ymd(purchase_date) ]
+
+# 3. extract the year of purchase 
+data[ , purchase_year := year(purchase_date) ]
+
+# obtain the 
+# 1. number of purchases per year
+# 2. average purchase amount per year
+# 3. total purchase amounts per year
+# data.table allows you to compute all that in one concatenated command 
+
+explore <- data[ , .( counter 	 = .N, 
+					  avg_amount = mean(purchase_amount),
+					  sum_amount = sum(purchase_amount) ), by = purchase_year ] %>%
+		   arrange( purchase_year )
+
+ggplot( explore, aes( purchase_year, sum_amount ) ) + 
+geom_bar( stat = "identity", fill = "lightblue", color = "black" ) + 
+scale_y_continuous( label = comma ) + 
+ggtitle( "Total Purchase Amount 2005-2015" )
+# as you can see from the bar plot there's a nice and positive trend in terms 
+# of how much money is spent for this retailer 
+
+# using sql-like command
+# GROUP BY 1 :group by the first column regardless of what it's called
+library(sqldf)
+sqldf( "SELECT 	 purchase_year, 
+			     COUNT(purchase_year) AS 'counter',
+			     AVG(purchase_amount) AS 'avg_amount',
+			     SUM(purchase_amount) AS 'sum_amount'
+		FROM     data
+		GROUP BY 1" )
+
+# ----------------------------------------------------------------------------
+# 										Segmentation 
+# ----------------------------------------------------------------------------
+
+# 1. name the column variables :
[…]

### 2015-12-12 3_marketing_analysis/marketing_analysis.R

@@ -1,135 +1,0 @@
-# marketing analysis
-
-
-library(dplyr)
-library(scales)
-library(ggplot2)
-library(gridExtra)
-library(lubridate)
-library(data.table)
-setwd("/Users/ethen/Business-Analytics/3_marketing_analysis")
-
-# customer purchase record for a retailer
-data <- fread( "purchase.txt", sep = "\t", header = FALSE  )
-
-# ----------------------------------------------------------------------------
-# 										Exploring around, not included  
-# ----------------------------------------------------------------------------
-
-# 1. name the column variables :
-# there are three columns for this dataset
-# namely, customer id, purchase amount and date of purchase ( year-month-day ) 
-setnames( data, c( "customer_id", "purchase_amount", "purchase_date" ) )
-
-# 2. convert date_of_purchase column to date type
-data[ , purchase_date := ymd(purchase_date) ]
-
-# 3. extract the year of purchase 
-data[ , purchase_year := year(purchase_date) ]
-
-# obtain the 
-# 1. number of purchases per year
-# 2. average purchase amount per year
-# 3. total purchase amounts per year
-# data.table allows you to compute all that in one concatenated command 
-
-explore <- data[ , .( counter 	 = .N, 
-					  avg_amount = mean(purchase_amount),
-					  sum_amount = sum(purchase_amount) ), by = purchase_year ] %>%
-		   arrange( purchase_year )
-
-ggplot( explore, aes( purchase_year, sum_amount ) ) + 
-geom_bar( stat = "identity", fill = "lightblue", color = "black" ) + 
-scale_y_continuous( label = comma ) + 
-ggtitle( "Total Purchase Amount 2005-2015" )
-# as you can see from the bar plot there's a nice and positive trend in terms 
-# of how much money is spent for this retailer 
-
-# using sql-like command
-# GROUP BY 1 :group by the first column regardless of what it's called
-library(sqldf)
-sqldf( "SELECT 	 purchase_year, 
-			     COUNT(purchase_year) AS 'counter',
-			     AVG(purchase_amount) AS 'avg_amount',
-			     SUM(purchase_amount) AS 'sum_amount'
-		FROM     data
-		GROUP BY 1" )
-
-# ----------------------------------------------------------------------------
-# 										Segmentation 
-# ----------------------------------------------------------------------------
-
[…]

### 2015-12-12 4_titanic/titanic.R

@@ -1,281 +1,0 @@
-# kaggle titanic  
-# https://www.kaggle.com/c/titanic/data
-
-library(mice)
-library(caret)
-library(dplyr)
-library(rpart)
-library(party)
-library(stringr)
-library(ggplot2)
-library(data.table)
-library(randomForest)
-setwd("/Users/ethen/Business-Analytics/4_titanic")
-
-data <- fread( "train.csv", stringsAsFactors = FALSE )
-# @Parch : Number of Siblings/Spouses Aboard
-# @Sibsp : Number of Parents/Children Aboard
-
-# percentage of people that have survived 
-prop.table( table( data$Survived ) )
-
-# unboard passenger's gender and survived distribution
-prop.table( with( data, table( Sex, Survived ) ), 1 )
-
-# confirmed that 74 percent of the female survived 
-
-# ------------------------------------------------------------------
-# 						Exploratory Analysis 
-# ------------------------------------------------------------------
-
-# 1. missing value (numeric)
-col_numeric <- which( sapply( data, is.numeric ) )
-
-Missing <- function(x) sum( is.na(x) ) / length(x) * 100
-
-missing_df <- apply( data %>% select(col_numeric), 2, Missing )
-missing_df <- data.frame( variable   = names(missing_df),
-						  percentage = missing_df ) %>%
-			  arrange( desc(percentage) )
-
-# visualize the percentage of missing values 
-ggplot( missing_df, aes( reorder( variable, -percentage ), percentage ) ) + 
-geom_bar( stat = "identity", fill = "skyblue3" ) + 
-labs( x = "variables", title = "Variables with Missing Values" )
-
-# count the percentage of columns that are missing for each observations
-# remove the observations that have too large of a proportion 
-table( apply( data %>% select(col_numeric), 1, Missing ) )
-
-# Embark also has missing values ! 
-
-# 2. Extract the title info from the Name information and replace the old one  
-
-# split the name on both comma and period symbol, remove the white space
-# at front
-data$Name <- lapply( data$Name, function(x)
-{
-	string <- strsplit( x, split = "[,.]" )[[1]][2] %>% str_trim()
-}) %>% unlist()
-
-# group rare factor variables together with Other  
-table(data$Name)
-
-data$Name[ !data$Name %in% c( "Master", "Miss", "Mr", "Mrs" ) ] <- "Other"
-
-
-
[…]

### 2015-12-15 2_marketing_analysis/marketing_analysis.Rmd
segmentation

@@
 author: "Ming-Yu Liu"
-date: "November 30, 2015"
+date: "December 20, 2015"
 output: 
@@
 # environment setting
+library(grid)
 library(dplyr)
@@
 # load data, reset column names and convert to date type
+# also create a year column for later convenience
 data <- fread( "purchase.txt", sep = "\t", header = FALSE  )
@@
 data[ , purchase_date := ymd(purchase_date) ] 
+data[ , purchase_year := year(purchase_date) ]
 data
@@
 Using a quick summary function on the dataset, we can see that it consists of customer purchase
-from January 2005 to the last day of 2015. The mean of the `purchase_year` tells that most purchases have been made during 2011 and the mean recency value is around 1632.
+from January 2005 to the last day of 2015. The mean of the `purchase_year` tells that most purchases have been made during 2011 and the average recency value is around 1632 days.
 
 Next, we'll compute the RFM value for each customer.
-Despite there's `r nrow(data)` rows in the dataset, there's actually only `r length( unique(data$customer_id) )` unique customers.
-Data.table allows you to compute all that in one concatenated command.
-After computing the value, histogram can be used to look at the value's distribution.
+Despite that that are `r nrow(data)` rows in the dataset, there's actually only `r length( unique(data$customer_id) )` unique customers for this entire dataset.
+To be explicit, we'll compute:
+
+- `avg_amount` The average amount spent by every customer to represent monetary value. 
+- `first_purchase` and `recency` The first time the customer bought something and the last time (closest to today) they bought something.
+- `frequency` How often did each customer purchased at this retail store.
+
+Data.table allows you to compute all that in one concatenated command. After computing the value, histogram can be used to look at the values' distribution.
 
@@
 rfm1 <- data[ , .( recency 	  	  = min(recency),
+				   first_purchase = max(recency),
 				   frequency  	  = .N, 
@@
 grid.arrange( ggplot( rfm1, aes( avg_amount ) ) + geom_histogram(),
-			  ggplot( rfm1, aes( frequency )  ) + geom_histogram(), ncol = 2 )
+			  ggplot( rfm1, aes( frequency )  ) + geom_histogram(), ncol = 2,
+			  top = textGrob( "Variable Distribution" ) )
 
@@
 
-`avg_amount`'s extremely skewed histogram tells us that most people spent very few for each purchase and probably requires some tuning.
-Because for example, someone who spends 15 dollars on average generates three times the revenue for the company compared with someone who only spends 5 dollars.
+`avg_amount`'s extremely skewed histogram tells us that most people spent very few for each purchase and probably requires some tuning if we're to use them for building a predictive model later on.
+We can think about this in an intuitive way.
+Someone who spends 15 dollars on average generates three times the revenue for the company compared with someone who only spends 5 dollars.
 But what about two customers who spent 310 and 320 dollars respectively?
-You'll notice the difference in purchase amounts for are 10 dollars for both occasions, meaningly that they will treated the same in distance computations if we do not do something about them.
+You'll notice the difference in purchase amounts for these two groups are 10 dollars for both occasions, meaningly that they will treated the same in distance computations if we do not do something about them.
 The same goes for `frequency`.
 
-Next, since our rfm variables are recorded in completely different units (days, dollars), to group customers into different segmentations, one thing we have to do is to transform our variables.
-Only by standardizing or normalizing the scale, will make them comparable.
-The complete transformation steps are listed in the following :
+Moving on to computing our segmentation.
+To perform this, we'll be using a set of rules.
+Customer dataset is changing at a continous bases; old customer disappear, new customers are acquired.
+Thus, using a statistical (machine-learning) segmentation method ( e.g.
+clustering algorithms ) for this kind of situatation may not be so cost-effective, as you will probably have to update the segmentation the moment you've created it.
+Another reason that statistical segementation may not be applicable for this circumstances is that perhaps customer behaviors are seasonal.
+The best segmentation solution from a statistical viewpoint will most likely vary if you run your analysis at different time point, making them uncomparable.
+To put it all together, how can you target your actions on one of the segment if the very definition of that segment keeps on fluctuating over time?
+For this reason, we'll be performing our customer segmentation by a set of simple rules and these rules are fixed for some period of time, update it only after a few quarters of years depending on the needs of the business.
 
-1. `customer_id` Remove it and store it elsewhere for now as they're meaningless in terms of segmentation.
-2. `avg_amount` and `frequency` Take logarithm value to make them more normally distributed.
-3. Scale each column to make them comparable.
+The indicator of determining which customer should receive more attention is based on how much they spent and how likely are they going to buy from us in the future.
+One basic criteria of determining whether customer will buy from us in the future is **recency**.
+Someone who has made a purchase a few weeks is far more likely to buy from the company again ( compared to someone who has made their lasy purchase many years ago ).
+
+Hence we'll first divide our customers into four main groups based on **recency**.
+We define a customer to be `active` if they have bought from us within the last year.
+`warm` if they have made a purchase between 1 to 2 years ago.
+`cold` is a customer has made a purchase between 2 to 3 years ago.
+For those who haven't purchased anything for more than three years, we will label them as `inactive`.
+( Of course, you should change these cutoff values for each segment depending on your needs ).
+
+After splitting our customers into four main segments, we'll also perform further segmenting on the `active` and `warm` segments.
+We'll take the customers from these segments and put the new ones, those with the history of making their first purchase within the same timespan into the segment of `new active` and the rest as `old active`.
+e.g.
+the active customer is defined as those who have made their last purchases within this year, if that customer made their first purchase within this year, a.k.a same timespan, then they would be classified as new.
+
+Next, we'll take a step further and split these two sub-groups into four sub-sub-groups based on how many money they've spent on average.
+For this cutoff value we decided to use 100 based on the 90 percent quantile.
+Again, tune this parameters based on your needs.
 
@@
 
-# store the original one  
-rfm2 <- copy(rfm1)
+quantile( rfm1$avg_amount, probs = seq( 0, 1, 0.1 ) )
+
+```
+
+This converts our segmentation for the original `active` segment to four segments `new active low`, `new active high`, `old active low` and `old active high`.
+We'll also apply the exact same idea to the `warm` segment.
+Resulting in a total of 10 segments.
+
+Hopefully all that wordy description will become much more clear if you looked at the code.
+
+The function [`Segmentation`][Segmentation] for creating this segmentation is not printed to conserve space.
+It takes the rfm data and creates manual segmentation based on the recency criteria first that splits customer's into inactive, cold, warm and active based on the three recency cutoff value.
+Then further splits the warm and active segment into four sub-groups each based on the amount cutoff value and whether they have made their first purchase within the same timespan with respect to the "baseline".
+This will classify each of the two original segments further into new high, new low, old high and old low.
+Input parameters:
+
+- `data` rfm segmentation data, must have five columns exactly named as customer_id, recency, first_purchase, frequency and avg_amount. The order of the column doesn't matter.
+- `inactive` Customers whose recency are higher than this value will be classified as inactive customers. Note that the unit of recency is "days".
+- `cold` Same functionality as above, except it's the cutoff value for cold customers.
+- `warm` Same functionality as above, cutoff value for splitting the warm and active customers.
+- `amount` Cutoff value for splitting the warm and active customers further into high and low average purchased amount.
+- The function returns the original rfm data and adds a new segment column to it. It also sorts the order of the rows by customer id.
+
+The next code chunk uses the function, prints out data and the proportional distribution of the segment column.
+
+```{r}
+
+# segmentation parameters 
+inactive <- 3 * 365
+cold 	 <- 2 * 365
+warm 	 <- 365 
+amount   <- 100
+
+source("/Users/ethen/Business-Analytics/2_marketing_analysis/code/marketing_functions.R")
+rfm1 <- Segmentation( data = rfm1, inactive = inactive, 
+					  cold = cold, warm = warm, amount = amount )
+rfm1
+
+# segment proportional distribution 
+round( table(rfm1$segment) / sum( table(rfm1$segment) ), 2 )
+
+```
+
+We can directly identify that around half of our customers are "inactive", meaning that they have not not purchased anything for at least three years after their last purchase.
+If we wish to identify the profile or characteristics of each segment, we can calculate the average number of each variable ( except for customer id of course ).
+
+```{r}
+
+rfm1[ , lapply( .SD, mean ), by = segment, .SDcols = -c("customer_id") ]
+
+```
+
+However, the most important usage of these segmentations is that you go back in time and check how many customers were in each customers a year or two years ago and see how things have evolved.
+By doing so, we'll be able to analyze trends such as are there increase in the number of high value customers?
+Or have you added more new customers than the year before, etc.
+Answers to these questions will evaluate the company's overall performance so far and see whether there's room for improvement.
+
+One caveat for doing this is that: The customers that appeared in your dataset today may not even existed a year or even a few weeks ago.
+Thus when we're using today's dataset to go back in time we have to make sure that we properly take care of those non-existent customers as we'll see later.
+
+The next chunck will obtain the rfm segmentation `rfm2`.
+This serves as our retrospective segmentation where we'll go back in time by *one year* and obtain the segmentation at that time using the dataset we have today.
+
+```{r}
+
+# retrospective customer segmentation 
+rfm2 <- data[ recency > 365, 
+			  .( recency 	  	= min(recency) - 365,
+				 first_purchase = max(recency) - 365,
+				 frequency  	= .N, 
+				 avg_amount 	= mean(purchase_amount) ), by = customer_id ]
+
+rfm2 <- Segmentation( data = rfm2, inactive = inactive, 
+					  cold = cold, warm = warm, amount = amount )
+rfm2
+
+# segment proportional distribution 
+round( table(rfm2$segment) / sum( table(rfm2$segment) ), 2 )
+
+```
+
+Now that we've obtain `rfm1` the segmentation for today and `rfm2` the segmentation for a year ago, the primary question we want to answer is: which segment should we concentrate on or in other words, which customer segmentation will be most beneficial to the future success of the company and are the ones that we should be allocating more resources on?
+Let's dive into that in the next code chunk.
+
+1. `revenue1` Aggregate each customer's total purchased amount for the current year.
+2.
+`actual2` Merge the previous year's segment with this year's revenue, this is essentially examining how did our customers from last year performed this year.
+If a customer has not purchased anything for the current year then they will not show up in the revenue1 data ( resulting in NA values ) and we'll replace these values with 0.
+3.
+Compute `revenue_average`: the average revenue that you would expect from each segment's customer in the next time period ( in our example, next year ).
+`revenue_total`: expected total revenue generated by each segment.
+
+```{r, fig.width=10, fig.height=6}
 
 # 1.
-customer_id <- rfm2$customer_id
-rfm2[ , customer_id := NULL ]
+revenue1 <- data[ purchase_year == 2015, 
+				  .( revenue = sum(purchase_amount) ), by = customer_id ]
+# 2.
+revenue2 <- merge( rfm2, revenue1, by = "customer_id", all.x = TRUE )
+revenue2[ is.na(revenue), revenue := 0 ]
 
-# 2. 
-rfm2[ , `:=`( avg_amount = log(avg_amount),
-		   	  frequency  = log(frequency) ) ]
+# 3. average revenue
+revenue_average <- revenue2[ , .( avg_revenue = mean(revenue) ), by = segment ] %>%
+				   arrange( desc(avg_revenue) )
 
-# 3.
-rfm2[ , names(rfm2) := lapply( .SD, scale ) ]
-rfm2
+# total revenue
+revenue_total <- revenue2[ , .( avg_revenue = sum(revenue) ), by = segment ] %>%
+				 arrange( desc(avg_revenue) )
+
+list( revenue_average, revenue_total )
 
 ```
+
+From the printed result, we can infer interesting facts about our customer segment.
+For the `revenue_average` data.table, each "`r revenue_average$segment[1]`" customer is worth approximately `r round(revenue_average$avg_revenue[1])` to our company ( or you can say contribute that much do our company if the previous statement sounded a bit rude ...
+).
+This is a lot higher than a "`r revenue_average$segment[ nrow(revenue_average) ]`" customer, which is at the bottom of the average revenue list.
+However, when we switched to the `revenue_total` data.table, it tells us that `r revenue_total$segment[1]` is the segment that generated the most revenue in total!
+If we looked back at our original `rfm2` segmentation ( scroll back up please ), we can also see that the customer size of this segment is around 20 percent and the second highest segment `r revenue_total$segment[1]` represents only about 3 percent of our entire customer.
+This is a good example where small segments in terms of number of customers that generates a huge share of revenues.
+And it's directly of telling us that customers should not be treated equally as there are just some that contributes more to the overall revenue.
+
+## Customer Life Time Value
+
+Let's envisage the following scenario.
+You're company lauched a new customer acquistion campaign including hard-discounts, online-keywords, e-mail campaigns etc.
+In the end, it only turned out to be a minor success.
+Sure, the company has in fact acquired new customers and generated new sales, but the sales generated does not even cover the cost of the campaign.
+The supervisor in charge of the campaign , however, argued that the newly acquired customers will remain loyal to the company and generate additional sales in the future.
+In a sense this campaign is said to be an investment for the future.
+Is this true?
+or is it just another excuse for the campaign's bad performance?
+
+This problem scenario is the whole point of measuring **Customer Life Time Value** (CLTV), to prove or disprove whether acquiring these new customers is a good investment.
+
+**Takeaways:**
+
+1. Use statistical segmentation when the segment you wish to obtain is a "oneshot" thing or use it as a exploratory tool when you're not famaliar with the overall objective.
+
+You can find all the source code [here](https://github.com/ethen8181/Business-Analytics/tree/master/2_marketing_analysis/code).
+marketing_analysis.R contains the documentation's general scripts and marketing_functions.R contains the functions that were sourced in and used with this documentation.
+
+If you happened to have any comments, feedbacks of any sort, Feel free to file them [here](https://github.com/ethen8181/Business-Analytics/issues)!
+
+## R Session Information
+
+```{r}
+
+sessionInfo()
+
+```
+
+[Segmentation]: https://github.com/ethen8181/Business-Analytics/tree/master/2_marketing_analysis/code/marketing_functions.R

### 2015-12-15 2_marketing_analysis/code/marketing_analysis.R

@@ -1,0 +1,189 @@
+# marketing analysis
+
+library(grid)
+library(dplyr)
+library(scales)
+library(ggplot2)
+library(gridExtra)
+library(lubridate)
+library(data.table)
+setwd("/Users/ethen/Business-Analytics/2_marketing_analysis")
+
+# customer purchase record for a retailer
+data <- fread( "purchase.txt", sep = "\t", header = FALSE )
+
+# ----------------------------------------------------------------------------
+# 										Exploring around, not included  
+# ----------------------------------------------------------------------------
+
+# 1. name the column variables :
+# there are three columns for this dataset
+# namely, customer id, purchase amount and date of purchase ( year-month-day ) 
+setnames( data, c( "customer_id", "purchase_amount", "purchase_date" ) )
+
+# 2. convert date_of_purchase column to date type
+data[ , purchase_date := ymd(purchase_date) ]
+
+# 3. extract the year of purchase 
+data[ , purchase_year := year(purchase_date) ]
+
+# obtain the 
+# 1. number of purchases per year
+# 2. average purchase amount per year
+# 3. total purchase amounts per year
+# data.table allows you to compute all that in one concatenated command 
+
+explore <- data[ , .( counter 	 = .N, 
+					  avg_amount = mean(purchase_amount),
+					  sum_amount = sum(purchase_amount) ), by = purchase_year ] %>%
+		   arrange( purchase_year )
+
+ggplot( explore, aes( purchase_year, sum_amount ) ) + 
+geom_bar( stat = "identity", fill = "lightblue", color = "black" ) + 
+scale_y_continuous( label = comma ) + 
+ggtitle( "Total Purchase Amount 2005-2015" )
+# as you can see from the bar plot there's a nice and positive trend in terms 
+# of how much money is spent for this retailer 
+
+# using sql-like command
+# GROUP BY 1 :group by the first column regardless of what it's called
+library(sqldf)
+sqldf( "SELECT 	 purchase_year, 
+			     COUNT(purchase_year) AS 'counter',
+			     AVG(purchase_amount) AS 'avg_amount',
+			     SUM(purchase_amount) AS 'sum_amount'
+		FROM     data
+		GROUP BY 1" )
+
+# ----------------------------------------------------------------------------
+# 										Segmentation 
+# ----------------------------------------------------------------------------
+
+# 1. name the column variables :
+# there are three columns for this dataset
[…]

### 2015-12-15 2_marketing_analysis/code/marketing_functions.R

@@ -1,0 +1,76 @@
+# marketing analysis
+
+
+# -------------------------------------------------------------------------------------------
+# [Segmentation]: 
+# Takes the rfm data and creates manual segmentation 
+# based on the recency criteria first that splits customer's into inactive, cold, 
+# warm and active based on the three recency cutoff value. Then further splits 
+# the warm and active segment into four sub-groups each based on the amount cutoff value and 
+# whether they have made their first purchase within the same timespan with respect to the "baseline". 
+# classifying each of the two original segments further into new high, new low, old high and old low.
+
+# @data 	: rfm segmentation data, must have five columns exactly named as  
+# 			  customer_id, recency, first_purchase, frequency and avg_amount, 
+# 			  the order of the column doesn't matter. 
+# @inactive : customers whose recency are higher than this value will be classified as 
+# 			  "inactive" customers. Note that the unit of recency is "days".
+# @cold		: same functionality as above, except it's the cutoff value for "cold" customers.
+# @warm 	: same functionality as above, cutoff value for "warm" and "active" customers.
+# @amount 	: cutoff value for splitting the warm and active customers further into
+# 			  high and low average purchased amount.
+# returns 	: original rfm data and adds a new segment column to it. It also sorts the order
+# 			  of the rows by customer id.  
+
+Segmentation <- function( data, inactive, cold, warm, amount )
+{
+	# segmentation criteria:
+	# four main groups, grouped by recency 
+	data[ recency > inactive, segment := "inactive" ]
+	data[ recency <= inactive & recency > cold, segment := "cold" ]
+	data[ recency <= cold & recency > warm, segment := "warm" ]
+	data[ recency <= warm, segment := "active" ]
+
+ 	# split the active and warm segment further into new and old customers 
+	# those that made their first purhase within a year, 
+	# with respecitve to the cold and warm baseline.
+	# and split it into low and high base on a specified average amount
+
+	# Note that the old and new is based on the cold criteria for the "warm" segment!
+	data[ segment == "warm" & first_purchase > cold & avg_amount < amount,
+		  segment := "old warm low" ]
+
+	data[ segment == "warm" & first_purchase > cold & avg_amount >= amount,
+		  segment := "old warm high" ]
+
+	data[ segment == "warm" & first_purchase <= cold & avg_amount < amount, 
+		  segment := "new warm low" ]
+
[…]

### 2015-12-15 2_marketing_analysis/marketing_analysis.R

@@ -1,309 +1,0 @@
-# marketing analysis
-
-library(dplyr)
-library(scales)
-library(ggplot2)
-library(gridExtra)
-library(lubridate)
-library(data.table)
-setwd("/Users/ethen/Business-Analytics/2_marketing_analysis")
-
-# customer purchase record for a retailer
-data <- fread( "purchase.txt", sep = "\t", header = FALSE )
-
-# ----------------------------------------------------------------------------
-# 										Exploring around, not included  
-# ----------------------------------------------------------------------------
-
-# 1. name the column variables :
-# there are three columns for this dataset
-# namely, customer id, purchase amount and date of purchase ( year-month-day ) 
-setnames( data, c( "customer_id", "purchase_amount", "purchase_date" ) )
-
-# 2. convert date_of_purchase column to date type
-data[ , purchase_date := ymd(purchase_date) ]
-
-# 3. extract the year of purchase 
-data[ , purchase_year := year(purchase_date) ]
-
-# obtain the 
-# 1. number of purchases per year
-# 2. average purchase amount per year
-# 3. total purchase amounts per year
-# data.table allows you to compute all that in one concatenated command 
-
-explore <- data[ , .( counter 	 = .N, 
-					  avg_amount = mean(purchase_amount),
-					  sum_amount = sum(purchase_amount) ), by = purchase_year ] %>%
-		   arrange( purchase_year )
-
-ggplot( explore, aes( purchase_year, sum_amount ) ) + 
-geom_bar( stat = "identity", fill = "lightblue", color = "black" ) + 
-scale_y_continuous( label = comma ) + 
-ggtitle( "Total Purchase Amount 2005-2015" )
-# as you can see from the bar plot there's a nice and positive trend in terms 
-# of how much money is spent for this retailer 
-
-# using sql-like command
-# GROUP BY 1 :group by the first column regardless of what it's called
-library(sqldf)
-sqldf( "SELECT 	 purchase_year, 
-			     COUNT(purchase_year) AS 'counter',
-			     AVG(purchase_amount) AS 'avg_amount',
-			     SUM(purchase_amount) AS 'sum_amount'
-		FROM     data
-		GROUP BY 1" )
-
-# ----------------------------------------------------------------------------
-# 										Segmentation 
-# ----------------------------------------------------------------------------
-
-# 1. name the column variables :
[…]

### 2015-12-16 2_marketing_analysis/marketing_analysis.Rmd
on hold

@@
 
-1. `revenue1` Aggregate each customer's total purchased amount for the current year.
+1. `revenue_per_customer` Aggregate each customer's total purchased amount for the current year.
 2.
-`actual2` Merge the previous year's segment with this year's revenue, this is essentially examining how did our customers from last year performed this year.
+`revenue_predicted` Merge the previous year's segment with this year's revenue, this is essentially examining how did our customers from last year performed this year.
 If a customer has not purchased anything for the current year then they will not show up in the revenue1 data ( resulting in NA values ) and we'll replace these values with 0.
 3.
-Compute `revenue_average`: the average revenue that you would expect from each segment's customer in the next time period ( in our example, next year ).
-`revenue_total`: expected total revenue generated by each segment.
+`revenue_summary`: The average revenue that you would expect from each segment's customer in the next time period ( in our example, next year ) and the expected total revenue generated by each segment.
 
@@
 # 1.
-revenue1 <- data[ purchase_year == 2015, 
+revenue_per_customer <- data[ purchase_year == 2015, 
 				  			  .( revenue = sum(purchase_amount) ), by = customer_id ]
 # 2.
-revenue2 <- merge( rfm2, revenue1, by = "customer_id", all.x = TRUE )
-revenue2[ is.na(revenue), revenue := 0 ]
+revenue_predicted <- merge( rfm2, revenue_per_customer, by = "customer_id", all.x = TRUE )
+revenue_predicted[ is.na(revenue), revenue := 0 ]
 
-# 3. average revenue
-revenue_average <- revenue2[ , .( avg_revenue = mean(revenue) ), by = segment ] %>%
-				   arrange( desc(avg_revenue) )
-
-# total revenue
-revenue_total <- revenue2[ , .( avg_revenue = sum(revenue) ), by = segment ] %>%
-				 arrange( desc(avg_revenue) )
-
-list( revenue_average, revenue_total )
+# 3. 
+revenue_summary <- revenue_predicted[ , .( avg_revenue 	 = mean(revenue),
+								  		   total_revenue = sum(revenue) ), by = segment ] %>%
+				   arrange( desc(total_revenue), desc( avg_revenue) )
+revenue_summary
 
@@
 From the printed result, we can infer interesting facts about our customer segment.
-For the `revenue_average` data.table, each "`r revenue_average$segment[1]`" customer is worth approximately `r round(revenue_average$avg_revenue[1])` to our company ( or you can say contribute that much do our company if the previous statement sounded a bit rude ...
+The average revenue column of the `revenue_summary` data.table tells us that each "`r revenue_summary$segment[2]`" customer is worth approximately `r round(revenue_summary$avg_revenue[2])` to our company ( or you can say contribute that much do our company if the previous statement sounded a bit rude ...
 ).
-This is a lot higher than a "`r revenue_average$segment[ nrow(revenue_average) ]`" customer, which is at the bottom of the average revenue list.
-However, when we switched to the `revenue_total` data.table, it tells us that `r revenue_total$segment[1]` is the segment that generated the most revenue in total!
-If we looked back at our original `rfm2` segmentation ( scroll back up please ), we can also see that the customer size of this segment is around 20 percent and the second highest segment `r revenue_total$segment[1]` represents only about 3 percent of our entire customer.
+This is a lot higher than a "`r revenue_summary$segment[ revenue_summary$segment == "inactive" ]`" customer.
+However, when we switched to the total revenue column, it tells us that `r revenue_summary$segment[1]` is the segment that generated the most revenue in total!
+If we looked back at our original `rfm2` segmentation ( scroll back up please ), we can also see that the customer size of this segment is around 20 percent and the second highest segment `r revenue_summary$segment[2]` represents only about 3 percent of our entire customer.
 This is a good example where small segments in terms of number of customers that generates a huge share of revenues.

### 2015-12-16 2_marketing_analysis/code/marketing_analysis.R

@@ -114,7 +114,7 @@ cold 	 <- 2 * 365
 warm 	 <- 365 
 amount   <- 100
 
-source("/Users/ethen/Business-Analytics/2_marketing_analysis/code/segmentation.R")
+source("/Users/ethen/Business-Analytics/2_marketing_analysis/code/marketing_functions.R")
 rfm1 <- Segmentation( data = rfm1, inactive = inactive, 
 					  cold = cold, warm = warm, amount = amount )
 
@@ -139,34 +139,59 @@ round( table(rfm2$segment) / sum( table(rfm2$segment) ), 2 )
 
 
 # revenue per customer 
-revenue1 <- data[ purchase_year == 2015, 
-				  .( revenue = sum(purchase_amount) ), by = customer_id ]
+revenue_per_customer <- data[ purchase_year == 2015, 
+				  			  .( revenue = sum(purchase_amount) ), by = customer_id ]
 
 # compare the previous year and see how they did this year 
-revenue2 <- merge( rfm2, revenue1, by = "customer_id", all.x = TRUE )
-revenue2[ is.na(revenue), revenue := 0 ]
+revenue_predicted <- merge( rfm2, revenue_per_customer, by = "customer_id", all.x = TRUE )
+revenue_predicted[ is.na(revenue), revenue := 0 ]
 
-# average revenue that you would expect from each segment's segment in the next year 
-revenue_average <- revenue2[ , .( avg_revenue = mean(revenue) ), by = segment ] %>%
-				   arrange( desc(avg_revenue) )
+# average revenue that you would expect from each segment's segment in the next year and 
+# expected total revenue generated by each segment
+revenue_summary <- revenue_predicted[ , .( avg_revenue 	 = mean(revenue),
+								  		   total_revenue = sum(revenue) ), by = segment ] %>%
+				   arrange( desc(total_revenue), desc( avg_revenue) )
 
-# expected total revenue generated by each segment 
-revenue_total <- revenue2[ , .( total_revenue = sum(revenue) ), by = segment ] %>%
-				 arrange( desc(total_revenue) )
-
-list( revenue_average, revenue_total )
-
-
-# visualize 
-ggplot( revenue_average, aes( reorder( segment, -avg_revenue ), avg_revenue ) ) + 
-geom_bar( stat = "identity" )
-
+revenue_summary
 
 # ----------------------------------------------------------------------------
 # 							Customer life time value   
 # ----------------------------------------------------------------------------
 
 
[…]

### 2016-01-05 3_AB_test/AB_test.Rmd
ab test 1

---
title: "A / B test"
author: "Ming-Yu Liu"
date: "January 11, 2016"
output: 
  html_document: 
    highlight: pygments
    number_sections: yes
    theme: cosmo
    toc: yes
---

## Overview

A / B testing is a general methodology used online to when you want to test out a new feature. What you're doing is you're going show one set of features, the control set ( your existing feature ) to one user group and another set, your experiment set ( your new feature ) to another user group and test how did these users respond differently so that you can determine which set of your feature is better.

Despite its useful functionality, there are still places where A / B testing isn't as useful. For example :

1. Testing out completely new experiences. An obvious way for thinking about this is if you're dramatically changing the user's experience. The user might prefer the old way of doing things or they've curious about the new features and they simply test out everything ( sometimes called the novelty effect ).
2. A / B testing can't tell you if you're missing something. Meaning it can tell you if A performs better B or vice versa, but it can't tell you that if you use C, then it will actually perform better than the former two.
3. Tesing out products that people rarely buy. e.g. cars, apartments. It might be too long before the user actually decides to take actions after seeing the information and you might be unaware of the actual motivation.

When launching a A / B test you need to ask yourself :

1. What is your baseline for comparison. It's essentially saying that you need to have clear control and metrics.
2. How much time you need in order to actually have your user adapt to the new experience.

## Examples

Suppose you're running an educational platform and your hypothesis is : Will changing the "Start Now" button from orange to pink increase how many students explore the platform's courses. So in this case the metric that's use to evaluate the change's performance is the click through probability ( Unique visitors who click the button / Unique visitors to page ). Note that it is often times impractical to use metrices such as total number of students that completed the course as it often takes weeks or months before a student can do that.

Next we will jot down the hypothesis that we wish to test out, in our case the our null and alternative hypothesis would be :

**The null hypothesis, H0: ** The experimental and control groups have the same probability of completing a checkout ( clicking the button ). Or equivalent to saying that the differences of the two groups' probability is 0.   
**The alternative hypothesis, H1: ** The two groups have different probability of completing a checkout.

## Designing the Test

Now that we've defined our hypothesis, the first question that comes into mind is how many tests do we need to run, or in a sense how long should the test last in order for us to make our decisions. To do that we can use a power analysis for two independent samples, which can be calculated using the `power.prop.test` function. Before doing that there are some basic parameters that we should be aware of :

1. **Significance Level: ** Denoted by $\alpha$, this threshold governs the chance of false positive. A significance level of 0.05 means that there's a 5 percent chance of false positive.
2. **Statistical power: ** This stands for $1 - \beta$ ( or so called sensitivity ), where $\beta$ represents the probability that you'll get a false negative. So a statistical power of 0.8 means that if there's is in fact a change, there's 80 percent that we'll detect it, which is equivalent to saying that there will be a 20 percent false negative.

The effect of picking a significance level of 0.05 and power of 0.8 means that we are 4 times more likely to obtain a false negative than a false positive. For A / B testing, we’re generally more concerned about getting a false positive : Making a change that doesn’t actually improve things than we are about not making a change at all, which is why we accept a greater likelihood of a false negative. The rule of thumb for $\alpha$ and $1 - \beta$ is 0.5 and 0.8 respectively.

Now suppose that our current baseline is 0.1 ( there's a 10 percent chance that people who saw the button will click it ). And we wish to detect a change of two percent in the click through rate ( This is already consider quite high for online experiment ).

Parameters :

- `baseline` Your current baseline solution. 
- `delta` Minimum detectable change, smallest effect that will be detected (1-β)% of the time. This parameter can also be referred to as the practical significance boundary.
- `power` Percent of the time the minimum effect size will be detected, assuming it exists.
- `sig_level` Percent of the time a difference will be detected, assuming one does NOT exist.

```{r}

# parameters
baseline  <- 0.1
delta 	  <- 0.02
power 	  <- 0.8
sig_level <- 0.05

result <- power.prop.test( p1 = baseline, p2 = baseline + delta, 
				 		   power = power, sig.level = sig_level,
				 		   alternative = "two.sided" )
result

```

The result shows that we need at least `r round(result$n)` sample size for each scenario to detect if there will actually be a 2 percent more-than-baseline click through probability.

How these parameters affect the sample size you need :

- `baseline` The higher the baseline click through probability ( but still less than 0.5 ), the larger the sample size you'll need. Since the probability is related to the standard deviation, where it reaches the maximum at 0.5.
- `delta` The smaller the change you wish to detect, the larger the sample size you'll need.
- `power` The higher the value means that that you wish to increase the confidence that you have in the result. Thus it means that you need a larger sample size.
- `sig_level` The smaller the value means that you wish to increase the confidence that you have in the result. Thus it means that you need a larger sample size.

## Analyze the Result

Suppose you have run the test and you've obtain the total number of sample sizes and the total number of successes for both groups. Given these variables we can use it to calculate whether the proportional change was due to variation or not. To do so, we'll calculate the confidence interval for the difference, which can be done by calculating the difference plus and minus the z score times the standard deviation. Math Formula :
$$p_{experiment} - p_{control} \pm Z \sqrt{ p_{pooled}( 1-p_{pooled} )( \frac{1}{n_{experiment}} + \frac{1}{n_{control}} ) }$$

Where you can calculate the $p_{pooled}$, the pooled probability by the sum of the number of successes for both groups divided by the sum of the total number of sample size for both groups.

[`ABTest`][ABTest] Function that calculates the confidence interval. Input parameters :

- `count_control` The number of successes. This is equivalent to the number of people that clicked the button for the control group ( your original feature ).
- `sizes_control` The total number of sample size for the control group.
- The same notion can be applied to the experiment's variable `count_experiment` and `sizes_experiment`.
- Returns : A data.frame consisted of the mean and the confidence interval of the result.

```{r}

# parameters 
count_control 	 <- 974
sizes_control 	 <- 10072
count_experiment <- 1242
sizes_experiment <- 9886

ABTest <- function( count_control = count_control, 
					sizes_control = sizes_control,
					count_experiment = count_experiment, 
					sizes_experiment = sizes_experiment )
{
	# probability of each group 
	p_control <- count_control / sizes_control
	p_experiment <- count_experiment / sizes_experiment

	# @p : pooled probability
	# @std_error pooled standard deviation (error ) 
	p <- ( count_control + count_experiment ) / ( sizes_experiment + sizes_control )
	std_error <- sqrt( p * ( 1 - p ) * ( 1 / sizes_control + 1 / sizes_experiment ) )

	# 95 percent confidence interval's z score = 1.96
	difference <- p_experiment - p_control
	confidence <- difference + c( -1, 1 ) * 1.96 * std_error

	return( data.frame( lower = confidence[1], 
						mean  = difference,
						upper = confidence[2] ) )
}

confidence <- ABTest( count_control = count_control, 
					  sizes_control = sizes_control,
					  count_experiment = count_experiment, 
					  sizes_experiment = sizes_experiment )

```

In order to launch a change, the change should be larger than the minimum detectable change that you wished to detect, or in other words it should be larger than your practical significance boundary. In our case, the value we've set was `r delta`. From the result above, we can denote that since even the lower bound of the confidence interval is larger than the value. Base on this fact, we'll definitely launch the newer version of the click button.

Different scenarios of the output :

```{r, warning=FALSE,message=FALSE}

# fixed artifical plot
# using delta = 0.02 as the minimum detectable boundary 
library(ggplot2)

scenario <- as.character(2:6)
lower <- c( -0.008, 0.01, -0.025, -0.005, 0.015   )
mean  <- c( 0.01, 0.015, 0.005, 0.025, 0.025 )
upper <- c( 0.018, 0.018, 0.03, 0.04, 0.03 )

examples <- data.frame( scenario, lower, mean, upper )
examples <- rbind( cbind( scenario = "1", confidence ), examples )
examples$scenario <- factor( examples$scenario, levels = as.character(6:1) )

ggplot( examples, aes( mean, scenario, color = scenario ) ) + 
geom_point() + 
geom_errorbarh( aes( xmin = lower, xmax = upper ), height = 0.1 ) + 
geom_vline( xintercept = 0, color = "black" ) + 
geom_vline( xintercept = delta, color = "blue", linetype = "dotted" ) + 
geom_vline( xintercept = -delta, color = "blue", linetype = "dotted" ) +
scale_color_discrete( breaks = as.character(1:6) ) +  
labs( title = "Different Scenarios of Confidence Interval",
	  x = "confidence interval" )

```

1. **Scenario 1:** The case where even the lower bound of the confidence interval lies above the practical significance boundary. Accept the change of the new feature.
2. **Scenario 2:** The lower bound of the confidence interval lies below 0 and the upper bound lies below the practical significance boundary. There's no statistically significant change from 0 ( the confidence interval includes 0 ) and that you're also confident that there's not a practical significance change. Given this it's not worth the effort to launch the change.
3. **Scenario 3:** The lower bound of the confidence interval lies above 0 and the upper end lies below the practical significance boundary. This implies that you're confident that there is a positive change, but it's not practically significant. In other words, you're confident that there was a change, but you don't care about the magnitude of the change.
4. **Scenario 4:** Both the lower and upper bound of the confidence interval lies beyond the practical significance boundary. This means that the new feature could cause users to increase by the minimum detectable change or it could be cuasing them to decrease by the minimum detectable change.  
5. **Scenario 5:** The point estimate is beyond the practical significant line, the lower bound of the confidence interval, however, overlaps 0. This means that this change is in fact the effect that you care about, but there's also a chance that there might not be a change at all.
6. **Scenario 6:** The point estimate is beyond the practical significant line and the lower bound of the confidence interval is greater than 0. This is a situation that indicates the change has a chance of being practically significant and not being practically significant.

For the last three scenario, scenario 4 - 6: If your confidence interval includes your practical significance boundary, would you be sure that the change should not be launched? After all, it's reasonably likely that there was an effect you care about. In these cases, you should run an additional test with greater power if you have the time. But sometimes, you'll have to make a decision even though there's an uncertainty about how real your result is.

**Sources: **

Source [code](https://github.com/ethen8181/machine-learning/blob/master/3_AB_test/AB_test.R) to the documentation.

Bugs are comments can be filed [here](https://github.com/ethen8181/Business-Analytics/issues).

## Reference

1. Determining the sample size for A / B test : https://signalvnoise.com/posts/3004-ab-testing-tech-note-determining-sample-size
2. Z-score distribution table : http://www.utdallas.edu/dept/abp/zscoretable.pdf

## R Session Information

```{r}

sessionInfo()

```

[ABTest]: https://github.com/ethen8181/machine-learning/blob/master/3_AB_test/AB_test.R

### 2016-01-06 3_AB_test/AB_test.Rmd
AB test 2

@@
     number_sections: yes
-    theme: cosmo
+    theme: united
     toc: yes
@@
 
+## Examples
+
 When launching a A / B test you need to ask yourself :
@@
 
-## Examples
-
-Suppose you're running an educational platform and your hypothesis is : Will changing the "Start Now" button from orange to pink increase how many students explore the platform's courses.
+So now, suppose you're running an educational platform and your hypothesis is : Will changing the "Start Now" button from orange to pink increase how many students explore the platform's courses.
 So in this case the metric that's use to evaluate the change's performance is the click through probability ( Unique visitors who click the button / Unique visitors to page ).
@@
 					  sizes_experiment = sizes_experiment )
+confidence
 
@@
 In our case, the value we've set was `r delta`.
-From the result above, we can denote that since even the lower bound of the confidence interval is larger than the value.
-Base on this fact, we'll definitely launch the newer version of the click button.
+Base on the result above, we can denote that since even the lower bound of the confidence interval is larger than the value, we'll definitely launch the newer version of the click button.
 
@@
 scenario <- as.character(2:6)
-lower <- c( -0.008, 0.01, -0.025, -0.005, 0.015   )
-mean  <- c( 0.01, 0.015, 0.005, 0.025, 0.025 )
-upper <- c( 0.018, 0.018, 0.03, 0.04, 0.03 )
+lower <- c( -0.008, 0.011, -0.025, -0.005, 0.015 )
+mean  <- c( 0.005, 0.014, 0.005, 0.025, 0.025 )
+upper <- c( 0.018, 0.017, 0.035, 0.055, 0.035 )
 
@@
 
+## Characterizing a Proper Metric
+
+After seeing an simple A / B test example, we'll take a step back and see how do come up with a well-defined metric.
+
+**Step1: Generate a high level metrices**
+
+A non-exhaustive list of techniques that might be able to help you come up with high level metrices that you might want to test out in your experiment.
+
+1. External Sources : For these types of data, do not *match* them with internal data that you might already have at hand. Instead, look at a time series of both your
+internally computed metric and the externally available one and see if the trends and seasonal variability line up.
+2. Gathering your own data : 
+    - User Experience Group.
+    - Focus Group.
+    - Retrospective Analysis.
+    - Survey.
+
+**Step2: Turning it into a well-defined metrices**
+
+Recall that in the examples above, we've chosen the high-level metric click-through probability.
+And this is defined by the unique number of people how click the button versus the number of users who visited the home page.
+
+But to actually use implement the definition, we'll also have to address some other questions.
+Such as : If the same user visits the page once and comes back a week or two later, do we still only want to count that once?
+Usually, it is preferred that we count these separately, thus we'll also need to specify a time period.
+
+So given that users are usually recorded by cookies online.
+One version of the fully-defined metric will be : For each minute ( choose any time interval that you prefer ), the number of cookies that clicked divided by the number of cookies that interacted with the page.
+
+Another way of implementing this is to calculate : For each minute the number of pageviews ( assign each pageview with an id ) with a click divided by the number of pageviews.
+
+**Step3: Segmenting and filtering your traffic**
+
+Creating segments and filtering out data can be useful for evaluating metric definitions, because, well, you can look at how the different definitions vary by segment.
+
+For example, you're looking at your metric of total active cookie users over time.
+And you see a spike in one of the timelines.
+After confirming that this is not caused by seasonal variation, we can look at different segment of our visitors to see if one of the segment is causing the spike.
+
+Suppose we have chosen segment to be platforms ( mobile and desktop ). Plot the rate and probability by platforms to determine
+
+## Others
+
+Some principles to consider when designing A / B test :
+
+1. Risk : What risk is the participant undertaking? The main threshold is whether the risk exceeds that of “minimal risk”. e.g. people's health.
+2. Benefits : What benefits might result from the study?
+3. Alternatives : What other choices do people have? In a sense are the people coerced to participate since they lack other alternatives to choose from.
+4.
+Data Sensitivity : What level of confidentiality can participants expect?
+Studies without personal data like names or simply data that can't be match to an individual are generally fine.
+Though a terms of service or a privacy policy is necessary, as it contains a clear statement of what data are you going to collect and how you use it.
+
 **Sources: **
 
-Source [code](https://github.com/ethen8181/machine-learning/blob/master/3_AB_test/AB_test.R) to the documentation.
+Source [code](https://github.com/ethen8181/Business-Analytics/blob/master/3_AB_test/AB_test.R) to the documentation.
 
-Bugs are comments can be filed [here](https://github.com/ethen8181/Business-Analytics/issues).
+Bugs or comments can be filed [here](https://github.com/ethen8181/Business-Analytics/issues).
 
@@
 
-[ABTest]: https://github.com/ethen8181/machine-learning/blob/master/3_AB_test/AB_test.R
+[ABTest]: https://github.com/ethen8181/Business-Analytics/blob/master/3_AB_test/AB_test.R

### 2016-01-15 3_AB_test/AB_test.Rmd
abtest done for now

@@
 ---
-title: "A / B test"
+title: "Online A / B test"
 author: "Ming-Yu Liu"
-date: "January 11, 2016"
+date: "January 15, 2016"
 output: 
@@
 
-## Overview
+# Overview
 
@@
 
-## Examples
+# Quick Example
 
@@
 
-1. What is your baseline for comparison. It's essentially saying that you need to have clear control and metrics.
-2. How much time you need in order to actually have your user adapt to the new experience.
+1. What is your hypothesis and what's the baseline for comparison. 
+2. What's the confidence level that you wish to have in your result ( This will also affect the number of samples ). 
+2. How many samples and time do you need in order to actually have your user adapt to the new experience.
 
@@
 **The alternative hypothesis, H1: ** The two groups have different probability of completing a checkout.
-
-## Designing the Test
 
@@
 
-## Analyze the Result
+# Analyze Quick Example's Result
 
@@
 
-	# 95 percent confidence interval's z score = 1.96
+	# 95 percent confidence interval's z score = 1.96, equivalent to 
+	# qnorm( 0.975 )
 	difference <- p_experiment - p_control
@@
 
-## Characterizing a Proper Metric
+# Characterizing a Proper Metric
 
-After seeing an simple A / B test example, we'll take a step back and see how do come up with a well-defined metric.
+After walking through a simple A / B test example, we'll take a step back and see how to come up with a well-defined metric.
 
-**Step1: Generate a high level metrices**
+## Generate a high level metrices
 
-A non-exhaustive list of techniques that might be able to help you come up with high level metrices that you might want to test out in your experiment.
+For generating a high level metric, a good place to start is to think about the overall business objective.
+Below are a non-exhaustive list of techniques that might be able to help you come up with high level metrices that you might want to test out in your experiment.
 
-1. External Sources : For these types of data, do not *match* them with internal data that you might already have at hand. Instead, look at a time series of both your
+1. **External Sources :** For these types of data, do not *match* them with internal data that you might already have at hand. Instead, look at a time series of both your
 internally computed metric and the externally available one and see if the trends and seasonal variability line up.
-2. Gathering your own data : 
-    - User Experience Group.
-    - Focus Group.
-    - Retrospective Analysis.
-    - Survey.
+2. **Gathering your own data :**
+    - **User Experience Group :** With user experience group, you go really deep with a few users.
+It can take form of observing users doing tasks or you ask users to self-document their behaviors.
+After collecting the feedback it's usually a good idea to follow it up with a retrospective analysis.
+Meaning that if you observe something in a UER study, you could then go looking for that pattern in your data logs to see if that observation bears out and is worth creating a metric about.
+    - **Focus Group :** This is basically where you recruit a bunch of users or potential users and bring them together for a group discussion or you can ask them questions.
+Though you run the risk of the answers get dominated by "loud voices".
+    - **Survey :** The main advantage of surveys is that you can typically reach thousands of users.
+The disadvantages are that the data is self-reported and users don’t always give full or truthful answers.
 
-**Step2: Turning it into a well-defined metrices**
+    Qualitative methods such as User Experience Group and Focus Group are much better suited for answering questions about why or how to fix a problem.
+Whereas quantitative methods like Surveys do a much better job answering how many and how much types of questions.
+Having such numbers from the survey helps prioritize resources and allows you to focus on the biggest issues at hand.
 
-Recall that in the examples above, we've chosen the high-level metric click-through probability.
-And this is defined by the unique number of people how click the button versus the number of users who visited the home page.
+## Turning it into a well-defined metrices
+
+Recall that in the examples above, we've chosen the high-level metric click-through probability, which is defined by the unique number of people how click the button versus the number of users who visited the page that the button was located.
 
@@
 
-**Step3: Segmenting and filtering your traffic**
+## Segmenting and filtering your traffic
 
@@
 
-For example, you're looking at your metric of total active cookie users over time.
-And you see a spike in one of the timelines.
+For example, you're looking at your metric of total active users over time and you see a spike in one of the timelines.
 After confirming that this is not caused by seasonal variation, we can look at different segment of our visitors to see if one of the segment is causing the spike.
+Suppose we have chosen segment to be geographic, it might just happen that we've identify a large proportion of the traffic is generated by a specific region and it might be best for us to dig deeper and understand why.
 
-Suppose we have chosen segment to be platforms ( mobile and desktop ). Plot the rate and probability by platforms to determine
+To develop an understanding about your data and system, we can plot different metrices together, such as the rate ( calculated by the number of clicks divided by number of pageviews ) and probability by different segments to determine the differences in the output.
 
-## Others
+## Summarizing Metric
 
-Some principles to consider when designing A / B test :
+Apart from click-through probability, there're four other common summarizations that might be useful.
 
-1. Risk : What risk is the participant undertaking? The main threshold is whether the risk exceeds that of “minimal risk”. e.g. people's health.
-2. Benefits : What benefits might result from the study?
-3. Alternatives : What other choices do people have? In a sense are the people coerced to participate since they lack other alternatives to choose from.
-4.
-Data Sensitivity : What level of confidentiality can participants expect?
-Studies without personal data like names or simply data that can't be match to an individual are generally fine.
-Though a terms of service or a privacy policy is necessary, as it contains a clear statement of what data are you going to collect and how you use it.
+1. **sums and counts** : e.g. how many users visited the homepage.
+2. **distributional metric** : e.g. quantiles. Mean age of users who completed the course ( though keed in mind that mean is sensitivity to outliers ).
+3.
+**probabilities and rates** : Recall the subtle differences between the definition of probabilities and rates here.
+Probability means that given a time interval, whether a person clicked on the button or not ( takes the number of 0 / 1 ) divided by the total number users that interacted with the page but did not click the button during that period of time.
+As for rate, it refers to calculating the total number of clicks divided by the total number of pageviews.
+4. **ratios.** e.g. the probability that a user clicks on a revenue generating click divided by any other click.
+
+## Measuring Sensitivity and Robustness
+
+It's the idea that you want to choose a metric that picks up the changes that you care about but robust against the changes that you don't care about, meaning that it doesn't move a lot when nothing interesting happens.
+It's an important thing to keep in mind when choosing the distributional metric as your summary metric.
+
+e.g.
+When testing out things like video latency, you plot different summary metric of comparable video's latency, such as the mean, 90 percentile.
+In theory there shouldn't be that much of a difference between these videos, if there is, then it means that the summary metric isn't robust enough.
+And suppose you increase the video quality across these comparable videos and the mean does not change at all, then this probably indicates that choosing the mean as the summary metric isn't sensitive enough.
+
+# Designing an Experiment
+
+Suppose you have chosen the suitable metrics to be used for the test, now it's time to apply that and work through the decisions you need to make when you're actually designing an experiment.
+This can be boiled down to three main part.
+
+1. Define what you mean as a "subject" in your experiment and control. In other words, what are the units in the population that you're going to be running the test on and comparing? 
+2. Choose the "populations". e.g. Everyone or only subjects in the U.S. ? You need to be sure that you're computing the metrices on equivalent populations ( This holds true for subject ).
+3. Determine the size of the experiment. This will also affect the duration of the experiment.
+
+## Define the Subject
+
+What you need to do is to decide how to assign events to either the control or the experiment.
+You can randomly assign every event, but if you have a user-visible change then this might not be a good idea.
+Since what will happen is : The user will see the change when they first visit the page and after reloading the page, the user will not see the change (leading to confusion).
+So for when conducting user-visible change, you want to assign people as oppose to events.
+
+The next question is how do we define a user? There're three commonly used categories, namely user id, anonymous id (cookie) and event.
+
+1.
+**user id :** e.g.
+Log in user names.
+Choosing this as the proxy for your user means that all the events that corresponds to the same user id are either in the control or experiment group, regardless of whether that user is switching between a mobile phone or desktop.
+This also means that if the user has not log in then he / she will neither be assgined to a control or experiment group.
+
+2.
+**anonymous id (cookie) :** The cookie is specific for a browser and device, thus if the user switches from Chrome to Firefox, they'll be assigned to a different cookie.
+Also note that users can clear the cookie, in which case the next time they visit the website they'll get assigned to a new cookie even if they're still using the same browser and device.
+For experiments that will be crossing the sign-in border, using cookie is preferred.
+e.g.
+Suppose you're changing the layout of the page or locations of the sign in bar then you should use a cookie.
+
+3. **event :** Should only be used when you're testing a non-user-visible change. e.g. page load time.
+
+## Define the Population
+
+If you think you can identify which population of your users will be affected by your experiment, you might want to target your experiment to that traffic ( e.g.
+changing features specific to one language's users ) so that the rest of the population won't dilute the effect.
+
+Next, depending on the problem you're looking at, you might want to use a **cohort** instead of a population.
+A cohort makes much more sense than looking at the entire population when testing out learning effects, examining user retention or anything else that requires the users to be established for some reason.
+
+A quick note on **cohort**.
+The gist of cohort analysis is basically putting your customers into buckets so you can track their behaviours over a period of time.
+The term cohort stands for a group of customers grouped by the timeline ( can be week, month ) where they first made a purchase ( can be a different action that's valuable to the business ).
+In a cohort, you can have roughly the same parameters in your two user group, which makes them more comparable.
+
+For example, you're an educational platform has an existing course that's already up and running.
+Some of the students have completed the course, some of them are midway through and there're students who have not yet started.
+If you want to change the structure of of one of the lessons to see if it improves the completion rate of the entire course and they started the experiment at time X.
+For students who have started before the experiment initiated they may have already finished the lesson already leading to the fact that they may not even see the change.
+So taking the whole population of students and running the experiment on them isn't what you want.
+Instead, you want to segment out the cohort, the group of customers, that started the lesson are the experiment was launched and split that into an experiment and control group.
+
+## Size and Duration
+
+You can size your experiment based on your significance level, power and minimum detectable change like we've mentioned in the example.
+But you can also decrease the size of your experiment by targeting experiment to specific traffic.
+Because you're only looking at a subset of your traffic, you might need a bigger change before it to matters to the business ( Recall that the smaller the change you wish to detect, the larger the sample size you’ll need ).
+On other note although this might not reduce the time frame of the experiment, but this allows other experiments to be run on the same time.
+
+After determining the size of the experiment it's time to translate that into a set of practical decisions including :
+
+1. When do I want to run the experiment.
+2.
+What fraction of the traffic are you going to send through the experiment.
+This will be closely tied to the duration of the experiment.
+There're some reasons that you might not want to run the experiment on all of your traffic to get the result faster.
+    - The first consideration might be you're just uncertained of how your users will react to the new feature, so you might want to test it out a bit before you get users blogging about it.
+To same notion applies to riskier cases, such as you're completely switching your backend system, if it doesn't work well, then the site might go down.
+    - Another consideration is that you may want to run the test a bit longer to make sure that the experiment captures the different user behaviour of weekdays, weekends.
+    - The final one is that you may be running multiple tests at the same time and if you wish to make them directly comparable ( different level or attributes of the same feature ), then the easiest way is to run them at the same time.
+So if there's a holiday during the timespan that you're running the experiment, then all of the experiments should be equally affected.
+
+# Analyzing Results
+
+## Sanity Checks
+
+Before jumping into analyze the result, we should check that the population size metric and any other metric that you don't expect to change shouldn't change while you're running the experiment, or in a sense that they're invariant and shouldn't differ between the experiment and control group.
+
+Checking invariants example :
+
+For instance, after running your experiment for a week, you've discovered that the total number of users assigned to the control group is 64454 and the total number of users assigned to the experiment group 61818.
+How would you figure out whether the difference is within expectation given that each user is randomly assigned to the control or experiment group with a probability of 0.5 ?
+
+This is equivalent to saying out of a total 126272 ( 64454 + 61818 ) users, is it surprising to see if 64454 users are assigned to the control group, which is essentially a binomial distribution.
+Knowing this information, we can construct a confidence interval to test if the number lies within the confidence interval.
+The confidence interval can be calculated by the mean plus and minus the z-score times the standard error.
+
+$$ mean \pm Z \sqrt{ np(1-p) } $$
+
+Where the mean is expected number of users in the control / experiment group, which is simply the total number of the two groups times p (0.5).
+And the standard error of a binomial distribution is $\sqrt{ np(1-p) }$.
+
+```{r}
+
+group1 <- 64454
+group2 <- 61818
+SanityCheck <- function( group1, group2 )
+{
+	n <- group1 + group2
+	confidence <- n * 0.5 + c( -1, 1 ) * 1.96 * sqrt( n * 0.5 * 0.5 ) 
+	return( confidence )
+}
+( sanity <- SanityCheck( group1, group2 ) )
+
+```
+
+The result shows that 64454 does not lie within the range of the computed confidence interval and therefore it indicates that cookies may not be split equally.
+
+When this kind of situation happens it's usually best to go back to the day by day data to get a better idea of what could be going wrong.
+One good thing is to check whether any particular day stands out, or it is just an overall pattern.
+If it is an overall pattern, then it is suggested that we should check if something went wrong with the experiment setup before proceeding on to analyzing the result.
+
+## Analyzing the Result
+
+One thing to keep in mind is that the goal of analyzing the result is not only about looking for statistical significance , but also about making a business decision about whether your experiment has favorably impacted your metric.
+So if you have an experiment that statistically improves the experience for thirty percent of the users but is neutral for everyone else, do you want to launch it as is, or do you want to try and make it better?
+
+**Some other takeaways: **
+
+A / B testing is an iterative process, some common mistakes that you want to avoid is :
+
+1.
+Falling into the trap of “We already tried that”.
+A hypothesis can be implemented in so many different ways.
+If you just do one headline test and say "we tried that," you’re really selling yourself short.
+2.
+Not testing continually or not retesting after months or years.
+Just because you tested a variation in the past doesn’t necessarily mean that those results are going to be valid a year or two from now.
+3.
+Optimizing for the top of the funnel, rather than the product.
+Understanding what the customers want so that you can make the product better.
+Ultimately, you can’t simply test your headlines and get people to like your product more.
 
@@
 
-## Reference
+# Reference
 
@@
 2. Z-score distribution table : http://www.utdallas.edu/dept/abp/zscoretable.pdf
+3. Additional techniques to generating metrices : https://storage.googleapis.com/supplemental_media/udacityu/3954679115/additional_techniques.pdf
+4. Pitfalls of A / B test : http://www.forbes.com/sites/sujanpatel/2015/10/29/how-to-do-ab-testing-right-and-avoid-the-most-common-mistakes-marketers-make/#2715e4857a0b54f535a44349
 
-## R Session Information
+# R Session Information
 

### 2016-01-15 3_AB_test/AB_test.R

@@ -48,7 +48,8 @@ ABTest <- function( count_control = count_control,
 	p <- ( count_control + count_experiment ) / ( sizes_experiment + sizes_control )
 	std_error <- sqrt( p * ( 1 - p ) * ( 1 / sizes_control + 1 / sizes_experiment ) )
 
-	# 95 percent confidence interval's z score = 1.96
+	# 95 percent confidence interval's z score = 1.96, equivalent to 
+	# qnorm( 0.975 )
 	difference <- p_experiment - p_control
 	confidence <- difference + c( -1, 1 ) * 1.96 * std_error
 
@@ -62,15 +63,17 @@ confidence <- ABTest( count_control = count_control,
 					  count_experiment = count_experiment, 
 					  sizes_experiment = sizes_experiment )
 
+# ------------------------------------------------------------------
+# different scenarios
 
 # fixed artifical plot
 # using delta = 0.02 as the minimum detectable boundary 
 library(ggplot2)
 
 scenario <- as.character(2:6)
-lower <- c( -0.008, 0.01, -0.025, -0.005, 0.015   )
-mean  <- c( 0.01, 0.015, 0.005, 0.025, 0.025 )
-upper <- c( 0.018, 0.018, 0.03, 0.04, 0.03 )
+lower <- c( -0.008, 0.011, -0.025, -0.005, 0.015 )
+mean  <- c( 0.005, 0.014, 0.005, 0.025, 0.025 )
+upper <- c( 0.018, 0.017, 0.035, 0.055, 0.035 )
 
 examples <- data.frame( scenario, lower, mean, upper )
 examples <- rbind( cbind( scenario = "1", confidence ), examples )
@@ -87,6 +90,38 @@ labs( title = "Different Scenarios of Confidence Interval",
 	  x = "confidence interval" ) 
 
 
+# ------------------------------------------------------------------
+#				Variance   
+# ------------------------------------------------------------------
+
+# mean 
+# might have to use a power.t.test to calculate the sample size needed 
+visit <- c( 87029, 113407, 84843, 104994, 99327, 92052, 60684 )
+mean(visit) + c( -1, 1 ) * 1.96 * sd(visit) / sqrt( length(visit) )
+
+
+group1 <- 64454
+group2 <- 61818
+SanityCheck <- function( group1, group2 )
+{
+	n <- group1 + group2
+	confidence <- n * 0.5 + c( -1, 1 ) * 1.96 * sqrt( n * 0.5 * 0.5 ) 
+	return( confidence )
+}
[…]

### 2016-02-14 3_AB_test/AB_test.Rmd
reorganize doc

@@
 
-1. **Significance Level: ** Denoted by $\alpha$, this threshold governs the chance of false positive. A significance level of 0.05 means that there's a 5 percent chance of false positive.
+1. **Significance Level:** Denoted by $\alpha$, this threshold governs the chance of false positive. A significance level of 0.05 means that there's a 5 percent chance of false positive.
 2.
-**Statistical power: ** This stands for $1 - \beta$ ( or so called sensitivity ), where $\beta$ represents the probability that you'll get a false negative.
-So a statistical power of 0.8 means that if there's is in fact a change, there's 80 percent that we'll detect it, which is equivalent to saying that there will be a 20 percent false negative.
+**Statistical power:** This stands for $1 - \beta$ ( or so called sensitivity ), where $\beta$ represents the probability that you'll get a false negative.
+So a statistical power of 0.8 means that if there's is in fact a change, there's 80 percent chance that we'll detect it, which is equivalent to saying that there will be a 20 percent false negative.
 

### 2016-02-15 1_finding_groups/finding_groups.Rmd
add sources
new readme

@@
 
-> This documentation assumes you're familiar with R and hierarchical clustering, thus technical details are not explained.
+> This documentation assumes you're familiar with hierarchical clustering.
 
@@
 
-- **volaility** : This can be measured by the [coefficient of variation](https://en.wikipedia.org/wiki/Coefficient_of_variation), which is simply the standard deviation of sales divided by the mean sales for each SKU.
+- **volatility** : This can be measured by the [coefficient of variation](https://en.wikipedia.org/wiki/Coefficient_of_variation), which is simply the standard deviation of sales divided by the mean sales for each SKU.
 
@@
 
-## Takeaways
+**Takeaways:**
 
@@
 There are different perspective (intuition, business sense, technical approaches) as to which variable should have or should not have been included in the dataset prior to running the clustering method.
-As clustering algorithms are so categorized under unsupervised algorithms, there're really no "correct" ouput, so choose your variables or features wisely !
-Suppose that for some reason you think that it's nonsense to exclude the `S` variable : the satisfaction level, like we did in this documentation, you should add it back and run the clustering, though you'll obtain similar grouping results for this dataset, and when you realize that whether you include a variable or not does not affect the result, you should exclude it as this lessen the calculation loading for the algorithm and your computer.
+As clustering algorithms are categorized under unsupervised algorithms, meaning that there is no way of measuring whether the ouput is "correct".
+Thus you should choose your input variables (features) wisely!
+Speaking of this, suppose that for some reason you think that it's nonsense to exclude the `S` variable : the satisfaction level, like we did in this documentation.
+You should add it back and run the clustering, though you'll obtain similar grouping results for this dataset, and when you realize that whether you included a variable or not does not affect the result, you should exclude it as this lessen the calculation loading for the algorithm and your computer.
 
@@
 We've heard like a billion times already that a picture is worth thousands of words.
-And more importantly, you should always ask yourself, given the results what are you going to do with it, how can this result help my business goals ?
+And more importantly, you should always ask yourself, given the results what are you going to do with it, how can this result help my business goals?
 
@@
 If you're interested in a more technical way of choosing the suitable number and evaluate the stability of your clustering result, you can refer to another documentation [here](http://ethen8181.github.io/machine-learning/clustering/clustering.html).
+
+**Sources:**
 

### 2016-02-16 articles/articles.Rmd
organize article

---
title: "Articles"
author: "Ming-Yu Liu"
output: html_document
---

- We need companies to start innovating instead of imitating success stories. [[Hustle](http://thehustle.co/snapchat-vs-spacex-battle-of-the-imitators-and-innovators)]
- My 4+ Years At Uber Taught Me These Key Lessons. 1. May the supply be with you ; 2. Don't listen, show them. [[Linkedin](https://www.linkedin.com/pulse/three-things-i-learned-during-4-years-uber-michael-pao)]
- Cohort analysis. [[Youtube](https://www.youtube.com/watch?v=NyhVdGmnh0I)]
- How to set goals: OKRs. [[Youtube](https://www.youtube.com/watch?v=mJB83EZtAjc)]
- Meetings that don’t suck. [[Youtube](https://www.youtube.com/watch?v=rU8nv38E0Qk)]
- Ted Talk : How to use data to make a hit TV show. [[Ted](http://www.ted.com/talks/sebastian_wernicke_how_to_use_data_to_make_a_hit_tv_show?utm_campaign=social&utm_medium=referral&utm_source=facebook.com&utm_content=talk&utm_term=technology)]

### 2016-03-16 3_AB_test/AB_test.Rmd
ggplot collection

@@
 	confidence <- n * 0.5 + c( -1, 1 ) * 1.96 * sqrt( n * 0.5 * 0.5 ) 
-	return( confidence )
+	return(confidence)
 }
@@
 4. Pitfalls of A / B test : http://www.forbes.com/sites/sujanpatel/2015/10/29/how-to-do-ab-testing-right-and-avoid-the-most-common-mistakes-marketers-make/#2715e4857a0b54f535a44349
+5. KissMetric Infographic on A / B testing : https://blog.kissmetrics.com/creating-effective-conversion-optimization-process/?utm_source=tofu-email&utm_medium=email
 

### 2016-03-16 articles/articles.Rmd
ggplot collection

@@
 author: "Ming-Yu Liu"
-output: html_document
+output: 
+  html_document: 
+    highlight: haddock
 ---
 
+- Wanted in College Graduates: Tolerance for Ambiguity. [[Linkedin Pulse](https://www.linkedin.com/pulse/wanted-college-graduates-tolerance-ambiguity-jeff-selingo)]
+- Know the Difference between Customer Service and Customer Experience.
+[[Harvard Business Review](https://hbr.org/sponsored/2016/03/know-the-difference-between-customer-service-and-customer-experience?utm_source=twitter&utm_medium=social&utm_campaign=harvardbiz)]
+- Two articles about saying no to p-values.
+[[R-bloggers](http://www.r-bloggers.com/after-150-years-the-asa-says-no-to-p-values/)][[R-bloggers](http://www.r-bloggers.com/its-not-the-p-values-fault-reflections-on-the-recent-asa-statement-relevant-r-resources/)]
+- The 9 Best Marketing And Sales Analytics You Should Know About.[[Data Science Central](http://www.datasciencecentral.com/profiles/blogs/the-9-best-marketing-and-sales-analytics-you-should-know-about)]
 - We need companies to start innovating instead of imitating success stories. [[Hustle](http://thehustle.co/snapchat-vs-spacex-battle-of-the-imitators-and-innovators)]

### 2016-03-16 ggplot2/plot.Rmd
ggplot collection

---
title: "Collection of ggplot2 Plots"
date: "`r Sys.Date()`"
author: "Ming-Yu Liu"
output:
  rmdformats::readthedown:
    highlight: haddock
---

# Avoiding Pie Charts

Date: 2016.3.16

The purpose of a pie chart is to show the relationship of parts out of a whole. With that being said they can still be really bad at the one thing they're ostensibly designed to do. Consider the following pie chart from the [Wall Street Journal](http://www.wsj.com/articles/what-data-scientists-do-all-day-at-work-1457921541) article on What Data Scientists Do All Day at Work.

<img src = "/Users/ethen/Business-Analytics/ggplot2/avoid_pie_charts/data_scientist.png" width = "400" height = "800" >

The problem with the plot is that, our original goal was to compare and contrast these six tasks. But at a glance, do you have any idea whether more time is spent on "Presenting Analysis" or "Data Cleaning"? Thus, if the intent was to primarily allow comparison of hours in-task, leaving some ability to compare the same time category across tasks then bar plots are probably the way to go.

Prepares the data `d`, which contains three columns.

- `Task` Different categories of task.
- `Hours` Amounts of time spent.
- `Percentage` Percentage of people that selected this answer.

```{r, message=FALSE, warning=FALSE}

library(scales)
library(ggplot2)
library(ggthemes)
library(data.table)
setwd("/Users/ethen/Business-Analytics/ggplot2/avoid_pie_charts")
d <- fread("pie.txt")
d <- melt( d, id.vars = "Task", 
		   variable.name = "Hours", value.name = "Percentage" )
head(d)

```

## Plots

**Alternative 1: Bar Plot**

Rotates the x axis's text and label the exact percentages as text right on the graph, which can be useful when you want to pick out a specific number to emphasize on it.

```{r, fig.height=7}

p1 <- ggplot( d, aes( x = Hours, y = Percentage ) ) + 
	  geom_bar( stat = "identity" ) + 
	  facet_wrap( ~ Task ) + 
	  xlab("Hours spent per week") + 
	  geom_text( aes( label = paste0( Percentage, "%" ), y = Percentage ),
			     vjust = 1.4, size = 3, color = "white" )
p1 + theme_bw() + 
theme( axis.text.x = element_text( angle = 90,  hjust = 1 ) )

```

**Alternative 2: Bar Plot**

Using the `theme_tufte` from the `ggthemes` package. The theme will drop all the borders, grids, and axis lines to maximize the data / ink ratio.

```{r, fig.height=7}

p1 + theme_tufte()

```

**Alternative 3: Stacked Bar Plot**

```{r}

ggplot( d, aes( x = Task, y = Percentage, fill = Hours ) ) + 
geom_bar( stat = "identity", position = "stack" ) +
coord_flip() +
scale_fill_brewer( palette = "YlGnBu" ) +
theme_minimal() + theme( legend.position = "bottom" )

```

**Alternative 4: Refined Version of Bar Plot**

Some notes on ggplot2's grammar.

- geom_bar's `size` controls the bar's border size.
- `expand` is a numeric vector of length two giving multiplicative and additive expansion constants. These constants ensure that the data is placed some distance away from the axes.
- `strip .background / .text` controls the title section for each facet.
- `panel.grid .minor / .major`  controls the grid lines in the plot.
- `panel.margin` contols the margin between facets.
- `as_labeller` the function that changes the strip's label without doesn't changing the underlying data. All you have to do is create a named character vector with the new label mapping to the original label name and call the function and pass it in to `facet_grid`'s `labeller` argument.

The amount of time spent on various tasks by surveyed non-managers in data-science positions. ( could be added to the plot as subtitles ). Although this isn’t actually going to tell us which tasks data scientists spend the most time on: we should do some kind of weighted measure to estimate the mean.

```{r, fig.height=7}

# refined x label and strip label
x_labels <- c( "<1 hr/\nweek", "1-4 hrs/\nweek", "1-3 hrs/\nday", "4+ hrs/\nday" )
label_names <- c( "Basic exploratory data analysis" = "Basic Exploratory\nData Analysis", 
				  "Data cleaning" = "Data\nCleaning", 
				  "Machine learning/statistics" = "Machine Learning,\nStatistics", 
				  "Creating visualizations" = "Creating\nVisualizations", 
				  "Presenting analysis" = "Presenting\nAnalysis", 
				  "Extract/transform/load" = "Extract,\nTransform, Load" )

ggplot( d, aes( x = Hours, y = Percentage / 100, fill = Hours ) ) +
geom_bar( stat = "identity", width = 0.75, color = "#2b2b2b", size = 0.05 ) + 
scale_y_continuous( labels = percent, limits = c( 0, 0.5 ) ) + 
scale_x_discrete( expand = c( 0, 1 ), labels = x_labels ) + 
scale_fill_manual( values = c( "#a6cdd9", "#d2e4ee", "#b7b079", "#efc750" ) ) +
facet_wrap( ~ Task, labeller = as_labeller(label_names) ) + 
labs( x = NULL, y = NULL, title = "Where Does the Time Go?" ) +
theme( strip.text = element_text( size = 12, color = "white", hjust = 0.5 ),
	   strip.background = element_rect( fill = "#858585", color = NA ),	   
	   panel.background = element_rect( fill = "#efefef", color = NA ),
	   panel.grid.major.x = element_blank(),
	   panel.grid.minor.x = element_blank(),
	   panel.grid.minor.y = element_blank(),
	   panel.grid.major.y = element_line( color = "#b2b2b2" ),
	   panel.margin.x = unit( 1, "cm" ),
	   panel.margin.y = unit( 0.5, "cm" ),
	   legend.position = "none" )

```

## Reference

1. Worst Chart In The World: http://www.businessinsider.com/pie-charts-are-the-worst-2013-6
2. How to replace a pie chart: http://varianceexplained.org/r/improving-pie-chart/
3. Take the Pies out of Pi Day: http://rud.is/b/2016/03/14/spinning-cycles-in-box-4-to-take-the-pies-out-of-pi-day/

# Next Chart

### 2016-03-16 3_AB_test/AB_test.R

@@ -106,29 +106,18 @@ SanityCheck <- function( group1, group2 )
 {
 	n <- group1 + group2
 	confidence <- n * 0.5 + c( -1, 1 ) * 1.96 * sqrt( n * 0.5 * 0.5 ) 
-	return( confidence )
+	return(confidence)
 }
 ( sanity <- SanityCheck( group1, group2 ) )
 
 
-
-
-
 # ----------------------------------------------------------------------------
-
 # rule of thumb
 # http://www.exp-platform.com/Documents/2014%20experimentersRulesOfThumb.pdf
 
 
-
 # -----------------------------------------------------------------------------
-
-
 # python a-b testing, bayesian 
 # http://blog.dominodatalab.com/ab-testing-with-hierarchical-models-in-python/
-# https://github.com/CamDavidsonPilon/Probabilistic-Programming-and-Bayesian-Methods-for-Hackers
 # https://econsultancy.com/blog/65755-using-data-science-with-a-b-tests-bayesian-analysis/
-
-
-
 

### 2016-03-16 ggplot2/avoid_pie_charts/pie.R

@@ -1,0 +1,78 @@
+# Avoiding Pie Charts
+library(scales)
+library(ggplot2)
+library(ggthemes)
+library(data.table)
+setwd("/Users/ethen/Business-Analytics/ggplot2/avoid_pie_charts")
+d <- fread("pie.txt")
+d <- melt( d, id.vars = "Task", 
+		   variable.name = "Hours", value.name = "Percentage" )
+
+
+p1 <- ggplot( d, aes( x = Hours, y = Percentage ) ) + 
+	  geom_bar( stat = "identity" ) + 
+	  facet_wrap( ~ Task ) + 
+	  xlab("Hours spent per week") + 
+	  geom_text( aes( label = paste0( Percentage, "%" ), y = Percentage ),
+			     vjust = 1.4, size = 3, color = "white" )
+p1 + theme_bw() + 
+theme( axis.text.x = element_text( angle = 90,  hjust = 1 ) )
+
+
+# remove chart junks using the theme tufte
+p1 + theme_tufte()
+
+
+# stacked bar plot 
+ggplot( d, aes( x = Task, y = Percentage, fill = Hours ) ) + 
+geom_bar( stat = "identity", position = "stack" ) +
+coord_flip() +
+scale_fill_brewer( palette = "YlGnBu" ) +
+theme_minimal() + theme( legend.position = "bottom" )
+
+ 
+# refined version of bar plot
+
+# 1. geom_bar's size controls the bar's border size
+# 2. expand:
+# A numeric vector of length two giving multiplicative and 
+# additive expansion constants. These constants ensure that the data is placed 
+# some distance away from the axes
+# 3. strip .background / .text: controls the title section for each facet
+# 4. panel.grid .minor / .major: controls the grid lines in the plot
+# 5. panel.margin: contols the margin between facets
+
+# http://stackoverflow.com/questions/3472980/ggplot-how-to-change-facet-labels
+x_labels <- c( "<1 hr/\nweek", "1-4 hrs/\nweek", "1-3 hrs/\nday", "4+ hrs/\nday" )
+label_names <- c("Basic exploratory data analysis" = "Basic Exploratory\nData Analysis", 
+				"Data cleaning" = "Data\nCleaning", 
+				"Machine learning/statistics" = "Machine Learning,\nStatistics", 
+				"Creating visualizations" = "Creating\nVisualizations", 
[…]

### 2016-03-16 visualization/visualization.Rmd
visualization collection

--- 2016-03-16 ggplot2/plot.Rmd
@@
 ---
-title: "Collection of ggplot2 Plots"
+title: "Collection of Visualization"
 date: "`r Sys.Date()`"
@@
 ---
+
+Articles related to producing informative visualization.
 
@@
 
-<img src = "/Users/ethen/Business-Analytics/ggplot2/avoid_pie_charts/data_scientist.png" width = "400" height = "800" >
+<img src = "/Users/ethen/Business-Analytics/visualization/avoid_pie_charts/data_scientist.png" width = "400" height = "800" >
 
@@
 library(data.table)
-setwd("/Users/ethen/Business-Analytics/ggplot2/avoid_pie_charts")
+setwd("/Users/ethen/Business-Analytics/visualization/avoid_pie_charts")
 d <- fread("pie.txt")
@@
 
+**Sources:** source code can be obtained [here](https://github.com/ethen8181/Business-Analytics/blob/master/ggplot2/avoid_pie_charts/pie.R).
+
 ## Reference

### 2016-03-16 visualization/avoid_pie_charts/pie.R

@@ -32,7 +32,6 @@ theme_minimal() + theme( legend.position = "bottom" )
 
  
 # refined version of bar plot
-
 # 1. geom_bar's size controls the bar's border size
 # 2. expand:
 # A numeric vector of length two giving multiplicative and 
@@ -69,10 +68,4 @@ theme( strip.text = element_text( size = 12, color = "white", hjust = 0.5 ),
 	   panel.margin.x = unit( 1, "cm" ),
 	   panel.margin.y = unit( 0.5, "cm" ),
 	   legend.position = "none" ) 
-
-
- 
-dat <- gather(dat, amount, value, -Task)
-dat$value <- dat$value / 100
-dat$amount <- factor(amount_trans[dat$amount], levels=amount_trans)
  

### 2016-04-01 articles/articles.Rmd
cohort analysis

--- 2016-02-16 articles/articles.Rmd
@@
 author: "Ming-Yu Liu"
-output: html_document
+output: 
+  html_document: 
+    highlight: haddock
+    toc: yes
 ---
 
+# Interviews and Careers
+
+- Land any job you want. [[blog](http://landanyjobyouwant.com/)]
+- The career advice no one tells you. [[Quartz](http://qz.com/640112/the-career-advice-no-one-tells-you/)]
+- Wanted in College Graduates: Tolerance for Ambiguity. [[Linkedin Pulse](https://www.linkedin.com/pulse/wanted-college-graduates-tolerance-ambiguity-jeff-selingo)]
+- How to set goals: OKRs. [[Youtube](https://www.youtube.com/watch?v=mJB83EZtAjc)]
+- Meetings that don’t suck. [[Youtube](https://www.youtube.com/watch?v=rU8nv38E0Qk)]
+
+# Analytics
+
+- Principles of good data analytics. [[blog](http://www.gregreda.com/2014/03/23/principles-of-good-data-analysis/)] 
+- Why I Ditched My Fitbit and What This Means for Analytics. [[Data Science Central](http://www.datasciencecentral.com/profiles/blogs/why-i-ditched-my-fitbit-and-what-this-means-for-analytics)]
+- Know the Difference between Customer Service and Customer Experience.
+[[Harvard Business Review](https://hbr.org/sponsored/2016/03/know-the-difference-between-customer-service-and-customer-experience?utm_source=twitter&utm_medium=social&utm_campaign=harvardbiz)]
+- Two articles about saying no to p-values.
+[[R-bloggers](http://www.r-bloggers.com/after-150-years-the-asa-says-no-to-p-values/)][[R-bloggers](http://www.r-bloggers.com/its-not-the-p-values-fault-reflections-on-the-recent-asa-statement-relevant-r-resources/)]
+- The 9 Best Marketing And Sales Analytics You Should Know About.[[Data Science Central](http://www.datasciencecentral.com/profiles/blogs/the-9-best-marketing-and-sales-analytics-you-should-know-about)]
+- Cohort analysis. [[Youtube](https://www.youtube.com/watch?v=NyhVdGmnh0I)]
+- Ted Talk : How to use data to make a hit TV show.
+[[Ted](http://www.ted.com/talks/sebastian_wernicke_how_to_use_data_to_make_a_hit_tv_show?utm_campaign=social&utm_medium=referral&utm_source=facebook.com&utm_content=talk&utm_term=technology)]
+
+# Business
+
+- Facebook is the new Excel. [[blog](http://alexmuir.com/facebook-is-the-new-excel)]
 - We need companies to start innovating instead of imitating success stories. [[Hustle](http://thehustle.co/snapchat-vs-spacex-battle-of-the-imitators-and-innovators)]
@@
 [[Linkedin](https://www.linkedin.com/pulse/three-things-i-learned-during-4-years-uber-michael-pao)]
-- Cohort analysis. [[Youtube](https://www.youtube.com/watch?v=NyhVdGmnh0I)]
-- How to set goals: OKRs. [[Youtube](https://www.youtube.com/watch?v=mJB83EZtAjc)]
-- Meetings that don’t suck. [[Youtube](https://www.youtube.com/watch?v=rU8nv38E0Qk)]
-- Ted Talk : How to use data to make a hit TV show.
-[[Ted](http://www.ted.com/talks/sebastian_wernicke_how_to_use_data_to_make_a_hit_tv_show?utm_campaign=social&utm_medium=referral&utm_source=facebook.com&utm_content=talk&utm_term=technology)]

### 2016-04-01 bayesian_statistics/bayesian_statistics.Rmd
cohort analysis

---
title: "Bayesian Statistics"
author: "Ming-Yu Liu"
date: "`r Sys.Date()`"
output:
  rmdformats::readthedown:
    highlight: pygments
---

# Beta Distribution

The **beta distribution** can be understood as representing a probability distribution of probabilities. That is, it represents all the possible values of a probability when we don’t know what that probability is. Let's consider the following example:

## Prior & Posterior

Anyone who follows baseball is familiar with batting averages - simply the number of times a player gets a base hit divided by the number of times he goes up at bat (so it’s just a percentage between 0 and 1). .266 is in general considered an average batting average, while .300 is considered an excellent one.

Imagine we have a baseball player, and we want to predict what his season-long batting average will be. You might say we can just use his batting average so far, but this will be a very poor measure at the start of a season! If a player goes up to bat once and gets a single, his batting average is briefly 1.000, while if he strikes out or walks, his batting average is 0.000. It doesn’t get much better if you go up to bat for another five or six times, you could get a lucky streak and get an average of 1.000, or an unlucky streak and get an average of 0, neither of which are a remotely good predictor of how you will bat that season.

Why is your batting average in the first few hits not a good predictor of your eventual batting average? When a player’s first at-bat is a strikeout, why do we predict that he’ll never get a hit all season? Because we’re going in with **prior expectations**. We know that in history, most batting averages over a season have hovered between something like .215 and .360, with some extremely rare exceptions on either side. We know that if a player gets a few strikeouts in a row at the start, that might indicate he’ll end up a bit worse than average, but we know he probably won’t deviate from that range.

Given our batting average problem, which can be represented with a binomial distribution (a series of successes and failures), the best way to represent these prior is with the **beta distribution**. It's saying, before we've seen the player take his first swing, what we roughly expect his batting average to be. The domain of the beta distribution is $(0, 1)$, just like a probability, so we already know we're on the right track, but the appropriateness of the beta for this task goes far beyond that.

We expect that the player's season-long batting average will be most likely around .27, but that it could reasonably range from .21 to .35. This can be represented with a **beta distribution** with parameters $\alpha=81$ and $\beta=219$. The two parameters were chosen based upon the fact that:

- The mean of the **beta distribution** is $\frac{\alpha}{\alpha+\beta}=\frac{81}{81+219}=.270$
- As you can see in the plot below, this distribution lies almost entirely within $(.2, .35)$, the reasonable range for a batting average.

```{r, message=FALSE, warning=FALSE}

library(ggplot2)
library(dplyr)

# synthesized data
# "\u03B1" unicode for the greek letter alpha
sim <- data.frame( a = c( 81, 82, 81 + 100 ),
				   b = c( 219, 219, 219 + 200 ) ) %>%
	   group_by( a, b ) %>%
	   do( data_frame( x = seq( 0, 1, by = .001 ), y = dbeta( x, .$a, .$b ) ) ) %>%
	   mutate( Parameters = paste0( "\u03B1 = ", a, ", \u03B2 = ", b ) ) %>%
	   ungroup() %>%
	   mutate( Parameters = factor( Parameters, levels = unique(Parameters) ) )

# plot of the distribution
PlotBeta <- function(sim)
{
	ggplot( sim, aes( x, y, color = Parameters ) ) + geom_line() +
	xlim( 0, .5 ) + ylab("Density of beta") + theme_bw()
}
PlotBeta( sim = filter( sim, a == 81 ) )

```

In the preceding plot, the x axis in a **beta distribution** density plot represents a player's batting average. Thus notice that in this case, not only is the y-axis a probability (or more precisely a probability density), but the x-axis is as well (batting average is just a probability of a hit, after all)! The **beta distribution** is representing a probability distribution of probabilities.

But here's why the **beta distribution** is so appropriate. Imagine the player gets a single hit. His record for the season is now "1 hit; 1 at bat." We have to then update our probabilities, we want to shift this entire curve over just a bit to reflect our new information. The new **beta distribution** will then become:

$$beta(\alpha_0 + hits, \beta_0 + misses)$$

Where $\alpha_0$ and $\beta_0$ are the parameters we started with, that is 81 and 219. Thus, in this case, $\alpha$ has increased by 1 (his one hit), while $\beta$ has not increased at all (no misses yet). That means our new distribution is $\beta(81+1, 219)$. Let's compare that to the original:

```{r, warning=FALSE}

# update 1 hit of 1 bat
PlotBeta( sim = filter( sim, a == 81 | a == 82 ) )

```

Notice that it has barely changed at all, that's because one extra hit doesn't really mean anything yet. However, the more the player hits over the course of the season, the more the curve will shift to accommodate the new evidence, and the furthermore more it will narrow based on the fact that we have more proof. Let's say halfway through the season he has been up to bat 300 times, hitting 100 out of those times. The new distribution would be $\beta(81 + 100, 219 + 200)$:

```{r, warning=FALSE}

# update 100 hit of 300 bat
PlotBeta( sim = sim )

```

Notice the curve is now both thinner and shifted to the right (higher batting average) than it used to be, since we now have a better sense of what the player's batting average is most likely to be.

One of the most interesting outputs of this formula is the expected value of the resulting **beta distribution**, which is basically your new estimate. Recall that the expected value of the **beta distribution** is $\frac{\alpha}{\alpha+\beta}$. Thus, after 100 hits of 300 real at-bats, the expected value of the new **beta distribution** is $\frac{82+100}{82+100+219+200}=.303$. Notice that it is lower than the naive estimate of $\frac{100}{100+200}=.333$, but higher than the estimate you started the season with ($\frac{81}{81+219}=.270$). Thus you can think of this formula is equivalent to adding a "head start" to the number of hits and non-hits of a player. You're saying "start him off in the season with 81 hits and 219 non hits on his record".

## Takeaways

- **Beta distribution** is only defined for values of x in the interval [0,1] and it is best for representing a probabilistic distribution of probabilities. Cases where we don’t know what a probability is in advance, but we have some reasonable guesses.
- A quick side note on the parameters for the **beta distribution**. As $\alpha$ gets bigger, the bulk of the distribution moves rightward over higher values of x, but as $\beta$ gets bigger, the bulk of the distribution moves leftward over lower values of x. And as $\alpha$ and $\beta$ get bigger together, the **beta distribution** gets narrower.

## Reference

1. Understanding the beta distribution http://varianceexplained.org/statistics/beta_distribution_and_baseball/

### 2016-04-01 2_marketing_analysis/code/marketing_analysis.R

@@ -151,7 +151,6 @@ revenue_predicted[ is.na(revenue), revenue := 0 ]
 revenue_summary <- revenue_predicted[ , .( avg_revenue 	 = mean(revenue),
 								  		   total_revenue = sum(revenue) ), by = segment ] %>%
 				   arrange( desc(total_revenue), desc( avg_revenue) )
-
 revenue_summary
 
 # ----------------------------------------------------------------------------

### 2016-04-01 bayesian_statistics/beta_distribution.R

@@ -1,0 +1,28 @@
+library(dplyr)
+library(ggplot2)
+
+# simulated data
+# "\u03B1" unicode for the greek letter alpha
+sim <- data.frame( a = c( 81, 82, 81 + 100 ),
+				   b = c( 219, 219, 219 + 200 ) ) %>%
+	   group_by( a, b ) %>%
+	   do( data_frame( x = seq( 0, 1, by = .001 ), y = dbeta( x, .$a, .$b ) ) ) %>%
+	   mutate( Parameters = paste0( "\u03B1 = ", a, ", \u03B2 = ", b ) ) %>%
+	   ungroup() %>%
+	   mutate( Parameters = factor( Parameters, levels = unique(Parameters) ) )
+
+# plot of the distribution
+PlotBeta <- function(sim)
+{
+	ggplot( sim, aes( x, y, color = Parameters ) ) + geom_line() +
+	xlim( 0, .5 ) + ylab("Density of beta") + theme_bw()
+}
+PlotBeta( sim = filter( sim, a == 81 ) )
+
+
+# update 1 hit of 1 bat
+PlotBeta( sim = filter( sim, a == 81 | a == 82 ) )
+
+
+# update 100 hit of 300 bat
+PlotBeta( sim = sim )

### 2016-04-01 notebook_format/notebook_format.py

@@ -1,0 +1,99 @@
+"""
+# -*- coding: utf-8 -*-
+
+"""
+
+from __future__ import (absolute_import, division, print_function,
+                        unicode_literals)
+
+from contextlib import contextmanager
+from IPython.core.display import HTML
+import json
+import matplotlib.pylab as pylab
+import matplotlib.pyplot as plt
+import numpy as np
+import os.path
+import sys
+import warnings
+
+# recent matplotlibs are raising deprecation warnings that
+# we don't worry about.
+warnings.filterwarnings("ignore")
+
+
+# sys.path.insert(0, './style') # allow us to import book_format
+
+
+def equal_axis():
+    pylab.rcParams['figure.figsize'] = 10,10
+    plt.axis('equal')
+
+
+def reset_axis():
+    pylab.rcParams['figure.figsize'] = 11, 4
+
+def set_figsize(x=11, y=4):
+    pylab.rcParams['figure.figsize'] = x, y
+
+
+@contextmanager
+def figsize(x=11, y=4):
+    """Temporarily set the figure size using 'with figsize(a,b):'"""
+
+    size = pylab.rcParams['figure.figsize']
+    set_figsize(x, y)
+    yield
+    pylab.rcParams['figure.figsize'] = size
+
+@contextmanager
+def numpy_precision(precision):
+	old = np.get_printoptions()['precision']
+	np.set_printoptions(precision=precision)
+	yield
+	np.set_printoptions(old)
+
+@contextmanager
+def printoptions(*args, **kwargs):
+    original = np.get_printoptions()
+    np.set_printoptions(*args, **kwargs)
+    yield
+    np.set_printoptions(**original)
+
+def _decode_list(data):
+    rv = []
+    for item in data:
+        if isinstance(item, unicode):
+            item = item.encode('utf-8')
+        elif isinstance(item, list):
+            item = _decode_list(item)
+        elif isinstance(item, dict):
+            item = _decode_dict(item)
+        rv.append(item)
+    return rv
+
+def _decode_dict(data):
+    rv = {}
+    for key, value in data.items():
+        if isinstance(key, unicode):
+            key = key.encode('utf-8')
+        if isinstance(value, unicode):
+            value = value.encode('utf-8')
+        elif isinstance(value, list):
+            value = _decode_list(value)
+        elif isinstance(value, dict):
+            value = _decode_dict(value)
+        rv[key] = value
+    return rv
+
+
[…]

### 2016-04-01 visualization/avoid_pie_charts/pie.R

@@ -8,7 +8,7 @@ d <- fread("pie.txt")
 d <- melt( d, id.vars = "Task", 
 		   variable.name = "Hours", value.name = "Percentage" )
 
-
+# faceted bar plot, x label 90 degrees rotation
 p1 <- ggplot( d, aes( x = Hours, y = Percentage ) ) + 
 	  geom_bar( stat = "identity" ) + 
 	  facet_wrap( ~ Task ) + 
@@ -34,9 +34,9 @@ theme_minimal() + theme( legend.position = "bottom" )
 # refined version of bar plot
 # 1. geom_bar's size controls the bar's border size
 # 2. expand:
-# A numeric vector of length two giving multiplicative and 
-# additive expansion constants. These constants ensure that the data is placed 
-# some distance away from the axes
+# 	 A numeric vector of length two giving multiplicative and 
+# 	 additive expansion constants. These constants ensure that the data is placed 
+# 	 some distance away from the axes
 # 3. strip .background / .text: controls the title section for each facet
 # 4. panel.grid .minor / .major: controls the grid lines in the plot
 # 5. panel.margin: contols the margin between facets
@@ -50,14 +50,14 @@ label_names <- c("Basic exploratory data analysis" = "Basic Exploratory\nData An
 				"Presenting analysis" = "Presenting\nAnalysis", 
 				"Extract/transform/load" = "Extract,\nTransform, Load" )
 
+# The amount of time spent on various tasks by surveyed non-managers in data-science positions.
 ggplot( d, aes( x = Hours, y = Percentage / 100, fill = Hours ) ) +
 geom_bar( stat = "identity", width = 0.75, color = "#2b2b2b", size = 0.05 ) + 
 scale_y_continuous( labels = percent, limits = c( 0, 0.5 ) ) + 
 scale_x_discrete( expand = c( 0, 1 ), labels = x_labels ) + 
 scale_fill_manual( values = c( "#a6cdd9", "#d2e4ee", "#b7b079", "#efc750" ) ) +
 facet_wrap( ~ Task, labeller = as_labeller(label_names) ) + 
-labs( x = NULL, y = NULL, 
-	  title = "Where Does the Time Go?\nThe amount of time spent on various tasks by surveyed non-managers in data-science positions.") +
+labs( x = NULL, y = NULL, title = "Where Does the Time Go?") +
 theme( strip.text = element_text( size = 12, color = "white", hjust = 0.5 ),
[…]

### 2016-04-05 finding_groups/finding_groups.Rmd
radar plot, notebook format

--- 2016-02-15 1_finding_groups/finding_groups.Rmd
@@
 ---
-title: "Business Analytics 1 : Finding Groups Within the Data"
+title: "Business Analytics - Finding Groups Within the Data"
 author: "Ming-Yu Liu"
-date: "November 11, 2015"
+date: "`r Sys.Date()`"
 output:
-  html_document: 
-    number_sections: yes
-    theme: united
-    toc: yes
+  rmdformats::readthedown:
+    highlight: pygments
 ---
@@
 
-Let's read in the dataset to get a feel of what it looks like, before going further. Oh, and the link to the datasets used in this documentation is located at the bottom of the report.
+Let's read in the dataset to get a feel of what it looks like, before going further. Oh, and the link to the datasets used in this documentation is located at the bottom of the documentation.
 
-```{r, message=FALSE}
+```{r, message=FALSE, warning=FALSE}
+library(caret)
+library(scales)
+library(ggradar)
+library(ggplot2)
+library(data.table)
 
-# environment setting
-library(dplyr)
-library(caret)
-library(ggplot2)
-library(reshape2)
-setwd("/Users/ethen/Business-Analytics/1_finding_groups")
+setwd("/Users/ethen/Business-Analytics/finding_groups")
 rm( list = ls() )
 
-data_sku <- read.csv("data/DATA_2.01_SKU.csv")
+# ----------------------------------------------------------
+#              example 1 : supply chain
+# ----------------------------------------------------------
+data_sku <- fread("data/DATA_2.01_SKU.csv")
 head(data_sku)
@@
 ggplot( data_sku, aes( CV, ADS ) ) + geom_point() +  
-labs( title = "SKU Example", y = "Average Daily Sales", x = "Coefficient of Variation" ) + 
 geom_hline( yintercept = 4, color = "blue" ) + 
-geom_vline( xintercept = .2, color = "blue" )
+geom_vline( xintercept = .2, color = "blue" ) + 
+labs( x = "Coefficient of Variation", y = "Average Daily Sales", 
+	  title = "SKU Example" )
 
@@
 cluster <- hclust( d, method = "ward.D" )
-( data_sku$groups  <- as.factor( cutree( cluster, k = 3 ) ) )
+data_sku[ ,  groups := as.factor( cutree( cluster, k = 3 ) ) ]
+head(data_sku)
 
@@
 ggplot( data_sku, aes( CV, ADS, color = groups ) ) + geom_point() + 
-labs( title = "SKU Example", y = "Average Daily Sales", x = "Coefficient of Variation" )
+labs( x = "Coefficient of Variation", y = "Average Daily Sales", 
+	  title = "SKU Example" )
 
@@
 
-data_hr <- read.csv("data/DATA_2.02_HR.csv")
-str(data_hr)
+data_hr <- fread("data/DATA_2.02_HR.csv")
+head(data_hr)
 
@@
 # this is used for latter plotting and applying clustering
-hr_scaled <- data.frame( scale(data_hr) )
+scaled <- scale(data_hr)
+data_hr_scaled <- data.table(scaled)
 
 # find variables that have correlations higher than 0.8 and print out variable names
-findCorrelation( cor(hr_scaled), cutoff = 0.8, names = TRUE )
+findCorrelation( cor(data_hr_scaled), cutoff = 0.8, names = TRUE )
 
@@
 
-```{r, message=FALSE}
+```{r, message=FALSE, warning=FALSE}
 
-ggplot( melt( hr_scaled[ , 2:3 ] ), aes( value, fill = variable ) ) + 
+ggplot( melt( data_hr_scaled[ , c( "LPE", "NP" ), with = FALSE ] ), 
+		aes( value, fill = variable ) ) + 
 geom_histogram( alpha = .4, position = "identity" )
@@
 
-d <- dist( hr_scaled[ , -1 ], method = "euclidean" ) 
+# clustering 
+d <- dist( data_hr_scaled[ , -1, with = FALSE ], method = "euclidean" ) 
 cluster <- hclust( d, method = "ward.D" )
-# dendogram
-# plot(cluster)
-
-data_hr$groups <- cutree( cluster, k = 4 )
+# plot(cluster) # dendogram
+data_hr_scaled[ ,  groups := as.factor( cutree( cluster, k = 4 ) ) ]
 
 # median aggregation 
-hr_agg <- aggregate( .~ groups, data = data_hr, FUN = median )
+hr_agg <- data_hr_scaled[ , lapply( .SD, median ), by = groups ]
 
-# count the cluster proportion size 
-tl <- table(data_hr$groups)
-hr_agg$proportion <- tl / sum(tl)
-
-# order by proportion 
-mutate( hr_agg, proportion = tl / sum(tl) ) %>%
-arrange( desc(proportion) )
+# order by cluster's proportion size
+t <- table(data_hr_scaled$groups)
+hr_agg[ , proportion := t / sum(t) ][ order(-proportion) ]
 
@@
 When presenting the upper result table, one visualization you can use is the radar chart.
-However since the current version of the ggplot2 library still does not support actual radar chart with a simple function call (as least not to my knowledge) and each variable of the data is recorded with different scales.
-Therefore, the visualization after using the max-min normalization method on the recorded score may look a bit hacky ( the `scale` function produces negative numbers and that does not sound very intuitive for radar charts, thus a different kind of normalization is used here ).
-As with this normalization, there will be values of 0.
+
+- To use the `ggradar` function you will need some extra fonts for it to work, see instructions [here](https://github.com/ricardo-bion/ggradar) to get it.
+Currently, it does not support facetting and the first column has to be the group (future work).
+- Each variable of the data is recorded with different scales.
+And normalizing them via the `scale` function produces negative numbers and that does not sound very intuitive for radar charts, thus a different kind of normalization is used here, namely the max-min normalization.
+With this method, you minus each column by its min value and divide it by the difference between the maximum and minimum value `( x - min(x) ) / ( max(x) - min(x) )`.
+Though the recorded score may look a bit odd as with this normalization, there will be values of 0.
 
@@
 
-Normalize <- function(x) ( x - min(x) ) / ( max(x) - min(x) )
+hr_agg2 <- hr_agg[ , -c( "groups", "proportion" ), with = FALSE ]
+hr_agg2 <- hr_agg2[ , lapply( .SD, rescale ) ]
+hr_agg2[ , groups := as.factor( paste0( "group", 1:nrow(hr_agg2) ) ) ]
+name <- colnames(hr_agg2)
+setcolorder( hr_agg2, c( "groups", name[ name != "groups" ] ) )
 
-# exclude the proportion, groups and store the groups
-hr_agg_2 <- hr_agg[ , c( -1, -8 ) ]
-groups <- hr_agg$groups
-
-# transform to long format for plotting
-hr_agg_melt <- apply( hr_agg_2, 2, Normalize ) %>%
-			   data.frame() %>%
-			   mutate( groups = groups ) %>%
-               melt( id.vars = "groups" )
-
-# radar chart
-ggplot( hr_agg_melt, aes( x = variable  ) ) + 
-geom_bar( aes( y = value, fill = groups ), stat = "identity" ) + 
-coord_polar() + 
-facet_wrap( ~ groups ) +
-ggtitle( "HR" ) + 
-theme( legend.position = "none",  
-	   axis.text.y = element_blank(), 
-	   axis.text.x = element_text( color = "black" ), 
-	   axis.ticks  = element_blank(),
-	   axis.title  = element_blank() )
+# radar plot
+ggradar(hr_agg2)
 
@@
 
-**Takeaways:**
+## Takeaways
 
@@
 This wraps up our example of where clustering algorithms can be used to improve your businesses efficiency.
-You can find all the code (finding_groups.R) and datasets (in the data folder) used in this documentation [here](https://github.com/ethen8181/Business-Analytics/tree/master/1_finding_groups).
+You can find all the code (finding_groups.R) and datasets (in the data folder) used in this documentation [here](https://github.com/ethen8181/Business-Analytics/tree/master/finding_groups).
 

### 2016-04-05 1_finding_groups/finding_groups.R

@@ -1,116 +1,0 @@
-#### finding similar groups
-
-# environment setting
-library(caret)
-library(dplyr)
-library(ggplot2)
-library(reshape2)
-
-setwd("/Users/ethen/Business-Analytics/1_finding_groups/data")
-rm( list = ls() )
-
-# ----------------------------------------------------------
-#              example 1 : supply chain
-# ----------------------------------------------------------
-
-data_sku <- read.csv("DATA_2.01_SKU.csv")
-head(data_sku)
-
-# figure 1
-ggplot( data_sku, aes( CV, ADS ) ) + 
-geom_point() + 
-labs( title = "SKU Example", y = "Average Daily Sales", x = "Coefficient of Variation" ) + 
-geom_hline( yintercept = 4, color = "blue" ) + 
-geom_vline( xintercept = .2, color = "blue" )
-
-# hierarchical clustering 
-d <- dist( scale(data_sku), method = "euclidean" )
-cluster <- hclust( d, method = "ward.D" )
-data_sku$groups  <- as.factor( cutree( cluster, k = 3 ) )
-
-# figure 2
-ggplot( data_sku, aes( CV, ADS, color = groups ) ) + 
-geom_point() + 
-labs( title = "SKU Example", y = "Average Daily Sales", x = "Coefficient of Variation" )
-
-# ----------------------------------------------------------
-#              example 2 : HR
-# ----------------------------------------------------------
-data_hr <- read.csv("DATA_2.02_HR.csv")
-str(data_hr)
-
-# correlation, method 1
-# convert the correlation of oneself to 0
-# look that which index is highly correlated, or i.e. often appears together
-M <- abs( cor(data_hr) )
-diag(M) <- 0 
-which( M > 0.8, arr.ind = TRUE )
-
-# correlation, method 2
-hr_scaled <- data.frame( scale(data_hr) )
-findCorrelation( cor(hr_scaled), cutoff = 0.8, names = TRUE )
-
-
-# normality plot 
-ggplot( melt( hr_scaled[ , 2:3 ] ), aes( value, fill = variable ) ) + 
-geom_histogram( alpha = .4, position = "identity" )
-
-
-# clustering 
-d <- dist( hr_scaled[ , -1 ], method = "euclidean" ) 
-cluster <- hclust( d, method = "ward.D" )
-plot(cluster)
-data_hr$groups <- cutree( cluster, k = 4 )
-
-# aggregation 
[…]

### 2016-04-05 finding_groups/finding_groups.R

@@ -1,0 +1,119 @@
+# finding similar groups
+
+# environment setting
+library(caret)
+library(scales)
+library(ggradar)
+library(ggplot2)
+library(data.table)
+rm( list = ls() )
+setwd("/Users/ethen/Business-Analytics/finding_groups")
+
+
+# ----------------------------------------------------------
+#              example 1 : supply chain
+# ----------------------------------------------------------
+data_sku <- fread("data/DATA_2.01_SKU.csv")
+head(data_sku)
+
+# figure 1
+ggplot( data_sku, aes( CV, ADS ) ) + geom_point() +  
+geom_hline( yintercept = 4, color = "blue" ) + 
+geom_vline( xintercept = .2, color = "blue" ) + 
+labs( x = "Coefficient of Variation", y = "Average Daily Sales", 
+	  title = "SKU Example" )
+
+# hierarchical clustering, with k (clustering number) = 3 
+d <- dist( scale(data_sku), method = "euclidean" )
+cluster <- hclust( d, method = "ward.D" )
+data_sku[ ,  groups := as.factor( cutree( cluster, k = 3 ) ) ]
+
+# figure 2
+ggplot( data_sku, aes( CV, ADS, color = groups ) ) + geom_point() + 
+labs( x = "Coefficient of Variation", y = "Average Daily Sales", 
+	  title = "SKU Example" )
+
+
+# ----------------------------------------------------------
+#              example 2 : HR
+# ----------------------------------------------------------
+data_hr <- fread("data/DATA_2.02_HR.csv")
+head(data_hr)
+
+# correlation, method 1
+# convert the correlation of oneself to 0
+# look that which index is highly correlated, or i.e. often appears together
+# M <- abs( cor(data_hr) )
+# diag(M) <- 0 
+# which( M > 0.8, arr.ind = TRUE )
+
+# correlation, method 2
+scaled <- scale(data_hr)
+data_hr_scaled <- data.table(scaled)
+findCorrelation( cor(data_hr_scaled), cutoff = 0.8, names = TRUE )
+
+
+# normality plot 
+ggplot( melt( data_hr_scaled[ , c( "LPE", "NP" ), with = FALSE ] ), 
+		aes( value, fill = variable ) ) + 
+geom_histogram( alpha = .4, position = "identity" )
+
+
+# clustering 
[…]

### 2016-04-05 notebook_format/formats.py

@@ -1,0 +1,29 @@
+import os
+import json
+import warnings
+import matplotlib
+from IPython.core.display import HTML
+
+"""
+def css_styling():
+    styles = open("../styles/custom.css", "r").read()
+    return HTML(styles)
+css_styling()
+
+
+s = json.load( open("../styles/bmh_matplotlibrc.json") )
+matplotlib.rcParams.update(s)
+"""
+
+def load_style(directory = ""):
+	# recent matplotlibs are raising deprecation warnings that
+	# we don't worry about.
+	warnings.filterwarnings("ignore")
+
+	# update the default matplotlib's formating
+	s = json.load( open( os.path.join( directory, "plot.json" ) ) )
+	matplotlib.rcParams.update(s)
+
+	# load the styles for the notebooks
+	styles = open( os.path.join( directory, "custom.css" ), "r" ).read()
+	return HTML(styles)

### 2016-04-05 notebook_format/notebook_format.py

@@ -1,99 +1,0 @@
-"""
-# -*- coding: utf-8 -*-
-
-"""
-
-from __future__ import (absolute_import, division, print_function,
-                        unicode_literals)
-
-from contextlib import contextmanager
-from IPython.core.display import HTML
-import json
-import matplotlib.pylab as pylab
-import matplotlib.pyplot as plt
-import numpy as np
-import os.path
-import sys
-import warnings
-
-# recent matplotlibs are raising deprecation warnings that
-# we don't worry about.
-warnings.filterwarnings("ignore")
-
-
-# sys.path.insert(0, './style') # allow us to import book_format
-
-
-def equal_axis():
-    pylab.rcParams['figure.figsize'] = 10,10
-    plt.axis('equal')
-
-
-def reset_axis():
-    pylab.rcParams['figure.figsize'] = 11, 4
-
-def set_figsize(x=11, y=4):
-    pylab.rcParams['figure.figsize'] = x, y
-
-
-@contextmanager
-def figsize(x=11, y=4):
-    """Temporarily set the figure size using 'with figsize(a,b):'"""
-
-    size = pylab.rcParams['figure.figsize']
-    set_figsize(x, y)
-    yield
-    pylab.rcParams['figure.figsize'] = size
-
-@contextmanager
-def numpy_precision(precision):
-	old = np.get_printoptions()['precision']
-	np.set_printoptions(precision=precision)
-	yield
-	np.set_printoptions(old)
-
-@contextmanager
-def printoptions(*args, **kwargs):
-    original = np.get_printoptions()
-    np.set_printoptions(*args, **kwargs)
-    yield
-    np.set_printoptions(**original)
-
-def _decode_list(data):
-    rv = []
-    for item in data:
-        if isinstance(item, unicode):
-            item = item.encode('utf-8')
-        elif isinstance(item, list):
-            item = _decode_list(item)
-        elif isinstance(item, dict):
-            item = _decode_dict(item)
-        rv.append(item)
-    return rv
-
-def _decode_dict(data):
-    rv = {}
-    for key, value in data.items():
-        if isinstance(key, unicode):
-            key = key.encode('utf-8')
-        if isinstance(value, unicode):
-            value = value.encode('utf-8')
-        elif isinstance(value, list):
-            value = _decode_list(value)
-        elif isinstance(value, dict):
-            value = _decode_dict(value)
-        rv[key] = value
-    return rv
-
-
[…]

### 2016-04-07 3_AB_test/AB_test.R
bayes ab test

@@ -115,9 +115,6 @@ SanityCheck <- function( group1, group2 )
 # rule of thumb
 # http://www.exp-platform.com/Documents/2014%20experimentersRulesOfThumb.pdf
 
-
-# -----------------------------------------------------------------------------
-# python a-b testing, bayesian 
-# http://blog.dominodatalab.com/ab-testing-with-hierarchical-models-in-python/
-# https://econsultancy.com/blog/65755-using-data-science-with-a-b-tests-bayesian-analysis/
+# determining sample size required
+# https://signalvnoise.com/posts/3004-ab-testing-tech-note-determining-sample-size
 

### 2016-04-11 bayesian_statistics/MCMC/MCMC.Rmd
mcmc

---
title: "Markov Chain Monte Carlo"
date: "`r Sys.Date()`"
author: "Ming-Yu Liu"
output:
  rmdformats::readthedown:
    highlight: pygments
---

# Learn About a Probability Distribution

Probability distributions are super-useful because they can define a range of possibilities we expect to observe if our models are true (if you're a frequentist) or the range of plausible values for our parameters (if you're a Bayesian).

Some probability distributions, like the Normal distribution and the Poisson distribution, are relatively straightforward to work with mathematically and we have lots of handy tools for working with them by hand and on the computer.

Other probability distributions over parameters in a complicated model are harder to work with. In this case we need more general tools. **Monte Carlo** methods work by taking samples from a distribution that's hard to study analytically, and then studying the samples instead.

## Example

Here's an ugly function that we could treat as a probability distribution. Note that, in addition to being wiggly, we don't actually know how to scale the y-axis so that the distribution sums to 1.

```{r, message=FALSE, warning=FALSE}

library(ggplot2)
library(ggthemes)

f <- function(x){
	( sin(.2 * x^3) + sin(x^2) / 5 + 2.5 + sin(10 * x) / 2 ) * dnorm( x, mean = 1 )
}

# boundary for the x axis
x_max <- 6
x_min <- -6

x <- seq( x_min, x_max, length.out = 500 )
y <- f(x)
data <- data.frame( x = x, y = y )

ggplot( data, aes( x, y ) ) + geom_line() + 
geom_ribbon( aes( ymin = 0, ymax = y ), alpha = 0.2 )

```

Even if you knew the equation that generated that thing, and you remembered everything you learned in calculus and statistics, you probably couldn't find its mean or its quantiles very easily. But if you had 10,000 samples from the distribution, you could estimate the distribution's mean instantly by taking the sample mean.

## Simple Monte Carlo: Rejection Sampling

To start off, let's look at a basic method called **rejection sampling**, the way it works the that we draw a box that completely surrounds our function. Then, we pick a point at random from our box. If it falls under our the area under the curve of the function, we say we accept this point and return it in the sample. If it falls above, we reject it and repeat the drawing process. We'll continue to do this again and again until you get the number of samples that you want. Let’s look at an example outcome of these drawings.

```{r, warning=FALSE, message=FALSE}

# throw N darts (uniformly distributed x-y pairs) at the plot above
N <- 10000
ceiling <- 1.5
xs <- runif( N, min = x_min, max = x_max )
ys <- runif( N, min = 0, max = ceiling )

# accept any samples that are below the function curve
accepted_samples <- xs[ ys <= f(xs) ]

ggplot( data.frame( x = accepted_samples ) ) + 
geom_histogram( aes( x = x, y = ..density.. ) ) + 
geom_line( data = data, aes( x, y ) )

# study the samples however you want
# mean(accepted_samples)
# quantile(accepted_samples, probs = seq(0, 1, length = 4 ) )

```

Intuitively, it should be clear that this works: wherever the curve is tallest, there's a higher probability that the darts will be retained.

That worked pretty well, but it was kind of inefficient, we ended up throwing away almost 90% of our sampling effort. Denoted in the following plot.

```{r}

# accepted and unaccepted points
df <- data.frame( xs = xs, ys = ys, accept = ys <= f(xs) )

# plot the accepted and unaccepted point
ggplot( df, aes( xs, ys ) ) + 
geom_point( aes( color = accept ), size = 0.5 ) + 
geom_line( data = data, aes( x, y ) )

```

Of course, this is a toy example. It turns out that you may end up throwing away exponentially more attempts per successful sample as the number of dimensions in your data goes up. So hardly anyone uses rejection sampling without a good reason -- that's where the Markov chains come in.

## Markov Chain Monte Carlo (MCMC)

Most of our discarded samples came from the tails of the distribution, where the acceptance probability is basically zero. We could increase the number of samples we keep by spending more time in the middle of the distribution, where the acceptance probabilities are much higher. But if all we did was concentrate our dart-throwing efforts there, then we'd end up with heavily biased samples. Meaning we won't find any probability mass in areas where we don't look. Also, even if we could correct for that issue, we won't always know where to concentrate our efforts. So we'll usually need something more automatic.

That's where the Markov chains come in. MCMC has exactly the same goal as our rejection sampler, but it has more sophisticated ways of deciding where to sample, which can be expressed at a high level as follows:

1. Start at current position.
2. Propose moving to a new position (investigate a point near the current position).
3. Accept/Reject the new position based on the position's adherence to the data and prior distributions (ask if the point likely came from the mountain).
4. If you accept: Move to the new position. Return to Step 1. Else: Do not move to new position. Return to Step 1.
5. After a large number of iterations, return all accepted positions.

If the current position of the MCMC algorithm is in an area of extremely low probability, which is often the case when the algorithm begins (typically at a random location in the space), the algorithm will move in positions that are likely not from the posterior but better than everything else nearby.

This way we move in the general direction towards the regions where the posterior distributions exist, and collect samples sparingly on the journey. Once we reach the posterior distribution, we can easily collect samples as they likely all belong to the posterior distribution. Notice that only the current position matters (new positions are investigated only near the current position). We can describe this property as memorylessness, i.e. the algorithm does not care how it arrived at its current position, only that it is there.

Using this algorithm, we can get a decent number of samples in a reasonable amount of time. Two things to note:

- Inference using the first few thousand points is a bad idea, as they are unrelated to the final distribution we are interested in, or in other words, the first few moves of the algorithm are not reflective of the posterior. Thus is it a good idea to discard those samples before using the samples for inference. We call this period before converge the **burn-in** period.
- The algorithm can be seen as a random "walk" around the space, thus the current position will exhibit some correlation with previous positions. This is both good and bad. We will always have correlation between current positions and the previous positions, but too much of it means we are not exploring the space well. The correlation can be solved, or at least reduced, by only returning to the user every nth sample. We call this **thinning**.

With all of that being said, let's see what happens if we try this out on a simpler distribution in 2D.

```{r}

lik <- function( x, y ){
	dnorm(x - 3) * dnorm(y - x + 2)
}

grid_values <- seq( x_min, x_max, length = 500 )
grid <- expand.grid( x = grid_values, y = grid_values )
z <- lik( grid$x, grid$y )

# prior probability plot
plot <- ggplot( data = grid, aes( x = x, y = y ) ) + 
		geom_raster( aes( fill = z ) ) + 
		scale_fill_gradient2() + coord_equal() +
		theme_tufte()
plot

```

When we setup a Bayesian inference problem with N unknowns, we are implicitly creating an N dimensional space for the **prior distributions** to exist in. Associated with the space is an additional dimension, which we can describe as the surface, or curve, that sits on top of the space, that reflects the **prior probability** of a particular point. The plots above visualize this. The more blue the color, the more **prior probability** is assigned to that location. Conversely, areas with whiter color represent that our priors assign very low probability to that location.

Let's take a random walk along X and Y, collecting samples in proportion to their probability. This is just like the rejection sampler, but we'll tend to throw our darts in the neighborhood of our last accepted proposal, and the acceptance rule is a bit more complicated.

The walk starts at some arbitrary point, specified by the user. Then the walk progresses by proposing a move to a new position in the parameter space and then deciding whether or not to accept the proposed move. The proposed move's distribution can take on many different forms, here, we'll consider the generic case in which the proposal distribution is a normal distribution, centered at the current position.

Having generated a proposed new position, the algorithm then decides whether or not to accept the proposal. The algorithm will accept the new proposal with probability:

$$p_{accept} = min \left( \frac{ p( \theta_{proposed} ) }{ p( \theta_{current} ) }, 1 \right)$$

Where $p(θ)$, simply represents the target distribution.

The next code chunck defines the `MCMC` function that draws sample points from the distribution, we'll also visualizes the path by drawing the trace of the sampling points.

```{r}

MCMC <- function( iter, burnin = 0, thin = 1 )
{
	# the list of the first argument for dimnames is to specify the row names,
	# sampling starts at a random point, here 0, 0 
	samples <- matrix( NA, nrow = iter, ncol = 2, 
					  dimnames = list( NULL, c( "x", "y" ) ) )
	samples[ 1, ] <- c( 0, 0 )

	for(i in 2:iter){

		sample <- samples[ i - 1, ]

		for(j in 1:thin){

			# propose a new sample point			
			proposal <- sample + rnorm( n = 2, mean = 0, sd = 1 )

			# compare its likelihood with the current position
			lik_old <- lik( sample["x"], sample["y"] )
			lik_new <- lik( proposal["x"], proposal["y"] )
			ratios  <- lik_new / lik_old

			# flip a coin and accept the 
			# new proposal with probability min( ratio, 1 ),
			# if you don't accept the proposal,
			# then just keep what you had in the last step,
			# meaning nothing changed.
			# note that 1 is evaluated as TRUE
			if( rbinom( 1, size = 1, prob = min( ratios, 1 ) ) )
				sample <- proposal
		}
		samples[ i, ] <- sample
	}

	return( data.frame( samples[ (burnin + 1):iter, ] )	)
}

TracePlot <- function( plot, samples ){
	plot + 
	geom_path( data = samples, aes( x, y ), color = "orange" ) + 
	geom_point( data = samples, aes( x, y ), size = 1 )
}

```

```{r}

# plot that shows the first few sampling points are not good ones
samples <- MCMC(iter = 50)
TracePlot( plot, samples )

# plot with burnin, discard the bad sampling points in the beginning
samples <- MCMC( iter = 250, burnin = 125 )
TracePlot( plot, samples )

# use thin to avoid correlation and 
# explore more of the sample space
samples <- MCMC( iter = 250, burnin = 125, thin = 2 )
TracePlot( plot, samples )

```

**End Note:**

The MCMC method is very useful, the intuition for the algorithm is that we can approximate the **posterior distribution** by generating a large sample of representative values and the larger the sample, the more accurate is our approximation.

## Reference

1. [The First Stop in Monte Carlo (Methods): Rejection Sampling](http://thexbar.me/2014/11/07/reject-sample/)
2. [mcmc tutorial](https://github.com/davharris/mcmc-tutorial)
3. [Probabilistic Programming and Bayesian Methods for Hackers Chapter 3](http://nbviewer.jupyter.org/github/CamDavidsonPilon/Probabilistic-Programming-and-Bayesian-Methods-for-Hackers/blob/master/Chapter3_MCMC/Chapter3.ipynb)

### 2016-04-11 bayesian_statistics/MCMC/MCMC.R

@@ -1,0 +1,130 @@
+# monte carlo markov chain
+library(ggplot2)
+library(ggthemes)
+
+f <- function(x){
+	( sin(.2 * x^3) + sin(x^2) / 5 + 2.5 + sin(10 * x) / 2 ) * dnorm( x, mean = 1 )
+}
+
+# # boundary for the x axis
+x_max <- 6
+x_min <- -6
+
+# for simply plotting the function
+ggplot( data.frame( x = c( x_max, x_min  ) ), aes(x) ) + 
+stat_function( fun = f ) + 
+
+x <- seq( x_min, x_max, length.out = 500 )
+y <- f(x)
+data <- data.frame( x = x, y = y )
+
+ggplot( data, aes( x, y ) ) + geom_line() + 
+geom_ribbon( aes( ymin = 0, ymax = y ), alpha = 0.2 )
+
+
+# -------------------------------------------------
+#   monte carlo rejection sampling
+# -------------------------------------------------
+
+# throw N darts (uniformly distributed x-y pairs) at the plot above
+N <- 10000
+ceiling <- 1.5
+xs <- runif( N, min = x_min, max = x_max )
+ys <- runif( N, min = 0, max = ceiling )
+
+# accept any samples that are below the function curve
+accepted_samples <- xs[ ys <= f(xs) ]
+
+ggplot( data.frame( x = accepted_samples ) ) + 
+geom_histogram( aes( x = x, y = ..density.. ) ) + 
+geom_line( data = data, aes( x, y ) )
+
+# study the samples however you want
+# mean(accepted_samples)
+# quantile(accepted_samples, probs = seq(0, 1, length = 4 ) )
+
+# accepted and unaccepted points
+df <- data.frame( xs = xs, ys = ys, accept = ys <= f(xs) )
+
+# plot the accepted and unaccepted point
+ggplot( df, aes( xs, ys ) ) + 
+geom_point( aes( color = accept ), size = 0.5 ) + 
+geom_line( data = data, aes( x, y ) )
+
+# -------------------------------------------------
+#   markov chain monte carlo 
+# -------------------------------------------------
+
+lik <- function( x, y ){
+	dnorm(x - 3) * dnorm(y - x + 2)
+}
+
+grid_values <- seq( x_min, x_max, length = 500 )
+grid <- expand.grid( x = grid_values, y = grid_values )
+z <- lik( grid$x, grid$y )
+
+# prior probability plot
[…]

### 2016-04-11 notebook_format/formats.py

@@ -4,17 +4,6 @@ import warnings
 import matplotlib
 from IPython.core.display import HTML
 
-"""
-def css_styling():
-    styles = open("../styles/custom.css", "r").read()
-    return HTML(styles)
-css_styling()
-
-
-s = json.load( open("../styles/bmh_matplotlibrc.json") )
-matplotlib.rcParams.update(s)
-"""
-
 def load_style(directory = ""):
 	# recent matplotlibs are raising deprecation warnings that
 	# we don't worry about.
@@ -27,3 +16,4 @@ def load_style(directory = ""):
 	# load the styles for the notebooks
 	styles = open( os.path.join( directory, "custom.css" ), "r" ).read()
 	return HTML(styles)
+

### 2016-04-12 bayesian_statistics/MCMC/MCMC.Rmd
mcmc

@@
 ---
-title: "Markov Chain Monte Carlo"
+title: "Markov Chain Monte Carlo (MCMC)"
 date: "`r Sys.Date()`"
@@
 		sample <- samples[ i - 1, ]
-
 		for(j in 1:thin){
@@
 	}
-
 	return( data.frame( samples[ (burnin + 1):iter, ] )	)
@@
 
+**Sources:**
+
+Code to the documentation can be found [here](https://github.com/ethen8181/Business-Analytics/blob/master/bayesian_statistics/MCMC/MCMC.R).
+
 ## Reference

### 2016-04-12 bayesian_statistics/MCMC/MCMC.R

@@ -81,8 +81,7 @@ MCMC <- function( iter, burnin = 0, thin = 1 )
 
 	for(i in 2:iter){
 
-		sample <- samples[ i - 1, ]
-		
+		sample <- samples[ i - 1, ]	
 		for(j in 1:thin){
 			
 			# propose a new sample point			
@@ -104,7 +103,6 @@ MCMC <- function( iter, burnin = 0, thin = 1 )
 		}
 		samples[ i, ] <- sample
 	}
-
 	return( data.frame( samples[ (burnin + 1):iter, ] )	)
 }
 

### 2016-04-12 bayesian_statistics/bayes/bayes.R

@@ -1,0 +1,120 @@
+library(dplyr)
+library(ggplot2)
+
+# simulated data
+# "\u03B1" unicode for the greek letter alpha
+sim <- data.frame( a = c( 81, 82, 81 + 100 ),
+				   b = c( 219, 219, 219 + 200 ) ) %>%
+	   group_by( a, b ) %>%
+	   do( data_frame( x = seq( 0, 1, by = .001 ), y = dbeta( x, .$a, .$b ) ) ) %>%
+	   mutate( Parameters = paste0( "\u03B1 = ", a, ", \u03B2 = ", b ) ) %>%
+	   ungroup() %>%
+	   mutate( Parameters = factor( Parameters, levels = unique(Parameters) ) )
+
+# plot of the distribution
+PlotBeta <- function(sim)
+{
+	ggplot( sim, aes( x, y, color = Parameters ) ) + geom_line() +
+	xlim( 0, .5 ) + ylab("Density of beta") + theme_bw()
+}
+PlotBeta( sim = filter( sim, a == 81 ) )
+
+
+# update 1 hit of 1 bat
+PlotBeta( sim = filter( sim, a == 81 | a == 82 ) )
+
+
+# update 100 hit of 300 bat
+PlotBeta( sim = sim )
+
+
+# http://varianceexplained.org/r/empirical_bayes_baseball/
+library(MASS)
+library(knitr)
+library(dplyr)
+library(tidyr)
+library(Lahman)
+library(ggplot2)
+
+
+# load the batting and pitching data from the Lahman package
+# the Master is used to get further details e.g. corresponding 
+# player name for the player id column in the Batting and Pitching data
+data(Master)
+data(Batting)
+data(Pitching)
+
+# prevent all warning and messages in code chunks,
+# and float numbers to print only three digits after the decimal point
+opts_chunk$set( warning = FALSE, message = FALSE )
+options( digits = 3 )
+
+# anti_join : return all rows from x, where there're no matching values in y
+# H = hits
+# AB = at bats
+career <- Batting %>% 
+		  filter( AB > 0 ) %>% 
+		  anti_join( Pitching, by = "playerID" ) %>%
+		  group_by(playerID) %>%
+		  summarize( H = sum(H), AB = sum(AB) ) %>%
+		  mutate( average = H / AB )
+
+# map the player name to player id
+career <- Master %>%
+		  tbl_df() %>%
[…]

### 2016-04-12 bayesian_statistics/beta_distribution.R

@@ -1,28 +1,0 @@
-library(dplyr)
-library(ggplot2)
-
-# simulated data
-# "\u03B1" unicode for the greek letter alpha
-sim <- data.frame( a = c( 81, 82, 81 + 100 ),
-				   b = c( 219, 219, 219 + 200 ) ) %>%
-	   group_by( a, b ) %>%
-	   do( data_frame( x = seq( 0, 1, by = .001 ), y = dbeta( x, .$a, .$b ) ) ) %>%
-	   mutate( Parameters = paste0( "\u03B1 = ", a, ", \u03B2 = ", b ) ) %>%
-	   ungroup() %>%
-	   mutate( Parameters = factor( Parameters, levels = unique(Parameters) ) )
-
-# plot of the distribution
-PlotBeta <- function(sim)
-{
-	ggplot( sim, aes( x, y, color = Parameters ) ) + geom_line() +
-	xlim( 0, .5 ) + ylab("Density of beta") + theme_bw()
-}
-PlotBeta( sim = filter( sim, a == 81 ) )
-
-
-# update 1 hit of 1 bat
-PlotBeta( sim = filter( sim, a == 81 | a == 82 ) )
-
-
-# update 100 hit of 300 bat
-PlotBeta( sim = sim )

### 2016-04-14 bayesian_statistics/bayes/bayes.Rmd
mcmc
bayesian statistics

--- 2016-04-01 bayesian_statistics/bayesian_statistics.Rmd
@@
 ---
+
+> Documentation reimplemented from [this series of posts](http://varianceexplained.org/r/bayesian_fdr_baseball/), with one major difference.
+The original posts uses the `dplyr` package when manipulating with data, whereas this one uses `data.table`.
+Code accompanying the documentation can be found [here](https://github.com/ethen8181/Business-Analytics/blob/master/bayesian_statistics/bayes/bayes.R).
+
+Which of these two proportions is higher: 4 out of 10, or 300 out of 1000? This sounds like a silly question. Obviously $4/10=.4$, which is greater than $300/1000=.3$!
+
+But suppose you were a baseball recruiter, trying to decide which of two potential players is a better batter based on how many hits they get.
+One has achieved 4 hits in 10 chances, the other 300 hits in 1000 chances.
+While the first player has a higher proportion of hits, it's not a lot of evidence: a typical player tends to achieve a hit around 27% of the time, and this player's $4/10$ could be due to luck.
+The second player, on the other hand, has a lot of evidence that he's an above-average batter.
+
+A lot of data takes the form of these success/total counts, where you want to estimate a "proportion of success" for each instance.
+When you work with pairs of successes/totals like this, you tend to get tripped up by the uncertainty in low counts.
+$1/2$ does not mean the same thing as $50/100$; nor does $0/1$ mean the same thing as $0/1000$.
+One approach is to filter out all cases that don't meet some minimum, but this isn't always an option: you're throwing away useful information.
+One approach to help make your estimate more accurate and practical is to use the **beta distribution** to represent your prior expectations, and update them based on new evidences.
 
@@
 library(ggplot2)
-library(dplyr)
+library(data.table)
 
-# synthesized data
+# set float numbers to print only three digits after the decimal point
+options( digits = 3 )
+
+# simulated data,
+# generate a sequence of numbers for each combination of a and b
+# to plot the probability density function.
 # "\u03B1" unicode for the greek letter alpha
-sim <- data.frame( a = c( 81, 82, 81 + 100 ),
-				   b = c( 219, 219, 219 + 200 ) ) %>%
-	   group_by( a, b ) %>%
-	   do( data_frame( x = seq( 0, 1, by = .001 ), y = dbeta( x, .$a, .$b ) ) ) %>%
-	   mutate( Parameters = paste0( "\u03B1 = ", a, ", \u03B2 = ", b ) ) %>%
-	   ungroup() %>%
-	   mutate( Parameters = factor( Parameters, levels = unique(Parameters) ) )
+sim <- data.table( a = c( 81, 82, 81 + 100 ),
+				   b = c( 219, 219, 219 + 200 ) )
+sim <- sim[ , .( x = seq( 0, 1, by = 0.002 ) ), by = .( a, b ) ]
+
+sim[ , `:=`( y = dbeta( x, a, b ),
+			 parameters = paste0( "\u03B1 = ", a, ", \u03B2 = ", b ) ) ]
+sim[ , parameters := factor( parameters, levels = unique(parameters) ) ]
+sim
+
+```
+
+```{r, message=FALSE, warning=FALSE}
 
@@
 {
-	ggplot( sim, aes( x, y, color = Parameters ) ) + geom_line() +
+	ggplot( sim, aes( x, y, color = parameters ) ) + geom_line() +
 	xlim( 0, .5 ) + ylab("Density of beta") + theme_bw()
 }
-PlotBeta( sim = filter( sim, a == 81 ) )
+PlotBeta( sim = sim[ a == 81, ] )
 
@@
 
-```{r, warning=FALSE}
+```{r, message=FALSE, warning=FALSE}
 
 # update 1 hit of 1 bat
-PlotBeta( sim = filter( sim, a == 81 | a == 82 ) )
+PlotBeta( sim = sim[ a %in% c( 81, 82 ), ] )
 
@@
 
-```{r, warning=FALSE}
+```{r, message=FALSE, warning=FALSE}
 
@@
 
-## Takeaways
+## Section Takeaways
 
@@
 
-## Reference
+# Empirical Bayes Estimation
 
-1. Understanding the beta distribution http://varianceexplained.org/statistics/beta_distribution_and_baseball/
+After introducing the **beta distribution**, we'll next see how it can be used in context with a very useful statistical method for estimating a large number of proportions, called **empirical Bayes estimation**.
+
+In the last section, we made some vague guesses about the distribution of batting averages across history.
+Here we'll apply **empirical Bayes estimation** to a real baseball dataset from the `Lahman` package, with the goal of improving our estimate of each player's batting average.
+
+## Working with Batting Averages
+
+In the dataset, we first filtered out pitchers (generally the weakest batters, who should be analyzed separately) using anti joins.
+We then summarized each player across multiple years to get their career Hits (H) and At Bats (AB), and batting average (H/AB).
+Finally, we added the player's first and last names to the dataset, so we could work with them rather than an identifier.
+
+```{r}
+
+# load the Batting and Pitching data from the Lahman package
+# the Master is used to get further details e.g. corresponding 
+# player name for the player id column in the Batting and Pitching data
+library(Lahman)
+data(Master)
+data(Batting)
+data(Pitching)
+
+master   <- data.table( Master  , key = "playerID" )
+pitching <- data.table( Pitching, key = "playerID" )
+batting  <- data.table( Batting , key = "playerID" )
+
+# ! stands for not join,
+# return all rows from x, where there're no matching values in y
+career <- batting[ AB > 0, ][!pitching]
+career <- career[ , .( H = sum(H), AB = sum(AB) ), by = playerID ]
+career[ , average := H / AB ]
+
+# map the player name to player id
+master <- master[ , .( playerID, nameFirst, nameLast ) ]
+career <- master[ , name := paste( nameFirst, nameLast ) ][career]
+career[ , `:=`( nameFirst = NULL, nameLast = NULL ) ]
+career
+
+```
+
+Let's list out the best batters in history. Well, here are the ones with the highest batting average:
+
+```{r}
+
+career[ head( order(-average) ), ]
+
+```
+
+Err, that's not really what we were looking for. These aren't the best batters, they're just the batters who went up once or twice and got lucky. How about the worst batters?
+
+```{r}
+
+career[ head( order(average) ), ]
+
+```
+
+Also not what we were looking for. That "average" is a really crummy estimate, they've just batted less than 10 times. Let's make a better one.
+
+## Estimate a Prior
+
+Let's look at the distribution of batting averages across players.
+For the sake of estimating the prior distribution, we'll filtered out all players that have fewer than 500 at-bats, since we'll get a better estimate from the less noisy cases.
+
+```{r}
+
+career_filtered <- career[ AB > 500, ]
+ggplot( career_filtered, aes(average) ) +
+geom_histogram( binwidth = .005 )
+
+```
+
+The first step of **empirical Bayes estimation** is to estimate a beta prior using this data.
+Estimating priors from the data you’re currently analyzing is not the typical Bayesian approach- usually you decide on your priors ahead of time.
+There’s a lot of debate and discussion about when and where it’s appropriate to use empirical Bayesian methods, but it basically comes down to how many observations we have: if we have a lot, we can get a good estimate that doesn’t depend much on any one individual.
+Empirical Bayes is an approximation to more exact Bayesian methods- and with the amount of data we have, it’s a very good approximation.
+
+So we know we want to fit the following model:
+
+$$X \sim \mbox{Beta}(\alpha_0,\beta_0)$$
+
+We just need to pick the $\alpha_0$ and $\beta_0$, which we call "hyper-parameters" of our model.
+There are many methods in R for fitting a probability distribution to data and you don't even have to use maximum likelihood if you have the [mean and variance](http://stats.stackexchange.com/questions/12232/calculating-the-parameters-of-a-beta-distribution-using-the-mean-and-variance) of the distribution.
+But here we'll use the `fitdistr` function from `MASS`.
+
+```{r}
+
+# fit beta distribution, you can specify any
+# starting parameter for shape1 and shape2 (alpha and beta),
+# the outputted warning does not matter
+m <- MASS::fitdistr( career_filtered$average, densfun = dbeta,
+					 start = list( shape1 = 70, shape2 = 200 ) )
+
+alpha0 <- m$estimate[1]
+beta0  <- m$estimate[2]
+
+```
+
+This comes up with $\alpha_0 = `r alpha0`$ and $\beta_0 = `r beta0`$. How well does this fit the data?
+
+```{r}
+
+ggplot(career_filtered) +
+geom_histogram( aes( x = average, y = ..density.. ), binwidth = .005 ) +
+xlab("Batting average") + 
+stat_function( fun = function(x) dbeta( x, alpha0, beta0 ), 
+			   color = "red", size = 1 )
+
+```
+
+Not bad!
+
+## Using the Prior
+
+Now when we look at any individual to estimate their batting average, we'll start with our overall prior, and update based on the individual evidence.
+This is as simple as adding $\alpha_0$ to the number of hits, and $\alpha_0 + \beta_0$ to the total number of at-bats.
+
+For example, consider our hypothetical batter from the introduction that went up 1000 times, and got 300 hits. We would estimate his batting average as:
+
+$$\frac{\alpha_0 + 300}{\alpha_0 + \beta_0 + 1000} = \frac{`r round(alpha0, 1)`+ 300}{`r round(alpha0, 1)` + `r round(beta0, 1)` + 1000}= `r (300 + alpha0) / (1000 + alpha0 + beta0)`$$
+
+How about the batter who went up only 10 times, and got 4 hits. We would estimate his batting average as:
+
+$$\frac{\alpha_0 + 4}{\alpha_0 + \beta_0 + 10} = \frac{`r round(alpha0, 1)` + 4}{`r round(alpha0, 1)` + `r round(beta0, 1)` + 10} = `r (4 + alpha0) / (10 + alpha0 + beta0)`$$
+
+Thus, even though $\frac{4}{10} > \frac{300}{1000}$, we would guess that the $\frac{300}{1000}$ batter is better than the $\frac{4}{10}$ batter!
+
+Performing this calculation for all the batters is simple enough:
+
+```{r}
+
+career[ , estimate := ( alpha0 + H ) / ( alpha0 + beta0 + AB ) ]
+
+```
+
+Now we can ask: who are the best and worst batters by this improved estimate?
+
+```{r}
+
+career[ head( order(-estimate) ), ]
+career[ head( order(estimate) ), ]
+
+```
+
+Notice that in each of these cases, **empirical Bayes estimates** didn’t simply pick the players who had 1 or 2 at-bats.
+It found players who batted well, or poorly, across a long career.
+What a load off our minds~ we can start using these **empirical Bayes estimates** in downstream analyses and algorithms, and not worry that we’re accidentally letting cases like 0/1 or 1/1 ruin everything.
+
+Let's see how **empirical Bayes estimates** changed all of the batting average estimates overall:
+
+```{r}
+
+ggplot( career, aes( average, estimate, color = AB ) ) + geom_point() + 
+geom_hline( yintercept = alpha0 / ( alpha0 + beta0 ), color = "red", lty = 2 ) + 
+labs( x = "Batting average", y = "Empirical Bayes batting average" ) +
+geom_abline( intercept = 0, slope = 1, color = "red" ) + 
+scale_color_gradient( trans = "log", breaks = 10^(1:4) )
+
+```
+
+The horizontal dashed red line marks $y = \frac{\alpha_0}{\alpha_0 + \beta_0} = `r sprintf("%.3f", alpha0 / (alpha0 + beta0))`$ - that's what we would guess someone's batting average was if we had no evidence at all.
+Notice that points above that line tend to move down towards it, while points below it move up.
+
+The diagonal red line marks $x=y$.
+Points that lie close to it are the ones that didn't get shrunk at al.
+Notice that they're the ones with the highest number of at-bats (the brightest blue): they have enough evidence that we're willing to believe the naive batting average estimate.
+
+This is why this process is sometimes called shrinkage: we've moved all our estimates towards the average.
+How much it moves these estimates depends on how much evidence we have: if we have very little evidence (4 hits out of 10) we move it a lot, if we have a lot of evidence (300 hits out of 1000) we move it only a little.
+
+## Section Takeaways
+
+Recall that there were two major steps in **empirical Bayes estimation**:
+
+1. Estimate the overall distribution of your data.
+2. Use that distribution as your prior for *updating* each average.
+
+Step 1 can be done once, "offline" - analyze all your data and come up with some estimates of your overall distribution.
+Step 2 is done for each new observation you’re considering, you can even apply thins to estimate the success of a post or an ad.
+
+And because we’re using the beta and the binomial, consider how easy that second step is.
+All we did was add one number to the successes, and add another number to the total.
+You can build that into your production system with a single line of code that takes nanoseconds to run.
+
+```
+# We hired a Data Scientist to analyze our Big Data
+# and all we got was this lousy line of code.
+
+float estimate = (successes + 78.7) / (total + 303.5);
+```
+
+# Understanding Credible Intervals
+
+Recall in the last section, we walked through the method of **empirical Bayes estimation**, a way to calculate useful proportions out of many pairs of success/total counts (e.g.
+0/1, 3/10, 235/1000).
+we used the example of estimating baseball batting averages based on $x$ hits in $n$ opportunities.
+If we run into a player with 0 hits in 2 chances, or 1 hit in 1 chance, we know we can't trust it, and this method uses information from the overall distribution to improve our guess.
+
+**Empirical Bayes estimation** gives us a single value for each player that can be reliably used as an estimate, but sometimes you want to know more than just our "best guess".
+Instead you wish to know how much uncertainty is present in our point estimate.
+We normally would use a binomial proportion confidence interval ( like a margin of error in a political poll, for more info about the calculation, refer to [here](http://onlinestatbook.com/2/estimation/proportion_ci.html) and [here](http://www.sigmazone.com/binomial_confidence_interval.htm) ), but this does not bring in information from our whole dataset.
+For example, the confidence interval for someone who hits 1 time out of 3 chances is:
+
+```{r}
+
+binom.test( x = 1, n = 3 )$conf.int
+
+```
+
+We can indeed be quite confident that that interval contains the true batting average...
+but from our knowledge of batting averages, we could have drawn a much tighter interval than that!
+There's no way that the player's real batting average is .1 or .9: it probably lies in the .2-.3 region that most other players' do.
+
+```{r}
+
+career
+
+```
+
+The end result of the last section gave us a *point estimate* for each proportion (the `estimate` column) and these new values tend to be pushed towards the overall mean (giving this the name "shrinkage"):
+
+This shrunken value is generally more useful than the raw estimate (the `average` column): we can use it to sort our data, or feed it into another analysis, without worrying too much about the noise introduced by low counts.
+But there's still uncertainty in the empirical Bayes estimate, and the uncertainty is very different for different players.
+We may want an interval of possible batting averages: one that will be wide for players we know very little about, and narrow for players with more information.
+Luckily, the Bayesian approach has a method to handle this.
+
+## Posterior distribution
+
+Consider that what we're really doing with **empirical Bayes estimation** is computing two new values, $\alpha_1$ and $\beta_1$, for each player.
+These are the posterior shape parameters for each distribution, after the prior (estimated on the whole dataset) has been updated.
+
+```{r}
+
+# alpha0 and beta0 is calculated in the last section
+career[ , `:=`( alpha1 = alpha0 + H, beta1 = beta0 + AB - H ) ]
+career
+
+```
+
+Using this information, we can visualize this posterior distribution for each player. Here, we'll just pick a few players from the 1998 Yankee lineup to illustrate the idea.
+
+```{r}
+
+# filter sample
+yankee_1998 <- c( "brosisc01", "jeterde01", "knoblch01", "martiti02", 
+				  "posadjo01", "strawda01", "willibe02" )
+career_yankee_1998 <- career[ playerID %in% yankee_1998, ]
+
+# create the x axis for the beta distribution's probability density function
+expand <- career_yankee_1998[ , .( x = seq( .18, .33, .0005 ) ), by = playerID ]
+yankee_beta <- career_yankee_1998[expand]
+yankee_beta[ , density := dbeta( x, alpha1, beta1 ) ]
+
+# visualize posterior beta
+ggplot( yankee_beta, aes( x, density, color = name ) ) + geom_line() +
+stat_function( fun = function(x) dbeta( x, alpha0, beta0 ),
+			   lty = 2, color = "black" )
+
+```
+
+The prior is shown as a dashed curve.
+All the other curves is our probability distribution of what the player’s batting average could be, after updating based on that player’s performance.
+That’s what we’re really estimating with this method: those posterior beta distributions.
+
+## Credible Intervals
+
+These density curves are nice, but can be hard to interpret visually, especially as the number of players increases, and it can't be summarized into a table or text.
+We'd instead prefer to create a credible interval, which says that some percentage (e.g.
+95%) of the posterior distribution lies within an particular region.
+Here's Derek Jeter's credible interval:
+
+```{r}
+
+# visualize the credible interval for one player
+jeter <- yankee_beta[ name == "Derek Jeter", ]
+
+# calculate the cumulative probability and
+# extract the ones that lies between the 95 percent credible interval
+p <- 0.95
+ci_low  <- ( 1 - p ) / 2
+ci_high <- p + ci_low
+
+jeter_pred <- jeter[ , cumulative := pbeta( x, alpha1, beta1 ) 
+				  ][ cumulative > ci_low & cumulative < ci_high ]
+
+# obtain the x coordinate of the 95 percent credible interval's 
+# endpoint to visualize the error bar
+jeter_low  <- qbeta( ci_low , jeter$alpha1[1], jeter$beta1[1] )
+jeter_high <- qbeta( ci_high, jeter$alpha1[1], jeter$beta1[1] )
+
+# credible interval plot
+ggplot( jeter, aes( x, density ) ) + geom_line( color = "blue" ) +
+xlim( .18, .34 ) + 
+geom_ribbon( data = jeter_pred, aes( ymin = 0, ymax = density ),
+			 alpha = .25, fill = "red" ) +
+stat_function( fun = function(x) dbeta( x, alpha0, beta0 ),
+			   lty = 2, color = "black" ) +
+geom_errorbarh( aes( xmin = jeter_low, xmax = jeter_high, y = 0 ), 
+				height = 3.5, color = "red" )
+```
+
+We can compute the edges of the interval quite easily for the rest of the players.
+
+```{r}
+# all the player's credible interval
+career_yankee_1998[ , `:=`( low  = qbeta( ci_low , alpha1, beta1 ),
+							high = qbeta( ci_high, alpha1, beta1 ) ) ]
+
+col_names <- colnames(career_yankee_1998)
+show <- col_names[ !col_names %in% c( "alpha1", "beta1", "estimate" ) ]
+
+# use kable from knitr for a more nicely formatted printed table
+knitr::kable( career_yankee_1998[ , show, with = FALSE ] )
+
+```
+
+And we can also view the intervals in a plot like this:
+
+```{r}
+
+ggplot( career_yankee_1998, aes( average, reorder( name, average ) ) ) +
+geom_point() +
+geom_errorbarh( aes( xmin = low, xmax = high ), height = 0.8 ) +
+geom_vline( xintercept = alpha0 / (alpha0 + beta0 ), color = "red", lty = 2 ) +
+xlab("Estimated batting average (95% interval)") +
+ylab("Player")
+
+```
+
+The vertical dashed red line is $\frac{\alpha_0}{\alpha_0 + \beta_0}$: the mean batting average across history (based on our beta fit).
+The earlier plot showing each posterior beta distribution communicated more information, but this is far more readable (depending on your purpose).
+
+## Credible Intervals versus Confidence Intervals
+
+Extract with the [wikipedia's](https://en.wikipedia.org/wiki/Credible_interval) definition on credible intervals.
+Credible intervals are analogous to confidence intervals in frequentist statistics, although they differ on a philosophical basis; Bayesian intervals treat their bounds as fixed and the estimated parameter as a random variable, whereas frequentist confidence intervals treat their bounds as random variables and the parameter as a fixed value.
+For a more detailed explanation, refer to this great [StackExchange post](http://stats.stackexchange.com/questions/2272/whats-the-difference-between-a-confidence-interval-and-a-credible-interval).
+
+But there’s also a very practical difference, in that credible intervals take prior information into account.
+If I take 20 random players and construct both confidence intervals (specifically a binomial proportion confidence interval) and posterior credible intervals for each, it could look something like this:
+
+```{r, message=FALSE, warning=FALSE}
+
+library(broom) # for tidying binom.test's output
+library(ggthemes)
+
+career[ , `:=`( low  = qbeta( ci_low , alpha1, beta1 ),
+				high = qbeta( ci_high, alpha1, beta1 ) ) ]
+
+# draw random 20 players
+set.seed(2015)
+some <- career[ sample.int( nrow(career), 20 ), ]
+some[ , name := paste0( name, " (", H, "/", AB, ")" ) ]
+
+# credible interval
+bayesian <- some[ , .( name, AB, estimate, low, high ) ]
+bayesian[ , method := "Credible" ]
+
+# confidence interval
+frequentist <- some[ , broom::tidy( binom.test( H, AB ) ), 
+					   by = .( playerID, name, AB ) ]
+frequentist <- frequentist[ , .( name, AB, estimate, 
+								 low = conf.low, high = conf.high ) ]
+frequentist[ , method := "Confidence"]
+
+combined <- rbind( bayesian, frequentist )
+ggplot( combined, aes( estimate, reorder( name, -AB ), color = method ) ) +
+geom_point() +
+geom_errorbarh( aes( xmin = low, xmax = high ) ) +
+geom_vline( xintercept = alpha0 / (alpha0 + beta0), color = "red", lty = 2 ) +
+labs( x = "Estimated batting average", y = "Player", color = "" ) +
+scale_color_tableau()
+
+```
+
+These are sorted in order of how many times a player went up to bat (thus, how much information we have about them).
+Notice that once there’s enough information, the **credible intervals** and **confidence intervals** are nearly identical.
+But for the 0/3 and 0/6 cases, the credible interval is much narrower.
+This is because **empirical Bayes estimate** brings in our knowledge from the full data, just as it did for the point estimate.
+
+# Bayesian False Discovery Rates
+
+We now have a **credible interval** for each player, including the lower and upper bounds for their batting average.
+But sometimes, rather than estimating a value or an interval, we’re looking to answer a yes or no question about each hypothesis, and thus classify them into two groups.
+For example, suppose we were constructing a Hall of Fame, where we wanted to include all players that have a batting average (chance of getting a hit) greater than .300 (note that this is just for illustration purpose, in real world, there are a lot of other, better metrics to judge a player by).
+We want to include as many players as we can, but we need to be sure that each belongs.
+
+The problem of hypothesis testing appears whenever we’re trying to identify candidates for future study.
+In this section we'll use the posterior distributions we’ve created to apply a Bayesian approach to a method usually associated with frequentist statistics, namely **false discovery rate control**.
+
+## Posterior Error Probabilities
+
+Consider the legendary player Hank Aaron.
+
+```{r}
+
+( hank_aaron <- career[ name == "Hank Aaron", ] )
+
+```
+
+We can see from the table that his career batting average is `r sprintf("%.4f", hank_aaron$average)`, but if we were to use this number to determine whether he should be permitted in our >.300 Hall of Fame, we're basing our decisions on his "true probability" of hitting.
+
+On the other hand, when Aaron's batting average is shrunken by the **empirical Bayes estimate**, we get an estimate of `r sprintf("%.4f", hank_aaron$estimate)`.
+We thus suspect that his true probability of hitting is higher than .300, but we're not necessarily certain because of the **credible interval**.
+Let's take a look at his posterior beta distribution:
+
+```{r}
+
+# hall of fame threshold
+threshold <- .3
+
+# posterior beta
+# merge the original data.table with the generated sequence ones,
+hank_aaron <- hank_aaron[ hank_aaron[ , .( x = seq( .27, .33, .0004 ) ), by = playerID ] ]
+hank_aaron[ , density := dbeta( x, alpha1, beta1 ) ]
+
+ggplot( hank_aaron, aes( x, density ) ) + geom_line() +
+geom_ribbon( aes( ymin = 0, ymax = density * ( x < threshold ) ),
+				  alpha = .1, fill = "red" ) +
+geom_vline( color = "red", lty = 2, xintercept = threshold )
+
+```
+
+We can see that there is a nonzero probability (shaded area) that his true probability of hitting is less than .3.
+We can calulate this with the cumulative distribution function (CDF) of the beta distribution, which in R is computed by the `pbeta` function:
+
+```{r}
+
+# posterior error probability (PEP)
+pbeta( threshold, hank_aaron$alpha1[1], hank_aaron$beta1[1] )
+
+```
+
+This probability that he doesn't belong in the Hall of Fame is called the **Posterior Error Probability**, or PEP.
+We could easily have calculated the probability Aaron does belong, which we would call the Posterior Inclusion Probability, or PIP.
+(Note that $\mbox{PIP} = 1-\mbox{PEP}$) The reason we chose to measure the PEP rather than the PIP will become clear in the next section.
+
+It's equally straightforward to calculate the PEP for every player, just like we calculated the **credible intervals** for each player in the last section.
+
+```{r}
+
+# PEP for every player
+career[ , PEP := pbeta( threshold, alpha1, beta1 ) ]
+
+```
+
+What does the distribution of the PEP look like across players?
+
+```{r}
+
+# histogram of PEP
+ggplot( career, aes(PEP) ) + geom_histogram( binwidth = .02 ) +
+xlab("Posterior Error Probability (PEP)") + xlim( 0, 1 )
+
+```
+
+Unsurprisingly, for most players, it’s almost certain that they don’t belong in the hall of fame: we know that their batting averages are way below .300.
+If they were included, it is almost certain that they would be an error.
+In the middle are the borderline players: the ones where we’re not sure.
+And down there close to 0 are the rare but proud players who we’re (almost) certain belong in the hall of fame.
+
+The PEP is closely related to the estimated batting average:
+
+```{r}
+
+# PEP and estimate batting average
+ggplot( career, aes( estimate, PEP, color = AB ) ) + geom_point( size = 1 ) +
+xlab("(Shrunken) batting average estimate") +
+ylab("Posterior Error Probability (PEP)") +
+geom_vline( color = "red", lty = 2, xintercept = threshold ) +
+scale_colour_gradient( trans = "log", breaks = 10^(1:5) )
+
+```
+
+Notice that crossover point: to have a PEP less than 50%, you need to have a shrunken batting average greater than .3.
+That's because the shrunken estimate is the center of our posterior beta distribution (the "over/under" point).
+If a player's shrunken estimate is above .3, it's more likely than not that their true average is as well.
+And the players we're not sure about (PEP $\approx$ .5) have batting averages very close to .300.
+
+Notice also the relationship between the number of at-bats (the amount of evidence) and the PEP.
+If a player's shrunken batting average is .28, but he hasn't batted (low AB) many times, it is still possible his true batting average is above .3, because his **credible interval** is wide.
+However, if the player with .28 has a high AB (light blue), the **credible interval** becomes thinner, we become confident that his batting average's true probability is under .3, and the PEP goes up to 1.
+
+## False Discovery Rate
+
+Now we want to set some threshold for inclusion in our Hall of Fame.
+This criterion is up to us: what kind of goal do we want to set?
+There are many options and here's a propose one: let's try to include as many players as possible, while ensuring that no more than 5% of the Hall of Fame was mistakenly included.
+Put another way, we want to ensure that if you're in the Hall of Fame, the probability you belong there is at least 95%.
+
+This criterion is called **false discovery rate (FDR) control**.
+It's particularly relevant in scientific studies, where we might want to come up with a set of candidates (e.g.
+genes, countries, individuals) for future study.
+There's nothing special about 5%: if we wanted to be more strict, we could choose the same policy, but change our desired FDR to 1% or .1%.
+Similarly, if we wanted a broader set of candidates to study, we could set an FDR of 10% or 20%.
+
+Let's start with the easy cases. Who are the players with the lowest posterior error probability?
+
+```{r}
+
+# ranking PEP
+career <- career[ order(PEP), ]
+by_PEP <- career[ , rank := 1:nrow(career) 
+			   ][ , .( rank, name, H, AB, estimate, PEP ) ]
+
+head( by_PEP, 10 )
+
+```
+
+These players are a no-brainer for our Hall of Fame: there’s basically no risk in including them. But suppose we instead tried to include the top 100. What do the 90th-100th players look like?
+
+```{r}
+
+by_PEP[ 90:100, ]
+
+```
+
+OK, so these players are borderline. We would guess that their career batting average is greater than .300, but we aren't as certain.
+
+So let's say we chose to take the top 100 players for our Hall of Fame (thus, cut it off at `r by_PEP$name[100]`).
+What would we predict the false discovery rate to be?
+That is, what fraction of these 100 players would be falsely included?
+Well, we know the PEP of each of these 100 players, which is the probability that that individual player is a false positive, we can simply add up these probabilities to get the expected value (the average) of the total number of false positives.
+
+```{r}
+
+# top 100 players false positive rate
+top_players <- career[ 1:100, ]
+sum(top_players$PEP)
+
+```
+
+This means that of these 100 players, we expect that about `r sum(top_players$PEP)` of them are false discoveries.
+Now, we don’t know which four or five players we are mistaken about!
+(If we did, we could just kick them out of the hall).
+But we can make predictions about the players in aggregate.
+Here, we can see that taking the top 100 players would get pretty close to our goal of FDR = 5%.
+
+Note that we’re calculating the FDR as `r sum(top_players$PEP)` / `r nrow(top_players)` (number of players). Thus, we’re really computing the mean PEP: the average Posterior Error Probability.
+
+```{r}
+
+mean(top_players$PEP)
+
+```
+
+We can experiment with many thresholds to get our desired **false discovery rate**, but it’s even easier just to compute them all at once, by computing the cumulative mean of all the (sorted) posterior error probabilities.
+
+## Q-values
+
+Notice that the name of the **false discovery rate**'s cumulative mean is **qvalue**. The q-value is convenient because we can say "to control the FDR at X%, collect only hypotheses where $q < X$".
+
+```{r}
+
+# PEP's cumulative mean for meeting the false discovery rate
+career[ , qvalue := cumsum(PEP) / 1:nrow(career) ]
+
+# number of hall of famers for 5% false discovery rate
+hall_of_fame <- career[ qvalue < .05, ]
+
+```
+
+This ends up with `r nrow(hall_of_fame)` players in the Hall of Fame. If we wanted to be more careful about letting players in, we'd simply set a stricter (smaller) q-value threshold.
+
+It’s also useful to look at how many players would be included at various thresholds:
+
+```{r}
+
+# numbers of players included at various q-value cutoff
+ggplot( career[ qvalue < .25, ], aes( qvalue, rank ) ) + geom_line() +
+labs( x = "q-value cutoff", y = "Number of players included" )
+
+```
+
+This shows that you could include 200 players in the Hall of Fame, but at that point you’d expect that about 25% of them would be incorrectly included.
+On the other side, you could create a hall of fame with just 50 players and be very confident that all of them have a batting probability of .300.
+
+It's worth emphasizing the difference between measuring an individual's **posterior error probability** and the **q-value**, which is the **false discovery rate** of a group including that player.
+Hank Aaron has a PEP of `r pbeta( threshold, hank_aaron$alpha1[1], hank_aaron$beta1[1] )`, but he can be included in the Hall of Fame while keeping the FDR below 5%.
+If this is surprising, imagine that you were instead trying to keep the average height above 6'0".
+You would start by including all players taller than 6'0", but could also include some players who were 5'10" or 5'11" while preserving your average.
+Similarly, we simply need to keep the average PEP of the players below 5%.
+
+## Takeaways
+
+During this modeling process, we’ve made an enormous simplification by assuming that all batting averages are drawn from a single distribution.
+In reality, we’d expect that it depends on some known factors.
+For instance, the distribution of batting averages has changed over time.
+Ideally, we’d want to estimate a different Beta prior for each decade.
+Similarly, we could estimate separate priors for each team, a separate prior for pitchers, and so on.
+One useful approach to this is the Bayesian hierarchical modeling, which is another story.
+
+# Reference
+
+1. [Understanding the beta distribution](http://varianceexplained.org/statistics/beta_distribution_and_baseball/)
+2. [Understanding empirical Bayes estimation](http://varianceexplained.org/r/empirical_bayes_baseball/)
+3. [Understanding Credible Intervals](http://varianceexplained.org/r/credible_intervals_baseball/)
+4. [Understanding the Bayesian Approach to False Discovery Rates](http://varianceexplained.org/r/bayesian_fdr_baseball/)

### 2016-04-14 bayesian_statistics/bayes/bayes.R

@@ -1,120 +1,282 @@
-library(dplyr)
+# bayesian statistics
 library(ggplot2)
+library(data.table)
 
-# simulated data
+# set float numbers to print only three digits after the decimal point
+options( digits = 3 )
+
+# -----------------------------------------------------------------------
+# beta distribution
+# http://varianceexplained.org/statistics/beta_distribution_and_baseball/
+# -----------------------------------------------------------------------
+
+# simulated data,
+# generate a sequence of numbers for each combination of a and b
+# to plot the probability density function.
 # "\u03B1" unicode for the greek letter alpha
-sim <- data.frame( a = c( 81, 82, 81 + 100 ),
-				   b = c( 219, 219, 219 + 200 ) ) %>%
-	   group_by( a, b ) %>%
-	   do( data_frame( x = seq( 0, 1, by = .001 ), y = dbeta( x, .$a, .$b ) ) ) %>%
-	   mutate( Parameters = paste0( "\u03B1 = ", a, ", \u03B2 = ", b ) ) %>%
-	   ungroup() %>%
-	   mutate( Parameters = factor( Parameters, levels = unique(Parameters) ) )
+sim <- data.table( a = c( 81, 82, 81 + 100 ),
+				   b = c( 219, 219, 219 + 200 ) )
+sim <- sim[ , .( x = seq( 0, 1, by = 0.002 ) ), by = .( a, b ) ]
+
+sim[ , `:=`( y = dbeta( x, a, b ),
+			 parameters = paste0( "\u03B1 = ", a, ", \u03B2 = ", b ) ) ]
+sim[ , parameters := factor( parameters, levels = unique(parameters) ) ]
+
 
 # plot of the distribution
 PlotBeta <- function(sim)
 {
-	ggplot( sim, aes( x, y, color = Parameters ) ) + geom_line() +
+	ggplot( sim, aes( x, y, color = parameters ) ) + geom_line() +
 	xlim( 0, .5 ) + ylab("Density of beta") + theme_bw()
 }
-PlotBeta( sim = filter( sim, a == 81 ) )
-
+PlotBeta( sim = sim[ a == 81, ] )
 
 # update 1 hit of 1 bat
-PlotBeta( sim = filter( sim, a == 81 | a == 82 ) )
-
+PlotBeta( sim = sim[ a %in% c( 81, 82 ), ] )
 
 # update 100 hit of 300 bat
 PlotBeta( sim = sim )
 
 
+# -----------------------------------------------------------------------
+# empirical bayes
[…]

### 2016-04-14 bayesian_statistics/bayes/bayes_dplyr.R

@@ -1,0 +1,300 @@
+# bayesian statistics
+library(tidyr)
+library(dplyr)
+library(knitr)
+library(broom)
+library(ggplot2)
+
+# set float numbers to print only three digits after the decimal point
+options( digits = 3 )
+
+# -----------------------------------------------------------------------
+# beta distribution
+# http://varianceexplained.org/statistics/beta_distribution_and_baseball/
+# -----------------------------------------------------------------------
+
+# simulated data
+# "\u03B1" unicode for the greek letter alpha
+sim <- data.frame( a = c( 81, 82, 81 + 100 ),
+				   b = c( 219, 219, 219 + 200 ) ) %>%
+	   group_by( a, b ) %>%
+	   do( data_frame( x = seq( 0, 1, by = .001 ), y = dbeta( x, .$a, .$b ) ) ) %>%
+	   mutate( parameters = paste0( "\u03B1 = ", a, ", \u03B2 = ", b ) ) %>%
+	   ungroup() %>%
+	   mutate( parameters = factor( parameters, levels = unique(parameters) ) )
+
+# plot of the distribution
+PlotBeta <- function(sim)
+{
+	ggplot( sim, aes( x, y, color = parameters ) ) + geom_line() +
+	xlim( 0, .5 ) + ylab("Density of beta") + theme_bw()
+}
+PlotBeta( sim = filter( sim, a == 81 ) )
+
+
+# update 1 hit of 1 bat
+PlotBeta( sim = filter( sim, a == 81 | a == 82 ) )
+
+
+# update 100 hit of 300 bat
+PlotBeta( sim = sim )
+
+
+# -----------------------------------------------------------------------
+# empirical bayes
+# http://varianceexplained.org/r/empirical_bayes_baseball/
+# -----------------------------------------------------------------------
+
+# load the Batting and Pitching data from the Lahman package
+# the Master is used to get further details e.g. corresponding 
+# player name for the player id column in the Batting and Pitching data
+library(Lahman)
+data(Master)
+data(Batting)
+data(Pitching)
+
+# anti_join : return all rows from x, where there're no matching values in y
+# H = hits
+# AB = at bats
+career <- Batting %>% 
+		  filter( AB > 0 ) %>% 
+		  anti_join( Pitching, by = "playerID" ) %>%
+		  group_by(playerID) %>%
[…]

### 2016-04-15 bayesian_statistics/MCMC/MCMC.R
to do mcmc

@@ -126,3 +126,13 @@ samples <- MCMC( iter = 250, burnin = 125, thin = 2 )
 TracePlot( plot, samples )
 
 
+# - [Gibbs sampling with for uninitiated](http://www.umiacs.umd.edu/~resnik/pubs/gibbs.pdf)
+
+# mcmc for dummies
+# http://twiecki.github.io/blog/2015/11/10/mcmc-sampling/
+# http://www.people.fas.harvard.edu/~plam/teaching/methods/mcmc/mcmc_print.pdf
+# http://stats.stackexchange.com/questions/165/how-would-you-explain-markov-chain-monte-carlo-mcmc-to-a-layperson
+
+
+
+

### 2016-04-18 articles/articles.Rmd
upload
organize folder

@@
 
+- Doing Data Science Right — Your Most Common Questions Answered.
+[[Review](http://firstround.com/review/doing-data-science-right-your-most-common-questions-answered/?imm_mid=0e29cf&cmp=em-data-na-na-newsltr_20160413)]
 - Principles of good data analytics. [[blog](http://www.gregreda.com/2014/03/23/principles-of-good-data-analysis/)] 
@@
 [[Linkedin](https://www.linkedin.com/pulse/three-things-i-learned-during-4-years-uber-michael-pao)]
+
+# Learn Data Science
+
+- The MBA Data Science Toolkit: 8 resources to go from the spreadsheet to the command line.
+[[Medium](https://medium.com/@dmca/the-mba-data-science-toolkit-8-resources-to-go-from-the-spreadsheet-to-the-command-line-cbb59ea82144#.pjph91n3m)]

### 2016-04-18 bayesian_statistics/MCMC/MCMC.Rmd
to do mcmc
organize folder

@@
 
+> To follow along, code to the documentation can be found [here](https://github.com/ethen8181/Business-Analytics/blob/master/bayesian_statistics/MCMC/MCMC.R).
+
 # Learn About a Probability Distribution
@@
 Some probability distributions, like the Normal distribution and the Poisson distribution, are relatively straightforward to work with mathematically and we have lots of handy tools for working with them by hand and on the computer.
-
 Other probability distributions over parameters in a complicated model are harder to work with.
 In this case we need more general tools.
-**Monte Carlo** methods work by taking samples from a distribution that's hard to study analytically, and then studying the samples instead.
+
+*Markov Chain Monte Carlo (MCMC)* is a class of methods in which we can simulate draws (taking samples) that are slightly dependent and are approximately from a (posterior) distribution.
+We then take those draws and calculate quantities of interest for the (posterior) distribution.
+In Bayesian statistics, there are generally two MCMC algorithms that we use: the **Gibbs Sampler** and the **Metropolis-Hastings algorithm**.
+
+In short, MCMC methods work by taking samples from a distribution that's hard to study analytically, and then study on their samples instead.
 
@@
 It turns out that you may end up throwing away exponentially more attempts per successful sample as the number of dimensions in your data goes up.
-So hardly anyone uses rejection sampling without a good reason -- that's where the Markov chains come in.
+So hardly anyone uses rejection sampling without a good reason.
 
@@
 
-That's where the Markov chains come in.
+That's where MCMC methods comes in.
 MCMC has exactly the same goal as our rejection sampler, but it has more sophisticated ways of deciding where to sample, which can be expressed at a high level as follows:
@@
 We call this **thinning**.
+e.g.
+when **thinning** by 2, only sample points at step 1, 3, 5, 7 and so on, are retained.
+Note that although the thinned sample points tend to be less correlated than the original ones, they can also contain less information and it takes longer to run (reduces sampling efficiency).
 
@@
 
+# Metropolis-Hastings algorithm (will elaborate on this later)
 MCMC <- function( iter, burnin = 0, thin = 1 )
@@
 
-			# propose a new sample point			
+			# propose a new sample point (will elaborate on this later)			
 			proposal <- sample + rnorm( n = 2, mean = 0, sd = 1 )
@@
 
-**End Note:**
-
 The MCMC method is very useful, the intuition for the algorithm is that we can approximate the **posterior distribution** by generating a large sample of representative values and the larger the sample, the more accurate is our approximation.
 
-**Sources:**
+## Gibbs Sampling
 
-Code to the documentation can be found [here](https://github.com/ethen8181/Business-Analytics/blob/master/bayesian_statistics/MCMC/MCMC.R).
+The MCMC method we've introduced above is also known as the *Metropolis-Hastings algorithm*.
+It is very general and broadly applicable, but one drawback is that your proposed distribution has to be properly tuned to the posterior distribution.
+Recall that we proposed a new random sample point using the a normal distribtuion with the mean of 1 and standard deviation of 1 `rnorm( ..., mean = 0, sd = 1 )`.
+If this random sample point generator's distribution is too wide or too narrow, a large proportion of our proposed sample point will still be rejected.
+Or more intuitively, the effectiveness of our random walk will farely low.
+
+Given this situation, there's an alternative sampling algorithm called **Gibbs Samplinig**. Unlike the **Metropolis Sampling**, with **Gibbs Samplinig**  all proposed jumps are accepted.
+
+## Takeaways
+
+- If you can decompose your probability distribution into conditional probabilities, then **Gibbs Samplinig** is preferable over **Metropolis Sampling**.
 
@@
 [Probabilistic Programming and Bayesian Methods for Hackers Chapter 3](http://nbviewer.jupyter.org/github/CamDavidsonPilon/Probabilistic-Programming-and-Bayesian-Methods-for-Hackers/blob/master/Chapter3_MCMC/Chapter3.ipynb)
+
+## R Session Information
+
+```{r}
+
+sessionInfo()
+
+```

### 2016-04-18 marketing_analysis/marketing_analysis.Rmd
organize folder

--- 2015-12-16 2_marketing_analysis/marketing_analysis.Rmd
@@
 author: "Ming-Yu Liu"
-date: "December 20, 2015"
+date: "`r Sys.Date()`"
 output:
-  html_document: 
-    highlight: haddock
-    number_sections: yes
-    theme: united
-    toc: yes
+  rmdformats::readthedown:
+    highlight: pygments
 ---
-
-## Background Information
 
@@
 
-## Segmentation
+# Segmentation
 
@@
 library(data.table)
-setwd("/Users/ethen/Business-Analytics/2_marketing_analysis")
+setwd("/Users/ethen/Business-Analytics/marketing_analysis")
 
@@
 
-source("/Users/ethen/Business-Analytics/2_marketing_analysis/code/marketing_functions.R")
+source("/Users/ethen/Business-Analytics/marketing_analysis/code/marketing_functions.R")
 rfm1 <- Segmentation( data = rfm1, inactive = inactive, 
@@
 
-## Customer Life Time Value
+**Section Takeaways:** Use statistical segmentation when the segment you wish to obtain is a "oneshot" thing or use it as a exploratory tool when you're not famaliar with the overall objective.
+
+# Customer Life Time Value
 
@@
 
-**Takeaways:**
+**Sources:**
 
-1. Use statistical segmentation when the segment you wish to obtain is a "oneshot" thing or use it as a exploratory tool when you're not famaliar with the overall objective.
-
-You can find all the source code [here](https://github.com/ethen8181/Business-Analytics/tree/master/2_marketing_analysis/code).
+You can find all the source code [here](https://github.com/ethen8181/Business-Analytics/tree/master/marketing_analysis/code).
 marketing_analysis.R contains the documentation's general scripts and marketing_functions.R contains the functions that were sourced in and used with this documentation.
@@
 
-## R Session Information
+# R Session Information
 

### 2016-04-18 bayesian_statistics/MCMC/MCMC.R

@@ -71,12 +71,13 @@ plot <- ggplot( data = grid, aes( x = x, y = y ) ) +
 plot
 
 
+# Metropolis-Hastings algorithm (will elaborate on this later)
 MCMC <- function( iter, burnin = 0, thin = 1 )
 {
 	# the list of the first argument for dimnames is to specify the row names,
 	# sampling starts at a random point, here 0, 0 
 	samples <- matrix( NA, nrow = iter, ncol = 2, 
-					  dimnames = list( NULL, c( "x", "y" ) ) )
+					   dimnames = list( NULL, c( "x", "y" ) ) )
 	samples[ 1, ] <- c( 0, 0 )
 
 	for(i in 2:iter){
@@ -84,7 +85,7 @@ MCMC <- function( iter, burnin = 0, thin = 1 )
 		sample <- samples[ i - 1, ]	
 		for(j in 1:thin){
 			
-			# propose a new sample point			
+			# propose a new sample point (will elaborate on this later)				
 			proposal <- sample + rnorm( n = 2, mean = 0, sd = 1 )
 
 			# compare its likelihood with the current position
@@ -126,7 +127,16 @@ samples <- MCMC( iter = 250, burnin = 125, thin = 2 )
 TracePlot( plot, samples )
 
 
-# - [Gibbs sampling with for uninitiated](http://www.umiacs.umd.edu/~resnik/pubs/gibbs.pdf)
+# for downloading jags
+# http://mcmc-jags.sourceforge.net/
+# install.packages( c( "rjags", "runjags" ) )
+# install.packages("runjags")
+
+
+
+
+# Gibbs sampling with for uninitiated (paper), with latent dirichlet allocation
+# http://www.umiacs.umd.edu/~resnik/pubs/gibbs.pdf
 
 # mcmc for dummies
 # http://twiecki.github.io/blog/2015/11/10/mcmc-sampling/

### 2016-04-21 articles/articles.Rmd
bayes done for now

@@
 title: "Articles"
+date: "`r Sys.Date()`"
 author: "Ming-Yu Liu"
 output:
-  html_document: 
-    highlight: haddock
-    toc: yes
+  rmdformats::readthedown:
+    highlight: pygments
 ---
@@
 
+- Using Machine Learning to Predict Customer Behaviour [[Data Science Central](http://www.datasciencecentral.com/profiles/blogs/using-machine-learning-to-predict-customer-behaviour)]
+    - Used with Complaints Management, Customer Upsell and Customer Retention.
+Questions include How to anticipate and manage the dissatisfaction of the customers?', 'How to sell more and better to existing customers?' and 'How to keep the most profitable customers at an appropriate cost?'
+    - Action: You identify a target behaviour, in our case, it is a specific customer action, like filing a complaint/not filing a complaint, accepting or rejecting an upsell offer, cancelling / not cancelling a service.
+Then perform supervised learning
 - Doing Data Science Right — Your Most Common Questions Answered.

### 2016-04-21 bayesian_statistics/MCMC/MCMC.Rmd
bayes done for now

@@
 We then take those draws and calculate quantities of interest for the (posterior) distribution.
-In Bayesian statistics, there are generally two MCMC algorithms that we use: the **Gibbs Sampler** and the **Metropolis-Hastings algorithm**.
+Here, we'll bw attempting to give you an intuition behind one branch of the MCMC sampling, namely the **Metropolis–Hastings Algorithm**.
 
@@
 
-## Markov Chain Monte Carlo (MCMC)
+## MCMC - Metropolis–Hastings Algorithm
 
@@
 ```
-
-When we setup a Bayesian inference problem with N unknowns, we are implicitly creating an N dimensional space for the **prior distributions** to exist in.
-Associated with the space is an additional dimension, which we can describe as the surface, or curve, that sits on top of the space, that reflects the **prior probability** of a particular point.
-The plots above visualize this.
-The more blue the color, the more **prior probability** is assigned to that location.
-Conversely, areas with whiter color represent that our priors assign very low probability to that location.
 
@@
 
-## Gibbs Sampling
-
-The MCMC method we've introduced above is also known as the *Metropolis-Hastings algorithm*.
-It is very general and broadly applicable, but one drawback is that your proposed distribution has to be properly tuned to the posterior distribution.
-Recall that we proposed a new random sample point using the a normal distribtuion with the mean of 1 and standard deviation of 1 `rnorm( ..., mean = 0, sd = 1 )`.
-If this random sample point generator's distribution is too wide or too narrow, a large proportion of our proposed sample point will still be rejected.
-Or more intuitively, the effectiveness of our random walk will farely low.
-
-Given this situation, there's an alternative sampling algorithm called **Gibbs Samplinig**. Unlike the **Metropolis Sampling**, with **Gibbs Samplinig**  all proposed jumps are accepted.
-
-## Takeaways
-
-- If you can decompose your probability distribution into conditional probabilities, then **Gibbs Samplinig** is preferable over **Metropolis Sampling**.
-
 ## Reference

### 2016-04-21 bayesian_statistics/MCMC/MCMC.R

@@ -126,23 +126,15 @@ TracePlot( plot, samples )
 samples <- MCMC( iter = 250, burnin = 125, thin = 2 )
 TracePlot( plot, samples )
 
+# -----------------------------------------------------------------------
+# not implemented
 
-# for downloading jags
-# http://mcmc-jags.sourceforge.net/
-# install.packages( c( "rjags", "runjags" ) )
-# install.packages("runjags")
-
-
-
-
-# Gibbs sampling with for uninitiated (paper), with latent dirichlet allocation
-# http://www.umiacs.umd.edu/~resnik/pubs/gibbs.pdf
-
-# mcmc for dummies
+# mcmc for dummies, the first one is in python
 # http://twiecki.github.io/blog/2015/11/10/mcmc-sampling/
 # http://www.people.fas.harvard.edu/~plam/teaching/methods/mcmc/mcmc_print.pdf
-# http://stats.stackexchange.com/questions/165/how-would-you-explain-markov-chain-monte-carlo-mcmc-to-a-layperson
 
 
-
+# not related, for downloading jags
+# http://mcmc-jags.sourceforge.net/
+# install.packages( c( "rjags", "runjags" ) )
 

### 2016-04-21 bayesian_statistics/MCMC/gibbs.py

@@ -1,0 +1,17 @@
+
+"""
+
+Gibbs Sampling
+The MCMC method we've introduced above is also known as the *Metropolis-Hastings algorithm*. It is very general and broadly applicable, but one drawback is that your proposed distribution has to be properly tuned to the posterior distribution. Recall that we proposed a new random sample point using the a normal distribtuion with the mean of 1 and standard deviation of 1 `rnorm( ..., mean = 0, sd = 1 )`. If this random sample point generator's distribution is too wide or too narrow, a large proportion of our proposed sample point will still be rejected. Or more intuitively, the effectiveness of our random walk will farely low.
+Given this situation, there's an alternative sampling algorithm called **Gibbs Samplinig**. Unlike the **Metropolis Sampling**, with **Gibbs Samplinig**  all proposed jumps are accepted.
+
+Takeaways
+If you can decompose your probability distribution into conditional probabilities, then **Gibbs Samplinig** is preferable over **Metropolis Sampling**.
+
+
+Gibbs sampling with for uninitiated (paper), with latent dirichlet allocation
+http://www.umiacs.umd.edu/~resnik/pubs/gibbs.pdf
+https://theclevermachine.wordpress.com/2012/11/19/a-gentle-introduction-to-markov-chain-monte-carlo-mcmc/
+http://nbviewer.jupyter.org/github/cs109/content/blob/master/labs/lab7/GibbsSampler.ipynb
+
+"""

### 2016-04-21 bayesian_statistics/ab_test/abtest.py

@@ -1,0 +1,131 @@
+import pymc as pm
+import numpy as np
+import matplotlib.pyplot as plt
+from pymc.Matplot import plot as mcplot
+
+
+def acf( x, lag = 100 ):
+    """
+    autocorrelation function that calculates the 
+    autocorrelation for series x from 1 to the user-specified lag
+    
+    Parameter
+    ---------
+    x : 1d-array, size ( iteration - burn-in iteration, )
+        pymc's random sample
+        
+    lag : int, default 100
+        maximum lagging number
+        
+    Return
+    ------
+    corr : list, size (lag)
+        autocorrelation for each lag
+    
+    Reference
+    ---------
+    http://stackoverflow.com/questions/643699/how-can-i-use-numpy-correlate-to-do-autocorrelation
+    """
+    # np.corrcoef here returns a 2 * 2 matrix, either [ 0, 1 ] or [ 1, 0 ]
+    # will be the actual autocorrelation between the two series,
+    # the diagonal is just the autocorrelation between each series itself
+    # which is 1
+    corr = [ np.corrcoef( x[k:], x[:-k] )[ 0, 1 ] for k in range( 1, lag ) ]
+    return corr
+
+
+def effective_sample( iteration, burnin, corr ):
+    """
+    calculate the effective sample size of the mcmc,
+    note that the calculation is based on the fact that
+    the autorrelation plot appears to have converged
+    
+    Parameters
+    ----------
+    iteration : int
+        number of iteration of the mcmc
+    
+    burnin : int
+        burn-in iteration of the mcmc
+    
+    corr : list
+        list that stores the autocorrelation for different lags
+    
+    Returns
+    -------
+    the effective sample size : int
+    """
+    for index, c in enumerate(corr):
+        if c < 0.05:
+            i = index
+            break
+    
+    return ( iteration - burnin ) / ( 1 + 2 * np.sum( corr[:i + 1] ) )
+
+
+@pm.stochastic( dtype = np.float64 )
+def beta_priors( value = [ 1.0, 1.0 ] ):  
+    a, b = value
+    # outside of the support of the distribution
+    if a <= 0 or b <= 0:
+        return -np.inf
+    else:
+        return np.log( np.power( ( a + b ), -2.5 ) )
+
+
+def bayesian_ab_test( beta_priors, trials, successes
+					  iteration, burnin, thin ):
[…]

### 2016-04-22 ab_test/frequentist_ab_test.Rmd
ab test done for now

--- 2016-03-16 3_AB_test/AB_test.Rmd
@@
 author: "Ming-Yu Liu"
-date: "January 15, 2016"
+date: "`r Sys.Date()`"
 output:
-  html_document: 
+  rmdformats::readthedown:
     highlight: pygments
-    number_sections: yes
-    theme: united
-    toc: yes
 ---
 
-# Overview
+> [Source code](https://github.com/ethen8181/Business-Analytics/blob/master/ab_test/frequentist_ab_test.R) to follow along the documentation.
 
@@
 
-# Characterizing a Proper Metric
-
-After walking through a simple A / B test example, we'll take a step back and see how to come up with a well-defined metric.
-
-## Generate a high level metrices
-
-For generating a high level metric, a good place to start is to think about the overall business objective.
-Below are a non-exhaustive list of techniques that might be able to help you come up with high level metrices that you might want to test out in your experiment.
-
-1. **External Sources :** For these types of data, do not *match* them with internal data that you might already have at hand. Instead, look at a time series of both your
-internally computed metric and the externally available one and see if the trends and seasonal variability line up.
-2. **Gathering your own data :**
-    - **User Experience Group :** With user experience group, you go really deep with a few users.
-It can take form of observing users doing tasks or you ask users to self-document their behaviors.
-After collecting the feedback it's usually a good idea to follow it up with a retrospective analysis.
-Meaning that if you observe something in a UER study, you could then go looking for that pattern in your data logs to see if that observation bears out and is worth creating a metric about.
-    - **Focus Group :** This is basically where you recruit a bunch of users or potential users and bring them together for a group discussion or you can ask them questions.
-Though you run the risk of the answers get dominated by "loud voices".
-    - **Survey :** The main advantage of surveys is that you can typically reach thousands of users.
-The disadvantages are that the data is self-reported and users don’t always give full or truthful answers.
-
-    Qualitative methods such as User Experience Group and Focus Group are much better suited for answering questions about why or how to fix a problem.
-Whereas quantitative methods like Surveys do a much better job answering how many and how much types of questions.
-Having such numbers from the survey helps prioritize resources and allows you to focus on the biggest issues at hand.
-
-## Turning it into a well-defined metrices
-
-Recall that in the examples above, we've chosen the high-level metric click-through probability, which is defined by the unique number of people how click the button versus the number of users who visited the page that the button was located.
-
-But to actually use implement the definition, we'll also have to address some other questions.
-Such as : If the same user visits the page once and comes back a week or two later, do we still only want to count that once?
-Usually, it is preferred that we count these separately, thus we'll also need to specify a time period.
-
-So given that users are usually recorded by cookies online.
-One version of the fully-defined metric will be : For each minute ( choose any time interval that you prefer ), the number of cookies that clicked divided by the number of cookies that interacted with the page.
-
-Another way of implementing this is to calculate : For each minute the number of pageviews ( assign each pageview with an id ) with a click divided by the number of pageviews.
-
-## Segmenting and filtering your traffic
-
-Creating segments and filtering out data can be useful for evaluating metric definitions, because, well, you can look at how the different definitions vary by segment.
-
-For example, you're looking at your metric of total active users over time and you see a spike in one of the timelines.
-After confirming that this is not caused by seasonal variation, we can look at different segment of our visitors to see if one of the segment is causing the spike.
-Suppose we have chosen segment to be geographic, it might just happen that we've identify a large proportion of the traffic is generated by a specific region and it might be best for us to dig deeper and understand why.
-
-To develop an understanding about your data and system, we can plot different metrices together, such as the rate ( calculated by the number of clicks divided by number of pageviews ) and probability by different segments to determine the differences in the output.
-
-## Summarizing Metric
-
-Apart from click-through probability, there're four other common summarizations that might be useful.
-
-1. **sums and counts** : e.g. how many users visited the homepage.
-2. **distributional metric** : e.g. quantiles. Mean age of users who completed the course ( though keed in mind that mean is sensitivity to outliers ).
-3.
-**probabilities and rates** : Recall the subtle differences between the definition of probabilities and rates here.
-Probability means that given a time interval, whether a person clicked on the button or not ( takes the number of 0 / 1 ) divided by the total number users that interacted with the page but did not click the button during that period of time.
-As for rate, it refers to calculating the total number of clicks divided by the total number of pageviews.
-4. **ratios.** e.g. the probability that a user clicks on a revenue generating click divided by any other click.
-
-## Measuring Sensitivity and Robustness
-
-It's the idea that you want to choose a metric that picks up the changes that you care about but robust against the changes that you don't care about, meaning that it doesn't move a lot when nothing interesting happens.
-It's an important thing to keep in mind when choosing the distributional metric as your summary metric.
-
-e.g.
-When testing out things like video latency, you plot different summary metric of comparable video's latency, such as the mean, 90 percentile.
-In theory there shouldn't be that much of a difference between these videos, if there is, then it means that the summary metric isn't robust enough.
-And suppose you increase the video quality across these comparable videos and the mean does not change at all, then this probably indicates that choosing the mean as the summary metric isn't sensitive enough.
-
-# Designing an Experiment
-
-Suppose you have chosen the suitable metrics to be used for the test, now it's time to apply that and work through the decisions you need to make when you're actually designing an experiment.
-This can be boiled down to three main part.
-
-1. Define what you mean as a "subject" in your experiment and control. In other words, what are the units in the population that you're going to be running the test on and comparing? 
-2. Choose the "populations". e.g. Everyone or only subjects in the U.S. ? You need to be sure that you're computing the metrices on equivalent populations ( This holds true for subject ).
-3. Determine the size of the experiment. This will also affect the duration of the experiment.
-
-## Define the Subject
-
-What you need to do is to decide how to assign events to either the control or the experiment.
-You can randomly assign every event, but if you have a user-visible change then this might not be a good idea.
-Since what will happen is : The user will see the change when they first visit the page and after reloading the page, the user will not see the change (leading to confusion).
-So for when conducting user-visible change, you want to assign people as oppose to events.
-
-The next question is how do we define a user? There're three commonly used categories, namely user id, anonymous id (cookie) and event.
-
-1.
-**user id :** e.g.
-Log in user names.
-Choosing this as the proxy for your user means that all the events that corresponds to the same user id are either in the control or experiment group, regardless of whether that user is switching between a mobile phone or desktop.
-This also means that if the user has not log in then he / she will neither be assgined to a control or experiment group.
-
-2.
-**anonymous id (cookie) :** The cookie is specific for a browser and device, thus if the user switches from Chrome to Firefox, they'll be assigned to a different cookie.
-Also note that users can clear the cookie, in which case the next time they visit the website they'll get assigned to a new cookie even if they're still using the same browser and device.
-For experiments that will be crossing the sign-in border, using cookie is preferred.
-e.g.
-Suppose you're changing the layout of the page or locations of the sign in bar then you should use a cookie.
-
-3. **event :** Should only be used when you're testing a non-user-visible change. e.g. page load time.
-
-## Define the Population
-
-If you think you can identify which population of your users will be affected by your experiment, you might want to target your experiment to that traffic ( e.g.
-changing features specific to one language's users ) so that the rest of the population won't dilute the effect.
-
-Next, depending on the problem you're looking at, you might want to use a **cohort** instead of a population.
-A cohort makes much more sense than looking at the entire population when testing out learning effects, examining user retention or anything else that requires the users to be established for some reason.
-
-A quick note on **cohort**.
-The gist of cohort analysis is basically putting your customers into buckets so you can track their behaviours over a period of time.
-The term cohort stands for a group of customers grouped by the timeline ( can be week, month ) where they first made a purchase ( can be a different action that's valuable to the business ).
-In a cohort, you can have roughly the same parameters in your two user group, which makes them more comparable.
-
-For example, you're an educational platform has an existing course that's already up and running.
-Some of the students have completed the course, some of them are midway through and there're students who have not yet started.
-If you want to change the structure of of one of the lessons to see if it improves the completion rate of the entire course and they started the experiment at time X.
-For students who have started before the experiment initiated they may have already finished the lesson already leading to the fact that they may not even see the change.
-So taking the whole population of students and running the experiment on them isn't what you want.
-Instead, you want to segment out the cohort, the group of customers, that started the lesson are the experiment was launched and split that into an experiment and control group.
-
-## Size and Duration
-
-You can size your experiment based on your significance level, power and minimum detectable change like we've mentioned in the example.
-But you can also decrease the size of your experiment by targeting experiment to specific traffic.
-Because you're only looking at a subset of your traffic, you might need a bigger change before it to matters to the business ( Recall that the smaller the change you wish to detect, the larger the sample size you’ll need ).
-On other note although this might not reduce the time frame of the experiment, but this allows other experiments to be run on the same time.
-
-After determining the size of the experiment it's time to translate that into a set of practical decisions including :
-
-1. When do I want to run the experiment.
-2.
-What fraction of the traffic are you going to send through the experiment.
-This will be closely tied to the duration of the experiment.
-There're some reasons that you might not want to run the experiment on all of your traffic to get the result faster.
-    - The first consideration might be you're just uncertained of how your users will react to the new feature, so you might want to test it out a bit before you get users blogging about it.
-To same notion applies to riskier cases, such as you're completely switching your backend system, if it doesn't work well, then the site might go down.
-    - Another consideration is that you may want to run the test a bit longer to make sure that the experiment captures the different user behaviour of weekdays, weekends.
-    - The final one is that you may be running multiple tests at the same time and if you wish to make them directly comparable ( different level or attributes of the same feature ), then the easiest way is to run them at the same time.
-So if there's a holiday during the timespan that you're running the experiment, then all of the experiments should be equally affected.
-
-# Analyzing Results
-
-## Sanity Checks
+# Sanity Checks
 
@@
 
-The result shows that 64454 does not lie within the range of the computed confidence interval and therefore it indicates that cookies may not be split equally.
+The result shows that 64454 does not lie within the range of the computed 95 percent confidence interval and therefore it indicates that cookies may not be split equally.
 
@@
 
-## Analyzing the Result
-
-One thing to keep in mind is that the goal of analyzing the result is not only about looking for statistical significance , but also about making a business decision about whether your experiment has favorably impacted your metric.
-So if you have an experiment that statistically improves the experience for thirty percent of the users but is neutral for everyone else, do you want to launch it as is, or do you want to try and make it better?
-
-**Some other takeaways: **
+**Some other takeaways:**
 
@@
 
-**Sources: **
-
-Source [code](https://github.com/ethen8181/Business-Analytics/blob/master/3_AB_test/AB_test.R) to the documentation.
-
-Bugs or comments can be filed [here](https://github.com/ethen8181/Business-Analytics/issues).
-
 # Reference
 
-1. Determining the sample size for A / B test : https://signalvnoise.com/posts/3004-ab-testing-tech-note-determining-sample-size
-2. Z-score distribution table : http://www.utdallas.edu/dept/abp/zscoretable.pdf
-3. Additional techniques to generating metrices : https://storage.googleapis.com/supplemental_media/udacityu/3954679115/additional_techniques.pdf
-4. Pitfalls of A / B test : http://www.forbes.com/sites/sujanpatel/2015/10/29/how-to-do-ab-testing-right-and-avoid-the-most-common-mistakes-marketers-make/#2715e4857a0b54f535a44349
-5. KissMetric Infographic on A / B testing : https://blog.kissmetrics.com/creating-effective-conversion-optimization-process/?utm_source=tofu-email&utm_medium=email
+1. [Determining the sample size for A / B test](https://signalvnoise.com/posts/3004-ab-testing-tech-note-determining-sample-size)
+2. [Z-score distribution table](http://www.utdallas.edu/dept/abp/zscoretable.pdf)
+3. [Additional techniques to brainstorm metrices](https://storage.googleapis.com/supplemental_media/udacityu/3954679115/additional_techniques.pdf) 
+4. [Pitfalls of A / B test](http://www.forbes.com/sites/sujanpatel/2015/10/29/how-to-do-ab-testing-right-and-avoid-the-most-common-mistakes-marketers-make/#2715e4857a0b54f535a44349)
+5. [KissMetric Infographic on A / B testing](https://blog.kissmetrics.com/creating-effective-conversion-optimization-process/?utm_source=tofu-email&utm_medium=email)
 

### 2016-05-06 articles/articles.Rmd
bandits version 1

@@
 
+- Labels are for cans not for people. [[Facebook Videos](https://www.facebook.com/morokcom/videos/237500609921393/)]
+- Practical Frameworks for Beating Burnout [[First Round Review](http://firstround.com/review/practical-frameworks-for-beating-burnout/?ct=t(How_Does_Your_Leadership_Team_Rate_12_3_2015))]
 - Land any job you want. [[blog](http://landanyjobyouwant.com/)]

### 2016-05-06 ab_tests/abtest.py

@@ -1,3 +1,4 @@
+# %matplotlib inline
 import pymc as pm
 import numpy as np
 import matplotlib.pyplot as plt

### 2016-05-06 bandits/bandits.py

@@ -1,0 +1,197 @@
+# %matplotlib inline
+import numpy as np
+from scipy.stats import beta
+import matplotlib.pylab as plt
+from collections import namedtuple
+
+
+class BetaBandit(object):
+	"""
+	Thompson Sampling
+	
+	Parameters
+	----------
+	K : int
+		total number of arms
+	
+	prior_params : list of float length 2 tuple, default None, (optional)
+		each element of the list is a tuple, where each tuple
+		contains the alpha and beta parameter that represents the prior
+		beta distribution for each arm. If not supplied
+		it will assume that all arms's prior starts with an uniform distribution
+    
+	Attributes
+	----------
+	trials, success : int 1d-array, shape( K, )
+		stores the trials and success for each arm,
+		e.g. trial = [ 1, 1 ] and success = [ 0, 1 ] means 
+		that both arm has been pulled once and arm 1 has generated
+		the reward (clicked)
+	"""
+
+	def __init__( self, K, prior_params = None ):
+		if prior_params:
+			priors = namedtuple( "priors", [ "alpha", "beta" ] )
+			prior  = [ priors(*p) for p in prior_params ]
+			self.alphas = np.asarray([ p.alpha for p in prior ])
+			self.betas  = np.asarray([ p.beta  for p in prior ])
+		else:
+			self.alphas = np.ones(K)
+			self.betas  = np.ones(K)
+
+		self.trials  = np.zeros( K, dtype = int )
+		self.success = np.zeros( K, dtype = int )
+		
+	
+	def get_recommendation(self):
+		"""
+		for all arms, construct their beta distribution and
+		draw a random sample from it, then return the arm
+		with the maximum random sample
+		"""
+		theta = np.random.beta( self.alphas + self.success, 
+								self.betas + self.trials - self.success )
+		return np.argmax(theta)
+	
+	
+	def add_result( self, arm, converted ):
+		"""
+		override the trials and success array, the success array
+		will only be updated if it has generated a reward
+		"""
+		self.trials[arm] += 1
+		if converted:
+			self.success[arm] += 1
+
+
+def experiment( T, K = None, ctr = None, prior_params = None ):
+	"""
+	run the experiment for Thompson Sampling.
[…]

### 2016-05-06 bandits/bandits.py
bandits done

@@ -153,18 +153,19 @@ def experiment_plot( ctr, trials, success ):
 	return fig
 
 
-def plot_beta_dist( ctr, trials, success, alphas, betas, turn ):
+def plot_beta_dist( ctr, trials, success, alphas, betas, turns ):
 	"""
 	Pass in the ctr, trials and success, alphas, betas returned
 	by the `experiment` function and the number of turns 
 	and plot the beta distribution for all the arms in that turn
 	"""
+	subplot_num = len(turns) / 2
 	x = np.linspace( 0.001, .999, 200 )
 	fig = plt.figure( figsize = ( 14, 7 ) ) 
 
 	for idx, turn in enumerate(turns):
 
-		plt.subplot( len(turns) / 2, 2, idx + 1 )
+		plt.subplot( subplot_num, 2, idx + 1 )
 
 		for i in range( len(ctr) ):
 			y = beta( alphas[i] + success[ turn, i ], 

### 2016-05-12 articles/articles.Rmd
scatter plot

--- 2016-04-18 articles/articles.Rmd
@@
 
+- 6 Best Practices for implementing In-store Personalization in 2015. [[beaconstac](http://blog.beaconstac.com/2015/07/6-best-practices-for-implementing-in-store-personalization-in-2015/)]
+    - Activating purchases: Say, Laura purchased a lemon yellow shirt at your retail store outlet located in New York.
+Now, a week later, you can send out an offer (via email and through a mobile app push notification) on a brown trouser that’ll go along with the shirt.
+The offer can be redeemed on your online store or a physical outlet.
+    - Pushing shoppers forward in the buying path: Say the same customer, Laura, actually went on to browse handbags online after she received the email offer.
+As soon she walks into your physical retail store, you can send a  notification through her mobile app, asking if she wishes to touch and feel the handbags that she browsed online.
+- How Big Data is changing e-commerce for good [[Future of Commerce](http://www.the-future-of-commerce.com/2014/12/15/how-big-data-is-changing-e-commerce-for-good/)]
+    - The top three things that retailers needs to know: how effective their pricing is, when shoppers are most active and what items they buy together (promoting products that shoppers often buy together can boost your average order value and help products move more quickly).
+- Do data scientists need to be domain experts to deliver good analytics?
+[[Data Science Central](http://www.datasciencecentral.com/profiles/blogs/do-data-scientists-need-to-be-domain-experts-to-deliver-good)]
+    - Often, good data scientists become subject experts just by playing with the data and asking questions to domain experts about the data anomalies.
+Through their questioning they will cross-pollinate the subject matter experts so the team as a whole is stronger.
+    - Machine learning brings a fresh perspective that leads to new insights and no prior domain knowledge can potentially be advantageous, especially in overcoming long standing domain bias.
+- Using Machine Learning to Predict Customer Behaviour [[Data Science Central](http://www.datasciencecentral.com/profiles/blogs/using-machine-learning-to-predict-customer-behaviour)]
+    - Used with Complaints Management, Customer Upsell and Customer Retention.
+Questions include How to anticipate and manage the dissatisfaction of the customers?', 'How to sell more and better to existing customers?' and 'How to keep the most profitable customers at an appropriate cost?'
+    - Action: You identify a target behaviour, in our case, it is a specific customer action, like filing a complaint/not filing a complaint, accepting or rejecting an upsell offer, cancelling / not cancelling a service.
+Then perform supervised learning
 - Doing Data Science Right — Your Most Common Questions Answered.

### 2016-05-12 articles/avoid_pie_charts/avoid_pie_charts.Rmd
scatter plot

--- 2016-03-16 ggplot2/plot.Rmd
@@
 ---
-title: "Collection of ggplot2 Plots"
+title: "Faceted Bar Charts"
 date: "`r Sys.Date()`"
@@
 
-# Avoiding Pie Charts
-
-Date: 2016.3.16
+> Source code can be obtained [here](https://github.com/ethen8181/Business-Analytics/blob/master/articles/avoid_pie_charts/avoid_pie_charts.R)
 
@@
 With that being said they can still be really bad at the one thing they're ostensibly designed to do.
-Consider the following pie chart from the [Wall Street Journal](http://www.wsj.com/articles/what-data-scientists-do-all-day-at-work-1457921541) article on What Data Scientists Do All Day at Work.
+Consider the following pie chart from the Wall Street Journal article on What Data Scientists Do All Day at Work.
 
-<img src = "/Users/ethen/Business-Analytics/ggplot2/avoid_pie_charts/data_scientist.png" width = "400" height = "800" >
+<img src = "/Users/ethen/Business-Analytics/articles/avoid_pie_charts/data_scientist.png" width = "400" height = "800" >
 
@@
 library(data.table)
-setwd("/Users/ethen/Business-Analytics/ggplot2/avoid_pie_charts")
-d <- fread("pie.txt")
+setwd("/Users/ethen/Business-Analytics/articles/avoid_pie_charts")
+d <- fread("avoid_pie_charts_data.txt")
 d <- melt( d, id.vars = "Task", 
@@
 ```
-
-## Plots
 
@@
 
-## Reference
+**Reference**
 
-1. Worst Chart In The World: http://www.businessinsider.com/pie-charts-are-the-worst-2013-6
-2. How to replace a pie chart: http://varianceexplained.org/r/improving-pie-chart/
-3. Take the Pies out of Pi Day: http://rud.is/b/2016/03/14/spinning-cycles-in-box-4-to-take-the-pies-out-of-pi-day/
-
-# Next Chart
+- [Worst Chart In The World](http://www.businessinsider.com/pie-charts-are-the-worst-2013-6)
+- [How to replace a pie chart](http://varianceexplained.org/r/improving-pie-chart/)
+- [Take the Pies out of Pi Day](http://rud.is/b/2016/03/14/spinning-cycles-in-box-4-to-take-the-pies-out-of-pi-day/)

### 2016-05-12 articles/nyt_scatter/nyt_scatter.Rmd
scatter plot

---
title: "Scatter Plot"
date: "`r Sys.Date()`"
author: "Ming-Yu Liu"
output:
  rmdformats::readthedown:
    highlight: haddock
---

> Simply focusing on creating a production ready scatter plot, without bothering about the interpretation. Source code can be obtained [here](https://github.com/ethen8181/Business-Analytics/blob/master/articles/nyt_scatter/nyt_scatter.R)

```{r, message=FALSE, warning=FALSE}

# 2016.5.12 you have to install the development version of ggplot2
# devtools::install_github("hadley/ggplot2") or subtitles won't work
library(tidyr)
library(dplyr)
library(readr)
library(scales)
library(ggplot2) 
setwd("/Users/ethen/Business-Analytics/articles/nyt_scatter")

file <- "nytimes_vote.tsv"
if( !file.exists(file) ){
	url <- "https://static01.nyt.com/newsgraphics/2016/04/21/undervote/ad8bd3e44231c1091e75621b9f27fe31d116999f/data.tsv"
	download.file( url, file )
}

# convert to more informative column name
df <- read_tsv(file)
df <- df %>% rename( "Someone else" = undervt, 
					 "Hilary Clinton" = clintonpct, 
					 "Bernie Sanders" = sanderspct )
df

df <- gather( select( df, -tvotes ), party, pct, -ratio, -fips )
df <- arrange( df, fips, ratio )
df <- df %>% mutate( 
	party = factor( party, levels = c("Hilary Clinton", "Bernie Sanders", "Someone else" ) )
)
df

```

```{r}

# 1. point shape = 21 is a doughnut circle
# 2. scale_fill_manual( name = "" ) gets rid of the legend title
# 3. theme's legend.key control the boxes around the legend's shape
fill_color <- c( 
	"Hilary Clinton" = "#5fa0d6",
	"Bernie Sanders" = "#83BC57",
	"Someone else" = "#d65454" 
)

ggplot( df, aes( x = ratio, y = pct ) ) + 
geom_point( aes( fill = party ), size = 3, alpha = 0.8, color = "white", shape = 21 ) + 
scale_fill_manual( name = "", values = fill_color ) +
theme_bw( base_family = "Arial Narrow" ) + 
scale_y_continuous( label = percent, limits = c( 0, 1.05 ) ) +
scale_x_continuous( limits = c( 0, 4.5 ), breaks = seq( 0, 4.5, 0.5 ) ) +
geom_text( data = data.frame( label = "↑ Share of 2016 primary vote"),
		   aes( x = 0, y = 1, label = label ), vjust = -1, hjust = 0, size = 3,
		   fontface = "bold", family = "Arial Narrow" ) + 
labs( x = "Ratio of registered Democrats to Obama voters →", 
	  y = NULL, title = "The Kinds of Places Sanders Beats Clinton",
	  subtitle = "Each dot on this chart represents the share of a county's vote for a candidate in the 2016 Democratic primary" ) +
theme( legend.key = element_blank(), 
	   legend.position = "top",
	   plot.title = element_text( face = "bold" ),
	   axis.ticks = element_blank(),
	   axis.text = element_text( size = 8 ),
	   axis.title.x = element_text( hjust = 1, face = "bold", size = 9 ),
	   panel.grid.minor = element_blank(),
	   panel.grid.major = element_line( linetype = "dotted", size = 0.5 ),
	   panel.border = element_blank(),
	   plot.margin = margin( t = 10, r = 10, b = 10, l = 10 ) )

```

**Reference**

- [Gist](https://gist.github.com/hrbrmstr/4632dc9cefc7566b412ec9945a6c4922)

### 2016-05-12 articles/nyt_scatter/nyt.scatter.R

@@ -1,0 +1,61 @@
+# 2016.5.12 you have to install the development version of ggplot2
+# devtools::install_github("hadley/ggplot2") or subtitles won't work
+library(tidyr)
+library(dplyr)
+library(readr)
+library(scales)
+library(ggplot2) 
+setwd("/Users/ethen/Business-Analytics/articles/nyt_scatter")
+
+
+file <- "nytimes_vote.tsv"
+if( !file.exists(file) ){
+	url <- "https://static01.nyt.com/newsgraphics/2016/04/21/undervote/ad8bd3e44231c1091e75621b9f27fe31d116999f/data.tsv"
+	download.file( url, file )
+}
+
+# convert to more informative column name
+df <- read_tsv(file)
+df <- df %>% rename( "Someone else" = undervt, 
+					 "Hilary Clinton" = clintonpct, 
+					 "Bernie Sanders" = sanderspct )
+
+df <- gather( select( df, -tvotes ), party, pct, -ratio, -fips )
+df <- arrange( df, fips, ratio )
+df <- df %>% mutate( 
+	party = factor( party, levels = c("Hilary Clinton", "Bernie Sanders", "Someone else" ) )
+)
+df
+
+# 1. point shape = 21 is a doughnut circle
+# 2. scale_fill_manual( name = "" ) gets rid of the legend title
+# 3. theme's legend.key control the boxes around the legend's shape
+fill_color <- c( 
+	"Hilary Clinton" = "#5fa0d6",
+	"Bernie Sanders" = "#83BC57",
+	"Someone else" = "#d65454" 
+)
+
+ggplot( df, aes( x = ratio, y = pct ) ) + 
+geom_point( aes( fill = party ), size = 3, alpha = 0.8, color = "white", shape = 21 ) + 
+scale_fill_manual( name = "", values = fill_color ) +
+theme_bw( base_family = "Arial Narrow" ) + 
+scale_y_continuous( label = percent, limits = c( 0, 1.05 ) ) +
+scale_x_continuous( limits = c( 0, 4.5 ), breaks = seq( 0, 4.5, 0.5 ) ) +
+geom_text( data = data.frame( label = "↑ Share of 2016 primary vote"),
+		   aes( x = 0, y = 1, label = label ), vjust = -1, hjust = 0, size = 3,
+		   fontface = "bold", family = "Arial Narrow" ) + 
+labs( x = "Ratio of registered Democrats to Obama voters →", 
+	  y = NULL, title = "The Kinds of Places Sanders Beats Clinton",
[…]

### 2016-05-24 articles/articles.Rmd
logic workflow
completely reorganized

@@
 
+- Reddit and Facebook Veteran On How to Troubleshoot Troublemakers.
+[[First Round Review](http://firstround.com/review/reddit-and-facebook-veteran-on-how-to-troubleshoot-troublemakers/?ct=t(How_Does_Your_Leadership_Team_Rate_12_3_2015))]
+    - Hermits. Get them to collaborate with a team.
+    - Nostalgia. Ask qualities that they missed most, and see if there're existing projects that could allow them to recapture that feeling.
+    - Trend Chaser. Ask them why this is the route their going to take? The potential has to worth more than the expense of experimenting.
 - Labels are for cans not for people. [[Facebook Videos](https://www.facebook.com/morokcom/videos/237500609921393/)]
@@
 
+- Data Pitfalls —the Startup Edition. [[Medium](https://medium.com/@mrogati/data-pitfalls-the-startup-edition-fe1eebcc39da#.si84k3dss)]
 - 6 Best Practices for implementing In-store Personalization in 2015. [[beaconstac](http://blog.beaconstac.com/2015/07/6-best-practices-for-implementing-in-store-personalization-in-2015/)]
@@
 
+- 'Amazon is not your friend'. [[Linkedin](https://www.linkedin.com/pulse/amazon-your-friend-caroline-fairchild)]
+    - The only way to survive is to build a brand that can grow without Amazon’s distribution.
+    - Amazon hasn’t proven that it can build inspirational brands.
+While they have mastered the ability to get customers commodity products, no one goes to Amazon to discover something they didn’t already know about.
 - Facebook is the new Excel. [[blog](http://alexmuir.com/facebook-is-the-new-excel)]

### 2016-05-24 bayesian_statistics/MCMC/MCMC.Rmd
bandits 1
completely reorganized

@@
 We then take those draws and calculate quantities of interest for the (posterior) distribution.
-Here, we'll bw attempting to give you an intuition behind one branch of the MCMC sampling, namely the **Metropolis–Hastings Algorithm**.
+Here, we'll be attempting to give you an intuition behind one branch of the MCMC sampling, namely the **Metropolis–Hastings Algorithm**.
 
@@
 
-			# propose a new sample point (will elaborate on this later)			
+			# propose a new sample point		
 			proposal <- sample + rnorm( n = 2, mean = 0, sd = 1 )

### 2016-05-24 bandits/bandits.py

@@ -33,8 +33,8 @@ class BetaBandit(object):
 		if prior_params:
 			priors = namedtuple( "priors", [ "alpha", "beta" ] )
 			prior  = [ priors(*p) for p in prior_params ]
-			self.alphas = np.asarray([ p.alpha for p in prior ])
-			self.betas  = np.asarray([ p.beta  for p in prior ])
+			self.alphas = np.array([ p.alpha for p in prior ])
+			self.betas  = np.array([ p.beta  for p in prior ])
 		else:
 			self.alphas = np.ones(K)
 			self.betas  = np.ones(K)

### 2016-05-24 marketing_analysis/code/marketing_analysis.R

@@ -153,61 +153,9 @@ revenue_summary <- revenue_predicted[ , .( avg_revenue 	 = mean(revenue),
 				   arrange( desc(total_revenue), desc( avg_revenue) )
 revenue_summary
 
-# ----------------------------------------------------------------------------
-# 							Customer life time value   
-# ----------------------------------------------------------------------------
 
-
-rfm3 <- data[ recency > 365, 
-			  .( recency 	  	= min(recency) - 365,
-				 first_purchase = max(recency) - 365,
-				 frequency  	= .N,
-				 total_amount   = sum(purchase_amount), 
-				 avg_amount 	= mean(purchase_amount) ), by = customer_id ]
-
-revenue_predicted <- merge( rfm3, revenue_per_customer, by = "customer_id", all.x = TRUE )
-revenue_predicted[ is.na(revenue), revenue := 0 ]
-
-# active column : whether they've spent money on the current period 
-revenue_predicted[ , active := as.numeric( revenue > 0 ) ]
-
-
-# likelihood whether they'll be active 
-formula <- paste( "active", 
-	   	   paste( "recency", "first_purchase", "frequency", 
-	   		      "avg_amount", sep = " + " ), sep = " ~ " ) %>% as.formula()
-model1 <- glm( formula, data = revenue_predicted )
-model1 <- update( model1, . ~ .-first_purchase )
-summary(model1)
-
-1- ( model1$deviance / model1$null.deviance )
-
-
-# active distribution 
-table( revenue_predicted$active )
-
-# predict the price for those that are active 
-active <- revenue_predicted[ active == 1, ]
-
-model2 <- lm( revenue ~ first_purchase + avg_amount + total_amount, data = active )
-summary(model2)
-
-# ----------------------------------------------------------------------------
-# 										Scoring Model  
-# ----------------------------------------------------------------------------
-
-# calculating the p-value by hand: coefficient / standard deviation of coefficient 
-# http://stats.stackexchange.com/questions/9715/how-to-set-up-and-estimate-a-multinomial-logit-model-in-r
-# https://www.researchgate.net/post/Significance_of_Regression_Coefficient
-# predict probability and amount 
-
-# predictors: information of the customers a year ago 
-# target variable: probability that the customer will buy something
-# and if they do, how much they will spend on it 
-
-
-# to predict the probability, we use the entire dataset as our calibration model
-# as for the amount, we use the customers that have had a purchase record 
-
[…]

### 2016-05-31 ab_tests/frequentist_ab_test.Rmd
fix doc
update ab testing

--- 2016-04-22 ab_test/frequentist_ab_test.Rmd
@@
 ---
-title: "Online A / B test"
+title: "Frequentist A/B test"
 author: "Ming-Yu Liu"
@@
 
-> [Source code](https://github.com/ethen8181/Business-Analytics/blob/master/ab_test/frequentist_ab_test.R) to follow along the documentation.
+> [Source code](https://github.com/ethen8181/Business-Analytics/blob/master/ab_tests/frequentist_ab_test.R) to follow along the documentation.
 
-A / B testing is a general methodology used online to when you want to test out a new feature.
-What you're doing is you're going show one set of features, the control set ( your existing feature ) to one user group and another set, your experiment set ( your new feature ) to another user group and test how did these users respond differently so that you can determine which set of your feature is better.
+# Getting Started
 
-Despite its useful functionality, there are still places where A / B testing isn't as useful. For example :
+## Hypothesis Testing
 
-1.
-Testing out completely new experiences.
-An obvious way for thinking about this is if you're dramatically changing the user's experience.
-The user might prefer the old way of doing things or they've curious about the new features and they simply test out everything ( sometimes called the novelty effect ).
-2.
-A / B testing can't tell you if you're missing something.
-Meaning it can tell you if A performs better B or vice versa, but it can't tell you that if you use C, then it will actually perform better than the former two.
-3.
-Tesing out products that people rarely buy.
-e.g.
-cars, apartments.
-It might be too long before the user actually decides to take actions after seeing the information and you might be unaware of the actual motivation.
+In order to understand classic frequentist A/B testing, you have to understand the basics of frequentist hypothesis testing.
+We’ll use a classic "intervention" example found in most industries as a motivation.
+Let’s say we want to conduct an experiment to test if a certain action helps prevent the occurrence of an adverse event.
+Depending on the business, this event could be fraud, attrition, downgrades, account closure, etc.
+The goal is to contact just enough people to decide if the action should be launched at full scale.
+
+**Step 1 - Create a null hypothesis and an alternative hypothesis**
+
+- Null hypothesis $(H_0)$: the event rates are the same for the test/experiment and control groups.
+- Alternative hypothesis $(H_1)$: the event rates are different
+
+Note that the event rate refers to the proportion of clients for whom the event is observed (e.g., fraud, attrition, downgrades, etc.).
+
+The null hypothesis is the hypothesis of no impact, while the alternative hypothesis is that the action will have an impact.
+To be explicit, this is a two-sided hypothesis, since we are testing if the event rate improved or worsened.
+
+**Step 2 - Draw a sample**
+
+Let’s say we draw a sample of N clients where 50% are exposed to the action and 50% are not. Note that the split is random, but does not have to be exactly 50/50. We’ll learn how to choose N later.
+
+**Step 3 - Calculate the test statistic**
+
+Select the appropriate test statistic and calculate its value from the data (we will call it ZZ).
+A test statistic is a single metric that can be used to evaluate the null hypothesis.
+For our test the underlying metric is a binary yes/no variable (event), which means that the appropriate test statistic is a test for differences in proportions.
+Math formula:
+
+$$Z = \frac{ p_{experiment} - p_{control} }
+{ \sqrt{ p_{pooled}( 1-p_{pooled} )( \frac{1}{n_{experiment}} + \frac{1}{n_{control}}) } }$$
+
+Where:
+
+- $p_{control}$ and $p_{experiment}$ are the event rates for the control and experiment groups, respectively.
+- $n_{control}$ and $n_{experiment}$ are the sample sizes for the control and experiment groups, respectively.
+- $p_{pooled}$ is the blended rate: $( x_{control} + x_{experiment} ) \big/ ( n_{control} + n_{experiment} )$ where $x_{control}$ and $x_{experiment}$ are event counts for the control and experiment groups, respectively.
+
+For testing of continuous variables, such as revenue, see the appendix of this post.
+
+**Step 4 - Reject or accept the null hypothesis**
+
+Based on the test statistic we can calculate the **p-value** which is probability of obtaining a result at least as "extreme" as the observed result by random chance, given that the null hypothesis is true.
+It is tempting to interpret the **p-value** as the probability of rejecting the null hypothesis when it is true, but technically speaking that is an incorrect interpretation, in fact this is what's called the Type I error (detail later).
+
+Here's an concrete example, suppose that a vaccine study produced a **p-value** of 0.04.
+This **p-value** indicates that if the vaccine had no effect, you’d obtain the observed difference or more in 4% of studies due to random sampling error.
+**p-value** address only one question: how likely are your data, assuming a true null hypothesis?
+It does not measure support for the alternative hypothesis!!!!
+Thus, while a low **p-value** indicates that your data are unlikely assuming a true null hypothesis, it can’t evaluate which of two competing cases is more likely:
+
+- The null is true but your sample was unusual.
+- The null is false.
+
+Determining which case is more likely requires subject area knowledge and replicate studies.
+
+If we seek to reject the null hypothesis, we want the **p-value** to be small and the typical threshold used is 5%.
+In other words, if the **p-value** is less than 5%, and the test group experienced a lower event rate than the control group, we conclude that the action worked.
+In other words, we are reasonably sure that there is something besides chance alone that gave us the observed data.
+The pre-chosen cutoff (5%) is also referred to as the **significance level** and plays an important role in determining the required sample size (detail later).
+
+Note that we do not have to directly calculate the p-value in order to reject or accept the null hypothesis.
+Alternatively, we can apply a cutoff directly to the test statistic (Z) based on its distribution and the chosen significance level.
+For example, for a two-sided test and a significance level of 5%, the cutoff corresponds to the upper and lower 2.5% on the standard normal distribution (normal distribution with a mean of 0 and a standard deviation of 1), which is 1.96.
+Hence we reject the null hypothesis if $\lvert Z \rvert > 1.96$.
+
+**Introducing the Power and the Significance Level**
+
+In the world of hypothesis testing, rejecting the null hypothesis when it is actually true is called a type 1 error.
+Committing a type 1 error is a false positive because we end up recommending something that does not work.
+
+Conversely, a type 2 error occurs when you accept the null hypothesis when it is actually false.
+This is a false negative because we end up sitting on our hands when we should have taken action.
+We need to consider both of these types of errors when choosing the sample size.
+
+Two important probabilities related to type 1 and type 2 error are:
+
+- **Significance level** governs the chance of a false positive.
+A significance level of 0.05 means that there is a 5% chance of a false positive.
+Choosing level of significance is an arbitrary task, but for many applications, a level of 5% is chosen, for no better reason than that it is conventional.
+- **Statistical power** represents the probability that you’ll get a false negative.
+A power of 0.80 means that there is an 80% chance that if there was an effect, we would detect it (or a 20% chance that we’d miss the effect).
+There are no formal standards for power, most researchers assess the power of their tests using 0.80 for adequacy.
+
+| Scenario       | $H_0$ is true | $H_0$ is false |
+|:--------------:|:-------------:|:-----------:|
+|  Accept $H_0$  |  Correct Decision  |   Type 2 Error (1 - power)   |
+|  Reject $H_0$  |  Type 1 Error (significance level) |   	Correct decision   |
+
+The concepts of power and significance level can seem somewhat convoluted at first glance.
+A good way to get a feel for the underlying mechanics is to plot the probability distribution of Z assuming that the null hypothesis is true.
+Then do the same assuming that the alternative hypothesis is true, and overlay the two plots.
+
+Consider the following example:
+
+- $H_0: p1 = p2$, $H_1: p 1> p2$. A one-sided test was chosen here for charting-simplicity.
+Our chosen significance level is 5%. The corresponding decision rule is $\lvert Z \rvert > 1.65$. The number (1.65) is the cutoff that corresponds to the upper 5% on the standard normal distribution.
+- Total sample size, N=5,000 (assume equal sample sizes for the control and experiment groups, meaning exactly 2,500 in each group).
+- Say we decide that we need to observe a difference of 0.02 (detailed later) in order to be satisfied that the intervention worked (i.e., $p_1 = 0.10$ and $p_2 = 0.08$).
+We will discuss how to make this decision later in the post.
+The desired difference of 0.02 under the alternative hypothesis corresponds to $Z = 2.47$ (using the formula for Z above).
+If you wish to convert the Z-score that is found to **p-value**, you can do it by one minus the associated probability.
+Also, for a two sided test we need to multiply the result by two.
+`2 * ( 1- pnorm( abs(Z) ) )`.
+
+```{r, message=FALSE, warning=FALSE}
+library(scales)
+library(ggplot2)
+library(data.table)
+
+PlotPower <- function( size, min_diff ){
+
+	size_a <- size_b <- size * 0.5 # size are assumed to be equal
+	p_a <- 0.08 # baseline
+	p_b <- p_a + min_diff
+	count_a  <- size_a * p_a
+	count_b  <- size_b * p_b
+	p_pooled <- ( count_a  + count_b ) / ( size_a + size_b )
+	Z <- ( p_b - p_a ) / sqrt( p_pooled * ( 1 - p_pooled ) * ( 1 / size_a + 1 / size_b ) )
+
+	# Z corresponds to the mean of the normal distribution
+	mean1 <- 0
+	mean2 <- Z
+
+	x <- seq( -4, 6, 0.1 ) # use for generating the x axis of the normal distribution
+	data <- data.frame( x = x, y1 = dnorm( x, mean1, 1 ), y2 = dnorm( x, mean2, 1 ) )
+
+	plot <- ggplot( data, aes( x = x ) ) +
+			geom_line( aes( y = y1, colour = 'H0 is true' ), size = 1.2 ) +
+			geom_line( aes( y = y2, colour = 'H1 is true' ), size = 1.2 ) +
+			geom_area( aes( y = y1, x = ifelse( x > 1.65, x, NA ) ), fill = 'black' ) +
+			geom_area( aes( y = y2, x = ifelse( x > 1.65, x, NA ) ), fill = 'blue', alpha = 0.3 ) +
+			labs( x = '', y = '', title = sprintf( 'p1 = %s, p2 = %s, size = %d', p_a, p_b, size ) ) + 
+			theme( legend.title = element_blank() ) +
+			scale_colour_manual( breaks = c( "H0 is true", "H1 is true" ), 
+								 values = c( "blue", "red" ) )
+	return(plot)
+}
+```
+
+```{r, message=FALSE, warning=FALSE}
+
+PlotPower( size = 5000, min_diff = 0.02 )
+
+```
+
+The shaded dark blue area denotes the significance region, while the the shaded blue area denotes the power.
+Note that if we pick a smaller N, or a smaller probability difference between the control and experiment group, the power drops (the shaded blue area decreases), meaning that if there’s is in fact a change, there’s lesser percent chance that we’ll detect it:
+
+```{r, message=FALSE, warning=FALSE}
+
+# smaller N
+PlotPower( size = 2500, min_diff = 0.02 )
+
+# smaller probability difference
+PlotPower( size = 5000, min_diff = 0.01 )
+
+```
+
+## Power Analysis
+
+Let’s say we've followed the rule of thumb and require the significance level to be 5% and the power to be 80%.
+This means we have now specified two key components of a power analysis.
+Our next task now is to find the sample size that meets these two criteria.
+To solve this, we also need to specify the detectable difference.
+The detectable difference is the level of impact we want to be able to detect with our test.
+
+Let’s go back to the definition of power: the power is the probability of rejecting the null hypothesis when it is false.
+Hence for us to calculate the power, we need to define what "false" means to us in the context of the study.
+In other words, how much impact, i.e., difference between test and control, do we need to observe in order to reject the null hypothesis and conclude that the action worked?
+
+Let’s consider two illustrative examples: if we think that an event rate reduction of, say, $10^{-10}$ is enough to reject the null hypothesis, then we need a very large sample size to get a power of 80%.
+This is pretty easy to deduce from the charts above: if the difference in event rates between test and control is a small number like $10^{-10}$, the null and alternative probability distributions will be nearly indistinguishable.
+Hence we will need to increase the sample size in order to move the alternative distribution to the right and gain power.
+Conversely, if we only require a reduction of 0.02 in order to claim success, we can make do with a much smaller sample size.
+
+> In sum, the smaller the detectable difference, the larger the required sample size.
+
+Here's how you would contact a power test. Note that the printed result is the sample size needed for each group!!
+
+```{r}
+
+baseline  <- 0.1  # baseline conversion rate 
+delta 	  <- 0.02 # minimum detectable boundary (practical significance boundary)
+power 	  <- 0.8  # sensitivity 
+sig_level <- 0.05 # specificity
+
+result <- power.prop.test( p1 = baseline, p2 = baseline - delta, 
+				 		   power = power, sig.level = sig_level,
+				 		   alternative = "two.sided" )
+round(result$n)
+
+```
+
+Unlike the significance level and the power, there are no plug-and-play values we can use for the detectable difference.
+The key is to define what "pay off" means for the study at hand, which depends on what the adverse event is a well as the cost of the action.
+Two guiding principles:
+
+- **Avoid wasteful sampling** Let’s say it takes an absolute difference of 0.02 between test and control in order for the treatment to pay off.
+In this case, aiming for a 0.01 detectable difference would just lead to more precision than we really need.
+Why have the ability to detect 0.01 if we don’t really care about a 0.01 difference?
+In many cases, sampling for unnecessary precision can be costly and a waste of time.
+- **Avoid missed opportunities** Conversely, if we are analyzing a sensitive metric where small changes can have a large impact e.g.
+email campaigns, we have to aim for a small detectable difference.
+If we choose an insufficient sample size, we may end up sitting on our hands and missing an opportunity (type 2 error).
+
+Hence, choosing the minimum detectable difference should be a cross-functional analysis/discussion between the data scientist and the business stakeholder.
+Once there is a viable range for the detectable difference, we can evaluate the sample size required for each option.
+For example, let’s say that $p1=0.10$ and we want the detectable difference to be between 0.01 and 0.03.
+Clearly, we’d rather be able to detect a difference of 0.01, but it may be too costly and hence we want to evaluate more conservative options as well.
+
+```{r}
+
+baseline  <- 0.1
+power 	  <- 0.8
+sig_level <- 0.05
+
+# calculate the the required sample size for reaching the 
+# range detectable difference (dd)
+dd <- seq( from = 0.01, to = 0.03, by = 0.0001 )
+result <- matrix( nrow = length(dd), ncol = 2 )
+result[ , 1 ] <- dd
+for( i in 1:length(dd) ){
+	result[ i, 2 ] <- power.prop.test( p1 = baseline, p2 = baseline - dd[i], 
+									   power = power, sig.level = sig_level,
+									   alternative = "two.sided" )$n
+}
+
+result <- data.table(result)
+setnames( result, c( 'dd', 'n' ) )
+
+ggplot(data = result, aes( x = dd, y = n ) ) +
+geom_line() + ylab("required sample size") + xlab("Detectable difference") + 
+scale_x_continuous( labels = comma )
+
+```
+
+From the graph, we can see that we need roughly 10x more observations to get a detectable difference of 0.01 compared to 0.03.
 
@@
 
-When launching a A / B test you need to ask yourself :
+When launching a A/B test you need to ask yourself:
 
 1. What is your hypothesis and what's the baseline for comparison. 
-2. What's the confidence level that you wish to have in your result ( This will also affect the number of samples ). 
+2. What's the confidence level that you wish to have in your result (This will also affect the number of samples ). 
 2. How many samples and time do you need in order to actually have your user adapt to the new experience.
@@
 
-**The null hypothesis, H0: ** The experimental and control groups have the same probability of completing a checkout ( clicking the button ).
+- $H_0$: The experimental and control groups have the same probability of completing a checkout ( clicking the button ).
 Or equivalent to saying that the differences of the two groups' probability is 0.
-**The alternative hypothesis, H1: ** The two groups have different probability of completing a checkout.
+- $H_1$: The two groups have different probability of completing a checkout.
 
@@
 To do that we can use a power analysis for two independent samples, which can be calculated using the `power.prop.test` function.
-Before doing that there are some basic parameters that we should be aware of :
-
-1. **Significance Level:** Denoted by $\alpha$, this threshold governs the chance of false positive. A significance level of 0.05 means that there's a 5 percent chance of false positive.
-2.
-**Statistical power:** This stands for $1 - \beta$ ( or so called sensitivity ), where $\beta$ represents the probability that you'll get a false negative.
-So a statistical power of 0.8 means that if there's is in fact a change, there's 80 percent chance that we'll detect it, which is equivalent to saying that there will be a 20 percent false negative.
-
-The effect of picking a significance level of 0.05 and power of 0.8 means that we are 4 times more likely to obtain a false negative than a false positive.
-For A / B testing, we’re generally more concerned about getting a false positive : Making a change that doesn’t actually improve things than we are about not making a change at all, which is why we accept a greater likelihood of a false negative.
-The rule of thumb for $\alpha$ and $1 - \beta$ is 0.5 and 0.8 respectively.
 
 Now suppose that our current baseline is 0.1 ( there's a 10 percent chance that people who saw the button will click it ).
-And we wish to detect a change of two percent in the click through rate ( This is already consider quite high for online experiment ).
+And we wish to detect a change of two percent in the click through rate (This is already consider quite high for online experiment).
 
-Parameters :
+Parameters:
 
@@
 
-How these parameters affect the sample size you need :
+Quick note on how these parameters affect the sample size you need:
 
@@
 
-# Analyze Quick Example's Result
+## Analyze Quick Example's Result
 
@@
 To do so, we'll calculate the confidence interval for the difference, which can be done by calculating the difference plus and minus the z score times the standard deviation.
-Math Formula :
+Math Formula:
+
 $$p_{experiment} - p_{control} \pm Z \sqrt{ p_{pooled}( 1-p_{pooled} )( \frac{1}{n_{experiment}} + \frac{1}{n_{control}} ) }$$
 
-Where you can calculate the $p_{pooled}$, the pooled probability by the sum of the number of successes for both groups divided by the sum of the total number of sample size for both groups.
+Note that this is the same formula from the first section, except now we're looking at the confidence interval.
 
-[`ABTest`][ABTest] Function that calculates the confidence interval. Input parameters :
+[`ABTest`][ABTest] Function that calculates the confidence interval. Input parameters:
 
-- `count_control` The number of successes. This is equivalent to the number of people that clicked the button for the control group ( your original feature ).
+- `count_control` The number of successes. This is equivalent to the number of people that clicked the button for the control group (your original feature).
 - `sizes_control` The total number of sample size for the control group.
@@
 
-	# @p : pooled probability
-	# @std_error pooled standard deviation (error ) 
+	# p: pooled probability
+	# std_error: pooled standard deviation (error) 
 	p <- ( count_control + count_experiment ) / ( sizes_experiment + sizes_control )
@@
 
-	# 95 percent confidence interval's z score = 1.96, equivalent to 
-	# qnorm( 0.975 )
+	# 95 percent confidence interval's z score = 1.96, equivalent to qnorm(0.975)
 	difference <- p_experiment - p_control
-	confidence <- difference + c( -1, 1 ) * 1.96 * std_error
+	confidence <- difference + c( -1, 1 ) * qnorm(0.975) * std_error
 
@@
 
-# Sanity Checks
-
-Before jumping into analyze the result, we should check that the population size metric and any other metric that you don't expect to change shouldn't change while you're running the experiment, or in a sense that they're invariant and shouldn't differ between the experiment and control group.
-
-Checking invariants example :
+## Sanity Checks
 
 For instance, after running your experiment for a week, you've discovered that the total number of users assigned to the control group is 64454 and the total number of users assigned to the experiment group 61818.
-How would you figure out whether the difference is within expectation given that each user is randomly assigned to the control or experiment group with a probability of 0.5 ?
+How would you figure out whether the difference is within expectation given that each user is randomly assigned to the control or experiment group with a probability of 0.5?
+It's usually good idea to check this.
 
-This is equivalent to saying out of a total 126272 ( 64454 + 61818 ) users, is it surprising to see if 64454 users are assigned to the control group, which is essentially a binomial distribution.
-Knowing this information, we can construct a confidence interval to test if the number lies within the confidence interval.
+This is equivalent to saying out of a total 126272 (64454 + 61818) users, is it surprising to see if 64454 users are assigned to the control group?
+This is essentially a binomial distribution, thus, knowing this information, we can construct a confidence interval to test if the number lies within the confidence interval.
 The confidence interval can be calculated by the mean plus and minus the z-score times the standard error.
@@
 {
+	# 95 percent confidence interval = qnorm(0.95 + (1-0.95) / 2 )
 	n <- group1 + group2
-	confidence <- n * 0.5 + c( -1, 1 ) * 1.96 * sqrt( n * 0.5 * 0.5 ) 
+	confidence <- n * 0.5 + c( -1, 1 ) * qnorm(0.975) * sqrt( n * 0.5 * 0.5 ) 
 	return(confidence)
@@
 
-**Some other takeaways:**
-
-A / B testing is an iterative process, some common mistakes that you want to avoid is :
-
-1.
-Falling into the trap of “We already tried that”.
-A hypothesis can be implemented in so many different ways.
-If you just do one headline test and say "we tried that," you’re really selling yourself short.
-2.
-Not testing continually or not retesting after months or years.
-Just because you tested a variation in the past doesn’t necessarily mean that those results are going to be valid a year or two from now.
-3.
-Optimizing for the top of the funnel, rather than the product.
-Understanding what the customers want so that you can make the product better.
-Ultimately, you can’t simply test your headlines and get people to like your product more.
-
 # Reference
 
-1. [Determining the sample size for A / B test](https://signalvnoise.com/posts/3004-ab-testing-tech-note-determining-sample-size)
-2. [Z-score distribution table](http://www.utdallas.edu/dept/abp/zscoretable.pdf)
-3. [Additional techniques to brainstorm metrices](https://storage.googleapis.com/supplemental_media/udacityu/3954679115/additional_techniques.pdf) 
-4. [Pitfalls of A / B test](http://www.forbes.com/sites/sujanpatel/2015/10/29/how-to-do-ab-testing-right-and-avoid-the-most-common-mistakes-marketers-make/#2715e4857a0b54f535a44349)
-5. [KissMetric Infographic on A / B testing](https://blog.kissmetrics.com/creating-effective-conversion-optimization-process/?utm_source=tofu-email&utm_medium=email)
+- [Stitchfix's Blog: So, You Need a Statistically Significant Sample?](http://multithreaded.stitchfix.com/blog/2015/05/26/significant-sample/)
+- [Z-score distribution table](http://www.utdallas.edu/dept/abp/zscoretable.pdf)
+- [Minitab's Blog: How to Correctly Interpret P Values](http://blog.minitab.com/blog/adventures-in-statistics/how-to-correctly-interpret-p-values)
 
@@
 
-[ABTest]: https://github.com/ethen8181/Business-Analytics/blob/master/3_AB_test/AB_test.R
+[ABTest]: https://github.com/ethen8181/Business-Analytics/blob/master/ab_tests/frequentist_ab_test.R

### 2016-05-31 bayesian_statistics/bayes/bayes.Rmd
to do mcmc
update ab testing

@@
 Credible intervals are analogous to confidence intervals in frequentist statistics, although they differ on a philosophical basis; Bayesian intervals treat their bounds as fixed and the estimated parameter as a random variable, whereas frequentist confidence intervals treat their bounds as random variables and the parameter as a fixed value.
-For a more detailed explanation, refer to this great [StackExchange post](http://stats.stackexchange.com/questions/2272/whats-the-difference-between-a-confidence-interval-and-a-credible-interval).
+For a quick review on confidence interval, refer to the beginning of this [Stitchfix blog post](http://multithreaded.stitchfix.com/blog/2015/02/12/may-bayes-theorem-be-with-you/).
 
@@
 4. [Understanding the Bayesian Approach to False Discovery Rates](http://varianceexplained.org/r/bayesian_fdr_baseball/)
+
+# R Session Information
+
+```{r}
+
+sessionInfo()
+
+```

### 2016-05-31 ab_tests/frequentist_ab_test.R

@@ -1,16 +1,84 @@
+library(scales)
+library(ggplot2)
+library(data.table)
 
-# campaign response rate testing
-# Problem definition:
+PlotPower <- function( size, min_diff ){
+	
+	size_a <- size_b <- size * 0.5 # size are assumed to be equal
+	p_a <- 0.08 # your current baseline solution
+	p_b <- p_a + min_diff
+	count_a  <- size_a * p_a
+	count_b  <- size_b * p_b
+	p_pooled <- ( count_a  + count_b ) / ( size_a + size_b )
+	Z <- ( p_b - p_a ) / sqrt( p_pooled * ( 1 - p_pooled ) * ( 1 / size_a + 1 / size_b ) )
 
-# we're sending out number of sales offers at one or more chosen price-points 
-# to different populations. 
-# we want to measure the response rate at each different price level
-# so that we can make a good choice of which price to continue with the future
-# sales campaigns
+	# Z corresponds to the mean of the normal distribution
+	mean1 <- 0
+	mean2 <- Z
+
+	x <- seq( -4, 6, 0.1 ) # use for generating the x axis of the normal distribution
+	data <- data.frame( x = x, y1 = dnorm( x, mean1, 1 ), y2 = dnorm( x, mean2, 1 ) )
+
+	plot <- ggplot( data, aes( x = x ) ) +
+			geom_line( aes( y = y1, colour = 'H0 is true' ), size = 1.2 ) +
+			geom_line( aes( y = y2, colour = 'H1 is true' ), size = 1.2 ) +
+			geom_area( aes( y = y1, x = ifelse( x > 1.65, x, NA ) ), fill = 'black' ) +
+			geom_area( aes( y = y2, x = ifelse( x > 1.65, x, NA ) ), fill = 'blue', alpha = 0.3 ) +
+			labs( x = '', y = '', title = sprintf( 'p1 = %s, p2 = %s, size = %d', p_a, p_b, size ) ) + 
+			theme( legend.title = element_blank() ) +
+			scale_colour_manual( breaks = c( "H0 is true", "H1 is true" ), 
+								 values = c( "blue", "red" ) )
+	return(plot)
+}
+
+PlotPower( size = 5000, min_diff = 0.02 )
+
+# smaller N
+PlotPower( size = 2500, min_diff = 0.02 )
+
+# smaller probability difference
+PlotPower( size = 5000, min_diff = 0.01 )
+
+
[…]

### 2016-06-02 ab_tests/frequentist_ab_test.R
add ab test for continous results later

@@ -155,8 +155,6 @@ labs( title = "Different Scenarios of Confidence Interval",
 	  x = "confidence interval" ) 
 
 
-
-
 # sanity checks
 group1 <- 64454
 group2 <- 61818
@@ -174,6 +172,6 @@ power.t.test(sig.level = a, d = es[i], sd = 1,
                                   alternative = 'two.sided', power = b)$n
 
 # ----------------------------------------------------------------------------
-# rule of thumb
+# rule of thumb for online testing, future reference
 # http://www.exp-platform.com/Documents/2014%20experimentersRulesOfThumb.pdf
 

### 2016-07-02 ab_tests/frequentist_ab_test.Rmd
data hygiene, increase rmarkdown font size

@@
 ---
+
+<style type="text/css">
+p{ /* Normal  */
+   font-size: 18px;
+}
+body{ /* Normal  */
+   font-size: 18px;
+}
+td {  /* Table  */
+   font-size: 14px;
+}
+h1 { /* Header 1 */
+ font-size: 32px;
+}
+h2 { /* Header 2 */
+ font-size: 26px;
+}
+h3 { /* Header 3 */
+ font-size: 22px;
+}
+code.r{ /* Code block */
+  font-size: 14px;
+}
+pre { /* Code block */
+  font-size: 14px
+}
+</style>
 

### 2016-07-02 articles/articles.Rmd
data hygiene, increase rmarkdown font size

@@
 
+- 5 Qualities You Need to Show in the Job Search.
+[[glassdoor](https://www.glassdoor.com/blog/qualities-show-off/?utm_source=newsletter&utm_medium=email&utm_content=Top5Jobs_US16&utm_campaign=May2016_US)]
+    - Leadership ability. e.g. As a result of an incentive program I spearheaded, our sales team surpassed its quarterly sales goal by 15 percent.
+    - Flexibility: keep track of times when you’ve had to assume multiple roles or responsibilities outside of your own.
+    - Coachability: Be open and honest about where you fall short and how you’re actively improving on those shortcomings.
+- From Developer to Manager - PyCon 2016. [[Youtube](https://www.youtube.com/watch?v=lzBm0r8CxhY)]
 - Reddit and Facebook Veteran On How to Troubleshoot Troublemakers.
@@
 
+- Check Yourself: 5 Things You Should Know About Data Science (Author Note). [[Blog](http://analyticsmadeskeezy.com/2012/11/05/check-yo-self-5-things-you-should-know-about-data-science-author-note/)]
+- Statistics for Hackers - PyCon 2016 [[Youtube]](https://www.youtube.com/watch?v=-7I7MWTX0gA)
+- The questions to ask before implementing any big data strategy. [[the guardian](http://www.theguardian.com/sap-partner-zone/2016/apr/28/questions-before-implementing-big-data-strategy)]
+    - What is the biggest challenge my business is facing right now and what do you do and don't know about your customer.
+and see how you can tweak your current strategy to make it more customer and data-driven.
 - Data Pitfalls —the Startup Edition. [[Medium](https://medium.com/@mrogati/data-pitfalls-the-startup-edition-fe1eebcc39da#.si84k3dss)]

### 2016-07-02 articles/data_hygiene/data_hygiene.Rmd
data hygiene, increase rmarkdown font size

---
title: "Data Hygiene Gotchas"
author: "Ming-Yu Liu"
date: "`r Sys.Date()`"
output:
  rmdformats::readthedown:
    highlight: pygments
---

<style type="text/css">
p{ /* Normal  */
   font-size: 18px;
}
body{ /* Normal  */
   font-size: 18px;
}
td {  /* Table  */
   font-size: 14px;
}
h1 { /* Header 1 */
 font-size: 32px;
}
h2 { /* Header 2 */
 font-size: 26px;
}
h3 { /* Header 3 */
 font-size: 22px;
}
code.r{ /* Code block */
  font-size: 14px;
}
pre { /* Code block */
  font-size: 14px
}
</style>

Feature Engineering and data cleaning are important. The biggest gains usually come from being smart about representing the data, rather than using some sort of complex algorithm.

# Data Sources

We should double check the data that we have in our hands.

**Including data from a period which is no longer valid:** Business strategy, process and system change happen frequently. Some of these changes might make historical data non-usable.

**Including incorrect information:** e.g. If we create a segmentation based on the customer's self stated income, they might under-stating or over-stating income.

**Include/Exclude outliers:** Outliers can skew inferences very significantly. While some modeling techniques are better at handling skew from outliers (e.g. tree-based algorithms), most of the techniques can not. We should find out why they're happening and treat them indepedently from the rest of the group, this makes our model more suited for general cases.

# Representing Timestamps

Time-stamp or datetime attributes are usually denoted by the EPOCH time. But in many applications, we need to transform that into multiple dimensions (Year, Month, Date, Hours, Minutes, Seconds), and some of that information might be unnecessary. Consider for example a supervised system that tries to predict traffic levels in a city as a function of Location+Time. In this case, trying to learn trends that vary by seconds would mostly be misleading. The year wouldn’t add much value to the model as well. Hours, day and month are probably the only dimensions we need. So when representing the time, try to ensure that our model does require all the numbers we're are providing it.

And don't forget Time Zones. If our data sources come from different geographical sources, do remember to normalize by time-zones if needed.

# Missing Values

Despite the fact that some modeling methods like tree-base algorithms can address missing values, others such as linear or logistic regression can not and they tend to drop records that contain one or more missing values from the selected set of predictor variable. Thus, if we want to evaluate the performance of different models, it makes sense to impute the missing values.

In case of missing numeric variables, it is common to impute it using a fixed value. Such as the mean, median is commonly used. In addition a categorical variable can be created for each predictor to indicate whether the value has been imputed or not. As for missing categorical variables, we can simply replaced it with the mode of the variable or create a new category indicating that the value is missing.

Sometimes, missing category values can convery important messages. A classific example is someone who is applying for loans and they non-reported their income category. Or maybe it shoudln't be missing and it's just something wrong with our data sources (e.g. database, or the way we're querying it).

# Categorical Variables

**Little variability** e.g. 99% of all the data points have the same category for one of the feauters. If we have a large amount of data, and there is reasonable number of data points (the notion of reasonable really depends on our judgement, at least 30 ...?), then including it in the model is a viable choice. Otherwise it makes sense to drop it (even better, understand why it occured).

The other way to address the issue is to combine the categorical variables that have few records. Though the combination of categories should have a sound logical basis, as opposed to being combined due to having a similar target variable.

- New data can't be predicted due to "unseen" category.

**Disguised as integers** A simple example would be a "color" feature that takes the value of {Red, Green, Blue}. When this occurs, One hot encode them, where we create transform the original color column into three column (since in this case it takes three distinct possible values) and have a binary value {0,1} indicating whether that data point contains that feature. Don't simply label them with numeric value. For example, the color feature might take one value from {1, 2, 3}, representing {Red, Green, Blue} respectively. First, for a mathematical model, this would mean that Red is somehow ‘more similar’ to Green than Blue (since |1-3| > |1-2|). Secondly, it would make statistical metrics (such as mean) meaningless. Both of which might be misleading to our model.

# Feature Interaction

This is where we "combine" two or more categorical attributes into a single one. This is extremely useful a technique, when certain features provides more information together than by themselves. A concrete/relatable example of a (possibly) good feature cross is something like (Latitude, Longitude). A common Latitude corresponds to so many places around the globe. Same goes for the Longitude. But once we combine Lat & Long buckets into discrete "blocks", they denote "regions" in a geography, with possibly similar properties for each one. Some other examples, includes computing the ratios beteen two features.

# Binning/Bucketing

Sometimes, it makes more sense to represent a numerical attribute as a categorical one. The idea is to reduce the noise endured by the learning algorithm, by assigning certain ranges of a numerical attribute to distinct ‘buckets’. Consider the problem of predicting whether a person owns a certain item of clothing or not. Age might definitely be a factor here. What is actually more pertinent, is the Age Group. So what we could do, is have ranges such as 1-10, 11-18, 19-25, 26-40, etc. Moreover, instead of decomposing these categories as in point 2, we could just use scalar values, since age groups that lie "closer by" do represent similar properties.

Bucketing makes sense when the feature can be divided into neat ranges, where all numbers falling in a range imply a common characteristic. It reduces overfitting in certain applications, where we don’t want our model to try and distinguish between values that are too close by – for example, we could club together all latitude values that fall in a city, if our property of interest is a function of the city as a whole. Binning also reduces the effect of possible data errors, by "rounding off" a given value to the nearest representative.

# Feature Scaling / Extraction / Selecting

**Scaling** A lot of times, certain features will have a higher "magnitude" than others. An example might be a person’s income – as compared to his age. In such cases, for models apart from tree-base models, it is in fact necessary that we scale all our attributes to comparable/equivalent ranges. This prevents our model from giving greater weightage to certain attributes as compared to others.

**Feature Extraction** Where we use algorithms that automatically generate a new set of features from our raw attributes. Dimensionality reduction methods such as SVD falls under this category.

**Feature Selecting** Some algorithms (lasso, tree-based algorithms) gives you information about which features are important than the others, meaning that they automatically select a subset of your original features so you can discard them when building other models. Here, you are not creating/modifying your current features, but rather pruning them to reduce noise or redundancy.

# Reference

- [Common data preparation mistakes and how to avoid them](http://www.analyticsvidhya.com/blog/2013/08/common-data-preparation-mistakes-avoid-them/)
- [Webinar: Enhance Predictive Modeling with Better Data Preparation](http://www.datasciencecentral.com/video/enhance-predictive-modeling-with-better-data-preparation)
- [Non-Mathematical Feature Engineering techniques for Data Science](https://codesachin.wordpress.com/2016/06/25/non-mathematical-feature-engineering-techniques-for-data-science/)
- [Simple Methods to deal with Categorical Variables in Predictive Modeling](http://www.analyticsvidhya.com/blog/2015/11/easy-methods-deal-categorical-variables-predictive-modeling/)

### 2016-07-02 finding_groups/finding_groups.Rmd
data hygiene, increase rmarkdown font size

@@
 ---
-title: "Business Analytics - Finding Groups Within the Data"
+title: "Finding Groups Within the Data"
 author: "Ming-Yu Liu"
-date: "`r Sys.Date()`"
+date: "2016-04-05"
 output:
@@
 
-> This documentation assumes you're familiar with hierarchical clustering.
+<style type="text/css">
+p{ /* Normal  */
+   font-size: 18px;
+}
+body{ /* Normal  */
+   font-size: 18px;
+}
+td {  /* Table  */
+   font-size: 14px;
+}
+h1 { /* Header 1 */
+ font-size: 32px;
+}
+h2 { /* Header 2 */
+ font-size: 26px;
+}
+h3 { /* Header 3 */
+ font-size: 22px;
+}
+code.r{ /* Code block */
+  font-size: 14px;
+}
+pre { /* Code block */
+  font-size: 14px
+}
+</style>
 
-## Background Information
+> You can find all the code and datasets (in the data folder) used in this documentation [here](https://github.com/ethen8181/Business-Analytics/tree/master/finding_groups).This documentation assumes you're familiar with hierarchical clustering.
 
-This documentation will go over two examples of how identifying groups within data allows business to improve efficiency.
+We will go over two examples of how identifying groups within data allows business to improve efficiency.
 Finding groups within our data allows us to strike the right balance between similarities and differences.
@@
 
-## Example 1 : Supply Chain Management
+# Example 1: Supply Chain Management
 
@@
 
-```{r}
+```{r, message=TRUE, warning=TRUE}
 
@@
 
-## Example 2 : Humance Resources
+# Example2 : Humance Resources
 
@@
 
-```{r, fig.height=8, fig.width=12 }
+```{r, fig.height=8, fig.width=12, message=FALSE, warning=FALSE}
 
@@
 
-## Takeaways
+**Takeaways**
 
@@
 
-**Sources:**
-
-This wraps up our example of where clustering algorithms can be used to improve your businesses efficiency.
-You can find all the code (finding_groups.R) and datasets (in the data folder) used in this documentation [here](https://github.com/ethen8181/Business-Analytics/tree/master/finding_groups).
-
-Any comments or suggestions is appreciated. That includes bugs or typos report. You can file these new issues [here](https://github.com/ethen8181/Business-Analytics/issues).
-
-## R Session Information
+# R Session Information
 

### 2016-07-02 marketing_analysis/code/marketing_analysis.R

@@ -1,161 +1,0 @@
-# marketing analysis
-
-library(grid)
-library(dplyr)
-library(scales)
-library(ggplot2)
-library(gridExtra)
-library(lubridate)
-library(data.table)
-setwd("/Users/ethen/Business-Analytics/2_marketing_analysis")
-
-# customer purchase record for a retailer
-data <- fread( "purchase.txt", sep = "\t", header = FALSE )
-
-# ----------------------------------------------------------------------------
-# 										Exploring around, not included  
-# ----------------------------------------------------------------------------
-
-# 1. name the column variables :
-# there are three columns for this dataset
-# namely, customer id, purchase amount and date of purchase ( year-month-day ) 
-setnames( data, c( "customer_id", "purchase_amount", "purchase_date" ) )
-
-# 2. convert date_of_purchase column to date type
-data[ , purchase_date := ymd(purchase_date) ]
-
-# 3. extract the year of purchase 
-data[ , purchase_year := year(purchase_date) ]
-
-# obtain the 
-# 1. number of purchases per year
-# 2. average purchase amount per year
-# 3. total purchase amounts per year
-# data.table allows you to compute all that in one concatenated command 
-
-explore <- data[ , .( counter 	 = .N, 
-					  avg_amount = mean(purchase_amount),
-					  sum_amount = sum(purchase_amount) ), by = purchase_year ] %>%
-		   arrange( purchase_year )
-
-ggplot( explore, aes( purchase_year, sum_amount ) ) + 
-geom_bar( stat = "identity", fill = "lightblue", color = "black" ) + 
-scale_y_continuous( label = comma ) + 
-ggtitle( "Total Purchase Amount 2005-2015" )
-# as you can see from the bar plot there's a nice and positive trend in terms 
-# of how much money is spent for this retailer 
-
-# using sql-like command
-# GROUP BY 1 :group by the first column regardless of what it's called
-library(sqldf)
-sqldf( "SELECT 	 purchase_year, 
-			     COUNT(purchase_year) AS 'counter',
-			     AVG(purchase_amount) AS 'avg_amount',
-			     SUM(purchase_amount) AS 'sum_amount'
-		FROM     data
-		GROUP BY 1" )
-
-# ----------------------------------------------------------------------------
-# 										Segmentation 
-# ----------------------------------------------------------------------------
-
-# 1. name the column variables :
[…]

### 2016-07-02 marketing_analysis/code/marketing_functions.R

@@ -1,76 +1,0 @@
-# marketing analysis
-
-
-# -------------------------------------------------------------------------------------------
-# [Segmentation]: 
-# Takes the rfm data and creates manual segmentation 
-# based on the recency criteria first that splits customer's into inactive, cold, 
-# warm and active based on the three recency cutoff value. Then further splits 
-# the warm and active segment into four sub-groups each based on the amount cutoff value and 
-# whether they have made their first purchase within the same timespan with respect to the "baseline". 
-# classifying each of the two original segments further into new high, new low, old high and old low.
-
-# @data 	: rfm segmentation data, must have five columns exactly named as  
-# 			  customer_id, recency, first_purchase, frequency and avg_amount, 
-# 			  the order of the column doesn't matter. 
-# @inactive : customers whose recency are higher than this value will be classified as 
-# 			  "inactive" customers. Note that the unit of recency is "days".
-# @cold		: same functionality as above, except it's the cutoff value for "cold" customers.
-# @warm 	: same functionality as above, cutoff value for "warm" and "active" customers.
-# @amount 	: cutoff value for splitting the warm and active customers further into
-# 			  high and low average purchased amount.
-# returns 	: original rfm data and adds a new segment column to it. It also sorts the order
-# 			  of the rows by customer id.  
-
-Segmentation <- function( data, inactive, cold, warm, amount )
-{
-	# segmentation criteria:
-	# four main groups, grouped by recency 
-	data[ recency > inactive, segment := "inactive" ]
-	data[ recency <= inactive & recency > cold, segment := "cold" ]
-	data[ recency <= cold & recency > warm, segment := "warm" ]
-	data[ recency <= warm, segment := "active" ]
-
- 	# split the active and warm segment further into new and old customers 
-	# those that made their first purhase within a year, 
-	# with respecitve to the cold and warm baseline.
-	# and split it into low and high base on a specified average amount
-
-	# Note that the old and new is based on the cold criteria for the "warm" segment!
-	data[ segment == "warm" & first_purchase > cold & avg_amount < amount,
-		  segment := "old warm low" ]
-
-	data[ segment == "warm" & first_purchase > cold & avg_amount >= amount,
-		  segment := "old warm high" ]
-
-	data[ segment == "warm" & first_purchase <= cold & avg_amount < amount, 
-		  segment := "new warm low" ]
-
[…]

### 2016-07-09 articles/articles.Rmd
calendar heatmap

@@
 
+<style type="text/css">
+p{ /* Normal  */
+   font-size: 18px;
+}
+body{ /* Normal  */
+   font-size: 18px;
+}
+td {  /* Table  */
+   font-size: 14px;
+}
+h1 { /* Header 1 */
+ font-size: 32px;
+}
+h2 { /* Header 2 */
+ font-size: 26px;
+}
+h3 { /* Header 3 */
+ font-size: 22px;
+}
+code.r{ /* Code block */
+  font-size: 14px;
+}
+pre { /* Code block */
+  font-size: 14px
+}
+</style>
+
+# Management
+
+- The Single Most Important Skill for a Data Scientist. [[Mango Solution](http://www.mango-solutions.com/wp/2015/06/the-single-most-important-skill-for-a-data-scientist/)]
+- How A/B Testing at LinkedIn, Wealthfront and eBay Made Me a Better Manager.
+[[First Round Review](http://firstround.com/review/how-a-b-testing-at-linkedin-wealthfront-and-ebay-made-me-a-better-manager/?ct=t(How_Does_Your_Leadership_Team_Rate_12_3_2015))]
+- How Constraints Create Space for Innovation. [[linkedin pulse](https://www.linkedin.com/pulse/how-constraints-create-space-innovation-ash-maurya)]
+- From Developer to Manager - PyCon 2016. [[Youtube](https://www.youtube.com/watch?v=lzBm0r8CxhY)]
+- Reddit and Facebook Veteran On How to Troubleshoot Troublemakers.
+[[First Round Review](http://firstround.com/review/reddit-and-facebook-veteran-on-how-to-troubleshoot-troublemakers/?ct=t(How_Does_Your_Leadership_Team_Rate_12_3_2015))]
+    - Hermits. Get them to collaborate with a team.
+    - Nostalgia. Ask qualities that they missed most, and see if there're existing projects that could allow them to recapture that feeling.
+    - Trend Chaser. Ask them why this is the route their going to take? The potential has to worth more than the expense of experimenting.
+- Practical Frameworks for Beating Burnout [[First Round Review](http://firstround.com/review/practical-frameworks-for-beating-burnout/?ct=t(How_Does_Your_Leadership_Team_Rate_12_3_2015))]
+- How to set goals: OKRs. [[Youtube](https://www.youtube.com/watch?v=mJB83EZtAjc)]
+- Meetings that don’t suck. [[Youtube](https://www.youtube.com/watch?v=rU8nv38E0Qk)]
+
 # Interviews and Careers
@@
     - Coachability: Be open and honest about where you fall short and how you’re actively improving on those shortcomings.
-- From Developer to Manager - PyCon 2016. [[Youtube](https://www.youtube.com/watch?v=lzBm0r8CxhY)]
-- Reddit and Facebook Veteran On How to Troubleshoot Troublemakers.
-[[First Round Review](http://firstround.com/review/reddit-and-facebook-veteran-on-how-to-troubleshoot-troublemakers/?ct=t(How_Does_Your_Leadership_Team_Rate_12_3_2015))]
-    - Hermits. Get them to collaborate with a team.
-    - Nostalgia. Ask qualities that they missed most, and see if there're existing projects that could allow them to recapture that feeling.
-    - Trend Chaser. Ask them why this is the route their going to take? The potential has to worth more than the expense of experimenting.
 - Labels are for cans not for people. [[Facebook Videos](https://www.facebook.com/morokcom/videos/237500609921393/)]
-- Practical Frameworks for Beating Burnout [[First Round Review](http://firstround.com/review/practical-frameworks-for-beating-burnout/?ct=t(How_Does_Your_Leadership_Team_Rate_12_3_2015))]
 - Land any job you want. [[blog](http://landanyjobyouwant.com/)]
@@
 - Wanted in College Graduates: Tolerance for Ambiguity. [[Linkedin Pulse](https://www.linkedin.com/pulse/wanted-college-graduates-tolerance-ambiguity-jeff-selingo)]
-- How to set goals: OKRs. [[Youtube](https://www.youtube.com/watch?v=mJB83EZtAjc)]
-- Meetings that don’t suck. [[Youtube](https://www.youtube.com/watch?v=rU8nv38E0Qk)]
 

### 2016-07-09 articles/avoid_pie_charts/avoid_pie_charts.Rmd
calendar heatmap

@@
 
-**Reference**
+# Reference
 

### 2016-07-09 articles/calendar_heatmaps/calendar_heatmaps.Rmd
fix link

---
title: "Calendar Heatmaps"
author: "Ming-Yu Liu"
date: "`r Sys.Date()`"
output:
  rmdformats::readthedown:
    highlight: pygments
---

<style type="text/css">
p{ /* Normal  */
   font-size: 18px;
}
body{ /* Normal  */
   font-size: 18px;
}
td {  /* Table  */
   font-size: 14px;
}
h1 { /* Header 1 */
 font-size: 32px;
}
h2 { /* Header 2 */
 font-size: 26px;
}
h3 { /* Header 3 */
 font-size: 22px;
}
code.r{ /* Code block */
  font-size: 14px;
}
pre { /* Code block */
  font-size: 14px
}
</style>

```{r, echo = FALSE}
knitr::opts_chunk$set(
  message = FALSE,
  warning = FALSE
)

```

> Source code can be obtained [here](https://github.com/ethen8181/Business-Analytics/blob/master/articles/calendar_heatmaps/calendar_heatmaps.R)

We'll use an attacker data with 200,000 "events". It has the `timestamp`, and the `source_country` ISO 3166-1 alpha-2 country code (which is the source of the attack) plus the `tz` time zone of the source IP address. Let's have a look:

```{r}

library(DT)
library(dplyr)
library(scales)
library(viridis)
library(ggplot2)
library(ggthemes)
library(gridExtra)
library(lubridate)
library(data.table)
setwd('/Users/ethen/Business-Analytics/articles/calendar_heatmaps')

attacks <- fread("data/eventlog.csv")
DT::datatable(attacks)

```

A quick exploratory data analysis on this attacker data is what were the "working hours" by country, and we can visualize this with a calendar heatmap. To create the heatmap, we’ll need the weekday and hour of each event, or as granular as you want to get.

```{r}

make_hr_wkday <- function( ts, sc, tz ){
	# convert each time with the appropriate timezone,
	# the timezone parameter, tz, only takes a single value,
	# then extract its weekdays and hour
	real_times <- ymd_hms( ts, tz = tz[1], quiet = TRUE )
	dt <- data.table( source_country = sc,
					  wkday = weekdays(real_times),
					  hour = hour(real_times) )
	return(dt)
}

# simply stores the name of the weekday
# convert weekday and hour into factor so they'll be ordered
# when plotting
weekday_levels <- levels( weekdays( 0, abbreviate = FALSE ) )
attacks <- attacks %>% 
		   group_by(tz) %>%
		   do( make_hr_wkday( .$timestamp, .$source_country, .$tz ) ) %>% 
		   ungroup() %>% 
		   mutate( wkday = factor( wkday, levels = weekday_levels ),
		   		   hour  = factor( hour, levels = 0:23 ) )
head(attacks)

```

Then we can simply group the count by hour and wkday and plot it, since we know that we have values for every combination there's no need to further preprocess the data.

```{r}

grouped <- attacks %>% count( wkday, hour ) %>% ungroup()

ggplot( grouped, aes( hour, wkday, fill = n ) ) + 
geom_tile( color = "white", size = 0.1 ) + 
theme_tufte( base_family = "Helvetica" ) + 
coord_equal() + 
scale_fill_viridis( name = "# of Events", label = comma ) + 
labs( x = NULL, y = NULL, title = "Events per weekday & time of day" ) +
theme( axis.ticks = element_blank(),
	   plot.title = element_text( hjust = 0.5 ),
	   legend.title = element_text( size = 8 ),
	   legend.text = element_text( size = 6 ) )

```

Some `ggplot2` explanation:

- `geom_tile()` will make tiles at each x and y location we've specified. We're also telling it to use a thin (0.1 units) white border to separate the tiles.
- `theme_tufte()` from the `ggthemes` package removes a lot of chart junk without having to do it manually.
- `coord_equal` ensures the plot will have a 1:1 aspect ratio (i.e. `geom_tile()`--which draws rectangles, will now draw squares).
- `scale_fill_viridis` from the `viridis` package gives us an awesome color scale.

That’s great, but what if we wanted the heatmap breakdown by country? We’ll can do this in two ways, first with each country’s heatmap using the same scale (using `facet`), then with each one using it’s own scale (using `grid.arrange`).

```{r}
events_by_country <- count( attacks, source_country ) %>% 
					 mutate( percent = percent( n / sum(n) ) ) %>%
					 arrange( desc(n) )
head(events_by_country)

top_country <- events_by_country$source_country[1:4]
top_country_attacks <- attacks %>%
					   filter( source_country %in% top_country ) %>% 
					   count( source_country, wkday, hour ) %>% 
					   ungroup() %>% 
					   mutate( source_country = factor( source_country, levels = top_country ) )

gg <- ggplot( top_country_attacks, aes( x = hour, y = wkday, fill = n ) ) + 
	  geom_tile( color = "white", size = 0.1 ) +
	  scale_fill_viridis( name = "# Events" ) + 
	  coord_equal() + 
	  facet_wrap( ~source_country, ncol = 2 ) +
	  labs( x = NULL, y = NULL, title = "Events per weekday & time of day by country\n" ) + 
	  theme_tufte( base_family = "Helvetica" ) + 
	  theme( axis.ticks = element_blank(),
			 axis.text = element_text( size = 8 ),
			 panel.border = element_blank(),
			 plot.title = element_text( hjust = 0.5 ),
			 strip.text = element_text( hjust = 0.5 ),
			 panel.margin = unit( 0.1, "cm" ),
	  		 legend.position = "bottom",
	  		 legend.title = element_text( size = 8 ),
	  		 legend.text = element_text( size = 6 ),
	  		 legend.key.size = unit( 0.4, "cm" ),
	  		 legend.key.width = unit( 1, "cm" ) )
gg

```

```{r}

plots <- lapply( top_country, function(x){

	subset_data <- top_country_attacks %>% filter( source_country == x )
	gg <- ggplot( subset_data, aes( x = hour, y = wkday, fill = n ) ) + 
		  geom_tile( color = "white", size = 0.1 ) +
		  scale_fill_viridis( name = "# Events" ) + 
		  scale_y_discrete( expand = c( 0, 0 ) ) +
		  coord_equal() + 
		  labs( x = NULL, y = NULL, title = x ) + 
		  theme_tufte( base_family = "Helvetica" ) + 
		  theme( axis.ticks = element_blank(),
				 axis.text = element_text( size = 7 ),
				 panel.border = element_blank(),
				 plot.title = element_text( hjust = 0.5 ),
				 strip.text = element_text( hjust = 0.5 ),
				 panel.margin = unit( 0.1, "cm" ),
		  		 legend.position = "bottom",
		  		 legend.title = element_text( size = 6 ),
		  		 legend.text = element_text( size = 6 ),
		  		 legend.key.size = unit( 0.4, "cm" ) )
	return(gg)
})

# specify the additional arguments to grid.arrange
# by adding it to the list that's going to be do.called
plots[['ncol']] = 2
do.call( grid.arrange, plots )

```

# Reference

- [Making faceted heatmaps with ggplot2](https://rud.is/b/2016/02/14/making-faceted-heatmaps-with-ggplot2/)

### 2016-07-09 articles/calendar_heatmaps/calendar_heatmaps.R

@@ -1,0 +1,126 @@
+library(DT)
+library(dplyr)
+library(scales)
+library(viridis)
+library(ggplot2)
+library(ggthemes)
+library(gridExtra)
+library(lubridate)
+library(data.table)
+
+setwd('/Users/ethen/facetedcountryheatmaps')
+
+
+attacks <- fread("data/eventlog.csv")
+DT::datatable(attacks)
+
+# preprocessing --------------------------------------------------------------
+make_hr_wkday <- function( ts, sc, tz ){
+	# convert each time with the appropriate timezone,
+	# the timezone parameter, tz, only takes a single value,
+	# then extract its weekdays and hour
+	real_times <- ymd_hms( ts, tz = tz[1], quiet = TRUE )
+	dt <- data.table( source_country = sc,
+					  wkday = weekdays(real_times),
+					  hour = hour(real_times) )
+	return(dt)
+}
+
+# simply stores the name of the weekday
+# convert weekday and hour into factor so they'll be ordered
+# when plotting
+weekday_levels <- levels( weekdays( 0, abbreviate = FALSE ) )
+attacks <- attacks %>% 
+		   group_by(tz) %>%
+		   do( make_hr_wkday( .$timestamp, .$source_country, .$tz ) ) %>% 
+		   ungroup() %>% 
+		   mutate( wkday = factor( wkday, levels = weekday_levels ),
+		   		   hour  = factor( hour, levels = 0:23 ) )
+head(attacks)
+
+# attacks --------------------------------------------------------------------
+
+# then we can simply group the count by hour and wkday
+# since we know that we have values for every combination
+# there's no need to further preprocess the data
+grouped <- attacks %>% count( wkday, hour ) %>% ungroup()
+
+ggplot( grouped, aes( hour, wkday, fill = n ) ) + 
+geom_tile( color = "white", size = 0.1 ) + 
+theme_tufte( base_family = "Helvetica" ) + 
+coord_equal() + 
+scale_fill_viridis( name = "# of Events", label = comma ) + 
+labs( x = NULL, y = NULL, title = "Events per weekday & time of day" ) +
+theme( axis.ticks = element_blank(),
+	   plot.title = element_text( hjust = 0.5 ),
+	   legend.title = element_text( size = 8 ),
+	   legend.text = element_text( size = 6 ) )
+
+
[…]

### 2016-07-27 ab_tests/frequentist_ab_test.Rmd
calendar heatmap
frequentist stats

@@
 
-Select the appropriate test statistic and calculate its value from the data (we will call it ZZ).
+Select the appropriate test statistic and calculate its value from the data (we will call it Z).
 A test statistic is a single metric that can be used to evaluate the null hypothesis.
@@
 ```{r, message=FALSE, warning=FALSE}
+
 library(scales)
@@
 
-The shaded dark blue area denotes the significance region, while the the shaded blue area denotes the power.
+The shaded dark blue area denotes the significance region, while the the shaded blue area denotes the power (note that it includes the shaded dark blue area).
 Note that if we pick a smaller N, or a smaller probability difference between the control and experiment group, the power drops (the shaded blue area decreases), meaning that if there’s is in fact a change, there’s lesser percent chance that we’ll detect it:
@@
 delta 	  <- 0.02 # minimum detectable boundary (practical significance boundary)
-power 	  <- 0.8  # sensitivity 
-sig_level <- 0.05 # specificity
+power 	  <- 0.8  # specificity, or true negative rate
+sig_level <- 0.05 # false positive rate
 
@@
 sig_level <- 0.05
-
 result <- power.prop.test( p1 = baseline, p2 = baseline + delta, 
@@
 Given these variables we can use it to calculate whether the proportional change was due to variation or not.
-To do so, we'll calculate the confidence interval for the difference, which can be done by calculating the difference plus and minus the z score times the standard deviation.
-Math Formula:
-
-$$p_{experiment} - p_{control} \pm Z \sqrt{ p_{pooled}( 1-p_{pooled} )( \frac{1}{n_{experiment}} + \frac{1}{n_{control}} ) }$$
-
-Note that this is the same formula from the first section, except now we're looking at the confidence interval.
-
-[`ABTest`][ABTest] Function that calculates the confidence interval. Input parameters:
 
@@
 - The same notion can be applied to the experiment's variable `count_experiment` and `sizes_experiment`.
-- Returns : A data.frame consisted of the mean and the confidence interval of the result.
 
@@
 
-ABTest <- function( count_control = count_control, 
-					sizes_control = sizes_control,
-					count_experiment = count_experiment, 
-					sizes_experiment = sizes_experiment )
-{
-	# probability of each group 
-	p_control <- count_control / sizes_control
-	p_experiment <- count_experiment / sizes_experiment
+result <- prop.test( c( count_control, count_experiment ), 
+					 c( sizes_control, sizes_experiment ) )
+result
 
-	# p: pooled probability
-	# std_error: pooled standard deviation (error) 
-	p <- ( count_control + count_experiment ) / ( sizes_experiment + sizes_control )
-	std_error <- sqrt( p * ( 1 - p ) * ( 1 / sizes_control + 1 / sizes_experiment ) )
+# do the computation ourselves to see that the confidence interval matches
+# compute the probability of each group and the standard error
+p1 <- count_control / sizes_control
+p2 <- count_experiment / sizes_experiment
+se <- sqrt( p1 * ( 1 - p1 ) / sizes_control + p2 * ( 1 - p2 ) / sizes_experiment )
 
-	# 95 percent confidence interval's z score = 1.96, equivalent to qnorm(0.975)
-	difference <- p_experiment - p_control
-	confidence <- difference + c( -1, 1 ) * qnorm(0.975) * std_error
-
-	return( data.frame( lower = confidence[1], 
-						mean  = difference,
-						upper = confidence[2] ) )
-}
-
-confidence <- ABTest( count_control = count_control, 
-					  sizes_control = sizes_control,
-					  count_experiment = count_experiment, 
-					  sizes_experiment = sizes_experiment )
-confidence
+# 95 percent confidence interval's z score
+conf_level <- 0.95
+zscore <- qnorm( conf_level + ( 1 - conf_level ) / 2 )
+conf_int <- abs(p2 - p1) + c( -1, 1 ) * zscore * se
+conf_int
 
@@
 
-```{r, warning=FALSE,message=FALSE}
+```{r, warning=FALSE, message=FALSE}
 
-# fixed artifical plot
-# using delta = 0.02 as the minimum detectable boundary 
-library(ggplot2)
-
-scenario <- as.character(2:6)
-lower <- c( -0.008, 0.011, -0.025, -0.005, 0.015 )
-mean  <- c( 0.005, 0.014, 0.005, 0.025, 0.025 )
-upper <- c( 0.018, 0.017, 0.035, 0.055, 0.035 )
-
+scenario <- as.character(1:6)
+lower <- c( conf_int[1], -0.008, 0.011, -0.025, -0.005, 0.015 )
+mean  <- c( abs(p2 - p1), 0.005, 0.014, 0.005, 0.025, 0.025 )
+upper <- c( conf_int[2], 0.018, 0.017, 0.035, 0.055, 0.035 )
 examples <- data.frame( scenario, lower, mean, upper )
-examples <- rbind( cbind( scenario = "1", confidence ), examples )
 examples$scenario <- factor( examples$scenario, levels = as.character(6:1) )
@@
 ```
-
-[ABTest]: https://github.com/ethen8181/Business-Analytics/blob/master/ab_tests/frequentist_ab_test.R

### 2016-07-27 articles/articles.Rmd
frequentist stats

@@
 }
-td {  /* Table  */
+td{  /* Table  */
    font-size: 14px;
 }
-h1 { /* Header 1 */
+h1{ /* Header 1 */
  font-size: 32px;
 }
-h2 { /* Header 2 */
+h2{ /* Header 2 */
  font-size: 26px;
 }
-h3 { /* Header 3 */
+h3{ /* Header 3 */
  font-size: 22px;
@@
 }
-pre { /* Code block */
+pre{ /* Code block */
   font-size: 14px
@@
 
+- The Pivotal Stories Every Startup Leader Should Be Able to Tell. [[First Round Review](http://firstround.com/review/the-pivotal-stories-every-startup-leader-should-be-able-to-tell/)]
+- The Eisenhower Matrix Productivity Tool – As A Trello Board.
+[[Trello](http://blog.trello.com/eisenhower-matrix-productivity-tool-trello-board/?utm_source=newsletter&utm_medium=email&utm_campaign=july-2016-newsletter-2)]
+- All About OKRs: How to Set Them, Achieve Them, and Track Them in Trello.
+[[Trello](http://blog.trello.com/okrs-set-achieve-track-trello/?utm_source=newsletter&utm_medium=email&utm_campaign=july-2016-newsletter-2)]
 - The Single Most Important Skill for a Data Scientist. [[Mango Solution](http://www.mango-solutions.com/wp/2015/06/the-single-most-important-skill-for-a-data-scientist/)]
@@
 
+- Data science tips from reddit
+The Top Five Data Science Tips From r/DataScience.
+[[Blog](http://www.datadependence.com/2016/07/best-data-science-tips-reddit/?utm_content=buffer076f2&utm_medium=social&utm_source=twitter.com&utm_campaign=buffer)]
+- Data scientists mostly just do arithmetic and that’s a good thing. [[Blog](https://m.signalvnoise.com/data-scientists-mostly-just-do-arithmetic-and-that-s-a-good-thing-c6371885f7f6#.dly3ud86c)]
+- The one machine learning concept you need to know. [[Blog](http://www.sharpsightlabs.com/one-concept-machine-learning/)]
+    - When you’re doing machine learning (specifically, supervised learning), you’re essentially using computational techniques to find the underlying relationships between the input and the output.
 - Check Yourself: 5 Things You Should Know About Data Science (Author Note). [[Blog](http://analyticsmadeskeezy.com/2012/11/05/check-yo-self-5-things-you-should-know-about-data-science-author-note/)]
@@
 
+- The New 'Must Haves' for Mobile Apps to Succeed Today. [[First Round Review](http://firstround.com/review/the-new-must-haves-for-mobile-apps-to-succeed-today/)]
 - 'Amazon is not your friend'. [[Linkedin](https://www.linkedin.com/pulse/amazon-your-friend-caroline-fairchild)]
@@
 
+- The Mathematics of Machine Learning. [[Blog]](http://datascience-africa.org/2016/07/mathematics-machine-learning/)
 - The MBA Data Science Toolkit: 8 resources to go from the spreadsheet to the command line.

### 2016-07-27 articles/calendar_heatmaps/calendar_heatmaps.Rmd
frequentist stats

@@
 
-library(DT)
 library(dplyr)
@@
 attacks <- fread("data/eventlog.csv")
-DT::datatable(attacks)
+head(attacks)
 
@@
 
+This type of calendar heatmap can also be used to detect patterns in IT support logs, the chart will then show the number of tickets raised (day-wise).
+We can then detect patterns to see commonly raised issues and figure out ways to address it.
+
 # Reference
@@
 - [Making faceted heatmaps with ggplot2](https://rud.is/b/2016/02/14/making-faceted-heatmaps-with-ggplot2/)
+- [Big Data Insights - IT Support Log Analysis](http://costaleconomist.blogspot.tw/2015/12/big-data-insights-it-support-log.html)

### 2016-07-27 frequentist_statistics/frequentist_statistics.Rmd
frequentist stats

---
title: "Frequentist Statistic"
author: "Ming-Yu Liu"
date: "`r Sys.Date()`"
output:
  rmdformats::readthedown:
    highlight: pygments
---

<style type="text/css">
p{ /* Normal  */
   font-size: 18px;
}
body{ /* Normal  */
   font-size: 18px;
}
td {  /* Table  */
   font-size: 14px;
}
h1 { /* Header 1 */
 font-size: 32px;
}
h2 { /* Header 2 */
 font-size: 26px;
}
h3 { /* Header 3 */
 font-size: 22px;
}
code.r{ /* Code block */
  font-size: 14px;
}
pre { /* Code block */
  font-size: 14px
}
</style>

```{r, echo=FALSE, message=FALSE, warning=FALSE}

library(knitr)
opts_chunk$set( message = FALSE, warning = FALSE )

```

# Statistic Inference Basics

## Central Limit Theorem (CLT)

Let’s say you are studying the population of beer drinkers in the US. You’d like to understand the mean age of those people but you don’t have time to survey the entire US population. Instead of surveying the whole population, you collect one sample of 100 beer drinkers in the US. With this data, you are able to calculate an arithmetic mean. Maybe for this sample, the mean age is 35 years old. Say you collect another sample of 100 beer drinkers. For that new sample, the mean age is 39 years old. As you collect more and more means of those samples of 100 beer drinkers, you get what is called a sampling distribution. The sampling distribution is the distribution of the samples mean. In this example, 35 and 39 would be two observations in that sampling distribution.

The theorem states that that if we collect "a large number" of different samples mean from the population, the sampling distribution, the distribution of the samples mean you collected, will approximately take the shape of a bell curve around the population mean no matter what the orginal population distribution is. This shape is also known as a normal distribution. Don’t get the statement wrong. The CLT is not saying that any population will have a normal distribution. It is saying that the sampling distribution will.

To be a little more specific, the sampling distribution of the mean is nearly normal centered at the population mean, with standard error equal to the population standard deviation divided by the square root of the sample size. In the description above, the rule of thumb for the term "a large number" is 30, but may need more if the population distribution is extremely skewed. And at most 10% of the population should be sampled so that we can make sure that our samples are independent of one another.

Knowing that the sampling distribution will take the shape of a normal distribution is what makes the theorem so powerful, as it is the foundation of concepts such as confidence intervals and margins of error in frequentist statistics.

## Confidence Interval

A plausible range of values for the population parameter is called a confidence interval. Example:

> The General Social Survey, the GSS, is a sociological survey used to collect data on demographic characteristics and attitudes of residents of the United States. In 2010, the survey collected responses from 1,154 U.S. residents. Based on the survey results, a 95% confidence interval for the average number of hours Americans have to relax or pursue activities that they enjoy after an average workday, was found to be 3.53 to 3.83 hours.

The general definition of a 95% confidence interval means that 95% of confidence intervals, **created based on random samples of the same size from the same population will contain the true population parameter**.

Thus in the context of the problem above, the correct interpretation of the **confidence interval** will be: If we collected many random samples of the same size (1,154 Americans) and calculated a confidence interval for average number of hours Americans spend relaxing after a work day, then we would expect 95% of the intervals to contain the true population average number. Or We are 95% confident that the average number of hours Americans (note that this is refering to the population) spend relaxing after a work day is between 3.53 to 3.83.

The interval is computed using the formula:

$$\text{point estimate} \pm z * SE $$

Where

- $z$ is called the **critival value** and it corresponds to the **confidence level** that we chose.
- $SE$ represents the **standard error**. Generally the standard error for a point estimate is estimated from the data and computed using a formula. For example, the standard error for the sample mean is $\frac{s}{ \sqrt{n} }$, where $s$ is the standard deviation and $n$ is the number of samples.
- The value $z * SE$ is called the **margin of error**.
- Note that this constructing confidence intervals framework holds true for doing can be easily adapted for any estimator that has a nearly normal sampling distribution. e.g. sample mean, two sample mean, sample proportion and two sample proportion (we'll later see). All we have to do this is change the way that we're calculating the standard error.

Suppose we took many samples and built a confidence interval from each sample using the equation point estimate plus or minus 1.96 (the **critical value**) times the standard error. Then about 95% of those confidence intervals would be expected to contain the true population mean, the $\mu$ (mu).

The 1.96 is so called the **confidence level** and it is something we choose as oppose to calculate. Commonly used confidence levels in practice are 90%, 95%, 98%, and 99%. Changing the confidence level simply means adjusting the value of the **critical value** in the confidence interval formula. We can compute the **critical value** using `qnorm`.

```{r}

# e.g. the positive and negative critical value for 95%
cl <- 0.95
c( -1, 1 ) * qnorm( ( 1 - cl ) / 2 )

```

Given these information we can also compute our confidence interval. Example1:

> A sample of 50 college students were asked, how many exclusive relationships they've been in so far? The students in the sample had an average of 3.2 exclusive relationships, with a standard deviation of 1.74. In addition, the same distribution was only slightly skewed to the right. Estimate the true number of exclusive relationships based on this sample using a 95% confidence interval.

So we're given that 50 college students are in the sample, so N is equal to 50. They had an average of 3.2 exclusive relationships. So the sample mean is 3.2 and the sample standard deviation is 1.74.

```{r}

3.2 + c( 1, -1 ) * qnorm( ( 1 - 0.95 ) / 2 ) * 1.74 / sqrt(50)

```

This means that we are 95% confident. That college students on average have been in 2.72 to 3.68 exclusive relationships.

Example2:

> Say we are studying the American beer drinkers and we want to know the average age of the US beer drinker population. We hire a firm to conduct a survey on 100 random American beer drinkers. From that sample, we get the following  results: 
> n (sample size): 100
> Standard Deviation of Age: 15
> Arithmetic Mean of Age: 40
> And the question is, what is the probability that the mean age of the US beer drinker population is between 38 and 42?

```{r}

# standard error
se <- 15 / sqrt(100)

# z-score is computed as ( value - mean ) / standard error
upper <- ( 42 - 40 ) / se
lower <- ( 38 - 40 ) / se
pnorm(upper) - pnorm(lower)

```

The probability that the mean age of the US beer drinker population is between 38 and 42 is around 81%.

---

Note that the higher the confidence interval, the larger the critical value. Hence the width of the confidence interval also gets wider. Thus if we want to be very certain that we capture the true population parameter, then we raise the confidence level to get a wider interval. Another way of think about this is using the width of the area that captures the middle 95 or 99% of the distribution.

![](images/confidence_interval.png)

The middle 99% will inevitably span a larger area. And hence the 99% confidence interval is going to be wider. So if we want to increase accuracy, we can increase the confidence level. But this might come at a cost.

Say you're watching the weather forecast, and you're told that the next day the low is negative 20 degrees Fahrenheit, and the high is positive 110 degrees Fahrenheit. Is this accurate? Most likely, yes. Tomorrow's temperature is probably going to be somewhere between negative 20 and positive 110. However, is it informative? or in other words, is it precise? Not really. Based on this weather report it would be near impossible to figure out what to wear tomorrow, or what really to expect in terms of the weather. As the confidence level increases, the width of the confidence interval increases as well which then increases the accuracy. However, the precision goes down.

Then how can we get the best of both worlds? Is it possible to get higher precision and higher accuracy? Well, of course, yes. The way is to increase the sample size. If we increase our sample size, that's going to shrink our standard error and our margin of error. And therefore we can still remain at a high confidence level while not necessarily needing to increase the width of the confidence interval as well.

## Required Sample Size

Given a target margin of error (denoted as ME), confidence level, and information on the variability of the sample or the population, we can determine the required sample size to achieve the desired margin of error. We do this by plugging known values into the equation of the margin of error and then rearranging things to solve for the unknown and the sample size.

$$
\text{ME} = z * \frac{s}{ \sqrt{n} } \rightarrow 
n = \left( \frac{z * s}{ME} \right) ^2
$$

> Suppose a group of researchers want to test the possible effect of an epilepsy medication taken by pregnant mothers on the cognitive development of their children. As evidence, they want to estimate the IQs of three-year-old children born to mothers who were on this medication during their pregnancy. Previous studies suggest that the standard deviation of IQ scores of three-year-old children is 18 points. How many such children should the researches sample in order to obtain a 90% confidence interval with a margin of error less than or equal to four points?

```{r}

# 1.65 is the critical value, z, for 90%
( 1.65 * 18 / 4 ) ^ 2

```

However, since we can't really have 0.13 of a person, we're going to need to round this number. Since 55.13 is the minimum required sample size, regardless of the value of the decimal, we always want to round up the number. Therefore, we need at least 56 such children in the sample to obtain a maximum margin of error of four points.

Another critical thing that we can observe from the formula is that: The previous desired margin of error was four and suppose the new desired margin of error is now two points.

```{r}

# 1.65 is the critical value, z, for 90%
( 1.65 * 18 / 2 ) ^ 2

```

Therefore, to go from four to two, we need to quadruple our sample size. This is important for two reasons. You are going to find yourself in many cases that if we want increase our sample size to get more accurate results. This is indeed true. However, we can see that in order to start really seeing gains from an increased sample size, you may need to really, really increase your sample size. And that's going to take resources.

## Hypothesis Testing

We start with a null hypothesis that we usually call it $H_0$ (H naught) and that represents our status quo. And we also have an alternative hypothesis our $H_1$ that represents the question that we wish to answer, in other word, what we're testing for. We conduct a hypothesis test under the assumption that the null hypothesis is true, either via simulation, or using theoretical methods that rely on the central limit theorem.

If the test results suggest that the data do not provide convincing evidence for the alternative hypothesis, we stick with the null hypothesis. If they do, then we reject the null hypothesis in favor of the alternative.

Recall from the confidence interval section, we were given the example:

> A sample of 50 college students were asked, how many exclusive relationships they've been in so far? The students in the sample had an average of 3.2 exclusive relationships, with a standard deviation of 1.74. In addition, the same distribution was only slightly skewed to the right. 
> Estimate the true number of exclusive relationships based on this sample using a 95% confidence interval.

```{r}

# We're able to compute the confidence interval using the following formula
3.2 + c( 1, -1 ) * qnorm( ( 1 - 0.95 ) / 2 ) * 1.74 / sqrt(50)

```

This means that we are 95% confident. That college students on average have been in 2.72 to 3.68 exclusive relationships.

So the next question is:

> Based on this confidence interval, do these data support the hypothesis that college students on average have been in more than 3 exclusive relationships?

We start with setting our hypothesis. Our null hypothesis, $H_0$, is $\mu = 3$. College students have been in three exclusive relationships on average. Our alternative hypothesis, $H_1$, is $\mu > 3$. College students have been in more than three exclusive relationships on average. Note that the hypothesis are always about the population parameters and never about the sample.

Our interval spans from around 2.7 to 3.7, and the null value of 3 is actually included in the interval. The interval says that any value within it could conceivably be the true population mean. Therefore we cannot reject the null hypothesis in favor of the alternative. This is a quick and dirty approach for hypothesis testing. However, it doesn't tell us the likelihood of certain outcomes under the null hypothesis. In other words, it does not tell us the **p-value**.

The **p-value** is defined as **the probability of obtaining the observed or more extreme outcome, given that the null hypothesis is true (not the probability that the alternative hypthesis is true)**. So in context of the data we've been working with, this is the probability of x bar being greater than 3.2, greater than the observed average, given that the true population mean is 3.

```{r}
# standard error = standard deviation / square root of the sample size
se <- 1.74 / sqrt(50)

# z score
z <- ( 3.2 - 3 ) / se
pvalue <- pnorm( z, lower.tail = FALSE )
pvalue

```

The p-value for this test is around 0.209.

If the p-value is low, and by low we mean lower than the significance level $\alpha$, which we usually set at 0.05, but we can certainly change it as well, we say that it would be very unlikely to observe the data if the null hypothesis were true, and therefore we reject the null hypothesis. If on the other hand the p-value is higher than $\alpha$, we say that it is indeed likely to observe the data even if the null hypothesis were true. And hence, we would not reject the null hypothesis.

So back to our original question. Our null hypothesis was that college students on average have three exclusive relationships, versus, the alternative was that that number was something greater than 3.

In our case there is a 20.9% chance that a random sample of 50 college students would still yield a sample mean of 3.2 or higher. Since this is a pretty high probability, we think that a sample mean of 3.2 or more exclusive relationships is likely to happen simply by chance.

So in this case we fail to reject the null hypothesis and say that even though we observed a sample mean slightly above 3, there is not enough evidence to reject the null hypothesis.

> A small p-value indicates the data provides strong evidence against the null hypothesis.

If we're doing two-sided hypothesis test, we can say that the significance level and the confidence level are complements of each other. Think about the most commonly used significance level, 5%, and think about the most commonly used confidence level, 95%. It is not a coincidence that the sum of those two numbers adds up to one. Consider the following graph:

<img src='images/two_sided.png' width='80%'>

We have a two-sided hypothesis test where we have an alpha of 0.05. This means that at each tail of the normal curve, we can afford to have a tail area of 0.025, so that the total of those tail areas add up to 5%. And usually, when we're thinking about confidence intervals, we're always thinking about the middle whatever percent of the distribution. So a 95% confidence level means that we're interested in the middle 95% of the normal curve.

# Inference For Means

## Comparing Two Means

> We're conducting a distracted eaters study where we compare the average snack consumption of distracted and non-distracted eaters post-lunch. We have a sample of 44 volunteer patients, which are randomized into two equally-sized groups. One group played solitaire on the computer while eating, and the other group was asked to eat their lunch without any distractions. Both groups were provided the same amount of lunch and offered the same amount of biscuits to snack on afterwards. The researchers measured the snack consumption of subjects in each group. The study reports average snack consumption levels for both groups, as well as the standard deviations (provided in the next R code chunk).

Suppose we want to estimate how much more, or less, distracted eaters snack compared to non-distracted eaters. We would use a confidence interval for this, which always takes the form of the point estimate plus or minus a margin of error, where:

- The point estimate is the **difference between the two sample averages**, $\bar{x_1} - \bar{x_2}$.
- The margin of error can be calculated as a critical value ($t_{df}$) times the standard error. Where:
    - the standard error can be computed by $\sqrt{ \frac{s_1^2}{n_1} + \frac{s_2^2}{n_2} }$.
    - the degree of freedom (df) for the t score: $min( n_1 - 1, n_2 - 1 )$, the lower value of the two sample sizes minus 1.

We can also conduct a hypothesis test to see if we should reject the $H_0$, which can be phrased as for the average snack consumption for those who eat with and without distraction; the difference between those two is 0.

```{r}

# 1 = solitare, 2 = no distraction
# x = mean, s = standard deviation, n = sample size
x1 <- 52.1
s1 <- 45.1
n1 <- 22
x2 <- 27.1
s2 <- 26.4
n2 <- 22

df <- min( n1 - 1, n2 - 1 )
critical <- qt( ( 1 - 0.95 ) / 2, df )
se <- sqrt( s1 ^ 2 / n2 + s2 ^ 2 / n2 )

# confidence interval
print( ( x1 - x2 ) + c( 1, -1 ) * critical * se )

# hypothesis test
tscore <- ( x1 - x2 ) / se
pt( tscore, df, lower.tail = FALSE ) * 2

```

The confidence interval for the average difference was 1.83 to 48.17 and the hypothesis test evaluating a different between the two means yielded a p-value of roughly 4%. Which means that we would reject the null hypothesis and conclude that these data do indeed provide convincing evidence that there is a difference between the average snack intake of distracted and non-distracted eaters.

The results of the confidence interval and the hypothesis test do in fact agree. We rejected the null hypothesis that set the difference between the two means equal to 0, and therefore this null value should not be included in our confidence interval and indeed it's not.

Note that the formula listed in this section is used when the **two groups are independent of one another (not paired)**. e.g. If we wish to compare pre- (beginning of semester) and post-test (end of semester) scores of students, then this will require a paired t-test.

## Comparing Multiple Means (ANOVA)

Sometimes we want to compare means across many groups. A new method to do this is called **analysis of variance (ANOVA)** and a new test statistic called F. ANOVA uses a single hypothesis test to check whether the means across many groups are equal:

- $H_0$: The mean outcome is the same across all groups. In statistical notation, $\mu_1 = \mu_2 = ... = \mu_k$ where $\mu_i$ represents the mean of the outcome for observations in category $i$.
- $H_A$: At least one mean is different.

In order to be able to reject the null hypothesis, we need to obtain a large F statisitc (a small p value), which requires that the variability between groups be much larger than the variability within groups.

List the conditions necessary for performing ANOVA

- the observations should be independent within and across groups.
- the data within each group are nearly normal.
- the variability across the groups is about equal.

> As an example we consider one of the data sets available with R relating to an experiment into plant growth. The purpose of the experiment was to compare the yields on the plants for a control group and two treatments of interest. The response variable was a measurement taken on the dried weight of the plants.

```{r}

library(ggplot2)

# built-in data PlantGrowth
plant_df <- PlantGrowth
plant_df$group <- factor( plant_df$group,
						  labels = c( 'Control', 'Treatment 1', 'Treatment 2' ) )

ggplot( plant_df, aes( x = group, y = weight ) ) + 
geom_boxplot()

```

Initial inspection of the data suggests that there are differences in the dried weight for the two treatments but it is not so clear cut to determine whether the treatments are actually different to the control group.

```{r}

anova_test <- aov( weight ~ group, data = plant_df )
summary(anova_test)

```

Here, the p-value is small, thus we reject the null hypothesis and say that we have sufficient evidence for the alternative. More explicitly, the data DO provide convincing evidence that at least one pair of population means are different from each other.

Note that ANOVA only answers whether there is evidence of difference in at least one pair of groups, it doesn't tell us which groups are different. Hence when we reject the null hypothesis in an ANOVA analysis, we might wonder, which of these groups have different means? To answer this question, we compare the means of each possible pair of groups. These comparisons can be accomplished using a two-sample t-test, but we need to use a **modified significance level**. Since if we were to conduct multiple pairwise comparisons, it is likely that we will eventually find a difference just by chance, even if there is no difference in the populations.

The **Bonferroni correction** suggests that a more stringent significance level is more appropriate for these multiple comparisons tests:

$$\alpha^* = \alpha / K$$

Where $K$ is the number of comparisons being considered. If there are k groups, then usually all possible pairs are compared and $K = \frac{k(k - 1)}{2}$.

```{r}

# the quickest way to do this "post-hoc" multiple
# comparison test
TukeyHSD(anova_test)

```

# Inference For Proportions

> We're told that 90% of all plant species are classified as angiosperms. These are flowering plants. If you were to randomly sample 200 plants from the list of all known plant species, what is the probability that at least 95% of the plants in your sample will be flowering plants?

The sampling distribution $\hat{p}$ taken from a sample of size $n$ from a population with a true proportion $p$, is nearly normal when:

1. The sample observations are independent
2. We expected to see at least 10 successes and 10 failures in our sample, i.e. $np$ and $n(1 - p)$. This is called the success-failure condition.

If these conditions are met, we can conclude that the sampling distribution of $\hat{p}$ is nearly normal with mean $p$ and standard error of $\sqrt{ \frac{ p ( 1 - p ) }{n} }$. **Note that it uses $p$ instead of $\hat{p}$** to compute the standard error. There is a slight difference of when to use which, and it is depicted in the following table:

![](images/p_versus_phat.png)

This is a important notion in proportion testing, because unlike doing inference for means, the point estimate $\mu$ does not appear in the calculation for the standard error. For the proportion testing, it does.

```{r}

n  <- 200
p  <- 0.9
se <- sqrt( p * ( 1 - p ) / n )

zscore <- ( 0.95 - p ) / se
pnorm( zscore, lower.tail = FALSE )

```

> A university newspaper is conducting a survey to determine what fraction of students support a $200 per year increase in fees to pay for a new football stadium. How big of a sample is required to ensure the margin of error is smaller than 0.04 using a 95% confidence level?

If we have an estimate of the population proportion $p$, perhaps from a similar survey, we could enter in that value and solve for $n$. If we have no such estimate, we must use some other value for $p$. It turns out that the margin of error is largest when $p$ is 0.5, so we typically use this worst case value if no estimate of the proportion is available. Recall that the formula for computing the margin of error for the proportion is:

$$\text{ME} = z *\sqrt{ \frac{ p ( 1 - p ) }{n} }$$

```{r}

me <- 0.03
p  <- 0.5
ci <- 0.95

# here, p and 1 - p is the same, thus we can use p squared
( qnorm( 0.95 + ( 1 - ci ) / 2 ) * p / me ) ^ 2

```

## Hypothesis Test

> A 2013 Pew Research poll found that 60% of 1,983 randomly sampled American adults believe in evolution. Does this provide convincing evidence that majority (more than 50%) of Americans believe in evolution?

Our alternative hypothesis should state $p$ is greater than 0.5. We are also given that sample proportion is 0.6. So, what we are checking to see is, does this particular sample yield convincing evidence that the majority of Americans believing in evolution.

```{r}

n <- 1983
p <- 0.5
se <- sqrt(  p * ( 1 - p ) / n )
zscore <- ( 0.6 - p ) / se
pnorm( zscore, lower.tail = FALSE )

```

Remember that p-value = P( observed or more extreme outcome | null hypthesis is true ). This shows that there is almost 0% chance of obtaining a random sample of 1983 Americans where 60% or more believe in evolution, if in fact 50% of the American believe in evolution.

## Chi-Square

In a **Chi-Square goodness of fit test, we evaluate the distribution of one categorical variable that has more than two levels**.

When dealing with counts and investigating how far the observed counts are from the expected counts, we use this new test statistic called the **chi-square** statistic.

$$\chi^2 = \sum_{i=1}^{k} \frac{(O_i - E_i)^2}{E_i}$$

It's calculated as the observed minus the expected for each cell squared divided by the expected counts and we want to sum this over all of the cells (levels of the categorical variable).

The chi-squared distribution has only one parameter, the degrees of freedom. It influences the shape, the center and the spread of the chi-square distribution. And for goodness of fit test, the degrees of freedom can be calculated as $k-1$, where k stands for the number of cells.

The conditions are exactly the same between the chi-square goodness of fit test and the chi-square test of independence as we'll later see.

- Independence, where we think about sampled observations having to be independent of each other. We can ensure this by random sampling or assignment. And each case should only contributes to one cell in the table
- We want to make sure that our sample size is less than 10% of our population.
- We want to make sure that each particular scenario or cell has at least five expected counts.

> In a county where jury selection is supposed to be random, a civil rights group sues the county, claiming racial disparities in jury selection. Distribution of ethnicities of people in the county who are eligible for jury duty based on census results and distribution of 2500 people who were selected for jury duty in the previous year are given in the table below.

![](images/chisq.png)

So we can see that in this population, we have 80.29% whites 12.06% blacks, 0.79% Native Americans, 2.92% Asians and Pacific islanders, and 3.94% other ethnicities. And for the observed count: 1,920 were white, 347 were black, 19 were native America, 84 were Asian and Pacific Islander, and 130 were categorized as other race or ethnicity.

We again start by setting up the hypothesis test:

- $H_0$: The observed counts of jurors from various race ethnicities follow the same ethnicity distribution in the population.
- $H_A$: The observed counts of jurors from various ethnicities, do not follow the same race ethnicity distribution in the population.

```{r}

obs <- c( 1920, 347, 19, 84, 130 )

# expected probability distribution
prob <- c( 0.8029, 0.1206, 0.0079, 0.0292, 0.0394 )
chisq.test( x = obs, p = prob )

```

In a **Chi-Square independent test, we evaluate the distribution of two categorical variable that has more than two levels**.

The computation for the independent test is exactly the same as the goodness of fit test (The expected count for the two way table is computed as *row total * column total / table total*). What is different, however, is how we calculate the degrees of freedom (df). Remember with the chi square goodness of fit test, the degrees of freedom was simply $k- 1$, with $k$ being the number of cells. In this case because we have a two-way table, we need to consider the number of levels for both of the categorical variables. So the degrees of freedom is calculated as the number of rows minus one times number of columns minus one. Denoted as $( r - 1 ) * ( c - 1 )$.

With a small p value, we reject the null hypothesis in favor of the alternative. Which means that these data provide convincing evidence that the two categorical variable are associated. Note that this type of analysis is not sufficient to deduce a causal relationship as there might be other confounding factors.

# Decision Errors And Powers

Just like court cases, hypothesis tests are not flawless. In the court system, innocent people sometimes are wrongly convicted and sometimes the guilty walk free. Similarly, we can make wrong decision in statistical hypothesis tests as well. The difference, though, is that we have the tools necessary to quantify how often we make errors in statistics, which is type I and type II errors. Consider the following truth versus decision contingency table:

![](images/type12.png)

If the null hypothesis is indeed true and you fail to reject it, you've done the right thing (denoted by the check mark). There's absolutely no reason to worry. Similarly, if the alternative hypothesis is true and you reject the null hypothesis in favor of the alternative, once again, you've done the right thing. The other two cells denotes:

- **Type I error** is rejecting the null hypothesis when the null hypothesis is actually true. So in other words, rejecting the null hypothesis when you should not have. 
- **Type II error** is failing to reject the null hypothesis when the alternative is true. In other words, it's failing to reject the null hypothesis when you shouldn't have.

> Consider a criminal trial: The null hypothesis says that the defendant is innocent. The alternative says that the defendant is guilty. Which type of error is being committed in the following circumstances? 
> 1. Declaring the defendant innocent when they are actually guilty.
> 2. Declaring the defendant guilty when they're actually innocent.

This means that we are failing to reject the null hypothesis when the alternative is actually true, which is a type II error. And the converse of that is declaring the defendant guilty when they're actually innocent. So in this case, we have rejected the null hypothesis in favor of the alternative when the null was true and that is the definition of a type one error.

As a general rule, we reject the null hypothesis when the p-value is less than .05. In other words, we set our significance level, our $\alpha$, to .05. This means that, for those cases where the null hypothesis is actually true, we do not want to incorrectly reject it more than 5% of those times.

In other words, when using a 5% **significance level**, there is about a 5% chance of making a type I error if the null hypothesis is true. This is why we prefer small values of alpha. Because if we were to increase $\alpha$ that might sound good because that would allow us to reject our null hypothesis more often which is usually what we are trying to do. However, we would also be inflating our type I error rate.

Returning back to our truth versus decision contingency table:

![](images/alpha_beta.png)

A type I error, is rejecting the null hypothesis when we shouldn't have. And the probability of doing so is $\alpha$, our **significance level**. The type II error is failing to reject the null hypothesis when we should have and the probability of doing so is labeled $\beta$ (Note that this is not something we get to set ahead of time during the hypothesis testing).

On the other hand, **power** is the probability of correctly rejecting the null hypothesis and the probability of doing so is $1 - \beta$. If the alternative hypothesis is true, there are one of two things you can do. You might commit a type II error, and the probability of doing so is $\beta$, or you might correctly reject the null hypothesis in favor of the alternative, and the probability of that, then, is going to be the complement of $\beta$, $1- \beta$. And that's what we call our **power**.

In general, our goal in hypothesis testing is to keep both $\alpha$ and $\beta$ low at the same time. But the likelihood of making a type I error and likelihood of making a type II error are actually inversely proportional. Hence, sometimes we have to choose between which one we are okay with being a little higher versus which one we really want to minimize as much as possible.

e.g. For the criminal trial example if the jury's mindset is: It's Better that 10 guilty persons escape than one innocent suffer, that means we want to minimize the type I error (Declaring the defendant guilty when they're actually innocent).

---

Another caveat is that, even if we have a small **effect size** (difference between the point estimate and the null value), which may not be practically significant. We can still find the statistically significant result simply by inflating our sample size. Thus, we should do some a-priori analysis to determine how many observations to collect. e.g.

```{r}

# x bar is 50
# standard deviation is 2
# H0: mu = 49.5, H1: mu > 49.5
x  <- 50
sd <- 2
mu <- 49.5

# consider different sample size
# first a sample size of 100
z1 <- ( x - mu ) / ( sd / sqrt(100) )
print(z1)

# then inflate it to 10000 and 
# see the difference between the z score,
# recall that the higher the z score the lower
# our p-value
z2 <- ( x - mu ) / ( sd / sqrt(10000) )
print(z2)

```

> A pharmaceutical company has developed a new drug for lowering blood pressure and they are preparing a clinical trial to test the drug's effectiveness. They recruit people who are taking a particular standard blood pressure medication, and half of the subjects are given the new drug, this is the treatment group. And the other half continued to take their medication through generic-looking pills to ensure blinding, this is our control group. Previously published studies suggest that the standard deviation of the patients' blood pressures will be about 12 millimeters of mercury. Suppose that the company researchers care about finding any effect on blood pressure that is 3 millimeters of mercury or larger cersus the standard medication. Then what is the minimum required sample size for a desired level of power 80%.

```{r}

power.t.test( power = 0.8, sig.level = 0.05, delta = 3, sd = 12 )

```

# Reference

- [The Theorem Every Data Scientist Should Know](http://www.jeannicholashould.com/the-theorem-every-data-scientist-should-know-2.html)
- [RBlogger: One-way Analysis of Variance](http://www.r-bloggers.com/one-way-analysis-of-variance-anova/)
- [Power and sample size calculation for t test in R](http://users.stat.umn.edu/~corbett/classes/3022/Power_and%20_sample_size_for_t_in_R.txt)
- [Coursera: Inference Statistics](https://www.coursera.org/learn/inferential-statistics-intro)

### 2016-07-27 ab_tests/frequentist_ab_test.R

@@ -44,9 +44,8 @@ PlotPower( size = 5000, min_diff = 0.01 )
 # Power test
 baseline  <- 0.1  # baseline conversion rate 
 delta 	  <- 0.02 # minimum detectable boundary (practical significance boundary)
-power 	  <- 0.8  # sensitivity 
-sig_level <- 0.05 # specificity 
-
+power 	  <- 0.8  # specificity, or true negative rate
+sig_level <- 0.05 # false positive rate
 result <- power.prop.test( p1 = baseline, p2 = baseline - delta, 
 				 		   power = power, sig.level = sig_level,
 				 		   alternative = "two.sided" )
@@ -54,8 +53,8 @@ round(result$n)
 
 
 # -----------------------------------------------------------------------------
-baseline  <- 0.1 # baseline rate
-power 	  <- 0.8  # sensitivity 
+baseline  <- 0.1
+power 	  <- 0.8 
 sig_level <- 0.05
 
 # calculate the the required sample size for reaching the 
@@ -82,68 +81,52 @@ scale_x_continuous( labels = comma )
 # ------------------------------------------------------------------
 
 baseline  <- 0.1  # baseline conversion rate 
-delta 	  <- 0.02 # minimum detectable boundary ( practical significance boundary )
-power 	  <- 0.8  # sensitivity 
-sig_level <- 0.05 # specificity 
-
-result <- power.prop.test( p1 = baseline, p2 = baseline + delta, 
-				 		   power = power, sig.level = sig_level,
-				 		   alternative = "two.sided" )
+delta 	  <- 0.02 # minimum detectable boundary (practical significance boundary)
+power 	  <- 0.8  # specificity, or true negative rate
+sig_level <- 0.05 # false positive rate
+result <- power.prop.test( p1 = baseline, p2 = baseline - delta, 
+                           power = power, sig.level = sig_level,
+                           alternative = "two.sided" )
 round(result$n)
 
 
 # ------------------------------------------------------------------
 # Analyze the result 
 
-# parameters 
-count_control 	 <- 974
-sizes_control 	 <- 10072
+# parameters
+count_control <- 974
+sizes_control <- 10072
 count_experiment <- 1242
 sizes_experiment <- 9886
 
-ABTest <- function( count_control = count_control, 
-					sizes_control = sizes_control,
-					count_experiment = count_experiment, 
[…]

### 2016-07-27 articles/calendar_heatmaps/calendar_heatmaps.R

@@ -1,4 +1,3 @@
-library(DT)
 library(dplyr)
 library(scales)
 library(viridis)
@@ -7,12 +6,9 @@ library(ggthemes)
 library(gridExtra)
 library(lubridate)
 library(data.table)
-
-setwd('/Users/ethen/facetedcountryheatmaps')
-
-
+setwd('/Users/ethen/Business-Analytics/articles/calendar_heatmaps')
 attacks <- fread("data/eventlog.csv")
-DT::datatable(attacks)
+
 
 # preprocessing --------------------------------------------------------------
 make_hr_wkday <- function( ts, sc, tz ){

### 2016-08-04 articles/articles.Rmd
fix nbviewer format

@@
 
+- What factors can increase your data science salary?.
+[[Springboard](https://www.springboard.com/blog/highest-data-scientist-salary-possible/?utm_campaign=Data%2BElixir&utm_medium=email&utm_source=Data_Elixir_90)]
 - 5 Qualities You Need to Show in the Job Search.
@@
 
+- The Inconvenient Truth About Data Science. [[Linkedin Pulse](https://www.linkedin.com/pulse/inconvenient-truth-data-science-kamil-bartocha)]
 - Data science tips from reddit

### 2016-08-04 frequentist_statistics/frequentist_statistics.Rmd
fix nbviewer format

@@
 
-The general definition of a 95% confidence interval means that 95% of confidence intervals, **created based on random samples of the same size from the same population will contain the true population parameter**.
+The general definition of a 95% confidence interval means that **95% of confidence intervals, created based on random samples of the same size from the same population will contain the true population parameter**.
 

### 2016-08-23 articles/articles.Rmd
fix typos

@@
 
+- Cameron Davidson Pilon: Mistakes I've Made. [[Youtube](https://www.youtube.com/watch?v=VWtsTIbFXxA)]
+    - Always be aware of the sample sizes when computing aggregated data.
+    - Correlated data can lead to misinterpretation of the results.
+- Data Science: Beyond the Kaggle. [[Blog](http://datawrangl.com/2016/05/15/data-science-beyond-the-kaggle/)]
+- Mentality for hackathons. [[Medium](https://medium.com/data-engineering/modeling-madly-8b2c72eb52be#.q9ly8hxq2)]
+- The top 5 habits of a professional data scientist. [[OREILLY](https://www.oreilly.com/learning/theres-nothing-magical-about-learning-data-science?imm_mid=0e6c5a&cmp=em-data-na-na-newsltr_20160817)]
 - The Inconvenient Truth About Data Science. [[Linkedin Pulse](https://www.linkedin.com/pulse/inconvenient-truth-data-science-kamil-bartocha)]

### 2016-08-23 articles/data_hygiene/data_hygiene.Rmd
calendar heatmap
fix typos

@@
 **Little variability** e.g.
-99% of all the data points have the same category for one of the feauters.
+99% of all the data points have the same category for one of the features.
 If we have a large amount of data, and there is reasonable number of data points (the notion of reasonable really depends on our judgement, at least 30 ...?), then including it in the model is a viable choice.
@@
 Though the combination of categories should have a sound logical basis, as opposed to being combined due to having a similar target variable.
-
-- New data can't be predicted due to "unseen" category.
 
@@
 Both of which might be misleading to our model.
+
+**Unseen Category** Another thing we might be to check is whether the out test data contains "unseen" category during the modeling phase.
 
@@
 So what we could do, is have ranges such as 1-10, 11-18, 19-25, 26-40, etc.
-Moreover, instead of decomposing these categories as in point 2, we could just use scalar values, since age groups that lie "closer by" do represent similar properties.
+Moreover, instead of doing one hot encoding this these binned categories, we could just use scalar values, since age groups that lie "closer by" do represent similar properties.
 

### 2016-09-16 articles/data_hygiene/data_hygiene.Rmd
fix typos

@@
 
-Sometimes, missing category values can convery important messages.
+Sometimes, missing category values can convey important messages.
 A classific example is someone who is applying for loans and they non-reported their income category.
-Or maybe it shoudln't be missing and it's just something wrong with our data sources (e.g.
+Or maybe it shouldn't be missing and it's just something wrong with our data sources (e.g.
 database, or the way we're querying it).

### 2016-12-25 articles/articles.Rmd
fix typos
article update

@@
 
+- 2016.12.24 | Scope the Solution before Solving the Machine Learning. [[Blog](https://making.dia.co/scope-the-solution-before-solving-the-machine-learning-7e5ddb622733#.iwbgi17n1)]
+    1. what is the actual end goal
+    2. what type of outputs do I need
+    3. what type of models produces this output
+    4. what type of inputs would help me
+    5. are these inputs valid
+- The Most Practical Big Data Use Cases Of 2016.
+[[Forbes](http://www.forbes.com/sites/bernardmarr/2016/08/25/the-most-practical-big-data-use-cases-of-2016/?inf_contact_key=851fbbc780fa7ed1332d888a227af9856e3f8925ffc7f91b9e0f03dfa5587a98#4c9dd6697533)]
+    - Sales, Manufacture, Health Care
 - Cameron Davidson Pilon: Mistakes I've Made. [[Youtube](https://www.youtube.com/watch?v=VWtsTIbFXxA)]

### 2016-12-28 ab_tests/frequentist_ab_test.Rmd
rerun rmarkdown, add license

@@
 title: "Frequentist A/B test"
-author: "Ming-Yu Liu"
+author: "Ethen Liu"
 date: "`r Sys.Date()`"
@@
 
-PlotPower <- function( size, min_diff ){
+plot_power <- function(size, min_diff) {
 
@@
 	count_b  <- size_b * p_b
-	p_pooled <- ( count_a  + count_b ) / ( size_a + size_b )
-	Z <- ( p_b - p_a ) / sqrt( p_pooled * ( 1 - p_pooled ) * ( 1 / size_a + 1 / size_b ) )
+	p_pooled <- (count_a  + count_b) / (size_a + size_b)
+	Z <- (p_b - p_a) / sqrt( p_pooled * (1 - p_pooled) * (1 / size_a + 1 / size_b) )
 
@@
 
-	x <- seq( -4, 6, 0.1 ) # use for generating the x axis of the normal distribution
-	data <- data.frame( x = x, y1 = dnorm( x, mean1, 1 ), y2 = dnorm( x, mean2, 1 ) )
+	x <- seq(-4, 6, 0.1) # use for generating the x axis of the normal distribution
+	data <- data.frame( x = x, y1 = dnorm(x, mean1, 1), y2 = dnorm(x, mean2, 1) )
 
-	plot <- ggplot( data, aes( x = x ) ) +
+	plot <- ggplot( data, aes(x = x) ) +
 			geom_line( aes( y = y1, colour = 'H0 is true' ), size = 1.2 ) +
 			geom_line( aes( y = y2, colour = 'H1 is true' ), size = 1.2 ) +
-			geom_area( aes( y = y1, x = ifelse( x > 1.65, x, NA ) ), fill = 'black' ) +
-			geom_area( aes( y = y2, x = ifelse( x > 1.65, x, NA ) ), fill = 'blue', alpha = 0.3 ) +
-			labs( x = '', y = '', title = sprintf( 'p1 = %s, p2 = %s, size = %d', p_a, p_b, size ) ) + 
+			geom_area( aes( y = y1, x = ifelse(x > 1.65, x, NA) ), fill = 'black' ) +
+			geom_area( aes( y = y2, x = ifelse(x > 1.65, x, NA) ), fill = 'blue', alpha = 0.3 ) +
+			labs( x = '', y = '', title = sprintf('p1 = %s, p2 = %s, size = %d', p_a, p_b, size) ) + 
 			theme( legend.title = element_blank() ) +
-			scale_colour_manual( breaks = c( "H0 is true", "H1 is true" ), 
-								 values = c( "blue", "red" ) )
+			scale_colour_manual( breaks = c("H0 is true", "H1 is true"), 
+								 values = c("blue", "red") )
 	return(plot)
@@
 
-PlotPower( size = 5000, min_diff = 0.02 )
+plot_power(size = 5000, min_diff = 0.02)
 
@@
 # smaller N
-PlotPower( size = 2500, min_diff = 0.02 )
+plot_power(size = 2500, min_diff = 0.02)
 
 # smaller probability difference
-PlotPower( size = 5000, min_diff = 0.01 )
+plot_power(size = 5000, min_diff = 0.01)
 
@@
 # range detectable difference (dd)
-dd <- seq( from = 0.01, to = 0.03, by = 0.0001 )
-result <- matrix( nrow = length(dd), ncol = 2 )
-result[ , 1 ] <- dd
-for( i in 1:length(dd) ){
-	result[ i, 2 ] <- power.prop.test( p1 = baseline, p2 = baseline - dd[i], 
+dd <- seq(from = 0.01, to = 0.03, by = 0.0001)
+result <- matrix(nrow = length(dd), ncol = 2)
+result[, 1] <- dd
+for( i in 1:length(dd) ) {
+	result[i, 2] <- power.prop.test( p1 = baseline, p2 = baseline - dd[i], 
 									   power = power, sig.level = sig_level,
@@
 result <- data.table(result)
-setnames( result, c( 'dd', 'n' ) )
+setnames( result, c('dd', 'n') )
 
-ggplot(data = result, aes( x = dd, y = n ) ) +
+ggplot(data = result, aes(x = dd, y = n) ) +
 geom_line() + ylab("required sample size") + xlab("Detectable difference") + 
-scale_x_continuous( labels = comma )
+scale_x_continuous(labels = comma)
 
@@
 
-result <- prop.test( c( count_control, count_experiment ), 
-					 c( sizes_control, sizes_experiment ) )
+result <- prop.test( c(count_control, count_experiment), 
+					 c(sizes_control, sizes_experiment) )
 result
@@
 p2 <- count_experiment / sizes_experiment
-se <- sqrt( p1 * ( 1 - p1 ) / sizes_control + p2 * ( 1 - p2 ) / sizes_experiment )
+se <- sqrt( p1 * (1 - p1) / sizes_control + p2 * (1 - p2) / sizes_experiment )
 
@@
 conf_level <- 0.95
-zscore <- qnorm( conf_level + ( 1 - conf_level ) / 2 )
-conf_int <- abs(p2 - p1) + c( -1, 1 ) * zscore * se
+zscore <- qnorm( conf_level + (1 - conf_level) / 2 )
+conf_int <- abs(p2 - p1) + c(-1, 1) * zscore * se
 conf_int
@@
 
-ggplot( examples, aes( mean, scenario, color = scenario ) ) + 
+ggplot( examples, aes(mean, scenario, color = scenario) ) + 
 geom_point() + 
-geom_errorbarh( aes( xmin = lower, xmax = upper ), height = 0.1 ) + 
-geom_vline( xintercept = 0, color = "black" ) + 
-geom_vline( xintercept = delta, color = "blue", linetype = "dotted" ) + 
-geom_vline( xintercept = -delta, color = "blue", linetype = "dotted" ) +
+geom_errorbarh( aes(xmin = lower, xmax = upper), height = 0.1 ) + 
+geom_vline(xintercept = 0, color = "black") + 
+geom_vline(xintercept = delta, color = "blue", linetype = "dotted") + 
+geom_vline(xintercept = -delta, color = "blue", linetype = "dotted") +
 scale_color_discrete( breaks = as.character(1:6) ) +  
@@
 group2 <- 61818
-SanityCheck <- function( group1, group2 )
-{
+sanity_check <- function(group1, group2) {
 	# 95 percent confidence interval = qnorm(0.95 + (1-0.95) / 2 )
 	n <- group1 + group2
-	confidence <- n * 0.5 + c( -1, 1 ) * qnorm(0.975) * sqrt( n * 0.5 * 0.5 ) 
+	confidence <- n * 0.5 + c(-1, 1) * qnorm(0.975) * sqrt(n * 0.5 * 0.5) 
 	return(confidence)
 }
-( sanity <- SanityCheck( group1, group2 ) )
+( sanity <- sanity_check(group1, group2) )
 
@@
 
-# Reference
-
-- [Stitchfix's Blog: So, You Need a Statistically Significant Sample?](http://multithreaded.stitchfix.com/blog/2015/05/26/significant-sample/)
-- [Z-score distribution table](http://www.utdallas.edu/dept/abp/zscoretable.pdf)
-- [Minitab's Blog: How to Correctly Interpret P Values](http://blog.minitab.com/blog/adventures-in-statistics/how-to-correctly-interpret-p-values)
-
 # R Session Information
@@
 ```{r}
+devtools::session_info()
+```
 
-sessionInfo()
+# Reference
 
-```
+- [Blog: So, You Need a Statistically Significant Sample?](http://multithreaded.stitchfix.com/blog/2015/05/26/significant-sample/)
+- [Blog: How to Correctly Interpret P Values](http://blog.minitab.com/blog/adventures-in-statistics/how-to-correctly-interpret-p-values)
+- [Z-score distribution table](http://www.utdallas.edu/dept/abp/zscoretable.pdf)

### 2016-12-28 articles/articles.Rmd
rerun rmarkdown, add license

@@
 
+- 2016.12.27 | Crushed it! Landing a data science job. [[Blog](http://www.erinshellman.com/crushed-it-landing-a-data-science-job/)]
+- 2016.12.27 | http://treycausey.com/data_science_interviews.html. [[Blog](http://treycausey.com/data_science_interviews.html)]
 - What factors can increase your data science salary?.
@@
 
+- 2016.12.27 | Sharing Your Side Projects Online and Making Your Github the Best Résumé. [[Youtube](https://www.youtube.com/watch?v=uRul8QdYvqQ)]
+    - Include a license to the project. there's even a [website](http://choosealicense.com/) that walks you through the copying and pasting process
+    - [Binder](http://mybinder.org/) is a project that allows you to open up the jupyter notebook in a virual environment (specifying all the requirements in a requirements.txt file) and run it online without needing to download it
+- 2016.12.24 | WHAT I LEARNED RECREATING ONE CHART USING 24 TOOLS.
+[[Blog](https://source.opennews.org/en-US/articles/what-i-learned-recreating-one-chart-using-24-tools/?utm_campaign=Data%2BElixir&utm_medium=web&utm_source=Data_Elixir_110)]
 - 2016.12.24 | Scope the Solution before Solving the Machine Learning. [[Blog](https://making.dia.co/scope-the-solution-before-solving-the-machine-learning-7e5ddb622733#.iwbgi17n1)]

### 2016-12-28 frequentist_statistics/frequentist_statistics.Rmd
fix typo
rerun rmarkdown, add license

@@
 title: "Frequentist Statistic"
-author: "Ming-Yu Liu"
+author: "Ethen Liu"
 date: "`r Sys.Date()`"
@@
 library(knitr)
-opts_chunk$set( message = FALSE, warning = FALSE )
+opts_chunk$set(message = FALSE, warning = FALSE)
 
@@
 
-$$\text{point estimate} \pm z * SE $$
+$$\text{point estimate} \pm z * SE$$
 
@@
 cl <- 0.95
-c( -1, 1 ) * qnorm( ( 1 - cl ) / 2 )
+c(-1, 1) * qnorm( (1 - cl) / 2 )
 
@@
 
-3.2 + c( 1, -1 ) * qnorm( ( 1 - 0.95 ) / 2 ) * 1.74 / sqrt(50)
+3.2 + c(1, -1) * qnorm( (1 - cl) / 2 ) * 1.74 / sqrt(50)
 
@@
 
-# z-score is computed as ( value - mean ) / standard error
-upper <- ( 42 - 40 ) / se
-lower <- ( 38 - 40 ) / se
+# z-score is computed as (value - mean) / standard error
+upper <- (42 - 40) / se
+lower <- (38 - 40) / se
 pnorm(upper) - pnorm(lower)
@@
 # 1.65 is the critical value, z, for 90%
-( 1.65 * 18 / 4 ) ^ 2
+(1.65 * 18 / 4) ^ 2
 
@@
 # 1.65 is the critical value, z, for 90%
-( 1.65 * 18 / 2 ) ^ 2
+(1.65 * 18 / 2) ^ 2
 
@@
 # We're able to compute the confidence interval using the following formula
-3.2 + c( 1, -1 ) * qnorm( ( 1 - 0.95 ) / 2 ) * 1.74 / sqrt(50)
+3.2 + c(1, -1) * qnorm( (1 - 0.95) / 2 ) * 1.74 / sqrt(50)
 
@@
 # z score
-z <- ( 3.2 - 3 ) / se
-pvalue <- pnorm( z, lower.tail = FALSE )
+z <- (3.2 - 3) / se
+pvalue <- pnorm(z, lower.tail = FALSE)
 pvalue
@@
 If the p-value is low, and by low we mean lower than the significance level $\alpha$, which we usually set at 0.05, but we can certainly change it as well, we say that it would be very unlikely to observe the data if the null hypothesis were true, and therefore we reject the null hypothesis.
-If on the other hand the p-value is higher than $\alpha$, we say that it is indeed likely to observe the data even if the null hypothesis were true.
+If on the other hand the p-value is higher than $\alpha$, we say that it is indeed likely to observe the data if the null hypothesis were true.
 And hence, we would not reject the null hypothesis.
 
-So back to our original question.
-Our null hypothesis was that college students on average have three exclusive relationships, versus, the alternative was that that number was something greater than 3.
-
-In our case there is a 20.9% chance that a random sample of 50 college students would still yield a sample mean of 3.2 or higher.
+Back to our original question.
+Our null hypothesis was that college students on average have 3 exclusive relationships, and the alternative was that that number was something greater than 3.
+From the result we saw that there was a 20.9% chance that a random sample of 50 college students would still yield a sample mean of 3.2 or higher.
 Since this is a pretty high probability, we think that a sample mean of 3.2 or more exclusive relationships is likely to happen simply by chance.
@@
 > A small p-value indicates the data provides strong evidence against the null hypothesis.
+
+p-value is an extremely important concept in frequentist statistics.
+For those interested, this blog post so covers it with an pizza store example.
+[Blog: What Are P-Values?](https://prateekvjoshi.com/2013/12/07/what-are-p-values/)
 
@@
 
-df <- min( n1 - 1, n2 - 1 )
-critical <- qt( ( 1 - 0.95 ) / 2, df )
-se <- sqrt( s1 ^ 2 / n2 + s2 ^ 2 / n2 )
+df <- min(n1 - 1, n2 - 1)
+critical <- qt( (1 - 0.95) / 2, df )
+se <- sqrt(s1 ^ 2 / n2 + s2 ^ 2 / n2)
 
 # confidence interval
-print( ( x1 - x2 ) + c( 1, -1 ) * critical * se )
+print( (x1 - x2) + c(1, -1) * critical * se )
 
 # hypothesis test
-tscore <- ( x1 - x2 ) / se
-pt( tscore, df, lower.tail = FALSE ) * 2
+tscore <- (x1 - x2) / se
+pt(tscore, df, lower.tail = FALSE) * 2
 
@@
 
-```{r}
+```{r, message=FALSE, warning=FALSE}
 
@@
 plant_df$group <- factor( plant_df$group,
-						  labels = c( 'Control', 'Treatment 1', 'Treatment 2' ) )
+						  labels = c('Control', 'Treatment 1', 'Treatment 2') )
 
-ggplot( plant_df, aes( x = group, y = weight ) ) + 
+ggplot( plant_df, aes(x = group, y = weight) ) + 
 geom_boxplot()
@@
 
-anova_test <- aov( weight ~ group, data = plant_df )
+anova_test <- aov(weight ~ group, data = plant_df)
 summary(anova_test)
@@
 p  <- 0.9
-se <- sqrt( p * ( 1 - p ) / n )
+se <- sqrt( p * (1 - p) / n )
 
-zscore <- ( 0.95 - p ) / se
-pnorm( zscore, lower.tail = FALSE )
+zscore <- (0.95 - p) / se
+pnorm(zscore, lower.tail = FALSE)
 
@@
 # here, p and 1 - p is the same, thus we can use p squared
-( qnorm( 0.95 + ( 1 - ci ) / 2 ) * p / me ) ^ 2
+( qnorm( 0.95 + (1 - ci) / 2 ) * p / me ) ^ 2
 
@@
 p <- 0.5
-se <- sqrt(  p * ( 1 - p ) / n )
-zscore <- ( 0.6 - p ) / se
-pnorm( zscore, lower.tail = FALSE )
+se <- sqrt(  p * (1 - p) / n )
+zscore <- (0.6 - p) / se
+pnorm(zscore, lower.tail = FALSE)
 
@@
 
-obs <- c( 1920, 347, 19, 84, 130 )
+obs <- c(1920, 347, 19, 84, 130)
 
 # expected probability distribution
-prob <- c( 0.8029, 0.1206, 0.0079, 0.0292, 0.0394 )
-chisq.test( x = obs, p = prob )
+prob <- c(0.8029, 0.1206, 0.0079, 0.0292, 0.0394)
+chisq.test(x = obs, p = prob)
 
@@
 # first a sample size of 100
-z1 <- ( x - mu ) / ( sd / sqrt(100) )
+z1 <- (x - mu) / ( sd / sqrt(100) )
 print(z1)
@@
 # our p-value
-z2 <- ( x - mu ) / ( sd / sqrt(10000) )
+z2 <- (x - mu) / ( sd / sqrt(10000) )
 print(z2)
@@
 
-power.t.test( power = 0.8, sig.level = 0.05, delta = 3, sd = 12 )
+power.t.test(power = 0.8, sig.level = 0.05, delta = 3, sd = 12)
 
+```
+
+# R Session Information
+
+```{r}
+devtools::session_info()
 ```
@@
 
-- [The Theorem Every Data Scientist Should Know](http://www.jeannicholashould.com/the-theorem-every-data-scientist-should-know-2.html)
-- [RBlogger: One-way Analysis of Variance](http://www.r-bloggers.com/one-way-analysis-of-variance-anova/)
-- [Power and sample size calculation for t test in R](http://users.stat.umn.edu/~corbett/classes/3022/Power_and%20_sample_size_for_t_in_R.txt)
 - [Coursera: Inference Statistics](https://www.coursera.org/learn/inferential-statistics-intro)
+- [Blog: What Are P-Values?](https://prateekvjoshi.com/2013/12/07/what-are-p-values/)
+- [Blog: One-way Analysis of Variance](http://www.r-bloggers.com/one-way-analysis-of-variance-anova/)
+- [Blog: The Theorem Every Data Scientist Should Know](http://www.jeannicholashould.com/the-theorem-every-data-scientist-should-know-2.html)
+- [Notes: Power and sample size calculation for t test in R](http://users.stat.umn.edu/~corbett/classes/3022/Power_and%20_sample_size_for_t_in_R.txt)

### 2016-12-28 articles/avoid_pie_charts/avoid_pie_charts.Rmd
re-run charts with latest ggplot2

@@
 date: "`r Sys.Date()`"
-author: "Ming-Yu Liu"
+author: "Ethen Liu"
 output:
   rmdformats::readthedown:
-    highlight: haddock
+    highlight: pygments
 ---
@@
 > Source code can be obtained [here](https://github.com/ethen8181/Business-Analytics/blob/master/articles/avoid_pie_charts/avoid_pie_charts.R)
+
+# Pie Chart
 
@@
 - `panel.grid .minor / .major`  controls the grid lines in the plot.
-- `panel.margin` contols the margin between facets.
+- `panel.spacing` contols the margin between facets.
 - `as_labeller` the function that changes the strip's label without doesn't changing the underlying data.
@@
 	   panel.grid.major.y = element_line( color = "#b2b2b2" ),
-	   panel.margin.x = unit( 1, "cm" ),
-	   panel.margin.y = unit( 0.5, "cm" ),
+	   panel.spacing.x = unit( 1, "cm" ),
+	   panel.spacing.y = unit( 0.5, "cm" ),
 	   legend.position = "none" )
@@
 
+# R Session Information
+
+```{r}
+devtools::session_info()
+```
+
 # Reference
 
-- [Worst Chart In The World](http://www.businessinsider.com/pie-charts-are-the-worst-2013-6)
-- [How to replace a pie chart](http://varianceexplained.org/r/improving-pie-chart/)
-- [Take the Pies out of Pi Day](http://rud.is/b/2016/03/14/spinning-cycles-in-box-4-to-take-the-pies-out-of-pi-day/)
+- [Blog: Worst Chart In The World](http://www.businessinsider.com/pie-charts-are-the-worst-2013-6)
+- [Blog: How to replace a pie chart](http://varianceexplained.org/r/improving-pie-chart/)
+- [Blog: Take the Pies out of Pi Day](http://rud.is/b/2016/03/14/spinning-cycles-in-box-4-to-take-the-pies-out-of-pi-day/)

### 2016-12-28 articles/calendar_heatmaps/calendar_heatmaps.Rmd
re-run charts with latest ggplot2

@@
 title: "Calendar Heatmaps"
-author: "Ming-Yu Liu"
+author: "Ethen Liu"
 date: "`r Sys.Date()`"
@@
 attacks <- fread("data/eventlog.csv")
-head(attacks)
+knitr::kable( head(attacks) )
 
@@
 
-make_hr_wkday <- function( ts, sc, tz ){
+make_hr_wkday <- function(ts, sc, tz) {
 	# convert each time with the appropriate timezone,
@@
 
-# simply stores the name of the weekday
-# convert weekday and hour into factor so they'll be ordered
-# when plotting
-weekday_levels <- levels( weekdays( 0, abbreviate = FALSE ) )
+# convert weekday and hour into factor so they'll be ordered when plotting
+wkday_levels <- c('Sunday', 'Monday', 'Tuesday', 'Wednesday', 
+				  'Thursday', 'Friday', 'Saturday')
 attacks <- attacks %>% 
@@
 		   ungroup() %>% 
-		   mutate( wkday = factor( wkday, levels = weekday_levels ),
-		   		   hour  = factor( hour, levels = 0:23 ) )
-head(attacks)
+		   mutate( wkday = factor(wkday, levels = wkday_levels),
+		   		   hour  = factor(hour, levels = 0:23) )
+
+knitr::kable( head(attacks) )
 
@@
 
-grouped <- attacks %>% count( wkday, hour ) %>% ungroup()
+grouped <- attacks %>% count(wkday, hour) %>% ungroup()
 
-ggplot( grouped, aes( hour, wkday, fill = n ) ) + 
-geom_tile( color = "white", size = 0.1 ) + 
-theme_tufte( base_family = "Helvetica" ) + 
+ggplot( grouped, aes(hour, wkday, fill = n) ) + 
+geom_tile(color = "white", size = 0.1) + 
+theme_tufte(base_family = "Helvetica") + 
 coord_equal() + 
-scale_fill_viridis( name = "# of Events", label = comma ) + 
-labs( x = NULL, y = NULL, title = "Events per weekday & time of day" ) +
+scale_fill_viridis(name = "# of Events", label = comma) + 
+labs(x = NULL, y = NULL, title = "Events per weekday & time of day") +
 theme( axis.ticks = element_blank(),
-	   plot.title = element_text( hjust = 0.5 ),
-	   legend.title = element_text( size = 8 ),
-	   legend.text = element_text( size = 6 ) )
+	   plot.title = element_text(hjust = 0.5),
+	   legend.title = element_text(size = 8),
+	   legend.text = element_text(size = 6) )
 
@@
 
-gg <- ggplot( top_country_attacks, aes( x = hour, y = wkday, fill = n ) ) + 
-	  geom_tile( color = "white", size = 0.1 ) +
-	  scale_fill_viridis( name = "# Events" ) + 
+gg <- ggplot( top_country_attacks, aes(x = hour, y = wkday, fill = n) ) + 
+	  geom_tile(color = "white", size = 0.1) +
+	  scale_fill_viridis(name = "# Events") + 
 	  coord_equal() + 
 	  facet_wrap( ~source_country, ncol = 2 ) +
-	  labs( x = NULL, y = NULL, title = "Events per weekday & time of day by country\n" ) + 
-	  theme_tufte( base_family = "Helvetica" ) + 
+	  labs(x = NULL, y = NULL, title = "Events per weekday & time of day by country\n") + 
+	  theme_tufte(base_family = "Helvetica") + 
 	  theme( axis.ticks = element_blank(),
-			 axis.text = element_text( size = 8 ),
+			 axis.text = element_text(size = 8),
 			 panel.border = element_blank(),
-			 plot.title = element_text( hjust = 0.5 ),
-			 strip.text = element_text( hjust = 0.5 ),
-			 panel.margin = unit( 0.1, "cm" ),
+			 plot.title = element_text(hjust = 0.5),
+			 strip.text = element_text(hjust = 0.5),
+			 panel.margin = unit(0.1, "cm"),
 	  		 legend.position = "bottom",
-	  		 legend.title = element_text( size = 8 ),
-	  		 legend.text = element_text( size = 6 ),
-	  		 legend.key.size = unit( 0.4, "cm" ),
-	  		 legend.key.width = unit( 1, "cm" ) )
+	  		 legend.title = element_text(size = 8),
+	  		 legend.text = element_text(size = 6),
+	  		 legend.key.size = unit(0.4, "cm"),
+	  		 legend.key.width = unit(1, "cm") )
 gg
@@
 
-plots <- lapply( top_country, function(x){
+plots <- lapply(top_country, function(x) {
 
-	subset_data <- top_country_attacks %>% filter( source_country == x )
-	gg <- ggplot( subset_data, aes( x = hour, y = wkday, fill = n ) ) + 
-		  geom_tile( color = "white", size = 0.1 ) +
-		  scale_fill_viridis( name = "# Events" ) + 
-		  scale_y_discrete( expand = c( 0, 0 ) ) +
+	subset_data <- top_country_attacks %>% filter(source_country == x)
+	gg <- ggplot( subset_data, aes(x = hour, y = wkday, fill = n) ) + 
+		  geom_tile(color = "white", size = 0.1) +
+		  scale_fill_viridis(name = "# Events") + 
+		  scale_y_discrete( expand = c(0, 0) ) +
 		  coord_equal() + 
-		  labs( x = NULL, y = NULL, title = x ) + 
-		  theme_tufte( base_family = "Helvetica" ) + 
+		  labs(x = NULL, y = NULL, title = x) + 
+		  theme_tufte(base_family = "Helvetica") + 
 		  theme( axis.ticks = element_blank(),
-				 axis.text = element_text( size = 7 ),
+				 axis.text = element_text(size = 7),
 				 panel.border = element_blank(),
-				 plot.title = element_text( hjust = 0.5 ),
-				 strip.text = element_text( hjust = 0.5 ),
-				 panel.margin = unit( 0.1, "cm" ),
+				 plot.title = element_text(hjust = 0.5),
+				 strip.text = element_text(hjust = 0.5),
+				 panel.margin = unit(0.1, "cm"),
 		  		 legend.position = "bottom",
-		  		 legend.title = element_text( size = 6 ),
-		  		 legend.text = element_text( size = 6 ),
-		  		 legend.key.size = unit( 0.4, "cm" ) )
+		  		 legend.title = element_text(size = 6),
+		  		 legend.text = element_text(size = 6),
+		  		 legend.key.size = unit(0.4, "cm") )
 	return(gg)
@@
 
+# R Session Information
+
+```{r}
+devtools::session_info()
+```
+
 # Reference
 
-- [Making faceted heatmaps with ggplot2](https://rud.is/b/2016/02/14/making-faceted-heatmaps-with-ggplot2/)
-- [Big Data Insights - IT Support Log Analysis](http://costaleconomist.blogspot.tw/2015/12/big-data-insights-it-support-log.html)
+- [Blog: Making faceted heatmaps with ggplot2](https://rud.is/b/2016/02/14/making-faceted-heatmaps-with-ggplot2/)
+- [Blog: Big Data Insights - IT Support Log Analysis](http://costaleconomist.blogspot.tw/2015/12/big-data-insights-it-support-log.html)

### 2016-12-28 articles/nyt_scatter/nyt_scatter.Rmd
re-run charts with latest ggplot2

@@
 
-# 2016.5.12 you have to install the development version of ggplot2
-# devtools::install_github("hadley/ggplot2") or subtitles won't work
 library(tidyr)
@@
 file <- "nytimes_vote.tsv"
-if( !file.exists(file) ){
+if( !file.exists(file) ) {
 	url <- "https://static01.nyt.com/newsgraphics/2016/04/21/undervote/ad8bd3e44231c1091e75621b9f27fe31d116999f/data.tsv"
-	download.file( url, file )
+	download.file(url, file)
 }
@@
 df <- read_tsv(file)
-df <- df %>% rename( "Someone else" = undervt, 
+df <- df %>% rename("Someone else" = undervt, 
 					"Hilary Clinton" = clintonpct, 
-					 "Bernie Sanders" = sanderspct )
-df
+					"Bernie Sanders" = sanderspct)
 
-df <- gather( select( df, -tvotes ), party, pct, -ratio, -fips )
-df <- arrange( df, fips, ratio )
+df <- gather( select(df, -tvotes), party, pct, -ratio, -fips )
+df <- arrange(df, fips, ratio)
 df <- df %>% mutate( 
-	party = factor( party, levels = c("Hilary Clinton", "Bernie Sanders", "Someone else" ) )
+	party = factor( party, levels = c("Hilary Clinton", "Bernie Sanders", "Someone else") )
 )
@@
 
-ggplot( df, aes( x = ratio, y = pct ) ) + 
-geom_point( aes( fill = party ), size = 3, alpha = 0.8, color = "white", shape = 21 ) + 
-scale_fill_manual( name = "", values = fill_color ) +
+ggplot( df, aes(x = ratio, y = pct) ) + 
+geom_point( aes(fill = party), size = 3, alpha = 0.8, color = "white", shape = 21 ) + 
+scale_fill_manual(name = "", values = fill_color) +
 theme_bw( base_family = "Arial Narrow" ) + 
-scale_y_continuous( label = percent, limits = c( 0, 1.05 ) ) +
-scale_x_continuous( limits = c( 0, 4.5 ), breaks = seq( 0, 4.5, 0.5 ) ) +
-geom_text( data = data.frame( label = "↑ Share of 2016 primary vote"),
-		   aes( x = 0, y = 1, label = label ), vjust = -1, hjust = 0, size = 3,
+scale_y_continuous( label = percent, limits = c(0, 1.05) ) +
+scale_x_continuous( limits = c(0, 4.5), breaks = seq(0, 4.5, 0.5) ) +
+geom_text( data = data.frame(label = "↑ Share of 2016 primary vote"),
+		   aes(x = 0, y = 1, label = label), vjust = -1, hjust = 0, size = 3,
 		   fontface = "bold", family = "Arial Narrow" ) + 
@@
 	   legend.position = "top",
-	   plot.title = element_text( face = "bold" ),
+	   plot.title = element_text(face = "bold"),
 	   axis.ticks = element_blank(),
-	   axis.text = element_text( size = 8 ),
-	   axis.title.x = element_text( hjust = 1, face = "bold", size = 9 ),
+	   axis.text = element_text(size = 8),
+	   axis.title.x = element_text(hjust = 1, face = "bold", size = 9),
 	   panel.grid.minor = element_blank(),
-	   panel.grid.major = element_line( linetype = "dotted", size = 0.5 ),
+	   panel.grid.major = element_line(linetype = "dotted", size = 0.5),
 	   panel.border = element_blank(),
-	   plot.margin = margin( t = 10, r = 10, b = 10, l = 10 ) )
+	   plot.margin = margin(t = 10, r = 10, b = 10, l = 10) )
 
+```
+
+# R Session Information
+
+```{r}
+devtools::session_info()
 ```
@@
 
-- [Gist](https://gist.github.com/hrbrmstr/4632dc9cefc7566b412ec9945a6c4922)
+- [Gist: nytscatter.r](https://gist.github.com/hrbrmstr/4632dc9cefc7566b412ec9945a6c4922)

### 2016-12-28 finding_groups/finding_groups.Rmd
re-run charts with latest ggplot2

@@
 title: "Finding Groups Within the Data"
-author: "Ming-Yu Liu"
+author: "Ethen Liu"
 date: "2016-04-05"
@@
 
-ggplot( data_sku, aes( CV, ADS ) ) + geom_point() +  
-geom_hline( yintercept = 4, color = "blue" ) + 
-geom_vline( xintercept = .2, color = "blue" ) + 
+ggplot( data_sku, aes(CV, ADS) ) + geom_point() +  
+geom_hline(yintercept = 4, color = "blue") + 
+geom_vline(xintercept = .2, color = "blue") + 
 labs( x = "Coefficient of Variation", y = "Average Daily Sales", 
@@
 
-```{r}
+```{r, results='hide'}
 
@@
 data_sku[ ,  groups := as.factor( cutree( cluster, k = 3 ) ) ]
-head(data_sku)
 
@@
 
-ggplot( data_sku, aes( CV, ADS, color = groups ) ) + geom_point() + 
+head(data_sku)
+ggplot( data_sku, aes(CV, ADS, color = groups) ) + geom_point() + 
 labs( x = "Coefficient of Variation", y = "Average Daily Sales", 
@@
 In this module this number is chosen mostly by common sense and business criteria.
-If you're interested in a more technical way of choosing the suitable number and evaluate the stability of your clustering result, you can refer to another documentation [here](http://ethen8181.github.io/machine-learning/clustering/clustering.html).
 
@@
 ```{r}
-
-sessionInfo()
-
+devtools::session_info()
 ```

### 2016-12-28 articles/calendar_heatmaps/calendar_heatmaps.R

@@ -11,7 +11,7 @@ attacks <- fread("data/eventlog.csv")
 
 
 # preprocessing --------------------------------------------------------------
-make_hr_wkday <- function( ts, sc, tz ){
+make_hr_wkday <- function(ts, sc, tz) {
 	# convert each time with the appropriate timezone,
 	# the timezone parameter, tz, only takes a single value,
 	# then extract its weekdays and hour
@@ -22,16 +22,15 @@ make_hr_wkday <- function( ts, sc, tz ){
 	return(dt)
 }
 
-# simply stores the name of the weekday
-# convert weekday and hour into factor so they'll be ordered
-# when plotting
-weekday_levels <- levels( weekdays( 0, abbreviate = FALSE ) )
+# convert weekday and hour into factor so they'll be ordered when plotting
+wkday_levels <- c('Sunday', 'Monday', 'Tuesday', 'Wednesday', 
+				  'Thursday', 'Friday', 'Saturday')
 attacks <- attacks %>% 
 		   group_by(tz) %>%
 		   do( make_hr_wkday( .$timestamp, .$source_country, .$tz ) ) %>% 
 		   ungroup() %>% 
-		   mutate( wkday = factor( wkday, levels = weekday_levels ),
-		   		   hour  = factor( hour, levels = 0:23 ) )
+		   mutate( wkday = factor(wkday, levels = wkday_levels),
+		   		   hour  = factor(hour, levels = 0:23) )
 head(attacks)
 
 # attacks --------------------------------------------------------------------
@@ -39,18 +38,18 @@ head(attacks)
 # then we can simply group the count by hour and wkday
 # since we know that we have values for every combination
 # there's no need to further preprocess the data
-grouped <- attacks %>% count( wkday, hour ) %>% ungroup()
+grouped <- attacks %>% count(wkday, hour) %>% ungroup()
 
-ggplot( grouped, aes( hour, wkday, fill = n ) ) + 
-geom_tile( color = "white", size = 0.1 ) + 
-theme_tufte( base_family = "Helvetica" ) + 
+ggplot( grouped, aes(hour, wkday, fill = n) ) + 
+geom_tile(color = "white", size = 0.1) + 
+theme_tufte(base_family = "Helvetica") + 
 coord_equal() + 
-scale_fill_viridis( name = "# of Events", label = comma ) + 
[…]

### 2016-12-28 articles/nyt_scatter/nyt_scatter.R

@@ -1,5 +1,3 @@
-# 2016.5.12 you have to install the development version of ggplot2
-# devtools::install_github("hadley/ggplot2") or subtitles won't work
 library(tidyr)
 library(dplyr)
 library(readr)
@@ -9,21 +7,21 @@ setwd("/Users/ethen/Business-Analytics/articles/nyt_scatter")
 
 
 file <- "nytimes_vote.tsv"
-if( !file.exists(file) ){
+if( !file.exists(file) ) {
 	url <- "https://static01.nyt.com/newsgraphics/2016/04/21/undervote/ad8bd3e44231c1091e75621b9f27fe31d116999f/data.tsv"
-	download.file( url, file )
+	download.file(url, file)
 }
 
 # convert to more informative column name
 df <- read_tsv(file)
-df <- df %>% rename( "Someone else" = undervt, 
-					 "Hilary Clinton" = clintonpct, 
-					 "Bernie Sanders" = sanderspct )
+df <- df %>% rename("Someone else" = undervt, 
+					"Hilary Clinton" = clintonpct, 
+					"Bernie Sanders" = sanderspct)
 
-df <- gather( select( df, -tvotes ), party, pct, -ratio, -fips )
-df <- arrange( df, fips, ratio )
+df <- gather( select(df, -tvotes), party, pct, -ratio, -fips )
+df <- arrange(df, fips, ratio)
 df <- df %>% mutate( 
-	party = factor( party, levels = c("Hilary Clinton", "Bernie Sanders", "Someone else" ) )
+	party = factor( party, levels = c("Hilary Clinton", "Bernie Sanders", "Someone else") )
 )
 df
 
@@ -36,26 +34,26 @@ fill_color <- c(
 	"Someone else" = "#d65454" 
 )
 
-ggplot( df, aes( x = ratio, y = pct ) ) + 
-geom_point( aes( fill = party ), size = 3, alpha = 0.8, color = "white", shape = 21 ) + 
-scale_fill_manual( name = "", values = fill_color ) +
+ggplot( df, aes(x = ratio, y = pct) ) + 
+geom_point( aes(fill = party), size = 3, alpha = 0.8, color = "white", shape = 21 ) + 
+scale_fill_manual(name = "", values = fill_color) +
 theme_bw( base_family = "Arial Narrow" ) + 
-scale_y_continuous( label = percent, limits = c( 0, 1.05 ) ) +
-scale_x_continuous( limits = c( 0, 4.5 ), breaks = seq( 0, 4.5, 0.5 ) ) +
[…]

### 2016-12-31 articles/articles.Rmd
license badge

@@
 date: "`r Sys.Date()`"
-author: "Ming-Yu Liu"
+author: "Ethen Liu"
 output:
@@
 
+- 2016.12.29 | Data Science Newsletter list.
+[[Medium](https://medium.com/blend-your-data/a-four-minute-read-big-data-and-data-science-newsletter-list-to-get-you-informed-on-anything-data-669d8e2c4b09#.2lukcelgg)]
 - 2016.12.27 | Sharing Your Side Projects Online and Making Your Github the Best Résumé. [[Youtube](https://www.youtube.com/watch?v=uRul8QdYvqQ)]

### 2017-01-14 R/Rmarkdown/Rmarkdown.Rmd
add R programming folder

---
title: "Rmarkdown Quickstart"
author: "Ethen Liu"
date: "`r Sys.Date()`"
output: 
  rmdformats::readthedown:
    highlight: pygments
---

<style type="text/css">
p{ /* Normal  */
   font-size: 18px;
}
body{ /* Normal  */
   font-size: 18px;
}
td {  /* Table  */
   font-size: 14px;
}
h1 { /* Header 1 */
 font-size: 32px;
}
h2 { /* Header 2 */
 font-size: 26px;
}
h3 { /* Header 3 */
 font-size: 22px;
}
code.r{ /* Code block */
  font-size: 14px;
}
pre { /* Code block */
  font-size: 14px
}
</style>

## R Markdown

This is an R Markdown document. Markdown is a simple formatting syntax for authoring HTML, PDF, and MS Word documents. For more details on using R Markdown see <http://rmarkdown.rstudio.com>.

So with Markdown you can:

- Create inline code-like highlighting by putting `back-ticks around` it
- Create different headers by using the pound sign `# header name`
- Create bold characters by putting double asterisks around it, e.g. `**characters**`, or one asterisk around it for italic characters. e.g. `*characters*`
- Create blockquotes using the greater than symbol. e.g. `> quote here`
- Create links. e.g. [I'm an inline-style link](https://www.google.com)
- Create un-ordered bullet points by using `- bullet`, or use numbers to create ordered bullet points, `1. bullet`
- Write latex equations by putting dollar signs around it, e.g. typing `$\mu$`, will give you $\mu$. Put two dollars sign for block equations. e.g. typing `$$\mu$$`

$$\mu$$

For a more comprehensive list, consider this link. [Github: Markdown Here Cheatsheet](https://github.com/adam-p/markdown-here/wiki/Markdown-Here-Cheatsheet)

When you click the **Knit** button a document will be generated that includes both content as well as the output of any embedded R code chunks within the document.

Next, you can embed an R code chunk. What you do is you initiate the code chunk with three backticks and a r within a curly bracket. After the code, there will be a line with just three backticks.

```{r cars}
# we can generate code outputs
summary(cars)
```

```{r}
# or figures
plot(pressure)
```

Or more efficiently, there's a insert button on the Rstudio panel above where you can simply click and it will generate the template for you.

You can give each code chunk a name, like cars above. The name is optional; if included, each code chunk needs a distinct name. One reason for doing this, say we have a `.Rmd` file and we only want the source code of that file. We can do this by using `knitr::purl(.Rmd file path here)`, the purl function from the knitr package on this file, this will extract solely the source code and it will also tag it with the code chunk name if there is one. This makes it a bit easier to track for errors or make changes.

Notice that we're using `knitr::purl` to call the purl function from the knitr package instead of loading the library knitr and calling the function like `library(knitr) purl(.Rmd file here)`. The rationale is that: people might not be familiar with the `purl` function, by doing this, the others will know this is a function coming from the knitr package. And the other is, when loading a library, we're loading every functions available from that package. If we're just going to use 1 of them, consider using this method to avoid polluting the workspace.

```{r}

# call knitr::purl(.Rmd file path here)
# on this .Rmd file and notice the difference
test <- 1
test

```

A reproducible report should never manually write down numbers that are derived from the data. Instead we use inline code. e.g. To refer to the value of the test variable, instead of manually writing the test value is 1, we can write the test value is `r test`, created by wrapping a backtick around the **r** character, a space and the variable name. As shown below:

```{r, eval=FALSE}
`r test`
```

## Code Chunk Options

There are many parameters you can specify for the R code chunk. For example:

- `eval = FALSE` Display the r code without evaluating it
- `echo = FALSE` The code will not be shown in the final document, but the corresponding results/output would still be displayed
- `results = "hide"` To hide the results/output, but the code would still be displayed
- `include = FALSE` The code chunk will be evaluated, but neither the code nor its output displayed
- For figures, you’ll want to use options like `fig.width` and `fig.height` to control the size, the best way is simply try and error (try different numbers and see which works best), or use `fig.show = "hide"` to hide figures. And there is `fig.align='center'` to center the plot (the default is left-aligned)
- For figures, we can also specify `dev = 'svg'` to change the default graphics device from png to scalable vector graphics, which may look better on the web.
- `message = FALSE` and `warning = FALSE` to suppress message and warning. Some packages will print these out when loading them, use these two options to remove them from the output file

For these options, there will be autocompletion when you start typing them in the r code chunk, which makes life easier for you.

```{r, fig.align='center', fig.height=5, fig.width=5}
# a plot using center, and 5 for both height and width
plot(pressure)
```

We can set the options globally, so it applies to all code chunk by:

```{r setup}
# for all chode chunks all code will be displayed
knitr::opts_chunk$set(echo = TRUE)
```

We usually set this a the very top of the Rmarkdown file and set `echo = FALSE` so it doesn't get shown. There are a bunch of other options, listed here. [Blog: Chunk options and package options](http://yihui.name/knitr/options/#chunk_options)

## Displaying Tables

One other cool thing is we can display data.frame objects in a table like manner. Consider the differences between the outputs below

```{r}
head(mtcars)
```

```{r}
knitr::kable( head(mtcars) )
```

We can pass the `results='asis'` options to the code chunk to ensure that the raw table output isn’t processed furthur by knitr. The `kable` function includes several options to control the maximum number of digits for numeric columns, alignment, check the help page, `?knitr::kable`, for more details.

We can even create dynamic table if necessary using the `datatable` function from the `DT` package.

```{r}
DT::datatable(mtcars)
```

# Reproducible Report

### Session Information

Include "session info" in your document, preferably at the bottom: this lists the version of R that you’re using, plus all of the packages you’ve loaded. So people that wishes to reproduce your work will know which version of R, packages you're using and when you're running the report.

```{r}

# or use sessionInfo() that's included with R
# see which format do you prefer
devtools::session_info()

```

### Avoid Absolute Path

Try to avoid using path like `/Users/ethen/Desktop/` across your documents. If you must set the working directory, set it at the beginning of the report, and consider using `file.path` so they are independent of the system that you're working on (Windows, Mac, Linux). Or simply use `normalizePath`.

```{r}
file_path <- file.path('/Users', 'ethen', 'Desktop')

# alternative
file_path <- normalizePath('/Users/ethen/Desktop')
setwd(file_path)
```

## Bonus: Rmarkdown Themes

People have created different themes for Rmarkdown. e.g. [Github: prettydoc](http://yixuan.cos.name/prettydoc/). We can change the theme by adding extra information in the header. For example, we can change the header info into:

```r
---
title: "Week 1"
author: "Ethen Liu"
date: "`r Sys.Date()`"
output:
  prettydoc::html_pretty:
    theme: leonids
    highlight: github
---
```

For the prettydoc package, the theme option can take the following values
`cayman`, `tactile`, `architect`, `leonids`, `hpstr`. The code highlight option takes value of `github` and `vignette`. More may be added in the future.

## Reference

- [Blog: R Code Chunks](http://rmarkdown.rstudio.com/authoring_rcodechunks.html)
- [Blog: Knitr with R Markdown](http://kbroman.org/knitr_knutshell/pages/Rmarkdown.html)
- [Blog: Figures and tables](http://kbroman.org/knitr_knutshell/pages/figs_tables.html)
- [Blog: Chunk options and package options](http://yihui.name/knitr/options/#chunk_options)
- [Github: Markdown Here Cheatsheet](https://github.com/adam-p/markdown-here/wiki/Markdown-Here-Cheatsheet)

### 2017-01-14 R/data_table/data_table.Rmd
add R programming folder

---
title: "R's data.table"
author: "Ethen Liu"
date: "`r Sys.Date()`"
output: 
  rmdformats::readthedown:
    highlight: pygments
---

<style type="text/css">
p{ /* Normal  */
   font-size: 18px;
}
body{ /* Normal  */
   font-size: 18px;
}
td {  /* Table  */
   font-size: 14px;
}
h1 { /* Header 1 */
 font-size: 32px;
}
h2 { /* Header 2 */
 font-size: 26px;
}
h3 { /* Header 3 */
 font-size: 22px;
}
code.r{ /* Code block */
  font-size: 14px;
}
pre { /* Code block */
  font-size: 14px
}
</style>

> The **source of this tutorial**, including the example datasets, is available [here on GitHub](https://github.com/ethen8181/Business-Analytics/blob/master/R/data_table)

# SQL Joins

For joining data.tables, the basic syntax are:

JOIN type   | DT syntax         | data.table::merge() syntax
----------- | ----------------- | --------------------------
INNER       | X[Y, nomatch = 0] | merge(X, Y, all = FALSE)
LEFT OUTER  | Y[X]              | merge(X, Y, all.x = TRUE)
RIGHT OUTER | X[Y]              | merge(X, Y, all.y = TRUE)
FULL OUTER  | -                 | merge(X, Y, all = TRUE)

---

The example data consists of the following two tables:

```{r}

library(data.table)
setwd("/Users/ethen/Business-Analytics/R/data_table")
employees <- fread("data/employees.csv")
departments <- fread("data/departments.csv")
knitr::kable(employees, caption = "Table Employees")
knitr::kable(departments, caption = "Table Departments")

```

## Inner Join

*Figure linked from [Pinal Dave][Pinal]*  
![Figure1 from Pinal Dave](http://www.pinaldave.com/bimg/March09UG/inner%20join.jpg)

**the *INNER JOIN* returns the rows with a match in both tables**

```{r}

# inner join with merge
merge(employees, departments, by = "Department")

# inner join with keys
# set the keys of the tables to represent the by:
setkey(employees, Department)
setkey(departments, Department)
# equivalent to setkeyv( departments, "Department" )

# note that you only need to set the keys once
# you can confirm if it works with
key(employees)

# perform the join,
# nomatch 0 means no rows will be returned 
# from not matched rows from the Right table
employees[ departments, nomatch = 0 ]

```

As we'll see later, without anything else, `TABLE_X[TABLE_Y]` returns a **right outer join**; setting `nomatch=0` returns a **inner join**.

**note:**

> Employee #6, Ferdinand, has not been returned, as he's in a yet to be created department 21.
> Neither has department #14, Research, as there are no employees.

## Left Outer Join

*Figure linked from [Pinal Dave][Pinal]*  
![Figure2 from Pinal Dave](http://www.pinaldave.com/bimg/March09UG/left%20join.jpg)

**the *LEFT OUTER JOIN* returns all the rows from the left table, filling in matched columns (or NA) from the right table**

```{r}

# left join with merge, notice that the column's ordering is different
merge(employees, departments, by = "Department", all.x = TRUE)
departments[employees]

```

**note:**

> Employee #6, Ferdinand, has been returned with department name as NA.  
> Department #14, Research, has not been returned.

## Right Outer Join

*Figure linked from [Pinal Dave][Pinal]*  
![Figure2 from Pinal Dave](http://www.pinaldave.com/bimg/March09UG/right%20join.jpg)

**the *RIGHT OUTER JOIN* returns all the rows from the right table, filling in matched columns (or else NA) from the left table**

```{r}

# right join with merge
merge(employees, departments, by = "Department", all.y = TRUE)
employees[departments]

# we can also do a not matched with !
# returning the rows in the employees table that are not in the
# department table
employees[!departments]

```

## Full Outer Join

*Figure linked from [Pinal Dave][Pinal]*  
![Figure2 from Pinal Dave](http://www.pinaldave.com/bimg/March09UG/outer%20join.jpg)

**the *FULL OUTER JOIN* returns all the rows from both tables, filling in matched columns (or else NA)**

```{r}

merge(employees, departments, by = "Department", all = TRUE)

```

# Other Tips and Tricks

## `.SD` and `.N`

First `.SD` refers to the subset of data for each group. The example below gets the sum of y and z and the number of rows (using `.N`) in each group while grouped by x.

```{r}

DT <- data.table( x = sample(1:2, 7, replace = TRUE), 
				  y = sample(1:13, 7), 
				  z = sample(1:14, 7) ) 
DT[ , c( lapply(.SD, sum), .N ), by = x ]
```

## Key Indexing

Set the keys and select the matching rows without using `==`.

```{r}

DT <- data.table( A = letters[ c(2, 1, 2, 3, 1, 2, 3) ],
				  B = c(5, 4, 1, 9, 8, 8, 6), C = 6:12 )
setkey(DT, A, B)

# matches rows where A == "b"
DT["b"]

# matches rows where A == "b" or A == "c"
DT[ c("b", "c"), ]

```

Since we did not provide any values for the second key column "B", it just matches c( "b", "c" ) against the first key column "A" and returns all the matched rows. Consider the example below that matches both keys.

```{r}

DT[ .("b", 5), ]

```

"b" is first matched against the first key column "A". And within those matching rows, 5 is matched against the second key column "B" to obtain row indices where both "A" and "B" match the given values.

```{r}

# Subset all rows where just the second key column matches
DT[ .(unique(A), 5), ]

```

## Selecting Parts of a Filtered Group

Select the first row of the `b` and `c` groups using `mult`.

```{r}

DT[ c("b", "c"), mult = "first" ]

```

When i is a data.table, `DT[ i, j, by = .EACHI ]` evaluates j for the groups of DT that each row in i joins to. We call this grouping by each i. See difference with and without `.EACHI` below.

```{r}

# we can use [] for .SD to subset the data
DT[ c("b", "c"), .SD[ c(1, .N), ] ]
DT[ c("b", "c"), .SD[ c(1, .N), ], by = .EACHI ]

```

Notice that `.N` here is used to retrieve the last row.

## R Session Information

```{r}
sessionInfo()
```

## Reference

- [Joining data in R using data.table](https://rstudio-pubs-static.s3.amazonaws.com/52230_5ae0d25125b544caab32f75f0360e775.html)
- [Data analysis in R using data.table](https://rpubs.com/davoodastaraky/dataTable)

[Pinal]:http://blog.sqlauthority.com/2009/04/13/sql-server-introduction-to-joins-basic-of-joins/

### 2017-01-14 R/efficient_looping/efficient_looping.Rmd
add R programming folder

---
title: "Efficient Looping with R"
author: "Ethen Liu"
date: "`r Sys.Date()`"
output: 
  rmdformats::readthedown:
    highlight: pygments
---

<style type="text/css">
p{ /* Normal  */
   font-size: 18px;
}
body{ /* Normal  */
   font-size: 18px;
}
td {  /* Table  */
   font-size: 14px;
}
h1 { /* Header 1 */
 font-size: 32px;
}
h2 { /* Header 2 */
 font-size: 26px;
}
h3 { /* Header 3 */
 font-size: 22px;
}
code.r{ /* Code block */
  font-size: 14px;
}
pre { /* Code block */
  font-size: 14px
}
</style>

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Apply Family

`lapply`: list apply. When you want to apply a function to each element of a list and get a list back.

```{r}
x <- list(a = 1, b = 1:3, c = 10:100) 
lapply(x, length)
lapply(x, sum)
```

`sapply`: simplified apply. When you want to apply a function to each element of a list and get avector back.

```{r}
x <- list(a = 1, b = 1:3, c = 10:100) 
sapply(x, length)
sapply(x, sum)
```

But never use this function as there's a faster alternative below.

`vapply`: vector apply. It provides the same functionality as sapply, but we also explicitly state the type of data that it will return instead of letting R spend time figuring out. This slightly extra effort can speed up our code a bit.

```{r}
# everything returned should be
# an numeric of length 1, other commonly
# used type is character
x <- list(a = 1, b = 1:3, c = 10:100) 
vapply(x, length, numeric(1))

# we can specify a anonymous function for the function
vapply(x, function(i) {
    # the index i represents each element in the sequence
    summed <- sum(i, na.rm = TRUE)
    return(summed)
}, numeric(1))
```

`mapply`: When we have multiple data structures (e.g. vectors, lists) and we want to apply a function to each elements of all the input sequence.

```{r}
x1 <- c('Micheal', 'Kobe')
x2 <- c('Jordan', 'Bryant')

# we can also specify the function outside and
# pass it as an argument, the ..., means to
# accept any number of argument
paste2 <- function(...) paste(..., sep = '-')
mapply(paste2, x1, x2)

# note that this is just for illustration, 
# for the code above, we can simply do:
paste(x1, x2, sep = '-')

```

# Parallel Programming

When we say parllel programming, it's worth noting that there's a difference between the implementation of parallel code that runs locally on a single multi-processor (multi-cores) computer or on a cluster of computers, as well as implementation for Unix vs Windows system.

Here we'll illustrate the implementation of parallel code that runs locally on a single multi-processor (multi-cores) computer using the `doParallel` library. This package provides a parallel backend for the `foreach` package and represents a merger of `doMC` (works on Unix-like system) and `doSNOW` (works on Windows) packages.

Let’s say we have an N-core processor (on a local machine) and we want to use N-1 cores to run our program.

```{r, message=FALSE, warning=FALSE}
library(doParallel)

# we specify the number of cores/workers we want to use
n_cores <- detectCores() - 1
n_cores

# generate a toy function that
# simply generate the summary of a bunch of random numbers
summarize <- function(i) {
    summary( rnorm(100000) )
}

# the time difference between using n_cores and not using it
inputs <- 1:20
system.time({
    results <- mclapply(inputs, summarize, mc.cores = n_cores)
})

system.time({
    results <- lapply(inputs, summarize)
})
```

**Warning** From the documentation, `mclapply` relies on Unix system, thus it will most likely not work on Windows. Consider trying the following code on Windows:

```r
# explicitly create the cluster
cl <- makeCluster(n_cores)

# run the parallel code using parLapply and
# pass the created cluster as the first argument
system.time({
    results <- parLapply(cl, inputs, summarize)
})

# remember to stop the cluster once we're finished
stopCluster(cl)
```

### Caveat 1: Overhead

Sending tasks out to another processesor and have work to be done there takes some time and there's some work involved, thus we don't want to parallelizing things that already runs really fast sequentially. Especially in R where some functions can be vectorized.

```{r}
# overhead when parallelizing trivial tasks
inputs <- 1:10000
system.time({
    results <- mclapply(inputs, sqrt, mc.cores = n_cores)
})

system.time({
    results <- lapply(inputs, sqrt)
})
```

So when working on code parallelization it is important to keep in mind that for very short jobs, the parallelization overhead will diminish the parallelization benefits. If we're not sure if the parallelization of specific piece of code improves its speed, we can always check the execution time using `system.time()`.

## For Loop For Life

`lapply` family is cool, but it may be a bit constrained, we might have to constructed our code in a way that we're not used to. At the end of day, we just want to work with the for loops that exists in every popular programming languages. And it turns out there's a package called `foreach` that allows us to do just that:

`foreach` simply gives us a construct/template for writing for loops and it will take care of spreading the work inside the body of the loop across multiple processors/cores on our machine. The way it works is the following:

```{r}
# take a subset of the iris dataset, and train a bunch of
# logisitic regression on the same dataset
iris_subset <- iris[iris$Species != 'setosa',]
iris_subset$Species <- factor( iris_subset$Species, 
                               levels = c('versicolor', 'virginica') )
# number of models to train
trials <- 1000

# we first register a backend specifying the number
# of cores that we wish to use
registerDoParallel(cores = detectCores() - 1)

# parllelized foreach
system.time({
    result <- foreach(1:trials) %dopar% {
        # foreach loop behaves similar to a function, 
        # the last line represents the value that will be returned
        model <- glm(Species ~., data = iris_subset, family = 'binomial')
    }
})

# foreach also allows us to run things in a serial manner,
# to do so, we simply change the dopar to do
system.time({
    result <- foreach(1:trials) %do% {
        model <- glm(Species ~., data = iris_subset, family = 'binomial')
    }
})

```

`foreach` not only runs the code in parallel, but also merges and returns the results obtained by execution of each loop. By default foreach returns a list.

```{r}
head(result, 2)

# we can specify how we want it to
# combine the result through the .combine 
# argument, say we want it to return a vector
result <- foreach(1:trials, .combine = c) %dopar% {
    model <- glm(Species ~., data = iris_subset, family = 'binomial')
    deviance <- model$deviance
    return(deviance)
}
head(result, 2)
```

When writing `foreach`, remember to keep track of the expected variable/object that we wish to return and how variables/objects from each iteration should be combined.

There are other options for combining results, such as rbind and cbind.

### Caveat 2: Modifying Global State

When we pass in our data to run a parallel task, each processors will get a copy of the data, thus it's not going to modify the global/intial data in any way.

```{r, results='hide'}
# create a vector of 0s
# and modifying the element by
# multiplying every index by 2
x1 <- rep(0, times = 5)
for(i in 1:5) {
    x1[i] <- i * 2
}

# we try to do the same
# thing in a foreach loop
x2 <- rep(0, times = 5)
foreach(i = 1:5) %dopar% {
    x2[i] <- i * 2
}
```

```{r}
# the two different results
list(serial = x1, parallel = x2)
```

As each processor will get a copy of the vector that we've passed in, if we modify the data inplace, it will simply be modifying the copy instead of the original vector.

So the better way to do this is to simply have function that returns the information from the parallelized task.

```{r}
# the way to fix this, is simply return what
# we actually want
x3 <- foreach(i = 1:5, .combine = c) %dopar% {
    i * 2
}
x3
```

As we can see from this example, apart from parallelizing the body of a loop, `foreach` has the advantages over regular for loops when the purpose of the loop is to create a data structure such as a vector, list, or matrix. Since it removes the need to write extra code to initialize it.

### Nested Foreach

The `foreach` package also provides an option for nested for loop implementation using the `%:%` operator. The operator turns multiple `foreach` loops into a single loop, creating a single stream of tasks that can all be executed in parallel.

```{r}

# simple simulation function
sim <- function(a, b) {
    return(10 * a + b)
}

# loop over all combinations of the values that
# are stored in the two vectors, avec, bvec to 
# create the simulated data
avec <- 1:2
bvec <- 1:4

# for loop way
x <- matrix(0, length(avec), length(bvec))
for(j in 1:length(bvec)) {
    for(i in 1:length(avec)) {
        x[i, j] <- sim(avec[i], bvec[j])
    }
}
x

# foreach way, notice we don't
# put braces around the inner foreach loop
x <- foreach(b = bvec, .combine = 'cbind') %:%
    foreach(a = avec, .combine = 'c') %dopar% {
        sim(a, b)
    }
x
```

In the `foreach` example above, the inner loop returns the columns of the result matrix as vectors, which are then combined in the outer loop into a matrix.

Once we're done using the registered cluster, remember to explicity stop it. This ensures the "health" of future cluster objects.

```{r}
stopImplicitCluster()
```

In sum, there are  lots of naturally parallel problem that could benefit from writing parallelized code. Think about independent tasks such as cross validation or grid search. These computational expensive, and easily parallelizable for loops are a great place to start looking for where we can apply parallel programming.

Before we continue on to the next section, we'll do a simple programming exercise. The task is to write a loop that prints the number from 1 to 5. This is not a trick question. If the answer in your mind is something like the following:

```r
for(i in 1:5) {
    print(i)
}
```

Continue on to the next section to see a better alternative.

## Iterators

The issue with the code above is probably unnoticeable in the toy example above, since we're only looping through a small sequence. But be aware of what the `1:5` part is doing.

```{r}
1:5
```

It's generating a sequence of numbers from 1 to 5, and we're using the sequence to tell the for loop we want to loop 5 times. But what if we want to loop for 1000000 times? Do we really need to create 1000000 numbers to fill up our memory just to create the loop?

To solve this issue, we'll introduce what's called the **iterators**. An iterator is an object for moving through a container one item at a time. Although iterators are not part of the native R language, they are implemented in the `iterators` and `itertools` packages

```{r}
# note that the iterators library is
# also loaded when we load the doParallel library
library(iterators)
library(itertools)

# create a iterator of length 2,
# Here iteration is performed manually by calling nextElem()
# on the iterator, each call yields the next element in sequence
iter_count <- icount(2)
nextElem(iter_count)
nextElem(iter_count)

# when the sequence has been exhausted, we'll get a StopIteration error
# nextElem(iter_count)
```

We can also create iterators from an existing sequence.

```{r}
name <- c('Bob', 'Mary')
iter_name <- iter(name)
nextElem(iter_name)
nextElem(iter_name)
```

So another way to write the loop that prints the number from 1 to 5 using a iterator is as follows:

```{r}
# wrap the iterator with the ihasNext function,
# while the sequence has not been exhausted, print
# the next element
iter_count <- ihasNext( icount(5) )
while( hasNext(iter_count) ) {
    print( nextElem(iter_count) )
}
```

Unfortunately, the syntax above might be a bit combersome to write, but good news is it looks better with the apply family and foreach. The following syntax assumes that we're not just going to print the sequence but we're also going to do some other stuff in the loop and return the result.

```{r}
# iterators with apply
# using the multiplying each index by 2 as an example
x1 <- vapply( icount(5), function(i) {
    return(i * 2)
}, numeric(1) )

# iterators with foreach
x2 <- foreach(i = icount(5), .combine = c) %dopar% {
    return(i * 2)
}
list(apply_way = x1, foreach_way = x2)
```

There are a lot of useful functionalities in two packages, consider exploring the documentations at your free time. e.g. `ireadLines` creates an iterator to read lines from a file, which is useful when the dataset is extremely large and we wish to do some preprocessing with each line of the data instead of reading everything into memory at once.

## R Session Information

```{r}
devtools::session_info()
```

## Reference

- [Vignettes: Nesting Foreach Loops](ftp://cran.r-project.org/pub/R/web/packages/foreach/vignettes/nested.pdf)
- [Youtube: Parallel Programming in R and Python](https://www.youtube.com/watch?v=FIS_LsOzxYo)
- [Blog: Iterators in R](http://www.exegetic.biz/blog/2013/11/iterators-in-r/)
- [Blog: R apply functions](http://www.stepupanalytics.com/r-grouping-functions-sapply-vs-lapply-vs-apply-vs-tapply-vs-by-vs-aggregate)
- [Blog: A few thoughts on the existing code parallelization](http://www.vesnam.com/Rblog/existing-code-parallelization-yes-or-no/)

### 2017-01-14 R/tests_packages/test.Rmd
add R programming folder

---
title: "Testing And Packaging"
author: "Ethen Liu"
date: "`r Sys.Date()`"
output: 
  rmdformats::readthedown:
    highlight: pygments
---

<style type="text/css">
p{ /* Normal  */
   font-size: 18px;
}
body{ /* Normal  */
   font-size: 18px;
}
td {  /* Table  */
   font-size: 14px;
}
h1 { /* Header 1 */
 font-size: 32px;
}
h2 { /* Header 2 */
 font-size: 26px;
}
h3 { /* Header 3 */
 font-size: 22px;
}
code.r{ /* Code block */
  font-size: 14px;
}
pre { /* Code block */
  font-size: 14px
}
</style>

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Unit Testing

The best way to check whether a programmer's code is functioning properly is by writing unit tests. The basic workflow is to write the code, write tests, and check if the code passes the tests. Let's see how can do this with a simple example using the `testthat` package.

```{r, message=FALSE, warning=FALSE}
library(testthat)
setwd("/Users/ethen/Desktop/northwestern/russ/3_tests_packages")

compute_square_root <- function(number) {
    # an overly simple wrapper function around sqrt
    # for illustration purpose
    return( sqrt(number) )
}

test_that('Test sqrt: positive numeric', {
	expected <- 2
	actual <- compute_square_root(4)
	expect_equal(expected, actual)
})

```

Hopefully the syntax should be pretty straightforward. We use the `test_that` function to conduct a single unit test, and we first give a short description of what the test is about. Then inside this function we need to define two things: the result we expect, and the actual result that is returned by the function we wish to test. In the example above, we wish to test the function `compute_square_root`, we know that the square root of 4 is 2. Thus we would use `expect_equal` to say the result of the `compute_square_root(4)` should equal to 2.

There're a bunch of different functions apart from `expect_equal` that allow us to test for a variety of situations. For example: When we like to test whether taking the square root of a negative value will generate a warning.

```{r}
# outputs a warning
compute_square_root(-4)

# test that the function will actually generate a warning
test_that('Test sqrt: negative numeric', {
	expect_warning(compute_square_root(-4))
})
```

For more options, please refer to the documentation of `testthat`.

## Packages

Developing a package makes it easier for us to our code with colleagues or even simply prevent us from copying and pasting code that we wish to use across multiple projects. To start writing a package, the easiest way is to load up Rstudio and start a new project. We can do that by going under the File menu and choose New Directory > R package and to keep it simple the package name will simply be `myFirstPackage`. After doing so, we'll see a bunch new files and folders created inside our package folder. Note that our working directory is now set to where the package sits.

The most important folder for now is the R folder, which is the folder that will hold our .R source code files. Apart from that, Rstudio will also open up a sample `hello.R` script, which we can delete. Here we'll simply be editing it.

### Documentations

Usually, we would write comments to describe what our function does, what are its inputs and outputs. `roxygen2` is a package that turns these comments into documentation. For example:

```{r, eval = FALSE}

#' Function to compute the square root of a number
#'
#' @description This function is a wrapper around the sqrt function that
#' computes the square root of a number using a loop
#'
#' @param num the number whose square root is computed
#' 
#' @return square root of the input number
#'
#' @details The function only takes in positive values
#'
#' @references None
#'
#' @examples
#' library(myFirstPackage)
#' num <- 4
#' compute_square_root(num)
#' @export
compute_square_root <- function(number) {
    return( sqrt(number) )
}

```

The first difference with standard comments is that `roxygen2` type comments start with the `#'` symbol instead of simply the `#` symbol. The fist line is a short description of what the function does. Then, we can supply different keywords to it. Common ones include:

- `@description` for a longer description of what the function does
- `@param` that explains what each and every parameter represents
- `@return` explains the value that the function returns
- `@details` for more elaboration on the functionality of the function
- `@references` if we have any
- `@examples` for a concrete example of how to use the function
- `@export` for making the function accessible to the user. Sometimes it makes sense to not use `@export` if we want to have helper functions that are used by our other functions inside the package, and we don't want to make these functions accessible to the users.

After writing out all of these text before our function, we can then use the `roxygenise()` function from the `roxygen2` package to create the documentation files for our package.

```{r, eval=FALSE}
roxygen2::roxygenise()
```

We can now test building our package by either by clicking on the button named `Build and Reload` button which we can find inside the Build pane or by using the following keyboard shortcut: `CTRL-SHIFT-B` in Rstudio. After doing so, we can now load the library use the function inside and search for the help page of the function.

```{r, message=FALSE, warning=FALSE}
library(myFirstPackage)
compute_square_root(4)

# search for the documentation
?compute_square_root
```

### Testing Packages

Now that we know the basics of creating a package, let's apply what we know about unit testing and start testing out package. We'll simply call `devtools::use_testthat()` which will create the necessary folders to hold our tests as well as creating a `testthat.R` file with the code that gets called to run our tests. To elaborate on the necessary folders that the function created, what it's doing is creating a folder called `tests` in the root of our package and inside this tests folder create another folder, called `testthat`. The `testthat` folder will hold our unit tests. For example, inside the folder, we create a script called `test_square_root.R` and put the following code in it:

```{r}

library(testthat)
library(myFirstPackage)

test_that('Test sqrt: positive numeric', {
	expected <- 2
	actual <- compute_square_root(4)
	expect_equal(expected, actual)
})

test_that('Test sqrt: negative numeric', {
	expect_warning(compute_square_root(-4))
})

```

Now we can run our test by either going under the Build pane in Rstudio and click on Test Package or by simply clicking the keyboard shortcut `CTRL-SHIFT-T`. After doing so, we should see the following output:

![output](test_output.png)

Notice the two `.` above DONE. This means that two unit tests passed. If a unit test does not pass, then the `.` will be replaced by a number `1` and it will also output the test cases that failed and its corresponding message.

### Checking Coverage of Unit Tests

To check if our tests has covered all the functions in our package, we can run the following code:

```{r, eval=FALSE}
library(covr)
cov <- package_coverage()
shine(cov)
```

The line `shine(cov)` launches an interactive app inside our viewer pane:

![coverage](test_coverage.png)

 In our case, the coverage shows 100 percent since we did fully tested our function `compute_square_root` function inside the `hello.R` script. If we did not fully tested all our functions, then we can click in each of the files to see which line/part of the code did we not test.

### Github

Now that we created the skeleton of the package, we can share it by uploading it to Github. After doing so any R user can install your package with just executing the following command in R:

```r
devtools::install_github("username/packagename")

# e.g.
devtools::install_github("ethen8181/myFirstPackage")
```

For a more detailed introduction please refer to [R packages](http://r-pkgs.had.co.nz/), especially the section about [package's metadata](http://r-pkgs.had.co.nz/description.html) if the goal is to share the package with other users.

## R Session Information

```{r}
devtools::session_info()
```

## Reference

- [Functional programming and unit testing for data munging with R](http://www.brodrigues.co/fput/)

### 2017-01-14 R/data_table/data_table.R

@@ -1,0 +1,89 @@
+library(knitr)
+library(data.table)
+setwd("/Users/ethen/Business-Analytics/R/data_table")
+
+# -------------------------------------------------
+# 						joins
+# -------------------------------------------------
+
+employees <- fread("data/employees.csv")
+departments <- fread("data/departments.csv")
+knitr::kable(employees, caption = "Table Employees")
+knitr::kable(departments, caption = "Table Departments")
+
+
+# inner join with merge
+merge(employees, departments, by = "Department")
+
+# inner join with keys
+# set the keys of the tables to represent the by:
+setkey(employees, Department)
+setkey(departments, Department)
+# equivalent to setkeyv( departments, "Department" )
+
+# note that you only need to set the keys once
+# you can confirm if it works with
+key(employees)
+
+# perform the join,
+# nomatch 0 means no rows will be returned 
+# from not matched rows from the Right table
+employees[ departments, nomatch = 0 ]
+
+
+# left join with merge, notice that the column's ordering is different
+merge(employees, departments, by = "Department", all.x = TRUE)
+departments[employees]
+
+
+# right join with merge
+merge(employees, departments, by = "Department", all.y = TRUE)
+employees[departments]
+
+# we can also do a not matched with !
+# returning the rows in the employees table that are not in the
+# department table
+employees[!departments]
+
+
+# full outer join
+merge(employees, departments, by = "Department", all = TRUE)
+
+
+# -------------------------------------------------
+# 				 tips and tricks
+# -------------------------------------------------
+
+# Get the sum of y and z and 
+# the number of rows in each group while grouped by x
+DT <- data.table( x = sample(1:2, 7, replace = TRUE), 
+				  y = sample(1:13, 7), 
+				  z = sample(1:14, 7) ) 
+DT[ , c( lapply(.SD, sum), .N ), by = x ]
+
+# set the keys and
+# select the matching rows without using `==`.
+DT <- data.table( A = letters[ c(2, 1, 2, 3, 1, 2, 3) ],
+				  B = c(5, 4, 1, 9, 8, 8, 6), C = 6:12 )
+setkey(DT, A, B)
+
+# matches rows where A == "b"
+DT["b"]
+
+# matches rows where A == "b" or A == "c"
+DT[ c("b", "c"), ]
+
+# match both keys
+DT[ .("b", 5), ]
+
[…]

### 2017-01-14 R/efficient_looping/efficient_looping.R

@@ -1,0 +1,187 @@
+# efficient looping
+
+# -----------------------------------------------------------------------------------
+# Parallel Programming
+library(doParallel)
+
+# we specify the number of cores/workers we want to use
+n_cores <- detectCores() - 1
+n_cores
+
+# generate a toy function that
+# simply generate the summary of a bunch of random numbers
+summarize <- function(i) {
+    summary( rnorm(100000) )
+}
+
+# the time difference between using n_cores and not using it
+inputs <- 1:20
+system.time({
+    results <- mclapply(inputs, summarize, mc.cores = n_cores)
+})
+
+system.time({
+    results <- lapply(inputs, summarize)
+})
+
+
+# overhead;
+# when parallelizing trivial tasks
+inputs <- 1:10000
+system.time({
+    results <- mclapply(inputs, sqrt, mc.cores = n_cores)
+})
+
+system.time({
+    results <- lapply(inputs, sqrt)
+})
+
+
+# foreach;
+
+# take a subset of the iris dataset, and train a bunch of
+# logisitic regression on the same dataset
+iris_subset <- iris[iris$Species != 'setosa',]
+iris_subset$Species <- factor( iris_subset$Species, 
+                               levels = c('versicolor', 'virginica') )
+# number of models to train
+trials <- 1000
+
+# we first register a backend specifying the number
+# of cores that we wish to use
+registerDoParallel(cores = detectCores() - 1)
+
+# parllelized foreach
+system.time({
+    result <- foreach(1:trials) %dopar% {
+        model <- glm(Species ~., data = iris_subset, family = 'binomial')
+    }
+})
+
+# foreach also allows us to run things in a serial manner,
+# to do so, we simply change the dopar to do
+system.time({
+    result <- foreach(1:trials) %do% {
+        model <- glm(Species ~., data = iris_subset, family = 'binomial')
+    }
+})
+
+# by default foreach returns a list
+head(result, 2)
+
+# we can specify how we want it to
+# combine the result through the .combine 
+# argument, say we want it to return a vector
+result <- foreach(1:trials, .combine = c) %dopar% {
+    model <- glm(Species ~., data = iris_subset, family = 'binomial')
+    deviance <- model$deviance
+    return(deviance)
+}
+
+
+# modifying global state;
+
+# create a vector of 0s
[…]

### 2017-01-17 bayesian_statistics/MCMC/MCMC.Rmd
multiple testing

@@
 
-f <- function(x){
-	( sin(.2 * x^3) + sin(x^2) / 5 + 2.5 + sin(10 * x) / 2 ) * dnorm( x, mean = 1 )
+f <- function(x) {
+    ( sin(0.2 * x ^ 3) + sin(x ^ 2) / 5 + 2.5 + sin(10 * x) / 2 ) * dnorm(x, mean = 1)
 }
 
-# boundary for the x axis
+# # boundary for the x axis
 x_max <- 6
@@
 
-x <- seq( x_min, x_max, length.out = 500 )
+# for simply plotting the function
+ggplot( data.frame( x = c(x_max, x_min) ), aes(x) ) + 
+stat_function(fun = f)
+
+x <- seq(x_min, x_max, length.out = 500)
 y <- f(x)
-data <- data.frame( x = x, y = y )
-
-ggplot( data, aes( x, y ) ) + geom_line() + 
-geom_ribbon( aes( ymin = 0, ymax = y ), alpha = 0.2 )
+data <- data.frame(x = x, y = y)
+ggplot( data, aes(x, y) ) + geom_line() + 
+geom_ribbon( aes(ymin = 0, ymax = y), alpha = 0.2 )
 
@@
 ceiling <- 1.5
-xs <- runif( N, min = x_min, max = x_max )
-ys <- runif( N, min = 0, max = ceiling )
+xs <- runif(N, min = x_min, max = x_max)
+ys <- runif(N, min = 0, max = ceiling)
 
@@
 
-ggplot( data.frame( x = accepted_samples ) ) + 
-geom_histogram( aes( x = x, y = ..density.. ) ) + 
-geom_line( data = data, aes( x, y ) )
+ggplot( data.frame(x = accepted_samples) ) + 
+geom_histogram( aes(x = x, y = ..density..) ) + 
+geom_line( data = data, aes(x, y) )
 
-# study the samples however you want
+# we can then study the samples however we want
 # mean(accepted_samples)
-# quantile(accepted_samples, probs = seq(0, 1, length = 4 ) )
+# quantile(accepted_samples, probs = seq(0, 1, length = 4) )
 
@@
 # plot the accepted and unaccepted point
-ggplot( df, aes( xs, ys ) ) + 
-geom_point( aes( color = accept ), size = 0.5 ) + 
-geom_line( data = data, aes( x, y ) )
+ggplot( df, aes(xs, ys) ) + 
+geom_point( aes(color = accept), size = 0.5 ) + 
+geom_line( data = data, aes(x, y) )
 
@@
 
-## MCMC - Metropolis–Hastings Algorithm
+## MCMC - Markov chain Monte Carlo
 
@@
 
-lik <- function( x, y ){
+lik <- function(x, y) {
 	dnorm(x - 3) * dnorm(y - x + 2)
@@
 
-grid_values <- seq( x_min, x_max, length = 500 )
-grid <- expand.grid( x = grid_values, y = grid_values )
-z <- lik( grid$x, grid$y )
+grid_values <- seq(x_min, x_max, length = 500)
+grid <- expand.grid(x = grid_values, y = grid_values)
+z <- lik(grid$x, grid$y)
 
 # prior probability plot
-plot <- ggplot( data = grid, aes( x = x, y = y ) ) + 
-		geom_raster( aes( fill = z ) ) + 
+plot <- ggplot( data = grid, aes(x = x, y = y) ) + 
+		geom_raster( aes(fill = z) ) + 
 		scale_fill_gradient2() + coord_equal() +
@@
 
-# Metropolis-Hastings algorithm (will elaborate on this later)
-MCMC <- function( iter, burnin = 0, thin = 1 )
-{
+MCMC <- function(iter, burnin = 0, thin = 1) {
 	# the list of the first argument for dimnames is to specify the row names,
@@
 	samples <- matrix( NA, nrow = iter, ncol = 2, 
-					   dimnames = list( NULL, c( "x", "y" ) ) )
-	samples[ 1, ] <- c( 0, 0 )
+					   dimnames = list( NULL, c("x", "y") ) )
+	samples[1, ] <- c(0, 0)
 
-	for(i in 2:iter){
+	for(i in 2:iter) {
 
-		sample <- samples[ i - 1, ]
-		for(j in 1:thin){
+		sample <- samples[i - 1, ]	
+		for(j in 1:thin) {
 
-			# propose a new sample point		
-			proposal <- sample + rnorm( n = 2, mean = 0, sd = 1 )
+			# propose a new sample point (will elaborate on this later)				
+			proposal <- sample + rnorm(n = 2, mean = 0, sd = 1)
 
@@
 			# note that 1 is evaluated as TRUE
-			if( rbinom( 1, size = 1, prob = min( ratios, 1 ) ) )
+			if( rbinom( 1, size = 1, prob = min(ratios, 1) ) )
 				sample <- proposal
 		}
-		samples[ i, ] <- sample
+		samples[i, ] <- sample
 	}
-	return( data.frame( samples[ (burnin + 1):iter, ] )	)
+	return( data.frame(samples[ (burnin + 1):iter, ]) )
 }
 
-TracePlot <- function( plot, samples ){
+plot_trace <- function(plot, samples) {
 	plot + 
-	geom_path( data = samples, aes( x, y ), color = "orange" ) + 
-	geom_point( data = samples, aes( x, y ), size = 1 )
+	geom_path( data = samples, aes(x, y), color = "orange" ) + 
+	geom_point( data = samples, aes(x, y), size = 1 )
 }
@@
 samples <- MCMC(iter = 50)
-TracePlot( plot, samples )
+plot_trace(plot, samples)
 
 # plot with burnin, discard the bad sampling points in the beginning
-samples <- MCMC( iter = 250, burnin = 125 )
-TracePlot( plot, samples )
+samples <- MCMC(iter = 250, burnin = 125)
+plot_trace(plot, samples)
 
-# use thin to avoid correlation and 
+# use thinning to avoid correlation and
 # explore more of the sample space
-samples <- MCMC( iter = 250, burnin = 125, thin = 2 )
-TracePlot( plot, samples )
-
+samples <- MCMC(iter = 250, burnin = 125, thin = 2)
+plot_trace(plot, samples)
 ```
@@
 
-## Reference
-
-1. [The First Stop in Monte Carlo (Methods): Rejection Sampling](http://thexbar.me/2014/11/07/reject-sample/)
-2. [mcmc tutorial](https://github.com/davharris/mcmc-tutorial)
-3.
-[Probabilistic Programming and Bayesian Methods for Hackers Chapter 3](http://nbviewer.jupyter.org/github/CamDavidsonPilon/Probabilistic-Programming-and-Bayesian-Methods-for-Hackers/blob/master/Chapter3_MCMC/Chapter3.ipynb)
-
 ## R Session Information
@@
 ```{r}
+sessionInfo()
+```
 
-sessionInfo()
+## Reference
 
-```
+- [Blog: The First Stop in Monte Carlo (Methods): Rejection Sampling](http://thexbar.me/2014/11/07/reject-sample/)
+- [Github: mcmc tutorial](https://github.com/davharris/mcmc-tutorial)
+- [Notebook: Probabilistic Programming and Bayesian Methods for Hackers Chapter 3](http://nbviewer.jupyter.org/github/CamDavidsonPilon/Probabilistic-Programming-and-Bayesian-Methods-for-Hackers/blob/master/Chapter3_MCMC/Chapter3.ipynb)

### 2017-01-17 bayesian_statistics/MCMC/MCMC.R

@@ -2,8 +2,8 @@
 library(ggplot2)
 library(ggthemes)
 
-f <- function(x){
-	( sin(.2 * x^3) + sin(x^2) / 5 + 2.5 + sin(10 * x) / 2 ) * dnorm( x, mean = 1 )
+f <- function(x) {
+    ( sin(0.2 * x ^ 3) + sin(x ^ 2) / 5 + 2.5 + sin(10 * x) / 2 ) * dnorm(x, mean = 1)
 }
 
 # # boundary for the x axis
@@ -11,15 +11,14 @@ x_max <- 6
 x_min <- -6
 
 # for simply plotting the function
-ggplot( data.frame( x = c( x_max, x_min  ) ), aes(x) ) + 
-stat_function( fun = f ) + 
+ggplot( data.frame( x = c(x_max, x_min) ), aes(x) ) + 
+stat_function(fun = f)
 
-x <- seq( x_min, x_max, length.out = 500 )
+x <- seq(x_min, x_max, length.out = 500)
 y <- f(x)
-data <- data.frame( x = x, y = y )
-
-ggplot( data, aes( x, y ) ) + geom_line() + 
-geom_ribbon( aes( ymin = 0, ymax = y ), alpha = 0.2 )
+data <- data.frame(x = x, y = y)
+ggplot( data, aes(x, y) ) + geom_line() + 
+geom_ribbon( aes(ymin = 0, ymax = y), alpha = 0.2 )
 
 
 # -------------------------------------------------
@@ -29,64 +28,62 @@ geom_ribbon( aes( ymin = 0, ymax = y ), alpha = 0.2 )
 # throw N darts (uniformly distributed x-y pairs) at the plot above
 N <- 10000
 ceiling <- 1.5
-xs <- runif( N, min = x_min, max = x_max )
-ys <- runif( N, min = 0, max = ceiling )
+xs <- runif(N, min = x_min, max = x_max)
+ys <- runif(N, min = 0, max = ceiling)
 
 # accept any samples that are below the function curve
 accepted_samples <- xs[ ys <= f(xs) ]
 
-ggplot( data.frame( x = accepted_samples ) ) + 
-geom_histogram( aes( x = x, y = ..density.. ) ) + 
-geom_line( data = data, aes( x, y ) )
+ggplot( data.frame(x = accepted_samples) ) + 
+geom_histogram( aes(x = x, y = ..density..) ) + 
+geom_line( data = data, aes(x, y) )
 
-# study the samples however you want
[…]

### 2017-01-17 articles/data_hygiene.Rmd
data science advice

--- 2016-09-16 articles/data_hygiene/data_hygiene.Rmd
@@
 title: "Data Hygiene Gotchas"
-author: "Ming-Yu Liu"
+author: "Ethen Liu"
 date: "`r Sys.Date()`"
@@
 
-- [Common data preparation mistakes and how to avoid them](http://www.analyticsvidhya.com/blog/2013/08/common-data-preparation-mistakes-avoid-them/)
 - [Webinar: Enhance Predictive Modeling with Better Data Preparation](http://www.datasciencecentral.com/video/enhance-predictive-modeling-with-better-data-preparation)
-- [Non-Mathematical Feature Engineering techniques for Data Science](https://codesachin.wordpress.com/2016/06/25/non-mathematical-feature-engineering-techniques-for-data-science/)
-- [Simple Methods to deal with Categorical Variables in Predictive Modeling](http://www.analyticsvidhya.com/blog/2015/11/easy-methods-deal-categorical-variables-predictive-modeling/)
+- [Blog: Common data preparation mistakes and how to avoid them](http://www.analyticsvidhya.com/blog/2013/08/common-data-preparation-mistakes-avoid-them/)
+- [Blog: Non-Mathematical Feature Engineering techniques for Data Science](https://codesachin.wordpress.com/2016/06/25/non-mathematical-feature-engineering-techniques-for-data-science/)
+- [Blog: Simple Methods to deal with Categorical Variables in Predictive Modeling](http://www.analyticsvidhya.com/blog/2015/11/easy-methods-deal-categorical-variables-predictive-modeling/)

### 2017-01-17 articles/data_science_advice.Rmd
data science advice

---
title: "Data Science Advice"
author: "Ethen Liu"
date: "`r Sys.Date()`"
output:
  rmdformats::readthedown:
    highlight: pygments
---

<style type="text/css">
p { /* Normal  */
   font-size: 18px;
}
body { /* Normal  */
   font-size: 18px;
}
td {  /* Table  */
   font-size: 14px;
}
h1 { /* Header 1 */
 font-size: 32px;
}
h2 { /* Header 2 */
 font-size: 26px;
}
h3 { /* Header 3 */
 font-size: 22px;
}
code.r { /* Code block */
  font-size: 14px;
}
pre { /* Code block */
  font-size: 14px
}
</style>

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# General Advice

## Mentality

**Help the business first:** Stacking outputs as features to go for the best models is not exactly what we are paid for. Instead we should aim for generating a actionable solution or get get a minimum viable product first. This does not mean we shouldn’t work on improving the model, it's simply saying premature optimization is the root of all evil. When starting a project aim for the low hanging fruit first, there will come a time to improve our models or add complexity as needed.

**Expect and accept ignorance and mistakes:** There are many limits to what we can learn from data and mistakes can still occur. Admitting our mistake is a strength but it is not usually immediately rewarded. But proactively owning up to our mistakes will translate into credibility and will ultimately earn us respect with colleagues and leaders who are data-wise.

**Don’t fall for the hype:** Deep learning..., Distributed systems.... Most likely we will not need it and it will be a distraction. Use familiar tools and if it ain’t broken don’t fix it. Consider that using any project younger than a few year may result in inavoidable pains (breaking APIs, bugs). Allow ourselves only a couple of new pieces of technology per work project. This is not saying that we can't work or keep up with on these new and cool stuff as our side-project. After all we should constantly be open to learning new knowledge.

**Be skeptical and check for reproducibility:** As we work with the data, we must be a skeptic of the insights that we are gaining. When we have an interesting phenomenon we should ask both "What other data could I gather to validate how awesome this is?"" and "What could I find that would invalidate this phenomenon?". If we are building models of the data, you want those models to be stable across small perturbations in the underlying data. Using different time ranges or random sub-samples of our data will tell us how reliable/reproducible this model is.

**Be aware of Simpson's Paradox:** When performing analysis, we often slice the data and separate it into subgroups and look at the values of our metrics in those subgroups separately. For example, in web traffic analysis, we commonly slice along dimensions like mobile vs. desktop, browser, locale, etc. If the underlying phenomenon is likely to work differently across subgroups, we must slice the data to see if it is. Even if we do not expect a slice to matter, looking at a few slices for internal consistency gives us greater confidence that we are measuring the right thing. In some cases, a particular slice may have bad data. But the most important thing to be aware of when slicing the data to compare between groups is Simpson's Paradox. Simpson’s paradox says that when we combine all of the groups together, and look at the data in aggregate form, the correlation that we noticed before may reverse itself. This is most often due to lurking variables that have not been considered. Generally, if the relative amount of data in a slice is the same across our two groups, we can safely make a comparison. A concrete example of Simpson's Paradox can be found at the following link. [Blog: What Is Simpson's Paradox?](http://statistics.about.com/od/HelpandTutorials/a/What-Is-Simpsons-Paradox.htm)

## EDA Process

Exploratory data analysis can be broken down into 4 interrelated stages:

- **Make hypotheses and look for evidence:** Good data analysis will have a story to tell. To make sure it’s the right story, we need to know what we should see in the data if that hypothesis is true, then look for evidence that it’s wrong. One way of doing this is to ask ourselves, “What experiments would we run that would validate/invalidate the story we're am telling?” Even if we don’t/can’t do these experiments right now, it may give us ideas for future analyses.
- **Validation or Initial Data Analysis:** Was the data collected correctly and helps answer my question? This often goes under the name of “sanity checking”. For example, if manual testing of a feature was done, can I look at the logs of that manual testing? e.g. For a feature launched on mobile devices, do my logs claim the feature exists on desktops? If it’s a features of a product, try it out yourself.
- **Description:** What’s the objective interpretation of this data? For example, “The time page load to click (given there was a click) is larger by 1%”
- **Evaluation:** Given the description, does the data tell us that something good is happening for the user, for the company? For example, “Users find results faster” or “The quality of the clicks is higher.”

By separating these phases, we can more easily reach agreement with others. Description should be things that everyone can agree on from the data. Evaluation is likely to have much more debate because we are imbuing meaning and value to the data.

## Documentations

Having a "paper trail" is very important.

**Acknowledge and count filtering:** Almost every large data analysis starts by filtering the data in various stages. Maybe we want to consider only US users. Whatever the case may be, we must acknowledge and clearly specify what filtering you are doing.

**Ratios should have clear numerator and denominators:** Many interesting metrics are ratios of underlying measures. Unfortunately, there is often ambiguity of what the ratio is. When presenting or communicating ratios, be clear about it to prevent misinterpretation.

## Social & Communications

Make sure we share what we’ve learned and learned from others so we can all be better data scientists.

**Talk & Learn from colleagues:** Learn from software engineers. Chances are they know more about best practices, tooling and devops than we do as a data scientist. Learn from non-technical peers. They know how we can help them from a business perspective, may understand the data much better and will recognize data quality issues faster than we will. Our feature engineering work will benefit a lot from their insight. Talk to our peers. Learn about the history of projects started by our colleagues. Ask about issues they encountered along the way. Try to take different paths or revisit them if e.g. we now have much more data. When working on the project, it is also nice to look for peer reviews. Peer reviewer can provide feedbacks from a different standpoint, suggest better approaches and better sanity-checks than the consumers of our project can.

**Educate our consumers:** We will often be presenting our analysis and results to people who are not data experts. Part of our job is to educate them on how to interpret and draw conclusions from our analysis data. This is especially important when the analysis has a high risk of being misinterpreted. We are responsible for providing the context and a full picture of the analysis and not just the number a consumer asked for.

# Template For Problem Solving

It's all about reviewing the completed work, the current work in progress, planned work, specific annual metrics, performance gap.

### Review Current Status

- List key metrics you're tracking, where they're at, and compare with last few weeks (to measure how are thing trending).
- What did you learn last week or what was accomplished? And is everything on track? Are you going to make some adjustments (including solutions and evaluation metrics) based on your findings?

### List Out Top Problems/Experiments

- List and prioritize the top (new) problems / experiments.
- What do we want to solve / learn and why? When asking the why, it can be thought of what steps will be taken given the result?
- List out who's feeling the pain? (figure out who are your priorities, this may be tied to the why)

### For All Problems, List Corresponding Hypothesized Solution

- Written in the form "[Specfic action] will create [expected result]."
- List out why do you believe each solution will help solve the problem?
- List metrics (quantitative) or proof (qualitative) you'll use to measure whether or not the solutions are doing what you expected (solving the problem). Meaning how will we conclude that the experiment succeeded? If it's a metric, set goals for it!
- How long will it take for you to run the experiment / get the problem solved.
- (optional) Include what measures will indicate the experiment isn't safe to continue.

From then on, it's about leveraging your current data and resources, creating features and running tests (testing out the hypothesized solution) and let the resulting data frame the conversation.

# Presenting

### The Pyramid Approach For Presentations

The core of the Pyramid Approach is:

- Start with the answer first.
- Group and summarize your supporting arguments. 
- Then logically order your supporting ideas and create a storyboard.

Meaning that when an executive asked a question — "What should we do?" — you were to start your response with, “You should do X,” very crisply and directly. Only then, after you have answered the question, should you present your supporting reasons. Why? What's the benefit of this approach:

- First, you maximize your time with your audience. Executives are busy people, they are perpetually short on time and get impatient when they feel like someone isn’t getting to the point. They want to focus on the big picture — in this case the "answer" — and may not want to get bogged down by details. In some cases, the executive may already mentally be at the conclusion you want them to reach, in which case she will accept your recommendation and move on (without you having to go into the detailed supporting arguments).
- Second, Making a statement to your audience that tells them something (the answer) they don’t know, will automatically raise a question in their minds. Why? How? Is this true? Etc. The listener will more likely want to hear the supporting reasons to the statement.

The structure of the flow will be:

- **Situation:** Starting with an illustration of the situation will establish a certain time and place for the listener. Preferably a time and place the listener can relate to. This should be a neutral description with facts that your listeners will agree upon. e.g. The company's market is mature and doesn't offer growth opportunity anymore.
- **Complication:** The complication will illustrate a problem, thus creating a relevant issue. A certain sense of urgency or compelling reason to listen or even act. This is the desired change of the current solution e.g. The CEO would like to increase his profit by reducing his cost by 10%.
- **Question:** The question that is results from the complication. The will also be the start of a question-answer dialogue. e.g. How can the company reduce its cost by 10%.
- **Answer:** The question will be the main lead for your story and the answer your main topic.
- **Supporting Arguments:** Underneath the single thought/answer, you are to group and summarize the next level of supporting ideas and arguments. You want to ensure that the ideas you bring together under each group actually belong together (at the same level of importance) and that all the possibilities has been covered.
- **Create the StoryBoard:** Follow some logical structure. There are a few different ways of logically ordering the story: 1) Degree order: present supporting ideas in rank order of importance, most to least important. 2) Time order: if there is a sequence of events that form a cause-effect relationship, you should present the ideas in time order. e.g. Back to the how can the company reduce its cost by 10% question. The story is: We should outsource. Because competitor X is profitable, it has the same revenue as us but lower cost. Why? Because 50% of his core functions has been outsourced.

### Data Stories

By telling a data story you're essentially walking the audience through a series of facts (can be shown through data visualization) that leads them to some sort a conclusion.

These are the 7 most common data stories:

1. **Change over time.** How are things before, how is it now and will it be in the future (trend).
2. **Drill down.** Aggregrated value doeesn't tell the whole story e.g. Split down the measure geographically. You can start with a big picture and drill down to the nitty gritty details that see how the two differ. 
3. **Zoom out.** Simply the opposite of drilling down.
4. **Contrast.** Comparying different or opposite things together. e.g. Comparing the maximum value of a measure and see where they appear geographically and then showing where the min value occured.
5. **Intersection.** When does one value surpass the other, or vice versa. e.g. revenue of different product category.
6. **Factors.** The data can tell you something that is true (a measure is decreasing over time), but often times that leads to another question, which is why. e.g. if we see some measures decreasing, we will surely want to figure out why. This is when breaking a metric that we're looking at down into different building blocks can help (factors that contributes to the metric). Note that, a lot of times the current data at hand won't be able to tell you the answer.
7. **Outlier.** Boxplot can be useful here. Outliers can be very important message that tells you whether the data has quality issues.

## Reference

- [Webinar: Exploring 7 Kinds of Data Stories](http://www.datasciencecentral.com/video/exploring-7-kinds-of-data-stories)
- [SlideShare: The Pyramid Principle](http://www.slideshare.net/BySlidebooks/pyramid-principle)
- [Blog: The Pyramid Principle](https://medium.com/lessons-from-mckinsey/the-pyramid-principle-f0885dd3c5c7#.j4nz2qagr)
- [Blog: What Is Simpson's Paradox?](http://statistics.about.com/od/HelpandTutorials/a/What-Is-Simpsons-Paradox.htm)
- [Blog: 10 pieces of advice to beginner data scientists](https://shapescience.xyz/blog/10-pieces-of-advice-to-beginner-data-scientists/)
- [Blog: Practical advice for analysis of large, complex data sets](http://www.unofficialgoogledatascience.com/2016/10/practical-advice-for-analysis-of-large.html)

### 2017-01-19 articles/articles.Rmd
DS template

@@
 [[Blog](https://source.opennews.org/en-US/articles/what-i-learned-recreating-one-chart-using-24-tools/?utm_campaign=Data%2BElixir&utm_medium=web&utm_source=Data_Elixir_110)]
-- 2016.12.24 | Scope the Solution before Solving the Machine Learning. [[Blog](https://making.dia.co/scope-the-solution-before-solving-the-machine-learning-7e5ddb622733#.iwbgi17n1)]
-    1. what is the actual end goal
-    2. what type of outputs do I need
-    3. what type of models produces this output
-    4. what type of inputs would help me
-    5. are these inputs valid
 - The Most Practical Big Data Use Cases Of 2016.

### 2017-01-19 articles/data_science_advice.Rmd
remove git messages

@@
 One way of doing this is to ask ourselves, “What experiments would we run that would validate/invalidate the story we're am telling?” Even if we don’t/can’t do these experiments right now, it may give us ideas for future analyses.
+- **Validation of initial data analysis:** Was the data collected correctly and helps answer my question?
+This often goes under the name of “sanity checking”.
+For example, if manual testing of a feature was done, can I look at the logs of that manual testing?
+e.g.
+For a feature launched on mobile devices, do my logs claim the feature exists on desktops?
+If it’s a features of a product, try it out yourself.
 - **Validation or Initial Data Analysis:** Was the data collected correctly and helps answer my question?
@@
 - (optional) Include what measures will indicate the experiment isn't safe to continue.
+
+From then on, it's about leveraging your current data and resources, creating features and running tests (testing out the hypothesized solution) and let the resulting data frame the conversation.
+
+# Template For Data Science Project
+
+Please describe your general “rule of thumb” on how to engage a new data research problem. Is there any framework that you general follow? How do you breakdown the problem into actionable tasks?
+
+1.
+**Defining the end goal from a business perspective:** A good way of doing so is to ask what types of output will we need and what actions will we take after obtaining the result.
+For this step, it is also important to identify the current solution for the problem and why is it not good enough.
+Only by confirming the baseline can we evaluate the performance of latter solutions.
+If there’s a certain level of resource budget (e.g.
+time) then it should also be noted as well
+2.
+**Data Collecting and Preprocessing:** This step encompasses data extracting, cleaning and exploratory analysis.
+Feature engineering is an important aspect of the work and checking the data quality at hand is an important task to prevent bias for later study
+3.
+**Modeling and Evaluation:** There’s actually not a really clear gap between the modeling process and the former step of data preprocessing.
+As trying out different algorithms on the data will give you ideas of whether the data at hand is good enough to address the issue at hand and whether the models still require further tuning.
+The evaluation part should be quite simple, as the criteria should already be defined in the first stage
+4.
+**Presentation, Documentation and Deployment:** Once we have a model that meets our success criteria.
+Different types of audience will require different ways of presenting the information.
+But basically, the workflow and the interpretation should be documented in a way that’s clear to both a data scientist that’s also on the team but did not participate in the project and to whoever’s going to rely on it to make business decisions
+5. **Future Works:** No models are perfect. If it is a problem that’s worth the time, future works should be discussed to improve the result
 
@@
 - [Blog: Practical advice for analysis of large, complex data sets](http://www.unofficialgoogledatascience.com/2016/10/practical-advice-for-analysis-of-large.html)
+- [Blog: Scope the Solution before Solving the Machine Learning](https://making.dia.com/scope-the-solution-before-solving-the-machine-learning-7e5ddb622733#.pcr044mri)

### 2017-02-19 articles/articles.Rmd
correlation

@@
 
+- 2017.02.14 | Why machine learning is critical to multi-touch attribution.
+[[MarTech](https://martechtoday.com/machine-learning-critical-multi-touch-attribution-194976?utm_source=Social&utm_medium=Share&utm_term=MTT+article+Feb&utm_content=Machine+learning+and+attribution&utm_campaign=Awareness)]
+- 2017.02.05 | Unlearning descriptive statistics. [[Blog](http://debrouwere.org/2017/02/01/unlearning-descriptive-statistics/?utm_campaign=Revue%20newsletter&utm_medium=Newsletter&utm_source=revue)]
+    - Caveat: multimodal distributions have more than one central or typical value, and they are trickier to describe.
+If you look at how tall the average adult human is, you will find a bump around 165 cm and another around 175 cm.
+These local maxima (modes) are useful statistics, but often it pays to prod a little deeper and see why there's more than one peak in the first place.
+In the case of human height, the answer is obvious: a typical adult woman is roughly 165 cm tall, a typical adult man roughly 175 cm.
+Once you split the data by gender, the bimodal distribution disappears.
 - 2016.12.29 | Data Science Newsletter list.

### 2017-03-24 articles/articles.Rmd
update license

@@
 
+- Why Companies like Lyft, Uber, Postmates, Instacart etc Will Never Be Profitable. [Linkedin](https://www.linkedin.com/pulse/why-companies-like-lyft-uber-postmates-instacart-etc-never-fabode)
 - The New 'Must Haves' for Mobile Apps to Succeed Today. [[First Round Review](http://firstround.com/review/the-new-must-haves-for-mobile-apps-to-succeed-today/)]

### 2017-04-23 articles/articles.Rmd
fix typo in multiple testing

@@
 
+- 2017.04.02 | The Startup Founder’s Guide to Analytics. [Medium](https://thinkgrowth.org/the-startup-founders-guide-to-analytics-1d2176f20ac1)
 - 2017.02.14 | Why machine learning is critical to multi-touch attribution.
@@
 
+- How Data Science Is Transforming Health Care. [O'Reilly](https://www.oreilly.com/ideas/how-data-science-is-transforming-health-care)
+    - Just like targeted ads, we should have targeted health care
 - Why Companies like Lyft, Uber, Postmates, Instacart etc Will Never Be Profitable. [Linkedin](https://www.linkedin.com/pulse/why-companies-like-lyft-uber-postmates-instacart-etc-never-fabode)

### 2017-06-08 frequentist_statistics/frequentist_statistics.Rmd
minor typo

@@
 e.g.
-If we wish to compare pre- (beginning of semester) and post-test (end of semester) scores of students, then this will require a paired t-test.
+If we wish to compare pre- (beginning of semester) and post-test (end of semester) scores of students, then this will require a **paired t-test**.
+The following video explains it nicely.
+[Youtube: The Paired-Difference t Procedure](https://www.youtube.com/watch?v=TMRBGFIj0RM)
 
@@
 - [Blog: The Theorem Every Data Scientist Should Know](http://www.jeannicholashould.com/the-theorem-every-data-scientist-should-know-2.html)
+- [Youtube: The Paired-Difference t Procedure](https://www.youtube.com/watch?v=TMRBGFIj0RM)
 - [Notes: Power and sample size calculation for t test in R](http://users.stat.umn.edu/~corbett/classes/3022/Power_and%20_sample_size_for_t_in_R.txt)
